---
title: "WT_KO_Blebbistatin_brightfield"
output: html_notebook
---

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

```{r message=FALSE, warning=FALSE, include=FALSE}
library(pracma)
library(peakPick)
library(quantmod)
library(ggpmisc)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggfortify)
library(forecast)
library(tsibble)
library(broom)
library(tidyquant)
library(tsbox)
library("pheatmap")
library("RColorBrewer")
library(ggpmisc)
library(ggfortify)
library(viridis)
library(hrbrthemes)
library(data.table)
library(ggpubr)
library(easyGgplot2)
library(oce)
```

## load existing/saved environment (.RData)
```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(Path_1)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
load(file=saved_path)

```

## Read .csv files of data 170522|170724|170812|170821
```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("170522|170724|170812|170821", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("class_2", subfolder_list) ]

freq <- 0.2 
pixel <- 0.1609 # set pixel size
csv_list<- list()
df_list<- list()

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list[[n]] <- list.files(subfolder_list[n],pattern = "(.*)csv$", 
                     recursive = TRUE,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list[[n]], read.csv, header = TRUE))
  df <-df[, grep("Tip|Length",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  df <-  pracma::hampel(df, k=7)$y
  
    colnames(df)<- paste("Length", seq( from = 1, to = ncol(df) ),  sep = ".")
  
  df_list[[n]] <-df
  
}

names(df_list) <- subfolder_list  
```

# convert time resolution to frame/2-min
```{r}

for (n in 1:length(df_list))  
{
df <- df_list[[n]] %>% as.data.frame()
df$time <- seq(0,(nrow(df)-1)*5, by = 5)

newTime <- seq (0, (nrow(df)-1)*5) %>% as.data.frame()
names(newTime) <- "time"

df <- left_join(x=newTime,y=df,by=c("time")) %>% zoo::na.approx() %>% as.data.table()
df <- df [ time %in% seq(0, nrow(df), by = 2)]
df <- df[,time:=NULL] 

df_list[[n]] <- df
}

```

## Read .csv files of data 170515
```{r}

subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("170515", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("class_2", subfolder_list) ]

freq <- 0.2 
pixel <- 0.1063 # set pixel size
csv_list_170515<- list()
df_list_170515<- list()

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list_170515[[n]] <- list.files(subfolder_list[n],pattern = "(.*)csv$", 
                     recursive = TRUE,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list_170515[[n]], read.csv, header = TRUE))
   df <-df[, grep("Tip|Length",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  df <-  pracma::hampel(df, k=7)$y
  
    colnames(df)<- paste("Length", seq( from = 1, to = ncol(df) ),  sep = ".")
  
  df_list_170515[[n]] <-df
  
}

names(df_list_170515) <- subfolder_list  
```

# convert time resolution to frame/2-min
```{r}

for (n in 1:length(df_list_170515))  
{
df <- df_list_170515[[n]] %>% as.data.frame()
df$time <- seq(0,(nrow(df)-1)*5, by = 5)

newTime <- seq (0, (nrow(df)-1)*5) %>% as.data.frame()
names(newTime) <- "time"

df <- left_join(x=newTime,y=df,by=c("time")) %>% zoo::na.approx() %>% as.data.table()
df <- df [ time %in% seq(0, nrow(df), by = 2)]
df <- df[,time:=NULL] 

df_list_170515[[n]] <- df
}

```

## Read .csv files of data 151209
```{r}

subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("151209", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("class_2", subfolder_list) ]

freq <- 0.5
pixel <- 0.1063 # set pixel size
csv_list_151209<- list()
df_list_151209<- list()

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list_151209[[n]] <- list.files(subfolder_list[n],pattern = "(.*)csv$", 
                     recursive = TRUE,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list_151209[[n]], read.csv, header = TRUE))
   df <-df[, grep("Tip|Length",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  df <-  pracma::hampel(df, k=7)$y
  
    colnames(df)<- paste("Length", seq( from = 1, to = ncol(df) ),  sep = ".")
  
  df_list_151209[[n]] <-df
  
}

names(df_list_151209) <- subfolder_list  

```

# (optional) convert time resolution to frame/5-min (downsampling to match tempro-res to other data)
```{r}

for (n in 1:length(df_list_151209))  
{
df <- df_list_151209[[n]] %>% as.data.frame()
df$time <- seq(0,(nrow(df)-1)*2, by = 2)

newTime <- seq (0, (nrow(df)-1)*2) %>% as.data.frame()
names(newTime) <- "time"

df <- left_join(x=newTime,y=df,by=c("time")) %>% zoo::na.approx() %>% as.data.table()
df <- df [ time %in% seq(0, nrow(df), by = 5)]
df <- df[,time:=NULL] 

df_list_151209[[n]] <- df
}

```

```{r}
df_list <- c(df_list, df_list_170515, df_list_151209)
```

```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
save.image(file=saved_path)
```

## Import time-series neurite length data
```{r}

mts_stationary_list <- list()

freq = 0.2 # 0.5 or 0.2

for (n in 1:length(df_list))  {
 
  length_mat<-df_list[[n]] %>% as.data.table()
  length_mat$sum_length <- length_mat %>% rowSums() 
  
df <- length_mat %>% 
 forecast::ma(3) %>% 
  MTS::diffM(d=1) %>% as.data.frame()
  names(df) <- names(length_mat)
  
  df<-df %>% 
  ts(frequency = freq, start = c(0, 1))

  mts_stationary_list[[n]] <- df
  
}

names(mts_stationary_list)<- names(df_list)
```

## We convert the multiple times series data to appropriate dataframe for stack area plot
```{r}

selected_list <-mts_stationary_list
stacked_avg_list <- list()

for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  
  colNames <- colnames(selected_list[[m]][ ,selected_data])
  length_mat<- selected_list[[m]][, selected_data] %>%  forecast::ma(5) %>% as.data.frame() # add forecast::ma(5) if smoothening is wanted
  colnames(length_mat) <- colNames 
  
  mts_dim<- dim(length_mat)
  length_mat$Time <- as.numeric(rep(seq(0,mts_dim[1]-1)* 1/freq), times=mts_dim[2] )
  
  stacked_mat_list <-list()
    for (n in 1:mts_dim[2]){
     stacked_mat <- length_mat[ ,c(mts_dim[2]+1, n)]
     stacked_mat$neurite <- colnames(length_mat)[n]
     colnames(stacked_mat)[2]<-"Length"
     stacked_mat_list[[n]]<-stacked_mat
    }
  stacked_mat <-  data.table::rbindlist(stacked_mat_list)
 stacked_avg_list [[m]] <- stacked_mat
}

names( stacked_avg_list)<- names(df_list)
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
stacked_avg_list$`170522_blebb_KO/Class_2/Kymo/Kymo_170522_KO1-30_WT31-60_DMSO1-10-51-60_blebb11-20-41-50_cytoD21-30-31-40_003_11_R3D.dv_class_2.tif`%>%
  ggplot(aes(x = Time, y = Length, fill = neurite)) +
    geom_area(alpha=0.5 , size=.25, colour="white") +
    scale_fill_viridis(discrete = T) + # scale_fill_brewer(palette = "Set2") +
    theme_ipsum(grid = F, ticks = T) 
```

## Calculate neurite outgrowth/retraction events from the stationary/differential curves at different thresholds

```{r}
selected_list <- mts_stationary_list # mts_stationary_list or mts_avg_list

pos_neg_thres_list <- list()

res <- 0.1
fold<- 10

for (n in 1: fold){ 
  
  thres= n*res
  pos_neg_list <- list()
  
  for (m in 1:length(selected_list)) {
    
    selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
    length_mat<- selected_list[[m]][, selected_data] %>% as.data.frame()
    
    length_mat2 <- length_mat
    neurite_num <- ncol(length_mat)
  
    tab1 <- apply(length_mat, 1, function(x) sum(x > thres, na.rm = TRUE)) %>%  table() %>% as.data.table()
    tab1$act<- "extension"
    names(tab1)[1]<- "Num"
    tab1$Num <- tab1$Num %>% as.numeric()
    tab1$pct<- tab1$N/sum(tab1$N)
    tab1$rel_n <- tab1$Num/neurite_num
    tab1$Cell<- names(df_list)[m]
    
    tab2 <- apply(length_mat, 1, function(x) sum(x < -thres, na.rm = TRUE)) %>%  table() %>% as.data.table()
    tab2$act<- "retraction"
    names(tab2)[1]<- "Num"
    tab2$Num <- tab2$Num %>% as.numeric()
    tab2$pct<- tab2$N /sum(tab2$N)
    tab2$rel_n <- tab2$Num/neurite_num
    tab2$Cell<- names(df_list)[m]
  
    tab3 <- apply(length_mat, 1, function(x) sum(x < thres & x > -thres, na.rm = TRUE)) %>%  table() %>% as.data.table()
    tab3$act<- "pause"
    names(tab3)[1]<- "Num"
    tab3$Num <- tab3$Num %>% as.numeric()
    tab3$pct<- tab3$N/sum(tab3$N)
    tab3$rel_n <- tab3$Num/neurite_num
    tab3$Cell<- names(df_list)[m]
    
    tab4 <- rbind(tab1, tab2, tab3)
    tab4$Thres<- thres
    
    pos_neg_list[[m]] <- tab4
  }

  names(pos_neg_list)<- names(df_list)
  
  pos_neg_thres_list[[n]]<- assign(paste0("Thres_", thres, sep=""),pos_neg_list)
  names(pos_neg_thres_list)[n]<- paste0("Thres_", thres, sep="")
  
}
```

```{r message=FALSE, warning=FALSE}
summary_list <- list()

for (n in 1: fold){
  
  pos_neg_list <- pos_neg_thres_list[[n]]
  
  pos_neg_tab_all <- pos_neg_list[[1]] 
  
  for (m in 2:length(pos_neg_list)) {
    
    df2 <- pos_neg_list[[m]] 
    pos_neg_tab_all <- rbind(pos_neg_tab_all, df2) 
  
  }

  tab <- pos_neg_tab_all[act == "extension"] %>% group_by(Num) %>% rstatix::get_summary_stats(pct)
  tab$Thres<- n*res
  summary_list [[n]] <- tab 

}

names(summary_list) <-  names(pos_neg_thres_list)

df_summary_all <- summary_list [[1]]

for (n in 2:fold){
  df_summary_all <-  rbind (df_summary_all, summary_list [[n]])
  
}

```

```{r  message = FALSE, warning=FALSE, fig.width=10}
ggplot(df_summary_all, aes(x=Num, y=mean, group=Thres)) +
  
  geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd, fill=Thres),   alpha=0.05) + 
  geom_line(aes(colour=Thres), size=1, alpha=1) +
  
  scale_y_continuous(  breaks = scales::pretty_breaks(n = 4))+
  scale_x_continuous(name="Neurite_No", limits=c(0, 10), expand = c(0, 0), breaks = c(0:10))+
  
  scale_fill_viridis(discrete = F) + # scale_fill_brewer(palette = "Set2") +
  scale_color_viridis(discrete = F)+

  theme_linedraw()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12), 
        legend.position = "none") #for centering the text
```

```{r}
ggsave(
  "pause_freq2.pdf",
  plot = last_plot(),
  device = "pdf")
```

```{r}
pos_neg_list <- pos_neg_thres_list$Thres_0.2

pos_neg_tab_all <- pos_neg_list[[1]] 

for (m in 2:length(pos_neg_list)) {
  
  df2 <- pos_neg_list[[m]] 
  pos_neg_tab_all <- rbind(pos_neg_tab_all, df2) 

}

tab <- pos_neg_tab_all[act == "extension"]

df_summary <- tab %>% group_by(Num) %>% rstatix::get_summary_stats(pct)
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
removed <- pos_neg_tab_all[act == "extension" & Num == 0 & pct > 0.5] 
removed_cell<- removed$Cell
pos_neg_tab_all_removed<- pos_neg_tab_all[grep(paste(removed_cell,collapse="|"), pos_neg_tab_all$Cell , invert=TRUE) ,]
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
ggplot(data=tab, aes(x=Num, y=pct, group=Cell)) +
  geom_line((aes(color=Cell)))+
  geom_point((aes(color=Cell)))+ 
scale_x_continuous(name="Neurite_No", limits=c(0, 6), breaks = c(0,1,2,3,4,5,6))+
theme_linedraw()+
  
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12), 
        legend.position = "none") #for centering the text
```

## Smoothen the stationary (diff) profiles 
```{r  message = FALSE, warning=FALSE, fig.width=10}

selected_stationary <- mts_stationary_list[[2]]
selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_stationary))
selected_stationary <-  selected_stationary[, selected_data ]

selected_stationary_avg<- selected_stationary %>% forecast::ma(5) %>% as.ts()
colnames(selected_stationary_avg)<-colnames(selected_stationary)

plot_title<- paste("TS Plot of ", names(mts_stationary_list)[2], sep = "")
thres= 0.1
 
 selected_stationary_avg %>%  autoplot() +
  ggtitle (plot_title) +
  geom_line(aes(),size=0.25)  +
   geom_hline(yintercept=0, linetype="dashed", color = "red")+
   geom_hline(yintercept=thres, linetype="dashed", color = "blue")+
   geom_hline(yintercept=-thres, linetype="dashed", color = "blue")+
   scale_color_viridis(discrete=T)+
   theme_minimal()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold")) #for centering the text
```

## calculate Pos-Neg table from smoothened stationary (diff) profile
```{r  message = FALSE, warning=FALSE, fig.width=10}
mts_dim<- dim(selected_stationary)
pos_neg_df <- data.frame(matrix(ncol = mts_dim[2], nrow = mts_dim[1]))
colnames(pos_neg_df) <- selected_stationary %>% colnames()

for (n in 1:mts_dim[2])  {
    pos_neg_df[, n][selected_stationary[ , colnames(pos_neg_df)[n]] > thres] <- "Pos" # change threshold here
    pos_neg_df[, n][selected_stationary[ , colnames(pos_neg_df)[n]] < -thres] <- "Neg" # change threshold here
    }
  
#Count the growth and retraction events at each time point
pos_neg_df$Pos <- apply(pos_neg_df, 1, function(x) length(which(x=="Pos")))
pos_neg_df$Neg <- apply(pos_neg_df, 1, function(x) length(which(x=="Neg")))
pos_neg_df$na <- apply(pos_neg_df, 1, function(x) length(which(is.na(x))))

ggplot(pos_neg_df, aes(x=Pos)) + # change to "Neg" if retraction is of interest
 geom_histogram( colour="black", fill="white", binwidth=0.5)
```

## We smoothen the stationary profiles first, separete outrgwoth and retraction events, and sum the changes of all neurites. (Smoothening -> separate -> sum)
```{r warning=FALSE}

# Smoothen the stationary (diff) profiles and optionally filter out small length change with threshold

selected_list <- mts_stationary_list # mts_stationary_list or mts_avg_list

mts_stationary_pos_list <- list()
mts_stationary_neg_list <- list()
thres= 0.0 # change threshold here

# Positive peaks only (outgrowth event)
for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  length_mat<- selected_list[[m]][, selected_data ]
  
  length_mat_avg<- length_mat %>% forecast::ma(3) %>% as.ts() # Average smoothening
  colnames(length_mat_avg)<-colnames(length_mat)

  mts_dim<- dim(length_mat_avg)
    for (n in 1:mts_dim[2]){
      length_mat_avg[, n][length_mat_avg[ , colnames(length_mat_avg)[n]] < thres] <- 0 
    }
  
  length_mat_avg <- length_mat_avg %>% as.data.frame() 
  length_mat_avg$sum <- length_mat_avg  %>% rowSums()  # We sum up the extent of outgrowth at each time point
  mts_stationary_pos_list[[m]] <- length_mat_avg %>% ts(frequency = freq, start = c(0, 1)) 

}

names(mts_stationary_pos_list) <- names(mts_stationary_list)

# Negative peaks only (retraction event)
for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  length_mat<- selected_list[[m]][, selected_data ]
  length_mat_avg<- length_mat %>% forecast::ma(3) %>% as.ts() # Average smoothening
  colnames(length_mat_avg)<-colnames(length_mat)
  
  mts_dim<- dim(length_mat_avg)
    for (n in 1:mts_dim[2]){
      length_mat_avg[, n][length_mat_avg[ , colnames(length_mat_avg)[n]] > - thres] <- 0 
    }
  length_mat_avg <- length_mat_avg %>% as.data.frame() 
  length_mat_avg$sum <- length_mat_avg  %>% rowSums()  # We sum up the extent of retraction at each time point
  mts_stationary_neg_list[[m]] <- length_mat_avg %>% ts(frequency = freq, start = c(0, 1)) 
}

names(mts_stationary_neg_list) <- names(mts_stationary_list)
```

## extension and retraction lifetime (duration)
# Extension duration
```{r}
thres <- 0.00

list_1 <- list()

for (m in 1:length(mts_stationary_pos_list)){
  
  df <- mts_stationary_pos_list[[m]] %>% as.data.table()
  df <- df[, !"sum"] %>% as.data.frame()
  df<- df[, (colSums(df, na.rm = T)> 0)] %>% as.data.table()
  
  list_2 <-  list()
  
   for (n in 1: ncol(df)){
    
    df2 <- df[ ,..n]
    df2$time <- seq(0, (nrow(df2)-1), 1)
    df2 <- df2[complete.cases(df2), ] 
    
    df2$inside <- df2[ , 1] > thres
    df2$start  <- ! c(FALSE, df2$inside[-nrow(df2)]) & df2$inside
    df2$end    <- df2$inside & ! c(df2$inside[-1], FALSE)  # Not used
    df2 <- df2[ df2$inside, ]
    df2$group  <- cumsum(df2$start)  # Number the spells
    colnames(df2)[1] <- "velocity"
    
    df3 <- plyr::ddply( df2, "group", summarize,
                        start  = min(time),
                        end    = max(time),
                        length = length(velocity),
                        mean   = mean(velocity))
   
    df3$neurite <- names(df)[n]
    list_2[[n]] <- df3

   }
  
  df3<- list_2 %>% data.table::rbindlist()
  df3$cell<- names(mts_stationary_list)[m]
  df3$growth<- "extension"
  df3$cond <- names(mts_stationary_list)[m] %>% 
  gsub("G:/Brightfield_analysis/longterm-brightfield_Blebb_CytoD/", "", .) %>% 
  gsub("\\/.*","",.)
  
  list_1[[m]] <- df3

}

extension_duration_df <- list_1 %>% data.table::rbindlist()

```

# Retraction duration
```{r}

list_1 <- list()
for (m in 1:length(mts_stationary_neg_list)){
  
  df <- mts_stationary_neg_list[[m]] %>% as.data.table()
  df <- df[, !"sum"] %>% as.data.frame()
  df<- df[, (colSums(df, na.rm = T)< 0)] %>% as.data.table()
  list_2 <-  list()
  
   for (n in 1: ncol(df)){
     
    df2 <- df[ ,..n]
    df2$time <- seq(0, (nrow(df2)-1), 1)
    df2 <- df2[complete.cases(df2), ] 
    
    df2$inside <- df2[ , 1] < -thres
    df2$start  <- ! c(FALSE, df2$inside[-nrow(df2)]) & df2$inside
    df2$end    <- df2$inside & ! c(df2$inside[-1], FALSE)  # Not used
    df2 <- df2[ df2$inside, ]
    df2$group  <- cumsum(df2$start)  # Number the spells
    colnames(df2)[1] <- "velocity"
    
    df3 <- plyr::ddply( df2, "group", summarize,
                        start  = min(time),
                        end    = max(time),
                        length = length(velocity),
                        mean   = mean(velocity))
   
    df3$neurite <- names(df)[n]
    list_2[[n]] <- df3

   }
  
  df3<- list_2 %>% data.table::rbindlist()
  
  # add sample labeling
  df3$cell<- names(mts_stationary_list)[m]
  df3$growth<- "retraction"
  df3$cond <- names(mts_stationary_list)[m] %>% 
  gsub("G:/Brightfield_analysis/longterm-brightfield_Blebb_CytoD/", "", .) %>% 
  gsub("\\/.*","",.)
  
  list_1[[m]] <- df3

}

retraction_duration_df <- list_1 %>% data.table::rbindlist()

```

# Data points at cell level
```{r}
duration_df <- rbind(extension_duration_df , retraction_duration_df )

duration_df$length <- duration_df$length/freq
df_summary_duration <-duration_df %>% 
 group_by(cell, cond, growth) %>% 
  rstatix::get_summary_stats(length) %>% as.data.table()
```

```{r}
# add sample labeling for grouping and batch control
df_summary_duration$ID<- paste (df_summary_duration$cell, df_summary_duration$cond, df_summary_duration$growth,  sep = "_")
# df_summary_duration$neurite_ID<- paste (df_summary_duration$cell, df_summary_duration$neurite, df_summary_duration$cond, df_summary_duration$growth,  sep = "_")
df_summary_duration$group<- paste (df_summary_duration$cond, df_summary_duration$growth, sep = "_")
df_summary_duration$batch<- "_"
df_summary_duration[grep("170522", df_summary_duration$cell), ]$batch <- "170522"
df_summary_duration[grep("170724", df_summary_duration$cell), ]$batch <- "170724"
df_summary_duration[grep("170814", df_summary_duration$cell), ]$batch <- "170814"
df_summary_duration[grep("170821", df_summary_duration$cell), ]$batch <- "170821"
df_summary_duration[grep("170812", df_summary_duration$cell), ]$batch <- "170812"
df_summary_duration[grep("170515", df_summary_duration$cell), ]$batch <- "170515"
df_summary_duration[grep("151209", df_summary_duration$cell), ]$batch <- "151209"

df_summary_duration <- df_summary_duration[cond != "CytoD_WT"] #& batch !="151209"
```

```{r}
outliers <- df_summary_duration %>% 
  group_by(cond, growth) %>% 
 rstatix::identify_outliers(mean) 

df_summary_duration<- df_summary_duration[! (ID %in% outliers$ID)]

```

```{r}
duration_df %>% group_by(cell, cond, growth, neurite) %>% dplyr::summarise(n = n())
```

# generate pairwise comparison list
```{r}
my_comparisons <- df_summary_duration$group %>% unique() %>% as.character() %>% combn(2, simplify = FALSE) 

my_comparisons <- list(c("DMSO_WT_extension",  "Blebb_WT_extension"), c("DMSO_KO_extension",  "Blebb_KO_extension"), c("DMSO_WT_retraction",  "Blebb_WT_retraction"),  c("DMSO_KO_retraction",  "Blebb_KO_retraction"), c("DMSO_WT_extension",  "DMSO_KO_extension"), c("Blebb_WT_extension",  "Blebb_KO_extension"), c("DMSO_WT_retraction",  "DMSO_KO_retraction"), c("Blebb_WT_retraction",  "Blebb_KO_retraction"))
```

```{r message = FALSE, warning=FALSE, fig.width=10}
# my_comparisons <- list(c("axon", "minor"))
#pj <- position_jitter(seed = 1, width = .1, height = 0)
# df_summary_duration[batch !="151209"] 
df_summary_duration  %>%  mutate(group = fct_relevel(group, "DMSO_WT_extension",  "Blebb_WT_extension", "DMSO_KO_extension",  "Blebb_KO_extension", "DMSO_WT_retraction",  "Blebb_WT_retraction", "DMSO_KO_retraction",  "Blebb_KO_retraction")) %>% 
  
ggplot ( aes(x=group, y=mean, fill =cond) )+ 
  scale_fill_brewer(palette="Dark2") + 
  theme_classic()+
geom_violin(aes(), trim=F,  kernel ="gaussian", adjust=1)  +
 geom_boxplot(aes(), width=0.2, outlier.colour = NA, fill="white")+ 
 geom_jitter(aes(color = batch), shape=16, position=position_jitter(0.1))+ #aes(color = batch)
  
  # geom_line(aes(group = Image))+
  # geom_point() +
  #geom_path(aes(group = cell),alpha = 0.5)+
    
 scale_y_continuous(expand = c(0, 0), limits = c(4, 60), breaks = scales::pretty_breaks(n = 5))+ 
# geom_signif(comparisons = my_comparisons ,test = "wilcox.test",  color="black")+
 stat_compare_means(comparisons = my_comparisons, paired=F, method="wilcox.test", 
                     #label.y = c(7, 7.5, 8, 8.5, 9, 9.5, 10, 10.5), 
                     label = "{p.adj.format}{p.adj.signif}") +
  
  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold", angle = 90)) #for centering the text 
```

```{r}
ggsave(
  "ext_ret_duration_thres_0.pdf",
  plot = last_plot(),
  device = "pdf")
```

# for extension and retraction velocity
```{r}
Ext_velocity <- list()
Ret_velocity <- list()

for (m in 1:length(mts_stationary_list)){
    
df <- mts_stationary_pos_list[[m]] %>% as.data.frame()
df[ df == 0 ] <- NA
df1 <- colMeans(df, na.rm = T)  %>% as.data.frame()
colnames(df1) <- "mean_vel"
df1$median_vel <- df %>% as.matrix() %>% matrixStats::colMedians(na.rm = T)  # calculate the average velocity through the observation window. Or matrixStats::colMedians
df1$q90<- df %>% apply(2 , quantile , probs = 0.9 , na.rm = TRUE ) 

df1$cond <- names(mts_stationary_list)[m] %>% 
  gsub("G:/Brightfield_analysis/longterm-brightfield_Blebb_CytoD/", "", .) %>% 
  gsub("\\/.*","",.)

df1$cell <- names(mts_stationary_list)[m] 
df1$neurite <- rownames(df1)
Ext_velocity [[m]] <- df1

df <- mts_stationary_neg_list[[m]] %>% as.data.frame()
df [ df == 0 ] <- NA
df2 <- colMeans(df, na.rm = T)  %>% as.data.frame()
colnames(df2) <- "mean_vel"
df2$median_vel <- df %>% as.matrix() %>% matrixStats::colMedians(na.rm = T)  # calculate the average velocity through the observation window. Or matrixStats::colMedians
df2$q90<- df %>% apply(2 , quantile , probs = 0.1 , na.rm = TRUE ) # 10-tile insetad of 90-tile due to negative value

df2$cond <- names(mts_stationary_list)[m] %>% 
  gsub("G:/Brightfield_analysis/longterm-brightfield_Blebb_CytoD/", "", .) %>% 
  gsub("\\/.*","",.)

df2$cell <- names(mts_stationary_list)[m] 
df2$neurite <- rownames(df2)
Ret_velocity [[m]] <- df2
}
```

```{r}
Ext_velocity <- Ext_velocity %>% reduce(bind_rows)  %>% as.data.frame()
Ext_velocity$Growth <- "Extension"
Ret_velocity <- Ret_velocity %>% reduce(bind_rows)  %>% as.data.frame()
Ret_velocity$Growth <- "Retraction"

velocity_comb_df<- rbind(Ext_velocity, Ret_velocity) %>% as.data.table()
#names(velocity_comb_df)[1]<- "Velocity"
velocity_comb_df <- velocity_comb_df[neurite != "sum"]

velocity_comb_df$mean_vel <- velocity_comb_df$mean_vel %>% abs()
velocity_comb_df$median_vel <- velocity_comb_df$median_vel %>% abs()
velocity_comb_df$q90 <- velocity_comb_df$q90 %>% abs()
velocity_comb_df[mean_vel == "NaN"]$mean_vel <- 0
velocity_comb_df[median_vel == "NaN"]$median_vel <- 0

velocity_comb_df$batch<- "_"
velocity_comb_df[grep("170522", velocity_comb_df$cell), ]$batch <- "170522"
velocity_comb_df[grep("170724", velocity_comb_df$cell), ]$batch <- "170724"
velocity_comb_df[grep("170814", velocity_comb_df$cell), ]$batch <- "170814"
velocity_comb_df[grep("170821", velocity_comb_df$cell), ]$batch <- "170821"
velocity_comb_df[grep("170812", velocity_comb_df$cell), ]$batch <- "170812"
velocity_comb_df[grep("170515", velocity_comb_df$cell), ]$batch <- "170515"
velocity_comb_df[grep("151209", velocity_comb_df$cell), ]$batch <- "151209"

#saveRDS(velocity_comb_df, "KO_Bleb.rds")
```

# Median
```{r}
df_summary_median_vel <-velocity_comb_df %>% 
 group_by(cell, cond, Growth, batch) %>% 
  rstatix::get_summary_stats(median_vel) %>% as.data.table()
```

```{r}
df_summary_median_vel$ID<- paste (df_summary_median_vel$cell, df_summary_median_vel$cond, df_summary_median_vel$Growth,  sep = "_")
df_summary_median_vel$group<- paste (df_summary_median_vel$cond, df_summary_median_vel$Growth, sep = "_")
df_summary_mean_vel <- df_summary_mean_vel[cond != "CytoD_WT"] #  & batch !="151209"

outliers <- df_summary_median_vel %>% 
  group_by(Growth, cond) %>% 
 rstatix::identify_outliers(mean) 

df_summary_median_vel <- df_summary_median_vel[! (ID %in% outliers$ID)]
```

# Mean
```{r}
df_summary_mean_vel <-velocity_comb_df %>% 
 group_by(cell, cond, Growth, batch) %>% 
  rstatix::get_summary_stats(mean_vel) %>% as.data.table()
```

```{r}
df_summary_mean_vel$ID<- paste (df_summary_mean_vel$cell, df_summary_mean_vel$cond, df_summary_mean_vel$Growth,  sep = "_")
df_summary_mean_vel$group<- paste (df_summary_mean_vel$cond, df_summary_mean_vel$Growth, sep = "_")
df_summary_mean_vel <- df_summary_mean_vel[cond != "CytoD_WT"] # & batch !="151209"

outliers <- df_summary_mean_vel %>% 
   group_by(Growth, cond) %>% 
 rstatix::identify_outliers(mean) 

df_summary_mean_vel <- df_summary_mean_vel[! (ID %in% outliers$ID)]

```

# 90-tile
```{r}
df_summary_q90_vel <-velocity_comb_df %>% 
 group_by(cell, cond, Growth, batch) %>% 
  rstatix::get_summary_stats(q90) %>% as.data.table()
```

```{r}
df_summary_q90_vel $ID<- paste (df_summary_q90_vel $cell, df_summary_q90_vel $cond, df_summary_q90_vel $Growth,  sep = "_")
df_summary_q90_vel $group<- paste (df_summary_q90_vel $cond, df_summary_q90_vel $Growth, sep = "_")
df_summary_q90_vel  <- df_summary_q90_vel [cond != "CytoD_WT"] # & batch !="151209"

outliers <- df_summary_q90_vel  %>% 
   group_by(Growth, cond) %>% 
 rstatix::identify_outliers(mean) 

df_summary_q90_vel  <- df_summary_q90_vel [! (ID %in% outliers$ID)]

```

# generate pairwise comparison list
```{r}
my_comparisons <- df_summary_median_vel$group %>% unique() %>% as.character() %>% combn(2, simplify = FALSE) 

my_comparisons <- list(c("DMSO_WT_Extension",  "Blebb_WT_Extension"), c("DMSO_KO_Extension",  "Blebb_KO_Extension"), c("DMSO_WT_Retraction",  "Blebb_WT_Retraction"),  c("DMSO_KO_Retraction",  "Blebb_KO_Retraction"), c("DMSO_WT_Extension",  "DMSO_KO_Extension"), c("Blebb_WT_Extension",  "Blebb_KO_Extension"), c("DMSO_WT_Retraction",  "DMSO_KO_Retraction"), c("Blebb_WT_Retraction",  "Blebb_KO_Retraction"))
```

```{r message = FALSE, warning=FALSE, fig.width=10}

# my_comparisons <- list(c("axon", "minor"))

#pj <- position_jitter(seed = 1, width = .1, height = 0)
# df_summary_duration[batch !="151209"] 
df_summary_mean_vel  %>%  mutate(group = fct_relevel(group, "DMSO_WT_Extension",  "Blebb_WT_Extension", "DMSO_KO_Extension",  "Blebb_KO_Extension", "DMSO_WT_Retraction",  "Blebb_WT_Retraction", "DMSO_KO_Retraction",  "Blebb_KO_Retraction")) %>% 
ggplot ( aes(x=group, y=mean, fill =cond) )+ 
  scale_fill_brewer(palette="Dark2") + 
  theme_classic()+
geom_violin(aes(), trim=F,  kernel ="gaussian", adjust=1)  +
 geom_boxplot(aes(), width=0.2, outlier.colour = NA, fill="white")+ 
  geom_jitter(aes(color = batch), shape=16, position=position_jitter(0.1))+ #aes(color = batch)
 # geom_line(aes(group = Image))+
  
 #  geom_point() +
 # geom_path(aes(group = cell),alpha = 0.5)+
    
 scale_y_continuous(expand = c(0, 0), limits = c(-1, 3), breaks = scales::pretty_breaks(n = 7))+ 
# geom_signif(comparisons = my_comparisons ,test = "wilcox.test",  color="black")+
  stat_compare_means(comparisons = my_comparisons, paired=F, method="wilcox.test", 
                     #label.y = c(1.15, 1.2, 1.25, 1.3, 1.35, 1.4), 
                     label = "{p.adj.format}{p.adj.signif}") +
  
    theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text.x=element_text(size=12, face="bold", angle = 90)) #for centering the text 
```

```{r}
ggsave(
  "WT_KO_Blebb_vel_mean.pdf",
  plot = last_plot(),
  device = "pdf")

df_summary_mean_vel$cond %>% table()
```

```{r message = FALSE, warning=FALSE, fig.width=10}

# my_comparisons <- list(c("axon", "minor"))

#pj <- position_jitter(seed = 1, width = .1, height = 0)

df_summary_median_vel  %>%  mutate(group = fct_relevel(group, "DMSO_WT_Extension",  "Blebb_WT_Extension", "DMSO_KO_Extension",  "Blebb_KO_Extension", "DMSO_WT_Retraction",  "Blebb_WT_Retraction", "DMSO_KO_Retraction",  "Blebb_KO_Retraction")) %>% 
ggplot ( aes(x=group, y=mean, fill =cond) )+ 
  scale_fill_brewer(palette="Dark2") + 
  theme_classic()+
geom_violin(aes(), trim=F,  kernel ="gaussian", adjust=1)  +
 geom_boxplot(aes(), width=0.2, outlier.colour = NA, fill="white")+ 
  geom_jitter(aes(color = batch), shape=16, position=position_jitter(0.1))+
 # geom_line(aes(group = Image))+
  
 #  geom_point() +
 # geom_path(aes(group = cell),alpha = 0.5)+
    
 scale_y_continuous(expand = c(0, 0), limits = c(-0.2, 1), breaks = scales::pretty_breaks(n = 5))+ 
# geom_signif(comparisons = my_comparisons ,test = "wilcox.test",  color="black")+
  stat_compare_means(comparisons = my_comparisons, paired=F, method="wilcox.test", 
                     #label.y = c(1.15, 1.2, 1.25, 1.3, 1.35, 1.4), 
                     label = "{p.adj.format}{p.adj.signif}") +
  
    theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold", angle = 90)) #for centering the text 
```

# Time-lagged cross-corrleation of outgrowth and retraction profiles
```{r warning=FALSE}

length_ccf_list_2 <- list()

for (m in 1:length(mts_stationary_list)) {
  
  df_pos <- mts_stationary_pos_list[[m]]
  df_neg <- mts_stationary_neg_list[[m]]

   length_ccf_plot <-ccf ( df_pos[, "sum"],  
                            - df_neg[, "sum"], 
                            type = "correlation", 
                            na.action = na.pass, plot=F,
                            lag.max=100) 
  
  length_ccf_list_2[[m]] <- length_ccf_plot 

}

names(length_ccf_list_2) <-  names(mts_stationary_list)
    
length_ccf_2 <- length_ccf_list_2[[1]][["acf"]] %>% as.data.frame()
for (m in 2:length(length_ccf_list_2)) {
  
  length_ccf_2 <- cbind(length_ccf_2 , length_ccf_list_2[[m]][["acf"]])
}

colnames(length_ccf_2) <- names(length_ccf_list_2)
length_ccf_2$lag<- c(-100:100)

```

```{r}
length_ccf_DMSO_KO <- length_ccf_2[grepl('DMSO_KO', colnames(length_ccf_2))]
length_ccf_Blebb_KO <- length_ccf_2[!grepl('Blebb_KO', colnames(length_ccf_2))]
length_ccf_DMSO_WT <- length_ccf_2[grepl('DMSO_WT', colnames(length_ccf_2))]
length_ccf_Blebb_WT <- length_ccf_2[!grepl('Blebb_WT', colnames(length_ccf_2))]
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
selected_ccf <- length_ccf_Blebb_WT %>% as.data.table()  
selected_ccf$lag<- length_ccf_2$lag/freq
stacked_ccf_mat <- gather(selected_ccf, Cell, ccf, -lag)
df_summary <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf)
```

# We plot the overall outgrowth/retraction level over time
```{r  message = FALSE, warning=FALSE, fig.width=10}
selected_stationary_pos <- mts_stationary_pos_list[[18]][, "sum"] 
selected_stationary_neg <- mts_stationary_neg_list[[18]][, "sum"]
ts.union(selected_stationary_pos, selected_stationary_neg) %>%  autoplot() +
  geom_line(aes(),size=0.25)  +
   geom_hline(yintercept=0, linetype="dashed", color = "red")+
   geom_hline(yintercept=thres, linetype="dashed", color = "blue")+
   geom_hline(yintercept=-thres, linetype="dashed", color = "blue")+
  theme_ipsum(grid = F, ticks = T) + theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold"),
        legend.position = "none") #for centering the text
```

## Here we separate the outrgwoth and retraction events first, sum the changes of all neurites, and then do smoothening. (separate -> sum -> smoothening)
```{r}

selected_list <- mts_stationary_list 
mts_stationary_sum_list <- list()

thres= 0.0 # change threshold here

# Positive peaks only (outgrowth event)
for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]])) # Neurite length ("Length") only or include actin intensity (Integerated.intensity)
  length_mat_pos<- selected_list[[m]][, selected_data ]
  length_mat_neg<- selected_list[[m]][, selected_data ]
  
  mts_dim<- dim(length_mat_pos)
  
    for (n in 1:mts_dim[2]){
      length_mat_pos[, n][length_mat_pos[ , colnames(length_mat_pos)[n]] < 0] <- 0 
      length_mat_neg[, n][length_mat_neg[ , colnames(length_mat_neg)[n]] > 0] <- 0 
    }
  
  length_mat_pos <- as.data.frame(length_mat_pos)
  length_mat_neg <- as.data.frame(length_mat_neg)
   
  length_mat_pos$sum <- length_mat_pos  %>% rowSums()  
  length_mat_neg$sum <- length_mat_neg  %>% rowSums()  
  
  length_mat_pos <- length_mat_pos %>% ts(frequency = freq, start = c(0, 1))
  length_mat_neg <- length_mat_neg %>% ts(frequency = freq, start = c(0, 1))
  
  length_mat_sum <- ts.union(length_mat_pos[ ,"sum"], length_mat_neg[ ,"sum"]) 
  
  length_mat_sum <- length_mat_sum %>% forecast::ma(3) %>% as.ts() # Average smoothening
  colnames(length_mat_sum)<-c("Pos", "Neg")
  
  length_mat_sum [, "Pos"][length_mat_sum [, "Pos"] < thres] <- 0 
  length_mat_sum [, "Neg"][length_mat_sum [, "Neg"] > -thres] <- 0 
  
  mts_stationary_sum_list[[m]] <- length_mat_sum

}

names(mts_stationary_sum_list)<- names(mts_stationary_list )

```

# Time-lagged cross-corrleation of outgrowth and retraction profiles
```{r warning=FALSE}

length_ccf_list <- list()
w <- 100 # set max lag size

for (m in 1:length(mts_stationary_sum_list)) {
  
  selected_length_sta_sum <- mts_stationary_sum_list[[m]]

  length_ccf_plot <-ccf (selected_length_sta_sum[, "Pos"],  
                            -selected_length_sta_sum[, "Neg"], 
                            type = "correlation", 
                            na.action = na.pass, plot=F,
                            lag.max=w) 
  
  length_ccf_list[[m]] <- length_ccf_plot 
  
}

names(length_ccf_list) <-  names(mts_stationary_sum_list)

length_ccf <- length_ccf_list[[1]][["acf"]] %>% as.data.frame()
for (m in 2:length(length_ccf_list)) {
  
  length_ccf <- cbind(length_ccf , length_ccf_list[[m]][["acf"]])
}

colnames(length_ccf) <- names(length_ccf_list)

length_ccf$lag<- c(-w:w)
```

```{r}
length_ccf_Blebb_KO <- length_ccf[grepl('Blebb_KO', colnames(length_ccf))]
length_ccf_Blebb_WT <- length_ccf[grepl('Blebb_WT', colnames(length_ccf))]
length_ccf_DMSO_KO <- length_ccf[grepl('DMSO_KO', colnames(length_ccf))]
length_ccf_DMSO_WT <- length_ccf[grepl('DMSO_WT', colnames(length_ccf))]
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
selected_ccf <- length_ccf_DMSO_WT %>% as.data.table()  
selected_ccf$lag<- length_ccf$lag/freq
stacked_ccf_mat <- gather(selected_ccf, Cell, ccf, -lag)
df_summary <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf)
```

```{r}
ggsave(
  "KO_DMSO_ext_ret_corr.pdf",
  plot = last_plot(),
  device = "pdf")
```
# Time-lagged cross-corrleation of outgrowth and retraction profiles
# Identify peak and valley

## moving/rolling window cross-correlation
## windowed Time Lagged Cross Correlation (wTLCC) 
```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
save.image(saved_path)
```

```{r}
library(biwavelet)
library(WaveletComp)
# install.packages("WaveletComp")

length_mat<- mts_stationary_sum_list[[1]] %>%  tseries::na.remove() %>% as.data.frame() 
mts_dim<- dim(length_mat)
length_mat$Time <- as.numeric(rep(seq(0,mts_dim[1]-1)* 1/freq), times=mts_dim[2] )
length_mat <- length_mat[, c(3, 1, 2)]

ts_1 <- length_mat[, c(1,2)]
ts_2 <- length_mat[, c(1,3)]

d1d2.xwt <- xwt(ts_1,ts_1) # Compute cross-wavelet
d1d2.wtc <- wtc(ts_1,ts_1) # Compute cross-wavelet

par(oma = c(0, 0, 0, 1), mar = c(5, 4, 4, 5) + 0.1)
plot(d1d2.xwt, type = "power.corr.norm", plot.cb = TRUE, plot.phase = TRUE)

d1d2.xwt %>% plot.biwavelet(type = "power.corr.norm", plot.sig=T, plot.cb = TRUE, plot.phase = TRUE)

my.wc <- analyze.coherency(length_mat, my.pair = c(2,3), loess.span = 0, dt = 1, dj = 1/20, make.pval = TRUE, n.sim = 10)

wc.avg(my.wc, siglvl = 0.01, sigcol = "red", sigpch = 20,periodlab = "period (days)")

plot(d1d2.xwt, type = "power.corr.norm", plot.cb = TRUE, plot.phase = TRUE)
wc.image(my.wc, n.levels = 250, legend.params = list(lab = "cross-wavelet power levels"), timelab = "", periodlab = "period (days)")
```

## Package installation
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
