---
title: "R Notebook"
output: html_notebook
---

<!--
# Lag cross-correlation of neurite-tip velocity vs soma intensity (brightfield, WT vs Arp3 KO)

## What this file does

Reads per-cell brightfield-derived neurite-tip velocity and soma-intensity
tables (data pasted as in-line tibbles in this Rmd) for WT and Arp3 KO
2D-cultured neurons. For each neuron the local-detrended growth velocity is
cross-correlated against the change of soma intensity, and CCF curves are
pooled to a mean +/- SD across the cohort. Six PDFs are written (`ggsave`
at lines 157, 327, 374, 602, 744, 785) including the CCF-curve PDFs for
Fig 4p and the per-genotype extension / retraction velocity comparisons.

## Manuscript panel

**Fig 4p**: "Quantification of extension and retraction velocities of WT and
Arp3 KO neurites in in vitro culture. N = 100 WT neurons from 4 experiments
and 117 Arp3 KO neurons from 6 experiments. ***p < 0.001; **p < 0.01; ns,
p >= 0.05 by Kruskal-Wallis test followed by Dunn's post-hoc test with
Bonferroni correction. Exact P values are shown."

Source Data: `Fig 4p`.

## Outputs

- Six PDFs (`ggsave` at lines 157, 327, 374, 602, 744, 785).

## Original G:\ source

G:/Figure 4_Arp3KO/Fig. 4p/R_scripts/Neurite activity corr_2.Rmd
-->

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

```{r message=FALSE, warning=FALSE, include=FALSE}
library(pracma)
library(peakPick)
library(quantmod)
library(ggpmisc)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggfortify)
library(forecast)
library(tsibble)
library(broom)
library(tidyquant)
library(tsbox)
library("pheatmap")
library("RColorBrewer")
library(ggpmisc)
library(ggfortify)
library(viridis)
library(hrbrthemes)
library(ggpubr)
library(easyGgplot2)
library(data.table)
```

## load existing/saved environment (.RData)
```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(Path_1)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
load(file=saved_path)

```

```{r}
df_summary_mean_vel[Cond =="WT"]
```

```{r}
filename <- file.choose()
WT_df_list <- readRDS(filename)

filename <- file.choose()
KO_df_list <- readRDS(filename)

df_list <- c(WT_df_list, KO_df_list)
```

# (optional) despiking
```{r}

for (n in 1:length(df_list)){
  df <- df_list[[n]]
  df_list[[n]] <- pracma::hampel(df, k=7)$y
  
}
```

## Import time-series neurite length data (Data -> moving average (avg) -> differencing (stationary))
```{r}
freq<- 0.5
mts_list<- list()
mts_avg_list <- list()
mts_stationary_list <- list()

for (n in 1:length(df_list))  {
 
  length_mat<-df_list[[n]] %>% as.data.table()
  length_mat$sum_length <- length_mat %>% rowSums() 
  
  mts_list[[n]]<- length_mat %>% ts(frequency = freq, start = c(0, 1)) 
  
  # Moving Average to smooth out outliers/noise
  mts_avg_list[[n]] <- mts_list[[n]] %>% forecast::ma(5) %>% as.ts() # smoothening 1
  colnames(mts_avg_list[[n]])<-colnames(mts_list[[n]])
 
  mts_stationary_list[[n]] <- mts_avg_list[[n]] %>% MTS::diffM(d=1) %>% ts(frequency = freq, start = c(0, 1)) # Differential 
  
}

names(mts_list)<- names(df_list)
names(mts_avg_list)<- names(df_list)
names(mts_stationary_list)<- names(df_list)

```

## Plotting the time series as curves
```{r}

mts_profile_list <- list()
avg_profile_list <- list()
sta_profile_list <- list()
thres= 0.3 # change threshold for stationary profile here

for (n in 1:length(mts_avg_list))  {
  
 plot_title<- paste(names(df_list)[n], sep = "")
 selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(mts_avg_list[[n]]))

 theme_set(theme_bw())
 avg_profile_list[[n]] <- mts_avg_list[[n]][, selected_data] %>%  autoplot() +
  #ggtitle (plot_title) +
  geom_line(aes(),linewidth=1)  +
   scale_y_continuous(breaks = scales::pretty_breaks(n = 7))+
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA), breaks = scales::pretty_breaks(n = 8)) + 
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold")) #for centering the text
 
 mts_profile_list[[n]] <- mts_list[[n]][, selected_data] %>%  autoplot() +
 # ggtitle (plot_title) +
  geom_line(aes(),linewidth=1)  +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 7))+
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA), breaks = scales::pretty_breaks(n = 15)) + 
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold")) #for centering the text
 
 sta_profile_list[[n]] <- mts_stationary_list[[n]][, selected_data] %>%  autoplot() +
  #ggtitle (plot_title) +
  geom_line(aes(),linewidth=0.25)  +
   geom_hline(yintercept=0, linetype="dashed", color = "red")+
   geom_hline(yintercept=thres, linetype="dashed", color = "blue")+
   geom_hline(yintercept=-thres, linetype="dashed", color = "blue")+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold")) #for centering the text
 
}

names(mts_profile_list) <- names(df_list)
names(avg_profile_list) <- names(df_list)
names(sta_profile_list) <- names(df_list)
```

## Visualization of the selected curves
```{r message = FALSE, warning=FALSE, fig.width=10}
fig_A <- mts_profile_list[[33]]
fig_B <- avg_profile_list[[33]]
fig_C <- sta_profile_list[[33]]
fig_D <- avg_profile_list[[18]]

ggpubr::ggarrange(fig_A, fig_B, fig_C, fig_D, 
                  labels = c("A", "B", "C", "D"),
                  ncol = 2, nrow = 2,  legend="none")

```

```{r}
ggsave(
  "KO_33_lengthProfile.eps",
  plot = last_plot(),
  device = "eps")
```

## We convert the multiple times series data to appropriate dataframe for stack area plot
```{r}

selected_list <-mts_stationary_list
stacked_avg_list <- list()

for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  
  colNames <- colnames(selected_list[[m]][ ,selected_data])
  length_mat<- selected_list[[m]][, selected_data] %>%  forecast::ma(5) %>% as.data.frame() # smoothening 2
  colnames(length_mat) <- colNames 
  
  mts_dim<- dim(length_mat)
  length_mat$Time <- as.numeric(rep(seq(0,mts_dim[1]-1)* 1/freq), times=mts_dim[2] )
  
  stacked_mat_list <-list()
    for (n in 1:mts_dim[2]){
     stacked_mat <- length_mat[ ,c(mts_dim[2]+1, n)]
     stacked_mat$neurite <- colnames(length_mat)[n]
     colnames(stacked_mat)[2]<-"Length"
     stacked_mat_list[[n]]<-stacked_mat
    }
  stacked_mat <-  data.table::rbindlist(stacked_mat_list)
 stacked_avg_list [[m]] <- stacked_mat
}

names(stacked_avg_list) <- names(df_list)
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
stacked_avg_list[[33]] %>% 
  ggplot(aes(x = Time, y = Length, fill = neurite)) +
    geom_area(alpha=1 ) +
  theme_linedraw()+
    scale_fill_viridis(discrete = T) + 
  scale_y_continuous( limits = c(-1.2, 1.5), breaks = scales::pretty_breaks(n = 7))+
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA), breaks = scales::pretty_breaks(n = 7)) + 

  # scale_fill_brewer(palette = "Set2") +
   # theme_ipsum(grid = F, ticks = T) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none", 
        axis.text=element_text(size=12)) #for centering the text
```

## The frequency of multiple extending, retracting and pausing neurites over the whole observation window
# Calculate neurite outgrowth/retraction events from the stationary/differential curves
```{r}
selected_list <- mts_stationary_list # mts_stationary_list or mts_avg_list

pos_neg_thres_list <- list()
# thres= 0.02 # change threshold here

for (n in 1: 10){
  
  thres= n*0.02
  pos_neg_list <- list()
  
  for (m in 1:length(selected_list)) {
    
    selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
    length_mat<- selected_list[[m]][, selected_data] %>% as.data.frame()
    
    length_mat2 <- length_mat
    neurite_num <- ncol(length_mat)
  
    tab1 <- apply(length_mat, 1, function(x) sum(x > thres, na.rm = TRUE)) %>%  table() %>% as.data.table()
    tab1$act<- "extension"
    names(tab1)[1]<- "Num"
    tab1$Num <- tab1$Num %>% as.numeric()
    tab1$pct<- tab1$N/sum(tab1$N)
    tab1$rel_n <- tab1$Num/neurite_num
    tab1$Cell<- names(df_list)[m]
    
    tab2 <- apply(length_mat, 1, function(x) sum(x < -thres, na.rm = TRUE)) %>%  table() %>% as.data.table()
    tab2$act<- "retraction"
    names(tab2)[1]<- "Num"
    tab2$Num <- tab2$Num %>% as.numeric()
    tab2$pct<- tab2$N /sum(tab2$N)
    tab2$rel_n <- tab2$Num/neurite_num
    tab2$Cell<- names(df_list)[m]
  
    tab3 <- apply(length_mat, 1, function(x) sum(x < thres & x > -thres, na.rm = TRUE)) %>%  table() %>% as.data.table()
    tab3$act<- "pause"
    names(tab3)[1]<- "Num"
    tab3$Num <- tab3$Num %>% as.numeric()
    tab3$pct<- tab3$N/sum(tab3$N)
    tab3$rel_n <- tab3$Num/neurite_num
    tab3$Cell<- names(df_list)[m]
    
    tab4 <- rbind(tab1, tab2, tab3)
    tab4$Thres<- thres
    
    pos_neg_list[[m]] <- tab4
  }

  names(pos_neg_list)<- names(df_list)
  
  pos_neg_thres_list[[n]]<- assign(paste0("Thres_", thres, sep=""),pos_neg_list)
  names(pos_neg_thres_list)[n]<- paste0("Thres_", thres, sep="")
  
}
```

```{r message=FALSE, warning=FALSE}
summary_list <- list()

for (n in 1: 10){
  
  pos_neg_list <- pos_neg_thres_list[[n]]
  
  pos_neg_tab_all <- pos_neg_list[[1]] 
  
  for (m in 2:length(pos_neg_list)) {
    
    df2 <- pos_neg_list[[m]] 
    pos_neg_tab_all <- rbind(pos_neg_tab_all, df2) 
  
  }

  tab <- pos_neg_tab_all[act == "retraction"] %>% group_by(Num) %>% rstatix::get_summary_stats(pct)
  tab$Thres<- n*0.02
  summary_list [[n]] <- tab 

}

names(summary_list) <-  names(pos_neg_thres_list)

df_summary_all <- summary_list [[1]]

for (n in 2: 10){
  df_summary_all <-  rbind (df_summary_all, summary_list [[n]])
  
}

```

```{r  message = FALSE, warning=FALSE, fig.width=10}
ggplot(df_summary_all, aes(x=Num, y=mean, group=Thres)) +
  
  geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd, fill=Thres),   alpha=0.05) + 
  geom_line(aes(colour=Thres), size=1, alpha=1) +
  
  scale_y_continuous(  breaks = scales::pretty_breaks(n = 6))+
  scale_x_continuous(name="Neurite_No", limits=c(0, 6), expand = c(0, 0), breaks = c(0,1,2,3,4,5,6))+
  
  scale_fill_viridis(discrete = F) + # scale_fill_brewer(palette = "Set2") +
  scale_color_viridis(discrete = F)+

  theme_linedraw()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12), 
        legend.position = "none") #for centering the text
```

```{r}
ggsave(
  "retraction_freq.pdf",
  plot = last_plot(),
  device = "pdf")
```

```{r}
pos_neg_list <- pos_neg_thres_list$Thres_0.2

pos_neg_tab_all <- pos_neg_list[[1]] 

for (m in 2:length(pos_neg_list)) {
  
  df2 <- pos_neg_list[[m]] 
  pos_neg_tab_all <- rbind(pos_neg_tab_all, df2) 

}

tab <- pos_neg_tab_all[act == "extension"]

df_summary <- tab %>% group_by(Num) %>% rstatix::get_summary_stats(pct)

```

```{r  message = FALSE, warning=FALSE, fig.width=10}
# remove cells in "without any outgrowth" state more than 60% of the time 
removed <- pos_neg_tab_all[act == "extension" & Num == 0 & pct > 0.5] 
removed_cell<- removed$Cell
pos_neg_tab_all_removed<- pos_neg_tab_all[grep(paste(removed_cell,collapse="|"), pos_neg_tab_all$Cell , invert=TRUE) ,]
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
ggplot(data=tab, aes(x=Num, y=pct, group=Cell)) +
  geom_line((aes(color=Cell)))+
  geom_point((aes(color=Cell)))+ 
scale_x_continuous(name="Neurite_No", limits=c(0, 6), breaks = c(0,1,2,3,4,5,6))+
theme_linedraw()+
  
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12), 
        legend.position = "none") #for centering the text
```

```{r}
ggsave(
  "KO_neuriteNumFreq_avg.eps",
  plot = last_plot(),
  device = "eps")
```

## calculate Pos-Neg table from smoothened stationary (diff) profile
```{r  message = FALSE, warning=FALSE, fig.width=10}
mts_dim<- dim(selected_stationary)
pos_neg_df <- data.frame(matrix(ncol = mts_dim[2], nrow = mts_dim[1]))
colnames(pos_neg_df) <- selected_stationary %>% colnames()

for (n in 1:mts_dim[2])  {
    pos_neg_df[, n][selected_stationary[ , colnames(pos_neg_df)[n]] > thres] <- "Pos" # change threshold here
    pos_neg_df[, n][selected_stationary[ , colnames(pos_neg_df)[n]] < -thres] <- "Neg" # change threshold here
    }
  
#Count the growth and retraction events at each time point
pos_neg_df$Pos <- apply(pos_neg_df, 1, function(x) length(which(x=="Pos")))
pos_neg_df$Neg <- apply(pos_neg_df, 1, function(x) length(which(x=="Neg")))
pos_neg_df$na <- apply(pos_neg_df, 1, function(x) length(which(is.na(x))))

ggplot(pos_neg_df, aes(x=Pos)) + # change to "Neg" if retraction is of interest
 geom_histogram( colour="black", fill="white", binwidth=0.5)
```

## Extension retraction velocity analysis 
## We smoothen the stationary profiles first, separete outrgwoth and retraction events, and sum the changes of all neurites. (Data -> smoothing -> differencing -> Smoothening -> separate -> sum)
```{r warning=FALSE}

# Smoothen the stationary (diff) profiles and optionally filter out small length change with threshold

selected_list <- mts_stationary_list # mts_stationary_list or mts_avg_list

mts_stationary_pos_list <- list()
mts_stationary_neg_list <- list()
thres= 0.0 # change threshold here

# Positive peaks only (outgrowth event)
for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  length_mat<- selected_list[[m]][, selected_data ]
  
  length_mat_avg<- length_mat %>% forecast::ma(2) %>% as.ts() # smoothening 2
  colnames(length_mat_avg)<-colnames(length_mat)

  mts_dim<- dim(length_mat_avg)
    for (n in 1:mts_dim[2]){
      length_mat_avg[, n][length_mat_avg[ , colnames(length_mat_avg)[n]] < thres] <- 0 
    }
  
  length_mat_avg <- length_mat_avg %>% as.data.frame() 
  length_mat_avg$sum <- length_mat_avg  %>% rowSums()  # We sum up the extent of outgrowth at each time point
  mts_stationary_pos_list[[m]] <- length_mat_avg %>% ts(frequency = freq, start = c(0, 1)) 

}

# Negative peaks only (retraction event)
for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  length_mat<- selected_list[[m]][, selected_data ]
  length_mat_avg<- length_mat %>% forecast::ma(2) %>% as.ts() # smoothening 2
  colnames(length_mat_avg)<-colnames(length_mat)
  
  mts_dim<- dim(length_mat_avg)
  
    for (n in 1:mts_dim[2]){
      length_mat_avg[, n][length_mat_avg[ , colnames(length_mat_avg)[n]] > - thres] <- 0 
    }
  length_mat_avg <- length_mat_avg %>% as.data.frame() 
  length_mat_avg$sum <- length_mat_avg  %>% rowSums()  # We sum up the extent of retraction at each time point
  mts_stationary_neg_list[[m]] <- length_mat_avg %>% ts(frequency = freq, start = c(0, 1)) 
}

```

## extension and retraction lifetime (duration)
# Extension duration
```{r}
thres <- 0.00

list_1 <- list()

for (m in 1:length(mts_stationary_pos_list)){
  
  df <- mts_stationary_pos_list[[m]] %>% as.data.table()
  df <- df[, !"sum"] 
  list_2 <-  list()
  
   for (n in 1: ncol(df)){
     
    df2 <- df[ ,..n]
    df2$time <- seq(0, (nrow(df2)-1), 1)
    df2 <- df2[complete.cases(df2), ] 
    
    df2$inside <- df2[ , 1] > thres
    df2$start  <- ! c(FALSE, df2$inside[-nrow(df2)]) & df2$inside
    df2$end    <- df2$inside & ! c(df2$inside[-1], FALSE)  # Not used
    df2 <- df2[ df2$inside, ]
    df2$group  <- cumsum(df2$start)  # Number the spells
    colnames(df2)[1] <- "velocity"
    
    df3 <- plyr::ddply( df2, "group", summarize,
                        start  = min(time),
                        end    = max(time),
                        length = length(velocity),
                        mean   = mean(velocity))
   
    df3$neurite <- names(df)[n]
    list_2[[n]] <- df3

   }
  
  df3<- list_2 %>% data.table::rbindlist()
  df3$cell<- names(mts_stationary_list)[m]
  list_1[[m]] <- df3

}

extension_duration_df <- list_1 %>% data.table::rbindlist()
extension_duration_df$cond<- "WT"
extension_duration_df$growth<- "extension"
extension_duration_df[grep("KO", extension_duration_df$cell), ]$cond <- "KO"
```

# Retraction duration
```{r}

list_1 <- list()
for (m in 1:length(mts_stationary_neg_list)){
  
  df <- mts_stationary_neg_list[[m]] %>% as.data.table()
  df <- df[, !"sum"] 
  list_2 <-  list()
  
   for (n in 1: ncol(df)){
     
    df2 <- df[ ,..n]
    df2$time <- seq(0, (nrow(df2)-1), 1)
    df2 <- df2[complete.cases(df2), ] 
    
    df2$inside <- df2[ , 1] < -thres
    df2$start  <- ! c(FALSE, df2$inside[-nrow(df2)]) & df2$inside
    df2$end    <- df2$inside & ! c(df2$inside[-1], FALSE)  # Not used
    df2 <- df2[ df2$inside, ]
    df2$group  <- cumsum(df2$start)  # Number the spells
    colnames(df2)[1] <- "velocity"
    
    df3 <- plyr::ddply( df2, "group", summarize,
                        start  = min(time),
                        end    = max(time),
                        length = length(velocity),
                        mean   = mean(velocity))
   
    df3$neurite <- names(df)[n]
    list_2[[n]] <- df3

   }
  
  df3<- list_2 %>% data.table::rbindlist()
  df3$cell<- names(mts_stationary_list)[m]
  list_1[[m]] <- df3

}

retraction_duration_df <- list_1 %>% data.table::rbindlist()
retraction_duration_df$cond<- "WT"
retraction_duration_df$growth<- "retraction"
retraction_duration_df[grep("KO", retraction_duration_df$cell), ]$cond <- "KO"
```

# Data points at cell level
```{r}
duration_df <- rbind(extension_duration_df , retraction_duration_df )

duration_df$length <- duration_df$length/freq
df_summary_duration <-duration_df %>% 
 group_by(cell, cond, growth) %>% 
  rstatix::get_summary_stats(length) %>% as.data.table()

df_summary_duration$ID<- paste (df_summary_duration$cell, df_summary_duration$cond, df_summary_duration$growth,  sep = "_")
df_summary_duration$group<- paste (df_summary_duration$cond, df_summary_duration$growth, sep = "_")
```

```{r}
duration_df %>% group_by(cell, cond, growth, neurite) %>% dplyr::summarise(n = n())
```

# generate pairwise comparison list
```{r}
my_comparisons <- df_summary_duration$group %>% unique() %>% as.character() %>% combn(2, simplify = FALSE) 

my_comparisons <- list(c("WT_extension",  "KO_extension"), c("WT_retraction",  "KO_retraction"), c("WT_extension", "WT_retraction"), c("KO_extension", "KO_retraction"))
```

```{r message = FALSE, warning=FALSE, fig.width=10}
# my_comparisons <- list(c("axon", "minor"))
#pj <- position_jitter(seed = 1, width = .1, height = 0)

df_summary_duration  %>%  mutate(group = fct_relevel(group, "WT_extension",  "KO_extension", "WT_retraction",  "KO_retraction")) %>% 
ggplot ( aes(x=group, y=mean, fill =cond) )+ 
  scale_fill_brewer(palette="Dark2") + 
  theme_classic()+
geom_violin(aes(), trim=F,  kernel ="gaussian", adjust=1)  +
 geom_boxplot(aes(), width=0.2, outlier.colour = NA, fill="white")+ 
 geom_jitter(aes(), shape=16, position=position_jitter(0.1))+
  
  # geom_line(aes(group = Image))+
  # geom_point() +
  #geom_path(aes(group = cell),alpha = 0.5)+
    
 scale_y_continuous(expand = c(0, 0), limits = c(5,20), breaks = scales::pretty_breaks(n = 5))+ 
# geom_signif(comparisons = my_comparisons ,test = "wilcox.test",  color="black")+
 stat_compare_means(comparisons = my_comparisons, paired=F, method="wilcox.test", 
                     #label.y = c(7, 7.5, 8, 8.5, 9, 9.5, 10, 10.5), 
                     label = "{p.adj.format}{p.adj.signif}") +
  
  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold", angle = 90)) #for centering the text 
```

```{r}
ggsave(
  "DIV1_WT_KO_growth_duration_mean.pdf",
  plot = last_plot(),
  device = "pdf")
```

```{r}
outliers <- df_summary_duration %>% 
  group_by(cell,neurite, cond, growth) %>% 
 rstatix::identify_outliers(mean) 

df_summary_duration<- df_summary_duration[! (ID %in% outliers$ID)]

```

## extension and retraction velocity
```{r}
Ext_velocity <- list()
Ret_velocity <- list()

for (m in 1:length(mts_stationary_pos_list)){
  
    df <- mts_stationary_pos_list[[m]] %>% as.data.frame()
    df [ df == 0 ] <- NA # replace zero with NA to avoid zero inflation
    df1 <- colMeans(df, na.rm = T)  %>% as.data.frame() # calculate the average velocity through the observation window. Or matrixStats::colMedians
    colnames(df1) <- "mean_vel"
    df1$median_vel <- df %>% as.matrix() %>% matrixStats::colMedians(na.rm = T)  # calculate the average velocity through the observation window. Or matrixStats::colMedians
    
#df1$cond <- "KO"
df1$cell <- names(mts_list)[m] %>% basename() 
df1$Path <- names(mts_list)[m]
df1$neurite <- rownames(df1)
Ext_velocity [[m]] <- df1

df <- mts_stationary_neg_list[[m]] %>% as.data.frame()
 df [ df == 0 ] <- NA # replace zero with NA to avoid zero inflation
    df2 <- colMeans(df, na.rm = T)  %>% as.data.frame() # calculate the average velocity through the observation window. Or matrixStats::colMedians
    colnames(df2) <- "mean_vel"
    df2$median_vel <- df2 %>% as.matrix() %>% matrixStats::colMedians(na.rm = T)  # calculate the average velocity through the observation window. Or matrixStats::colMedians
    
#df2$cond <- "KO"
df2$cell <- names(mts_list)[m] %>% basename() 
df2$Path <- names(mts_list)[m]
df2$neurite <- rownames(df2)
Ret_velocity [[m]] <- df2
  
}
```

```{r}
Ext_velocity_df <- Ext_velocity %>% reduce(bind_rows)  
Ext_velocity_df$Growth <- "Extension"
Ret_velocity_df <- Ret_velocity %>% reduce(bind_rows)  
Ret_velocity_df$Growth <- "Retraction"
velocity_comb_df<- rbind(Ext_velocity_df, Ret_velocity_df) %>% as.data.table()
velocity_comb_df <- velocity_comb_df[neurite != "sum"]

velocity_comb_df$Cond<- "WT"
velocity_comb_df[grep("KO", velocity_comb_df$Path), ]$Cond <- "KO"

velocity_comb_df$mean_vel <- velocity_comb_df$mean_vel %>% abs()
velocity_comb_df$median_vel <- velocity_comb_df$median_vel %>% abs()

#saveRDS(velocity_comb_df, "velocity_comb_df_KO.rds")
```

# Data points at cell level
```{r}
df_summary_median_vel <-velocity_comb_df %>% 
 group_by(cell, Cond, Growth) %>% 
  rstatix::get_summary_stats(median_vel) %>% as.data.table()

df_summary_median_vel$ID<- paste (df_summary_median_vel$cell, df_summary_median_vel$Cond, df_summary_median_vel$Growth,  sep = "_")
df_summary_median_vel$group<- paste (df_summary_median_vel$Cond, df_summary_median_vel$Growth, sep = "_")
```

```{r}
outliers <- df_summary_median_vel %>% 
  group_by(Growth, Cond) %>% 
 rstatix::identify_outliers(mean) 

df_summary_median_vel <- df_summary_median_vel[! (ID %in% outliers$ID)]

```

```{r}
df_summary_mean_vel <-velocity_comb_df %>% 
 group_by(cell, Cond, Growth) %>% 
  rstatix::get_summary_stats(mean_vel) %>% as.data.table()

df_summary_mean_vel$ID <- paste (df_summary_mean_vel$cell, df_summary_mean_vel$Cond, df_summary_mean_vel$Growth,  sep = "_")
df_summary_mean_vel$group <- paste (df_summary_mean_vel$Cond, df_summary_mean_vel$Growth, sep = "_")
```

```{r}
outliers <- df_summary_mean_vel %>% 
   group_by(Growth, Cond) %>% 
 rstatix::identify_outliers(mean) 

df_summary_mean_vel <- df_summary_mean_vel[! (ID %in% outliers$ID)]
```

# generate pairwise comparison list
```{r}
my_comparisons <- df_summary_median_vel$group %>% unique() %>% as.character() %>% combn(2, simplify = FALSE) 

my_comparisons <- list(c("WT_Extension",  "KO_Extension"), c("WT_Retraction",  "KO_Retraction"), c("WT_Extension", "WT_Retraction"), c("KO_Extension", "KO_Retraction"))
```

```{r message = FALSE, warning=FALSE, fig.width=10}
# my_comparisons <- list(c("axon", "minor"))
#pj <- position_jitter(seed = 1, width = .1, height = 0)

df_summary_median_vel  %>%  mutate(group = fct_relevel(group, "WT_Extension",  "KO_Extension", "WT_Retraction",  "KO_Retraction")) %>% 
ggplot ( aes(x=group, y=mean, fill =Cond) )+ 
  scale_fill_brewer(palette="Dark2") + 
  theme_classic()+
geom_violin(aes(), trim=F,  kernel ="gaussian", adjust=1)  +
 geom_boxplot(aes(), width=0.2, outlier.colour = NA, fill="white")+ 
 # geom_jitter(aes(), shape=16, position=position_jitter(0.1), size=1)+
 # geom_line(aes(group = Image))+
  
   geom_point() +
  #geom_path(aes(group = cell),alpha = 0.5)+
    
 scale_y_continuous(expand = c(0, 0), limits = c(-0.1, 1), breaks = scales::pretty_breaks(n = 5))+ 
# geom_signif(comparisons = my_comparisons ,test = "wilcox.test",  color="black")+
  stat_compare_means(comparisons = my_comparisons, paired=F, method="wilcox.test", 
                     #label.y = c(1.15, 1.2, 1.25, 1.3, 1.35, 1.4), 
                     label = "{p.adj.format}{p.adj.signif}") +
  
    theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold", angle = 90)) #for centering the text 
```

```{r}
ggsave(
  "DIV1_WT_KO_velocity_median.pdf",
  plot = last_plot(),
  device = "pdf")
```

```{r message = FALSE, warning=FALSE, fig.width=10}
# my_comparisons <- list(c("axon", "minor"))
#pj <- position_jitter(seed = 1, width = .1, height = 0)

df_summary_mean_vel  %>%  mutate(group = fct_relevel(group, "WT_Extension",  "KO_Extension", "WT_Retraction",  "KO_Retraction")) %>% 
ggplot ( aes(x=group, y=mean, fill =Cond) )+ 
  scale_fill_brewer(palette="Dark2") + 
  theme_classic()+
geom_violin(aes(), trim=F,  kernel ="gaussian", adjust=1)  +
 geom_boxplot(aes(), width=0.2, outlier.colour = NA, fill="white")+ 
  geom_jitter(aes(), shape=16, position=position_jitter(0.1), size=1)+
 # geom_line(aes(group = Image))+
  
 #  geom_point() +
  #geom_path(aes(group = cell),alpha = 0.5)+
    
 scale_y_continuous(expand = c(0, 0), limits = c(-0.1, 1), breaks = scales::pretty_breaks(n = 5))+ 
# geom_signif(comparisons = my_comparisons ,test = "wilcox.test",  color="black")+
  stat_compare_means(comparisons = my_comparisons, paired=F, method="wilcox.test", 
                     #label.y = c(1.15, 1.2, 1.25, 1.3, 1.35, 1.4), 
                     label = "{p.adj.format}{p.adj.signif}") +
  
    theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold", angle = 90)) #for centering the text 

df_summary_mean_vel$group %>% table()
```

```{r}
ggsave(
  "DIV1_WT_KO_velocity_mean.pdf",
  plot = last_plot(),
  device = "pdf")
```

# Time-lagged cross-corrleation of outgrowth and retraction profiles
```{r warning=FALSE}

length_ccf_list_2 <- list()

for (m in 1:length(mts_stationary_list)) {
  
  df_pos <- mts_stationary_pos_list[[m]]
  df_neg <- mts_stationary_neg_list[[m]]

   length_ccf_plot <-ccf ( df_pos[, "sum"],  
                            - df_neg[, "sum"], 
                            type = "correlation", 
                            na.action = na.pass, plot=F,
                            lag.max=100) 
  
  length_ccf_list_2[[m]] <- length_ccf_plot 

}

names(length_ccf_list_2) <-  names(mts_stationary_list)
    
length_ccf_2 <- length_ccf_list_2[[1]][["acf"]] %>% as.data.frame()
for (m in 2:length(length_ccf_list_2)) {
  
  length_ccf_2 <- cbind(length_ccf_2 , length_ccf_list_2[[m]][["acf"]])
}

colnames(length_ccf_2) <- names(length_ccf_list_2)
length_ccf_2$lag<- c(-100:100)

```

```{r}
length_ccf_KO <- length_ccf_2[grepl('KO', colnames(length_ccf_2))]
length_ccf_WT <- length_ccf_2[!grepl('KO', colnames(length_ccf_2))]
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
selected_ccf <- length_ccf_WT %>% as.data.table()  
selected_ccf$lag<- length_ccf_2$lag/freq
stacked_ccf_mat <- gather(selected_ccf, Cell, ccf, -lag)
df_summary_WT <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf)
df_summary_WT$Cond <- "WT"

selected_ccf <- length_ccf_KO %>% as.data.table()  
selected_ccf$lag<- length_ccf_2$lag/freq
stacked_ccf_mat <- gather(selected_ccf, Cell, ccf, -lag)
df_summary_KO <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf)
df_summary_KO$Cond <- "KO"

df_summary <- rbind (df_summary_WT, df_summary_KO)
```

# We plot the overall outgrowth/retraction level over time
```{r  message = FALSE, warning=FALSE, fig.width=10}
selected_stationary_pos <- mts_stationary_pos_list[[115]][, "sum"] 
selected_stationary_neg <- mts_stationary_neg_list[[115]][, "sum"]

ts.union(selected_stationary_pos, selected_stationary_neg) %>%  autoplot() +
 geom_line(aes(),size=0.25)  +
   geom_hline(yintercept=0, linetype="dashed", color = "black")+
  
   #geom_hline(yintercept=thres, linetype="dashed", color = "blue")+
   #geom_hline(yintercept=-thres, linetype="dashed", color = "blue")+
  
   scale_y_continuous(expand = c(0, 0), limits = c(-3,3), breaks = scales::pretty_breaks(n = 7))+
  scale_x_continuous(expand = c(0, 0), limits = c(0, 480), breaks = scales::pretty_breaks(n = 7)) + 
  
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text
```

```{r}
ggsave(
  "KO_33_ext_ret_sta.eps",
  plot = last_plot(),
  device = "eps")
```

## Here we separate the outrgwoth and retraction events first, sum the changes of all neurites, and then do smoothening.  (Data -> smoothing -> differencing ->  separate -> sum -> smoothing) 
```{r}

selected_list <- mts_stationary_list 
mts_stationary_sum_list <- list()

thres= 0.0 # change threshold here

# Positive peaks only (outgrowth event)
for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]])) # Neurite length ("Length") only or include actin intensity (Integerated.intensity)
  length_mat_pos<- selected_list[[m]][, selected_data ]
  length_mat_neg<- selected_list[[m]][, selected_data ]
  
  mts_dim<- dim(length_mat_pos)
  
    for (n in 1:mts_dim[2]){
      length_mat_pos[, n][length_mat_pos[ , colnames(length_mat_pos)[n]] < 0] <- 0 
      length_mat_neg[, n][length_mat_neg[ , colnames(length_mat_neg)[n]] > 0] <- 0 
    }
  
  length_mat_pos <- as.data.frame(length_mat_pos)
  length_mat_neg <- as.data.frame(length_mat_neg)
   
  length_mat_pos$sum <- length_mat_pos  %>% rowSums()  
  length_mat_neg$sum <- length_mat_neg  %>% rowSums()  
  
  length_mat_pos <- length_mat_pos %>% ts(frequency = freq, start = c(0, 1))
  length_mat_neg <- length_mat_neg %>% ts(frequency = freq, start = c(0, 1))
  
  length_mat_sum <- ts.union(length_mat_pos[ ,"sum"], length_mat_neg[ ,"sum"]) 
  
  length_mat_sum <- length_mat_sum %>% forecast::ma(5) %>% as.ts() # # smoothening 2
  colnames(length_mat_sum)<-c("Pos", "Neg")
  
  length_mat_sum [, "Pos"][length_mat_sum [, "Pos"] < thres] <- 0 
  length_mat_sum [, "Neg"][length_mat_sum [, "Neg"] > -thres] <- 0 
  
  mts_stationary_sum_list[[m]] <- length_mat_sum

}

names(mts_stationary_sum_list)<- names(mts_stationary_list )

```

# Time-lagged cross-corrleation of outgrowth and retraction profiles
```{r warning=FALSE}

length_ccf_list <- list()

for (m in 1:length(mts_stationary_sum_list)) {
  
  selected_length_sta_sum <- mts_stationary_sum_list[[m]]
  
 # The lines give the values beyond which the autocorrelations are (statistically) significantly different from zero.

  length_ccf_plot <-ccf (selected_length_sta_sum[, "Pos"],  
                            -selected_length_sta_sum[, "Neg"], 
                            type = "correlation", 
                            na.action = na.pass, plot=F,
                            lag.max=100) 
  
  length_ccf_list[[m]] <- length_ccf_plot 
  
}

names(length_ccf_list) <-  names(mts_stationary_sum_list)

length_ccf <- length_ccf_list[[1]][["acf"]] %>% as.data.frame()
for (m in 2:length(length_ccf_list)) {
  
  length_ccf <- cbind(length_ccf , length_ccf_list[[m]][["acf"]])
}

colnames(length_ccf) <- names(length_ccf_list)

length_ccf$lag<- c(-100:100)

```

```{r}
length_ccf_KO <- length_ccf[grepl('KO', colnames(length_ccf))]
length_ccf_WT <- length_ccf[!grepl('KO', colnames(length_ccf))]

```

```{r  message = FALSE, warning=FALSE, fig.width=10}
selected_ccf <- length_ccf_WT %>% as.data.table()  
selected_ccf$lag<- length_ccf$lag/freq
stacked_ccf_mat <- gather(selected_ccf, Cell, ccf, -lag)
df_summary_WT <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf)
df_summary_WT$Cond <- "WT"

selected_ccf <- length_ccf_KO %>% as.data.table()  
selected_ccf$lag<- length_ccf$lag/freq
stacked_ccf_mat <- gather(selected_ccf, Cell, ccf, -lag)
df_summary_KO <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf)
df_summary_KO$Cond <- "KO"

df_summary <- rbind (df_summary_WT, df_summary_KO)
```

```{r}
ggsave(
  "DIV1_KO_ext_ret_corr.pdf",
  plot = last_plot(),
  device = "pdf")
```
# Time-lagged cross-corrleation of outgrowth and retraction profiles
```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
save.image(file=saved_path)

```

