---
title: "R Notebook"
output: html_notebook
---

<!--
# KIF5C1-560 Z-score intensity at extending / pausing / retracting neurite tips

## What this file does

Reads paired per-neurite tip-intensity CSVs (`Tip_actin*.csv`, `Tip_Kif5*.csv`
from per-cell subfolders; lines 68, 103, 128) for DIV-2 neurons co-expressing
Lifeact-mScarlet and KIF5C1-560-mNeonGreen. Each neurite's tip KIF5C1-560
intensity is converted to a per-neuron Z-score, classified by growth phase
(extending / pausing / retracting), and pooled across the cohort. Four PDFs
are written (`ggsave` at lines 644, 896, 1169, 1659), including the violin
plot of Z-scored KIF5C1-560 across the three growth phases.

## Manuscript panel

**ED Fig 3n**: "Normalized KIF5C1-560-mNeonGreen intensities as Z-scores at the
extending, the pausing and the retracting neurites of individual neurons.
N = 92 neurons from 3 experiments. ***p < 0.001; ns, p >= 0.05 by Two-tailed
paired Wilcoxon test with Holm correction."

Source Data: `ED Fig 3n`.

## Outputs

- Four PDFs (`ggsave` at lines 644, 896, 1169, 1659).

## Original G:\ source

G:/Figure S3_Arp3/Fig. S3n/Kif5C/Neurite activity corr.Rmd
-->

```{r message=FALSE, warning=FALSE, include=FALSE}
library(pracma)
#library(peakPick)
library(quantmod)
library(ggpmisc)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggfortify)
library(forecast)
library(tsibble)
library(broom)
library(tidyquant)
library(tsbox)
library("pheatmap")
library("RColorBrewer")
library(ggpmisc)
library(ggfortify)
library(viridis) # Default Color Maps from 'matplotlib' 
library(hrbrthemes)
library(data.table)
library(superheat)
```

## load existing/saved environment (.RData)
```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(Path_1)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
load(file=saved_path)
```

# save environment (.RData)
```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
save.image(file=saved_path)
```

## Read .csv files
```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
```

```{r}
pat_list <- list(Tip_actin="^Actin_tip(.*)csv$",  Neurite_actin= "^Actin_neurite(.*)csv$")
```

# Neurite length
```{r}
length_list<- list()
freq <- 0.5 # set frequency (1/interval)
pixel <- 0.325 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Tip_actin_(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Tip.point",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  colnames(df)<- paste("Length", seq( from = 1, to = ncol(df)  ),  sep = ".")
  
  length_list[[n]] <-df

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(length_list) <- gsub("Kymo_", "", subfolder_name_list)

length_list$DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif
```

# DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif and DIV2_LifeactmSC_Kif5mNG_03_37_R3D_D3D.dv_binned.tif are of the same neuron
```{r}
colnames(length_list$DIV2_LifeactmSC_Kif5mNG_03_37_R3D_D3D.dv_binned.tif ) <- c("Length.5")
length_list$DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif <-  cbind(length_list$DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif , length_list$DIV2_LifeactmSC_Kif5mNG_03_37_R3D_D3D.dv_binned.tif )
length_list <- length_list[-88]
```

# Actin intensity at neurite tip 
```{r}
Tip_actin_list<- list()

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Tip_actin(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, 1/10000)
  #df <-pracma::hampel(df, k=5)$y # remove jumping spikes
  
  colnames(df)<- paste("Tip_actin", seq( from = 1, to = dim(df)[2] ),  sep = ".")
  Tip_actin_list[[n]]<- df

}
names(Tip_actin_list) <- names(length_list)
```

# Kif5 intensity at neurite tip 
```{r}
Tip_Kif5_list<- list()

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Tip_Kif5(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  
  df <- df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, 1/10000)
  #df <-pracma::hampel(df, k=5)$y # remove jumping spikes
  
  colnames(df)<- paste("Tip_Kif5", seq( from = 1, to = dim(df)[2] ),  sep = ".")
  Tip_Kif5_list[[n]]<- df

}
names(Tip_Kif5_list) <- names(length_list)
```

# DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif and DIV2_LifeactmSC_Kif5mNG_03_37_R3D_D3D.dv_binned.tif are of the same neuron
```{r}
colnames(Tip_actin_list$DIV2_LifeactmSC_Kif5mNG_03_37_R3D_D3D.dv_binned.tif ) <- c("Tip_actin.5")
Tip_actin_list$DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif <-  cbind(Tip_actin_list$DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif , Tip_actin_list$DIV2_LifeactmSC_Kif5mNG_03_37_R3D_D3D.dv_binned.tif )
Tip_actin_list <- Tip_actin_list[-88]

colnames(Tip_Kif5_list$DIV2_LifeactmSC_Kif5mNG_03_37_R3D_D3D.dv_binned.tif ) <- c("Tip_Kif5.5")
Tip_Kif5_list$DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif <-  cbind(Tip_Kif5_list$DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif , Tip_Kif5_list$DIV2_LifeactmSC_Kif5mNG_03_37_R3D_D3D.dv_binned.tif )
Tip_Kif5_list <- Tip_Kif5_list[-88]
```

# Actin intensity at whole neurite
```{r}

Neurite_actin_list<- list()

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Neurite_actin(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, 1/10000)
  #df <-pracma::hampel(df, k=5)$y # remove jumping spikes
  colnames(df)<- paste("Neurite_actin", seq( from = 1, to = dim(df)[2] ),  sep = ".")
  
  Neurite_actin_list[[n]]<- df
 
}

names(Neurite_actin_list) <- gsub("Kymo_", "", subfolder_name_list)
```

# Kif5 intensity at whole neurite
```{r}

Neurite_Kif5_list<- list()

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Neurite_Kif5(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, 1/10000)
  #df <-pracma::hampel(df, k=5)$y # remove jumping spikes
  colnames(df)<- paste("Neurite_Kif5", seq( from = 1, to = dim(df)[2] ),  sep = ".")
  
  Neurite_Kif5_list[[n]]<- df
 
}

names(Neurite_Kif5_list) <- gsub("Kymo_", "", subfolder_name_list)
```

# DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif and DIV2_LifeactmSC_Kif5mNG_03_37_R3D_D3D.dv_binned.tif are of the same neuron
```{r}
colnames(Neurite_actin_list$DIV2_LifeactmSC_Kif5mNG_03_37_R3D_D3D.dv_binned.tif ) <- c("Neurite_actin.5")
Neurite_actin_list$DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif <-  cbind(Neurite_actin_list$DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif , Neurite_actin_list$DIV2_LifeactmSC_Kif5mNG_03_37_R3D_D3D.dv_binned.tif )
Neurite_actin_list <- Neurite_actin_list[-88]

colnames(Neurite_Kif5_list$DIV2_LifeactmSC_Kif5mNG_03_37_R3D_D3D.dv_binned.tif ) <- c("Neurite_Kif5.5")
Neurite_Kif5_list$DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif <-  cbind(Neurite_Kif5_list$DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif , Neurite_Kif5_list$DIV2_LifeactmSC_Kif5mNG_03_37_R3D_D3D.dv_binned.tif )
Neurite_Kif5_list <- Neurite_Kif5_list[-88]
```

# Actin density of the whole neurite ((tip + neurite shaft)/neurite length))
```{r}
Density_actin_list<- list()

for (n in 1:length(length_list))  {

  df <- (Neurite_actin_list[[n]])/(length_list[[n]])
  colnames(df)<- paste("density_actin", seq( from = 1, to = ncol(df) ),  sep = ".")
  df[sapply(df, is.infinite)] <- NA
  Density_actin_list[[n]]<- df
  
}

names(Density_actin_list) <- names(length_list)
```

# Kif5 density of the whole neurite ((tip + neurite shaft)/neurite length))
```{r}
Density_Kif5_list<- list()

for (n in 1:length(length_list))  {

  df <- (Neurite_Kif5_list[[n]])/(length_list[[n]])
  colnames(df)<- paste("density_Kif5", seq( from = 1, to = ncol(df) ),  sep = ".")
  df[sapply(df, is.infinite)] <- NA
  Density_Kif5_list[[n]]<- df
  
}

names(Density_Kif5_list) <- names(length_list)
```

# WT_DIV4_LAmScar_01Ctl_23paBlebb02_07_R3D.dv and WT_DIV4_LAmScar_01Ctl_23paBlebb02_06_R3D.dv are of the same neuron
```{r}
colnames(Density_actin_list$WT_DIV4_LAmScar_01Ctl_23paBlebb02_06_R3D.dv) <- c("density_actin.7", "density_actin.8")
Density_actin_list$WT_DIV4_LAmScar_01Ctl_23paBlebb02_07_R3D.dv <-  cbind(Density_actin_list$WT_DIV4_LAmScar_01Ctl_23paBlebb02_07_R3D.dv, Density_actin_list$WT_DIV4_LAmScar_01Ctl_23paBlebb02_06_R3D.dv)
Density_actin_list <- Density_actin_list[-117]
```

# DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif and DIV2_LifeactmSC_Kif5mNG_03_37_R3D_D3D.dv_binned.tif are of the same neuron
```{r}
colnames(Density_actin_list$DIV2_LifeactmSC_Kif5mNG_03_37_R3D_D3D.dv_binned.tif ) <- c("density_actin.5")
Density_actin_list$DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif <-  cbind(Density_actin_list$DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif , Density_actin_list$DIV2_LifeactmSC_Kif5mNG_03_37_R3D_D3D.dv_binned.tif )
Density_actin_list <- Density_actin_list[-88]

colnames(Density_Kif5_list$DIV2_LifeactmSC_Kif5mNG_03_37_R3D_D3D.dv_binned.tif ) <- c("density_Kif5.5")
Density_Kif5_list$DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif <-  cbind(Density_Kif5_list$DIV2_LifeactmSC_Kif5mNG_03_36_R3D_D3D.dv_binned.tif , Density_Kif5_list$DIV2_LifeactmSC_Kif5mNG_03_37_R3D_D3D.dv_binned.tif )
Density_Kif5_list <- Density_Kif5_list[-88]
```

```{r}
all_list <- list(length = length_list, 
                 tip_actin = Tip_actin_list,  
                 neurite_actin = Neurite_actin_list, 
                 Density_actin = Density_actin_list, 
                 tip_Kif5 = Tip_Kif5_list,  
                # tip_Kif5_frac = Tip_Kif5_frac_list, 
                 neurite_Kif5 = Neurite_Kif5_list, 
                 Density_Kif5 = Density_Kif5_list)

```

# filter for initial neurite length > 50 um
```{r}
length_50_list <- list()
tip_actin_50_list <- list()
neurite_actin_50_list <- list()
density_actin_50_list <- list()

tip_Kif5_50_list <- list()
neurite_Kif5_50_list <- list()
density_Kif5_50_list <- list()

selected_list <- length_list 

for (n in 1:length(selected_list)){
  
 df <- length_list[[n]] %>% as.data.table()
 selected_columns <- names(df)[df[1, ] > 50]
 length_50_list[[n]] <- df[,..selected_columns]
 
 df <- Tip_actin_list[[n]] %>% as.data.table()
 modified_vector <- gsub("Length", "Tip_actin",  selected_columns)
 tip_actin_50_list [[n]] <- df[, ..modified_vector]
 
 df <- Neurite_actin_list[[n]] %>% as.data.table()
 modified_vector <- gsub("Length", "Neurite_actin",  selected_columns)
 neurite_actin_50_list[[n]] <- df[,..modified_vector]
 
 df <- Density_actin_list[[n]] %>% as.data.table()
 modified_vector <- gsub("Length", "density_actin",  selected_columns)
 density_actin_50_list[[n]] <- df[,..modified_vector]
 
  df <- Tip_Kif5_list[[n]] %>% as.data.table()
 modified_vector <- gsub("Length", "Tip_Kif5",  selected_columns)
 tip_Kif5_50_list [[n]] <- df[, ..modified_vector]
 
 df <- Neurite_Kif5_list[[n]] %>% as.data.table()
 modified_vector <- gsub("Length", "Neurite_Kif5",  selected_columns)
 neurite_Kif5_50_list[[n]] <- df[,..modified_vector]
 
 df <- Density_Kif5_list[[n]] %>% as.data.table()
 modified_vector <- gsub("Length", "density_Kif5",  selected_columns)
 density_Kif5_50_list[[n]] <- df[,..modified_vector]
 
}
 
names(length_50_list) <- names(length_list) 
names(tip_actin_50_list) <- names(length_list) 
names(neurite_actin_50_list) <- names(length_list) 
names(density_actin_50_list) <- names(length_list) 

names(tip_Kif5_50_list) <- names(length_list) 
names(neurite_Kif5_50_list) <- names(length_list) 
names(density_Kif5_50_list) <- names(length_list) 
```

```{r}
length_50_list %>% discard(~ nrow(.x) == 0) -> length_50_list
tip_actin_50_list %>% discard(~ nrow(.x) == 0) -> tip_actin_50_list
neurite_actin_50_list %>% discard(~ nrow(.x) == 0) -> neurite_actin_50_list
density_actin_50_list %>% discard(~ nrow(.x) == 0) -> density_actin_50_list

tip_Kif5_50_list %>% discard(~ nrow(.x) == 0) -> tip_Kif5_50_list
neurite_Kif5_50_list %>% discard(~ nrow(.x) == 0) -> neurite_Kif5_50_list
density_Kif5_50_list %>% discard(~ nrow(.x) == 0) -> density_Kif5_50_list

all_50_list <- list(length = length_50_list, 
                 tip_actin = tip_actin_50_list,  
                 neurite_actin = neurite_actin_50_list, 
                 Density_actin = density_actin_50_list, 
                 tip_Kif5 = tip_Kif5_50_list,  
                 neurite_Kif5 = neurite_Kif5_50_list, 
                 Density_Kif5 = density_Kif5_50_list
                 
                 )
```

# Soma actin
```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
soma_list  <- subfolder_list [ grepl("Soma", subfolder_list) ]
somaActin_list<- list()

freq <- 0.5 # set frequency (1/interval)

# read csv files in the subfolders of subfolder list
for (n in 1:length(soma_list))  
{
  csv_list <- list.files(soma_list[n],pattern = "^soma_actin*(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  somaActin_list[[n]]  <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE)) %>% as.data.table()
}

soma_actin <- do.call("cbind", somaActin_list) %>% as.data.table()
```

# Soma Kif5
```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
soma_list  <- subfolder_list [ grepl("Soma", subfolder_list) ]
somaKif5_list<- list()

freq <- 0.5 # set frequency (1/interval)

# read csv files in the subfolders of subfolder list
for (n in 1:length(soma_list))  
{
  csv_list <- list.files(soma_list[n],pattern = "^soma_Kif5*(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  somaKif5_list[[n]]  <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE)) %>% as.data.table()
}

soma_Kif5 <- do.call("cbind", somaKif5_list) %>% as.data.table()
```

# Create stationary (differential) curve of all measurements
```{r}
all_sta_list <- list()
freq <- 0.5 # set frequency (1/interval)

for (m in 1:length(all_list)) {
  
  selected_list <- all_list[[m]]
  sta_list <- list()

  for (n in 1:length(selected_list))  {
 
  sta_list[[n]]<- selected_list[[n]] %>% 
    forecast::ma(5) %>%                                    # smoothing #1
    MTS::diffM(d=1) %>% 
    forecast::ma(3) %>% 
    ts(frequency = freq, start = c(0, 1))
  
  colnames(sta_list[[n]])<-colnames(selected_list[[n]])
  
  }
names(sta_list)<- names(selected_list)
all_sta_list[[m]]<- sta_list
}

names(all_sta_list)<- names(all_list )
```

# Create stationary (differential) curve of all measurements (for length filtered)
```{r}
all_50_sta_list <- list()

freq <- 0.5 # set frequency (1/interval)

for (m in 1:length(all_50_list)) {
  
  selected_list <- all_50_list[[m]]

  sta_list <- list()

  for (n in 1:length(selected_list))  {
 
  sta <- selected_list[[n]] %>% 
    forecast::ma(5) %>%                                    # smoothing #1
    MTS::diffM(d=1) %>% 
    forecast::ma(3) %>% 
    as.data.frame()
  
    colnames(sta) <- colnames(selected_list[[n]])
   sta_list [[n]] <- sta   %>%  ts(frequency = freq, start = c(0, 1))
  
  }

names(sta_list)<- names(selected_list)
  
all_50_sta_list[[m]]<- sta_list
  
}

names(all_50_sta_list)<- names(all_list )

```

# Prepare data for cross-correlation analysis at individual neurite level (for length filtered)
```{r}
all_50_sta_cell_list <- list()

for (k in 1:length(length_50_list)){
  
neurite_list <- list()
dt1 <- all_50_sta_list[[1]][[k]] %>% as.data.table()

    for (n in 1:ncol(dt1)){
      
     dt <- dt1[,..n]
          for (m in 2:7){
            
            dt2 <- all_50_sta_list[[m]][[k]] %>% as.data.table()
            dt<- cbind(dt, dt2[, ..n])
          }
    
    neurite_list[[n]] <- dt
    }

names(neurite_list)<- paste("Neurite_", 1:ncol(dt1), sep = "")
all_50_sta_cell_list[[k]]<- neurite_list

}
names(all_50_sta_cell_list) <- names(all_50_sta_list[[1]])
```

# subset the time series 
```{r}

for (m in 1:length(all_sta_list)) {
  
  selected_list <- all_sta_list[[m]]
  
  for (n in 1:length(selected_list)){
    
   selected_list[[n]] <- selected_list[[n]][1:150, ]
  }
  
  all_sta_list[[m]] <- selected_list
}

names(all_sta_list)<- names(all_list )

```

# Prepare data for cross-correlation analysis at individual neurite level
```{r}

library("data.table")

all_sta_cell_list <- list()

for (k in 1:length(length_list)){
  
neurite_list <- list()

dt1 <- all_sta_list[[1]][[k]] %>% as.data.table()
NC <- dt1 %>% ncol()

    for (n in 1:NC){
      
     dt <- dt1[,..n]
     
          for (m in 2:7){
            
            dt2 <- all_sta_list[[m]][[k]] %>% as.data.table()
            dt<- cbind(dt, dt2[, ..n])
              
          }
    
    neurite_list[[n]] <- dt
    
    }

names(neurite_list)<- paste("Neurite_", 1:NC, sep = "")

all_sta_cell_list[[k]]<- neurite_list

}

names(all_sta_cell_list) <- names(all_sta_list[[1]])

```

# cross-correlation with neurite length at individual neurite level
```{r warning=FALSE}
ccf_neurite_length_list <- list()
selected_list <- all_sta_cell_list # all_sta_cell_list or all_50_sta_cell_list
selected_list <- unlist(selected_list, recursive=FALSE)

for (n in 2:7){
  
    ccfPlot <- ccf (selected_list[[1]][ ,1],  
                      selected_list[[1]][ ,..n], 
                            type = "correlation", 
                            na.action = na.pass,
                            plot=F, 
                            lag.max=50) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf (selected_list[[m]][ ,1],  
                      selected_list[[m]][ ,..n], 
                            type = "correlation", 
                            na.action = na.pass,
                            plot=F, 
                            lag.max=50) 
  
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-50:50)
  ccf_neurite_length_list[[n-1]] <- ccf
}

names(ccf_neurite_length_list)<- c("Tip_actin",  "Neurite_actin",  "density_actin", "Tip_Kif5",  "Neurite_Kif5",  "density_Kif5")

```

# cross-correlation with neurite length at individual neurite level
```{r warning=FALSE}
ccf_neurite_TipActin_list <- list()
selected_list <- all_sta_cell_list # all_sta_cell_list or all_50_sta_cell_list
selected_list <- unlist(selected_list, recursive=FALSE)

for (n in 1:7){
  
    ccfPlot <- ccf (selected_list[[1]][ ,2],  
                      selected_list[[1]][ ,..n], 
                            type = "correlation", 
                            na.action = na.pass,
                            plot=F, 
                            lag.max=50) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf (selected_list[[m]][ ,2],  
                      selected_list[[m]][ ,..n], 
                            type = "correlation", 
                            na.action = na.pass,
                            plot=F, 
                            lag.max=50) 
  
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-50:50)
  ccf_neurite_TipActin_list[[n]] <- ccf
}

names(ccf_neurite_TipActin_list)<- c("Length", "Tip_actin",  "Neurite_actin",  "density_actin", "Tip_Kif5",  "Neurite_Kif5",  "density_Kif5")

```

```{r}
selected_ccf <- ccf_neurite_length_list$Tip_actin
#selected_ccf <- selected_ccf[ ,grep('DIV2', names(selected_ccf )) ] %>% as.data.table() # DIV1 or DIV2
selected_ccf$lag<- c(-50:50)/freq
stacked_ccf_mat <- selected_ccf %>% gather(Neurite, ccf, -lag) # select which CCF of measurements to show
df_summary <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf)
```

```{r}
ggsave(
  "Tip_actin_length.pdf",
  plot = last_plot(),
  device = "pdf")
```
# Here we separate the outrgwoth and retraction events first, sum the changes of all neurites, and then do smoothening. (separate -> sum -> smoothening)
```{r}
all_sta_PN_list <- list()

for (k in 1:length(all_sta_list)){
  
  selected_list <- all_sta_list[[k]]
  sta_PN_list <- list()
  
  thres= 0.0 # change threshold here
  
  for (m in 1:length(selected_list)) {
    
    DT <- selected_list[[m]] %>% as.data.frame()
   
    DT_pos<- DT
    DT_neg<- DT
  
    for (n in 1:dim(DT)[2]){
        DT_pos[, n][DT_pos[, n] < 0] <- 0 
        DT_neg[, n][DT_neg[, n] > 0] <- 0 
    }
    
    mat_sum <- data.frame(rowSums(DT_pos), -rowSums(DT_neg)) # add neg sign
    
   # mat_sum [, 1][mat_sum [, 1] < thres] <- 0 
   # mat_sum [, 2][mat_sum [, 2] > -thres] <- 0
    
    mat_sum <- mat_sum %>% forecast::ma(5)                    # smoothing #2
  
    colnames(mat_sum)<- c(paste("Pos_", names(all_list)[k], sep = ""), paste("Neg_", names(all_list)[k], sep = "")) # create column names
    sta_PN_list[[m]] <- mat_sum
  
  }
  
  names(sta_PN_list)<-  names(selected_list)
  
  all_sta_PN_list[[k]]<- sta_PN_list
  
}

names(all_sta_PN_list)<- names(all_list )
```

# Combine all stationary PN data of each videos, and add stationary soma data
```{r}
comb_sta_list<- list()

k <- all_sta_PN_list[[1]]%>% length()

for (m in 1:k){
  
  mat <- all_sta_PN_list[[1]][m] %>% as.data.frame()
  colnames(mat)<- sub('.*\\.', '', colnames(mat))
  
  for (n in 2:length(all_sta_PN_list)){
    
    mat2 <- all_sta_PN_list[[n]][m] %>% as.data.table()
    colnames(mat2)<- sub('.*\\.', '', colnames(mat2))
    
    mat <- data.frame(mat, mat2)
  }
  
  Dim <- mat %>% nrow()
  NM <- all_sta_PN_list[[1]][m] %>% names()
  mat3 <- mts_stationary_list[[NM]][1:Dim, c(2, 3)] %>% as.data.table() # soma data
  mat <- data.frame(mat, mat3)
  
 comb_sta_list[[m]]<- mat %>% ts(frequency = freq, start = c(0, 1)) 

}

names(comb_sta_list)<-   all_sta_PN_list[[1]] %>% names()
```

# cross-correlation of stationary PN data

```{r warning=FALSE}
selected_list <- comb_sta_list

ccf_PN_list<- vector(mode = "list", length = length(all_list)+1)

names(ccf_PN_list)<- c(names(all_list), "soma")

for (n in 1:5){
  
    ccfPlot <- ccf (selected_list[[1]][,(n*2-1)],  
                            selected_list[[1]][,(n*2)], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=50) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf (selected_list[[m]][,(n*2-1)],  
                            selected_list[[m]][,(n*2)], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=50) 
                            
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-50:50)
  ccf_PN_list[[n]] <- ccf
}

```

```{r warning=FALSE}
# cross-correlation with stationary Pos_length data

ccf_Poslen_list<- vector(mode = "list", length = dim(selected_list[[1]])[2]-2)

names(ccf_Poslen_list)<- colnames(selected_list[[1]])[-c(1,2)] 

for (n in 3:10){
  
    ccfPlot <- ccf (selected_list[[1]][,c("Pos_length")],  
                            selected_list[[1]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=50) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf (selected_list[[m]][,c("Pos_length")],  
                            selected_list[[m]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=50) 
  
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-50:50)
  ccf_Poslen_list[[n-2]] <- ccf
}
```

```{r warning=FALSE}
# cross-correlation with stationary Neg_length data

ccf_Neglen_list<- vector(mode = "list", length = dim(selected_list[[1]])[2]-2)

names(ccf_Neglen_list)<- colnames(selected_list[[1]])[-c(1,2)] 

for (n in 3:10){
  
    ccfPlot <- ccf (selected_list[[1]][,c("Neg_length")],  
                            selected_list[[1]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=50) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf (selected_list[[m]][,c("Neg_length")],  
                            selected_list[[m]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=50) 
  
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-50:50) # corresponds to "lag.max=50" set in ccf
  ccf_Neglen_list[[n-2]] <- ccf
}
```

```{r warning=FALSE}
# cross-correlation with soma actin

ccf_Actsoma_list<- vector(mode = "list", length = 8)

names(ccf_Actsoma_list)<- c(paste("Pos_", c("length", "tip_actin", "neurite_actin", "Density_actin"), sep=""), paste("Neg_", c("length","tip_actin", "neurite_actin", "Density_actin"), sep=""))

for (n in 1:8){
  
  df1 <- selected_list[[1]][ ,"soma_actin"]
  df2 <- selected_list[[1]][,grep(names(ccf_Actsoma_list)[n],  colnames(selected_list[[n]]))] 
  
    ccfPlot <- ccf (df1,  
                      df2, 
                      type = "correlation", 
                      na.action = na.pass,  plot=F, 
                      lag.max=50) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
     df1 <- selected_list[[m]][ ,"soma_actin"]
     df2 <- selected_list[[m]][,grep(names(ccf_Actsoma_list)[n],  colnames(selected_list[[n]]))]  
  
      ccfPlot_2 <- ccf (df1,  
                          df2, 
                          type = "correlation", 
                          na.action = na.pass,  plot=F, 
                          lag.max=50) 
      
      ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
      
      ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-50:50)
  ccf_Actsoma_list[[n]] <- ccf
}

ccf_all_list<-list(PN=ccf_PN_list, Poslen=ccf_Poslen_list, Neglen=ccf_Neglen_list, somaActin=ccf_Actsoma_list)

```

```{r}
selected_ccf <- ccf_all_list$PN$length

selected_ccf <-  selected_ccf [ , grep('DIV4', colnames(selected_ccf)) ]  %>% as.data.table()  # filter for DIV3 or DIV4 neuron only
  
selected_ccf$lag<- c(-50:50)/freq

stacked_ccf_mat <- selected_ccf %>% gather(Neurite, ccf, -lag) # select which CCF of measurements to show

df_summary <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf)
```

```{r}
ggsave(
  "PosNeg_velocity_corr_DIV4.eps",
  plot = last_plot(),
  device = "eps")
```
# analysis at whole-cell level
# Create time-series data of total neurite length, and actin intensity/density at neurite tip, shaft, and soma. 
```{r}
mts_list<- list()
mts_avg_list <- list()
mts_stationary_list <- list()

freq <- 0.5 # set frequency (1/interval)
pixel <- 0.325 # set pixel size

for (n in 1:length(length_list))  {
 
  mat <- length_list[[n]] %>% rowSums() %>% as.data.table() # total neurite length
  colnames(mat) <- "sum_length"
  
  # get soma intensity
  ColName <- names(Tip_actin_list)[n]

  Dim <- nrow(Tip_actin_list[[n]])
  mat$soma_actin <- soma_actin[1:Dim, ..ColName ] 
  mat$soma_Kif5 <- soma_Kif5[1:Dim, ..ColName ] 
  
  mat$sum_tip_actin <- Tip_actin_list[[n]] %>% rowSums()
  mat$sum_actin <-  Neurite_actin_list[[n]] %>% rowSums() # total actin intensity at the whole neurite
  mat$sum_neurite_actin <- (Neurite_actin_list[[n]] - Tip_actin_list[[n]]) %>% rowSums() 
  mat$sum_actin_density <- mat$sum_actin/mat$sum_length # actin density at the whole neurite

    mat$sum_tip_Kif5 <- Tip_Kif5_list[[n]] %>% rowSums()
  mat$sum_Kif5 <-  Neurite_Kif5_list[[n]] %>% rowSums() # total Kif5 intensity at the whole neurite
  mat$sum_neurite_Kif5 <- (Neurite_Kif5_list[[n]] - Tip_Kif5_list[[n]]) %>% rowSums() 
  mat$sum_Kif5_density <- mat$sum_Kif5/mat$sum_length # Kif5 density at the whole neurite

  mts_list[[n]]<- mat %>% ts(frequency = freq, start = c(0, 1)) 
  
  mts_avg_list[[n]] <- mts_list[[n]] %>% forecast::ma(5) %>% as.ts() # Moving Average #1
  colnames(mts_avg_list[[n]])<-colnames(mts_list[[n]])
  
  mts_stationary_list[[n]] <- mts_avg_list[[n]] %>% 
    MTS::diffM(d=1) %>% 
    forecast::ma(5) %>%                                             # Moving Average #2
    ts(frequency = freq, start = c(0, 1)) 
  
  colnames(mts_stationary_list[[n]])<-colnames(mts_list[[n]])
  
}

names(mts_list)<- names(Tip_actin_list)
names(mts_avg_list)<- names(Tip_actin_list)
names(mts_stationary_list)<- names(Tip_actin_list)
```

## We smoothen the stationary profiles first, separete outrgwoth and retraction events, and sum the changes of all neurites. (Smoothening -> separate -> sum)
```{r warning=FALSE}

# Smoothen the stationary (diff) profiles and optionally filter out small length change with threshold

selected_list <- all_sta_list$length 

mts_stationary_pos_list <- list()
mts_stationary_neg_list <- list()
thres= 0.0 # change threshold here

# Positive peaks only (outgrowth event)
for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  length_mat<- selected_list[[m]][, selected_data ]
  
  length_mat_avg<- length_mat %>% forecast::ma(5) %>% as.ts() # smoothening 1
  colnames(length_mat_avg)<-colnames(length_mat)

  mts_dim<- dim(length_mat_avg)
    for (n in 1:mts_dim[2]){
      length_mat_avg[, n][length_mat_avg[ , colnames(length_mat_avg)[n]] < thres] <- 0 
    }
  
  length_mat_avg <- length_mat_avg %>% as.data.frame() 
  length_mat_avg$sum <- length_mat_avg  %>% rowSums()  # We sum up the extent of outgrowth at each time point
  mts_stationary_pos_list[[m]] <- length_mat_avg %>% ts(frequency = freq, start = c(0, 1)) 

}

# Negative peaks only (retraction event)
for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  length_mat<- selected_list[[m]][, selected_data ]
  length_mat_avg<- length_mat %>% forecast::ma(5) %>% as.ts() # smoothening 1
  colnames(length_mat_avg)<-colnames(length_mat)
  
  mts_dim<- dim(length_mat_avg)
    for (n in 1:mts_dim[2]){
      length_mat_avg[, n][length_mat_avg[ , colnames(length_mat_avg)[n]] > - thres] <- 0 
    }
  length_mat_avg <- length_mat_avg %>% as.data.frame() 
  length_mat_avg$sum <- length_mat_avg  %>% rowSums()  # We sum up the extent of retraction at each time point
  mts_stationary_neg_list[[m]] <- length_mat_avg %>% ts(frequency = freq, start = c(0, 1)) 
}

```

# extension and retraction velocity
```{r}
Ext_velocity_WT <- list()
Ret_velocity_WT <- list()

for (m in 1:length(mts_stationary_pos_list)){
  
df1 <- mts_stationary_pos_list[[m]] %>% as.data.frame()
df1[ df1 == 0 ] <- NA
df1 <- colMeans(df1, na.rm = T)  %>% as.data.frame() # calculate the average velocity through the observation window
df1$cond <- "WT"
df1$cell <- names(mts_list)[m] %>% basename() 
df1$neurite <- rownames(df1)
df1 <- df1 %>% subset (df1$neurite !="sum")
df1$length <- length_list[[m]][1,] %>% as.data.table() # add the neurite length at the first time point for later axon-minor neurite separation
Ext_velocity_WT [[m]] <- df1

df2 <- mts_stationary_neg_list[[m]] %>% as.data.frame()
df2[ df2 == 0 ] <- NA
df2 <- colMeans(df2, na.rm = T)  %>% as.data.frame() %>% abs() # calculate the average velocity through the observation window
df2$cond <- "WT"
df2$cell <- names(mts_list)[m] %>% basename() 
df2$neurite <- rownames(df2)
df2 <- df2 %>% subset (df2$neurite !="sum")
df2$length <- length_list[[m]][1,] %>% as.data.table()
Ret_velocity_WT [[m]] <- df2
  
}
```

```{r}
Ext_velocity_WT_df <- Ext_velocity_WT %>% reduce(bind_rows)  %>% as.data.frame()
Ext_velocity_WT_df$Growth <- "Extension"
Ret_velocity_WT_df <- Ret_velocity_WT %>% reduce(bind_rows)  %>% as.data.frame()
Ret_velocity_WT_df$Growth <- "Retraction"

velocity_comb_df<- rbind(Ext_velocity_WT_df, Ret_velocity_WT_df) %>% as.data.table()

names(velocity_comb_df)[1]<- "Velocity"

#saveRDS(velocity_comb_df, "velocity_comb_df")
```

```{r}
velocity_comb_df_axon <- velocity_comb_df[length >50]
velocity_comb_df_minor <- velocity_comb_df[length <50]
```

# Scatter plot for neurite length vs velocity
```{r}
velocity_comb_df[Growth=="Extension"] %>% ggplot(aes(x=Velocity, y=length) ) +
  scale_x_continuous( expand = c(0, 0),  limits = c(-0.1, 2),  breaks = scales::pretty_breaks(n = 5)) +
  geom_point()
```

# Data points at cell level (axon + minor neurites)
```{r}

summary_WT_ext <-velocity_comb_df[cond == "WT" & Growth == "Extension"] %>% 
  group_by(cell) %>% 
  rstatix::get_summary_stats(Velocity)

summary_WT_ext$cond <- "WT_ext"

summary_WT_ret <-velocity_comb_df[cond == "WT" & Growth == "Retraction"] %>% 
  group_by(cell) %>% 
  rstatix::get_summary_stats(Velocity)

summary_WT_ret$cond <- "WT_ret"

df_summary <- rbind(summary_WT_ext, summary_WT_ret)
```

# Data points at cell level (axon only)
```{r}

summary_WT_ext <-velocity_comb_df_axon[cond == "WT" & Growth == "Extension"] %>% 
  group_by(cell) %>% 
  rstatix::get_summary_stats(Velocity)

summary_WT_ext$cond <- "WT_ext"

summary_WT_ret <-velocity_comb_df_axon [cond == "WT" & Growth == "Retraction"] %>% 
  group_by(cell) %>% 
  rstatix::get_summary_stats(Velocity)

summary_WT_ret$cond <- "WT_ret"

df_summary_axon <- rbind(summary_WT_ext, summary_WT_ret) %>% as.data.table()
df_summary_axon$axon <-"axon"

# remove outliers
WT_ext_axon <- df_summary_axon[cond=="WT_ext"]
outliers <- WT_ext_axon %>% rstatix::identify_outliers(mean) 
WT_ext_axon <-  WT_ext_axon[ !(cell %in% outliers$cell)]

WT_ret_axon <- df_summary_axon[cond=="WT_ret"]
outliers <- WT_ret_axon %>% rstatix::identify_outliers(mean) 
WT_ret_axon <-  WT_ret_axon[ !(cell %in% outliers$cell)]
```

# Data points at cell level (minor neurites only)
```{r}
summary_WT_ext <-velocity_comb_df_minor[cond == "WT" & Growth == "Extension"] %>% 
  group_by(cell) %>% 
  rstatix::get_summary_stats(Velocity)

summary_WT_ext$cond <- "WT_ext"

summary_WT_ret <-velocity_comb_df_minor[cond == "WT" & Growth == "Retraction"] %>% 
  group_by(cell) %>% 
  rstatix::get_summary_stats(Velocity)

summary_WT_ret$cond <- "WT_ret"

df_summary_minor <- rbind(summary_WT_ext, summary_WT_ret) %>% as.data.table()
df_summary_minor$axon <-"minor"

# remove outliers
WT_ext_minor <- df_summary_minor[cond=="WT_ext"]
outliers <- WT_ext_minor %>% rstatix::identify_outliers(mean) 
WT_ext_minor <-  WT_ext_minor[ !(cell %in% outliers$cell)]

WT_ret_minor <- df_summary_minor[cond=="WT_ret"]
outliers <- WT_ret_minor %>% rstatix::identify_outliers(mean) 
WT_ret_minor <-  WT_ret_minor[ !(cell %in% outliers$cell)]
```

```{r}
df_summary_axon_minor <- rbind(WT_ext_axon, WT_ret_axon, WT_ext_minor, WT_ret_minor )
```

```{r}
df_summary_axon_minor$group <- paste(df_summary_axon_minor$cond, df_summary_axon_minor$axon, sep = "_")
df <-df_summary_axon_minor$group %>% table() %>% as.data.frame()
my_comparisons <- df$. %>% as.character() %>% combn(2, simplify = FALSE) 
```

```{r message = FALSE, warning=FALSE, fig.width=8}
library(easyGgplot2)
library(ggpubr)

df_summary_axon_minor[grep('DIV2', df_summary_axon_minor$cell) ] %>% # filter for DIV3 or DIV4 neuron only
  mutate(group = fct_relevel(group, "WT_ext_axon", "WT_ext_minor", "WT_ret_axon", "WT_ret_minor")) %>% 
  ggplot2.violinplot(yName="mean", xName='group', groupName='group', 
                     addDot=TRUE, dotSize=1,dotPosition="jitter", jitter=0.2, 
                     trim=F,  kernel ="gaussian", adjust=0.5)+
    
  geom_boxplot(width=0.1, fill="white")  +
 # rotate_x_text(angle = 45)+  
  scale_y_continuous( expand = c(0, 0),  limits = c(-0.1, 1.5),  breaks = scales::pretty_breaks(n = 5)) +
  stat_compare_means(comparisons = my_comparisons, paired=F, method="wilcox.test", label.y = c(1.15, 1.2, 1.25, 1.3, 1.35, 1.4), label = "{p.adj.format}{p.adj.signif}") +
  
  #geom_hline(yintercept = mean(df_summary$median), linetype = 2)+ # Add horizontal line at base mean
  #stat_compare_means(label.y = 0.45) +        # Add global annova p-value
  #stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = ".all.") +

  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold"))
   
```

```{r}
ggsave(
  "Axon_minor_velocities_DIV3.eps",
  plot = last_plot(),
  device = "eps")

```

# cross-correlation with sum neurite length
```{r warning=FALSE}

selected_list <- mts_stationary_list

ccf_sum_list<- vector(mode = "list", length = ncol(selected_list[[1]])-1)

names(ccf_sum_list)<- colnames(selected_list[[1]])[-1]

for (n in 2:11){
  
    ccfPlot <- ccf(selected_list[[1]][,1],  
                            selected_list[[1]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=80) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf(selected_list[[m]][,1],  
                            selected_list[[m]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=80) 
  
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-80:80)
  ccf_sum_list[[n-1]] <- ccf
}
```

# cross-correlation with soma actin
```{r warning=FALSE}

selected_list <-mts_stationary_list

ccf_sumActsoma_list<- vector(mode = "list", length = ncol(selected_list[[1]]))

names(ccf_sumActsoma_list)<- colnames(selected_list[[1]])

for (n in 1:11){
  
    ccfPlot <- ccf(selected_list[[1]][,2],  
                            selected_list[[1]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=80) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf(selected_list[[m]][,2],  
                            selected_list[[m]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=80) 
  
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-80:80)
  ccf_sumActsoma_list[[n]] <- ccf
}
```

# cross-correlation with soma Kif5
```{r warning=FALSE}

selected_list <-mts_stationary_list

ccf_sumKif5soma_list<- vector(mode = "list", length = ncol(selected_list[[1]]))

names(ccf_sumKif5soma_list)<- colnames(selected_list[[1]])

for (n in 1:11){
  
    ccfPlot <- ccf(selected_list[[1]][,2],  
                            selected_list[[1]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=80) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf(selected_list[[m]][,2],  
                            selected_list[[m]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=80) 
  
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-80:80)
  ccf_sumKif5soma_list[[n]] <- ccf
}
```

```{r}
ccf_sum_all_list<-list(sumlength=ccf_sum_list,  somaActin=ccf_sumActsoma_list, somaKif5=ccf_sumKif5soma_list)

```

```{r}
selected_ccf <- ccf_sum_all_list$sumlength$soma_actin
#selected_ccf <- selected_ccf[ ,grep('DIV4', names(selected_ccf )) ] %>% as.data.table() # DIV3 or DIV4
selected_ccf$lag<- c(-80:80)/freq
stacked_ccf_mat <- selected_ccf %>% gather(Neurite, ccf, -lag) 
df_summary <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf, type = "common")
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
ggplot(df_summary, aes(x=lag, y=mean)) +
   geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd), fill="lightgrey")+
   geom_line(size=1, alpha=1) +
  
   stat_peaks(colour = "red", span = 15, ignore_threshold = 0.5) +
  stat_peaks(geom = "text", colour = "red", span = 15, ignore_threshold = 0.6, hjust = 2, angle = 0) +
  stat_valleys(colour = "lightblue", span = 13, ignore_threshold = 0.9) +
 stat_valleys(geom = "text", colour = "blue", span = 13, ignore_threshold = 0.9, hjust = -0.5, angle = 0) +
  
   theme_ipsum(grid = F, ticks = T)+ 
   geom_vline(xintercept=0, linetype="dashed", color = "black")+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
   scale_y_continuous(expand = c(0, 0), limits = c(-0.5 ,0.8),  breaks = scales::pretty_breaks(n = 6))+
   scale_x_continuous(expand = c(0, 0),  breaks = scales::pretty_breaks(n = 8), limits = c(-100,100)) +
  
  theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text
```
#################################################################################################################################
# Analysis with Jacobson 2006 approach
```{r}
# change of fraction of total protein intensity

Tip_Kif5_frac_list <- list()

for (n in 1:length(Tip_Kif5_list)){
  
  df <- Tip_Kif5_list[[n]]
  
  Tip_Kif5_frac_list[[n]] <- df/rowSums(df, na.rm = T)
}

names(Tip_Kif5_frac_list) <- names(Tip_Kif5_list)

```

```{r}
all_list_2 <- list(length = length_list, 
                 tip_Kif5 = Tip_Kif5_list)

```

# downsampling the time series data from frame/2 min to frame/30 min and the ncreate stationary time-series data
```{r}
# Create stationary (differential) curve of all measurements

all_sta_downsampling_list <- list()
freq <- 0.5 # set frequency (1/interval)

for (m in 1:length(all_list_2)) {
  
  selected_list <- all_list_2[[m]]
  sta_list <- list()

  for (n in 1:length(selected_list))  {
    
    df <- selected_list[[n]] 
    
    row_num <- nrow(df)
    #df <- df[seq(0, row_num , by=15), ] # define the subsampling freq
     df <- df[1:200, ] 
    new_freq <- freq/15 # define the subsampling freq
    
    df  <- df %>% 
    forecast::ma(5) %>%                                    # smoothing #1
    MTS::diffM(d=1) %>% 
    #forecast::ma(3) %>% 
    ts(frequency = freq, start = c(0, 1)) # or new_freq

    colnames(df) <- colnames(selected_list[[n]])
    sta_list[[n]] <- df
 
  }
names(sta_list)<- names(selected_list)
all_sta_downsampling_list[[m]]<- sta_list
}

names(all_sta_downsampling_list)<- names(all_list_2)
```

# Prepare data for cross-correlation analysis at individual neurite level
```{r}
library("data.table")

all_sta_cell_downsampling_list <- list()

for (k in 1:length(length_list)){
  
neurite_list <- list()

dt1 <- all_sta_downsampling_list[[1]][[k]] %>% as.data.table()
NC <- dt1 %>% ncol()

    for (n in 1:NC){
      
     dt <- dt1[,..n]
   
     dt2 <- all_sta_downsampling_list[[2]][[k]] %>% as.data.table()
     dt<- cbind(dt, dt2[, ..n])
     names(dt) <- c("Length", "Kif5_frac")
     dt$Neurite <-  paste("Neurite_", n,  sep = "")
     dt$Cell <- names(all_sta_list[[1]]) [k]
    
     neurite_list[[n]] <- dt
    }

all_sta_cell_downsampling_list[[k]]<- neurite_list %>% reduce(bind_rows)
}

names(all_sta_cell_downsampling_list) <- names(all_sta_list[[1]])
```

```{r}
all_sta_cell_downsampling_df <- all_sta_cell_downsampling_list %>% reduce(bind_rows) %>% as.data.table()
```

```{r}
  library(easyGgplot2)

all_sta_cell_downsampling_df %>% 
  ggplot2.violinplot(yName="Kif5_frac",  addDot=TRUE, dotSize=2,
      dotPosition="jitter", jitter=0.2, trim=F,  kernel ="gaussian", adjust=2)+
  geom_boxplot(width=0.1, fill="white")  
```

```{r}
all_sta_cell_downsampling_df %>% 
ggplot(aes(x = Length, y = Kif5_frac))  +
  geom_point(alpha=0.05 )+
    #scale_y_continuous(expand = c(0, 0), limits = c(-0.05 ,0.05),  breaks = scales::pretty_breaks(n = 4))+
   scale_x_continuous(expand = c(0, 0),  breaks = scales::pretty_breaks(n = 8), limits = c(-2,2)) +
  #stat_density_2d(aes(fill = after_stat(level)), geom="polygon")+
   geom_smooth(method=glm, color="black")
```
# cross-correlation with neurite length at individual neurite level
```{r warning=FALSE}
ccf_neurite_length_list <- list()
selected_list <- all_sta_cell_downsampling_list 
selected_list <- unlist(selected_list, recursive=FALSE) # all_sta_cell_list or all_50_sta_cell_list

for (n in 2:8){
  
    ccfPlot <- ccf (selected_list[[1]][ ,1],  
                      selected_list[[1]][ ,..n], 
                            type = "correlation", 
                            na.action = na.pass,
                            plot=F, 
                            lag.max=13) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf (selected_list[[m]][ ,1],  
                      selected_list[[m]][ ,..n], 
                            type = "correlation", 
                            na.action = na.pass,
                            plot=F, 
                            lag.max=13) 
  
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-13:13)
  ccf_neurite_length_list[[n-1]] <- ccf
}

names(ccf_neurite_length_list)<- c("Tip_actin",  "Neurite_actin",  "density_actin", "Tip_Kif5",  "Tip_Kif5_frac", "Neurite_Kif5",  "density_Kif5")

```

```{r}
selected_ccf <- ccf_neurite_length_list$Tip_Kif5_frac
#selected_ccf <- selected_ccf[ ,grep('DIV2', names(selected_ccf )) ] %>% as.data.table() # DIV1 or DIV2
selected_ccf$lag<- c(-13:13)/(freq/15)
stacked_ccf_mat <- selected_ccf %>% gather(Neurite, ccf, -lag) # select which CCF of measurements to show
df_summary <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf)
```

```{r  message = FALSE, warning=FALSE, fig.width=8}
ggplot(df_summary, aes(x=lag, y=mean)) +

  geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd), fill="lightgrey", alpha=1)+
  geom_line(size=1) +
  
  geom_vline(xintercept=0, linetype="dashed", color = "black")+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  
  stat_peaks(colour = "red", ignore_threshold = 0.9) +
  stat_peaks(geom = "text", colour = "red", vjust = -0.5, ignore_threshold = 0.9) +
  
  stat_valleys(colour = "blue", ignore_threshold = 0.9, span = 11)+
  stat_valleys(geom = "text", colour = "blue", angle = 45, vjust = 1.5, hjust = 1, ignore_threshold = 0.9, span = 11)+
  
  scale_y_continuous(expand = c(0, 0), limits = c(-0.4, 0.8), breaks = scales::pretty_breaks(n = 7))+
  scale_x_continuous(expand = c(0, 0), limits = c(-150, 150), breaks = scales::pretty_breaks(n = 8)) + 
  
  theme_ipsum(grid = F, ticks = T)+ 
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text
```

############################################################################################
# extract neurite growth velocity dataframe and add sample labels
```{r}
Pos_neg_list <- list()

selected_list <- all_sta_cell_list # all_sta_cell_downsampling_list

for (n in 1:length(all_sta_cell_list)){
  
  selected_cell <- all_sta_cell_list[[n]]
  #selected_cell_Arp3 <- all_list$tip_Arp3[[n]]
  
  df_1 <-selected_cell[[1]][,1] # select the first 50 frames only to reduce photobleaching at later time points ([1:100, ])
  df_1$Neurite <- names(df_1)
  names(df_1)[1]<-"Velocity"
  
  df_1$Cell <- names(all_sta_cell_list)[n]
  
  for (m in 2:length(selected_cell)){
    
  df_2 <- selected_cell[[m]][,1] # select the first 50 frames only to reduce photobleaching at later time points
 
  df_2$Neurite <- names(df_2)
  names(df_2)[1]<-"Velocity"
  df_2$Cell <- names(all_sta_cell_list)[n]
  
  df_1 <- rbind(df_1, df_2)
    
  }
  
  Pos_neg_list[[n]]<-df_1
  
}
```

# add extension-retraction (Positive-Negative) labels to velocity dataframe
```{r}
Pos_neg_label_list<- list()
threshold <- 0.4 # setup the length change threshold to define extension and retraction

for (n in 1:length(Pos_neg_list)){
  
  df<- Pos_neg_list[[n]]
  df$labels <- ifelse(df$Velocity> threshold, "Positive",
                    ifelse(df$Velocity < -threshold, "Negative",
                           "Pause"))
  
  Pos_neg_label_list [[n]]<- df
  
}
```

# extract selected protein intensity dataframe and add sample labels
```{r}
selected_list <- all_list$tip_Kif5

tip_Kif5CInt_list <- list()

for (n in 1:length(selected_list)){
  
  selected_cell <- selected_list [[n]] %>% as.data.frame()
  
  df_1 <-selected_cell[,1] %>% as.data.frame()
  df_1$Neurite <- names(selected_cell)[1]
  names(df_1)[1]<-"Intensity"
  df_1$Cell <- names(selected_list)[n]
  df_1 <- df_1[-1, ]
 
  for (m in 2:length(selected_cell)){
    
  df_2 <-selected_cell[,m]  %>% as.data.frame()
  df_2$Neurite <- names(selected_cell)[m]
  names(df_2)[1]<-"Intensity"
  df_2$Cell <- names(selected_list)[n]
  df_2 <- df_2[-1, ]
  
  df_1 <- rbind(df_1, df_2)  
  }
  
  tip_Kif5CInt_list[[n]]<-df_1 %>% as.data.frame()
  
}
```

# calculate average intensity at extending and retractin neurites in each neuron, and get the Z-score
```{r}
summary_list <- list()

for (n in 1: length(tip_Kif5CInt_list)){
  
  df <- tip_Kif5CInt_list[[n]]
  df$labels <- Pos_neg_label_list[[n]]$labels 
  
  df <- df %>% group_by(labels) %>% rstatix::get_summary_stats(Intensity) %>% as.data.table()
  df <- df[labels == "Negative" | labels == "Positive" | labels == "Pause"] # | labels == "Pause"
  df$Zscore <- (df$median-mean(df$median))/sd(df$median) # mean or median
  df$Cell<- names(all_list$tip_Kif5)[n]
  
  summary_list[[n]] <- df
}

names(summary_list) <- names(all_list$tip_Kif5)

summary_df<- summary_list %>% reduce(bind_rows) %>% as.data.table()

# summary_df<- summary_df[labels == "Negative" | labels == "Positive"]

```

```{r, message = FALSE, warning=FALSE, fig.width=8}
library("ggpubr")
my_comparisons <- list( c("Positive", "Negative"), c("Positive", "Pause"), c("Negative", "Pause"))

ggplot(summary_df[sd < 30] , aes(x=labels, y=Zscore, fill=labels)) + 
  geom_violin(trim=FALSE, 
              kernel ="gaussian", 
              adjust=0.8)+
  geom_boxplot(width=0.2, fill="white")  +
  geom_point(aes(fill=labels,group=Cell), position = position_dodge(0.2))+
  geom_line(aes(group=Cell), position = position_dodge(0.2)) +
  stat_compare_means(comparisons = my_comparisons, paired=T, method="wilcox.test")+
 # scale_y_continuous( expand = c(0, 0), limits = c(-0.2, 1.2), breaks = scales::pretty_breaks(n = 7))+
  
  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
  
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold")) #for centering the text 

# plotly::ggplotly(p)
```
```{r}
ggsave(
  "Tip_Kif5C_extension_retraction.eps",
  plot = last_plot(),
  device = "eps")

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
