---
title: "R Notebook"
output: html_notebook
---

<!--
# Actin-filament orientation distribution from cryo-EM tomograms (WT vs Arp3 KO)

## What this file does

Reads per-tomogram actin-filament orientation tables (angles relative to the
leading edge, exported from the segmentation + tracing pipeline; data pasted
into in-line tibbles in this Rmd). For each tomogram an orientation
histogram (or rose-density plot) is computed; orientations are then pooled
across the cohort (5 WT, 8 Arp3 KO tomograms) into the mean +/- SD curve.
Two PDFs are written (`ggsave` at lines 233, 249).

## Manuscript panel

**ED Fig 10e**: "Orientation distribution of actin filaments. N = 5 WT
tomograms of 4 neurons and 8 Arp3 KO tomograms of 8 neurons. Values are
plotted as mean +/- SD."

Source Data: `ED Fig 10e`.

## Outputs

- Two orientation-distribution PDFs (`ggsave` at lines 233, 249).

## Original G:\ source

G:/Figure S10_EM/Fig. S10f/R_scripts/actin_fila_quant.Rmd
-->

```{r message=FALSE, warning=FALSE, include=FALSE}
library(pracma)
library(peakPick)
library(quantmod)
library(ggpmisc)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggfortify)
library(forecast)
library(tsibble)
library(broom)
library(tidyquant)
library(tsbox)
library("pheatmap")
library("RColorBrewer")
library(ggpmisc)
library(ggfortify)
library(viridis) # Default Color Maps from 'matplotlib' 
library(hrbrthemes)
library(circular)
library("gridExtra") 
library("ggExtra") 

# install.packages("interactions") #the analysis and exploration of statistical interactions in the context of regression. 
```

## load existing/saved environment (.RData)
```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "actin_fila_quant.RData",  sep = "/")
load(file=saved_path)

```

```{r,  message = FALSE, warning=FALSE, fig.width=3}

df<- fila_angle_polarity_updated

colnames(df) %>% length()

ggplot(df, aes(x=WT_TS_05)) +  #KO_TS_04_180921, KO_TS_03_180814, WT_TS_01, WT_TS_02
 geom_histogram(aes(y=..density..), 
                colour="black", 
                fill="white", 
                binwidth=15) +
  
 geom_density(alpha=.2, fill="#FF6666")

ggplot(df, aes(x=WT_TS_05)) + 
             geom_histogram(colour="black", 
                            fill="white", 
                            binwidth=15) + 
  
              stat_bin(aes(y=..count.., label=..count..), 
                       geom="text",  
                       vjust=-.5, 
                       binwidth=15)
```

```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=5}
# ggplot_build(figA)$data[[1]]

binned_df <- last_plot()  %>% layer_data(1) # access the plot up to the most recent +

ggplot(binned_df, aes(x = x, y = ndensity)) +
  coord_polar(theta = "x", 
              start = 0, # Offset of starting point from 12 o'clock in radians. Offset is applied clockwise or anticlockwise depending on value of direction.
              direction = -1) + # 1, clockwise; -1, anticlockwise
  geom_bar(stat = "identity") +
  #geom_text(aes(y = y+5, label = y)) +
  theme(text = element_text(size=20),
        axis.text.x = element_text(angle=0, hjust=5)) +
  scale_x_continuous(limits=c(0,360), breaks=seq(0, 360, 30)) # scale_x_continuous(limits=c(-90,90),breaks=seq(-90, 90, 45))

```

# Batch process the angle data of all samples (df)
```{r}
polarity_list <- list()
histplot_list <- list()
roseplot_list <- list()
binned_list<-list()
for (n in 1:length(colnames(df))){
  
 histplot_list[[n]] <- ggplot(df[ ,n], aes_string(x=colnames(df)[n])) + 
             geom_histogram(colour="black", 
                            fill="white", 
                            binwidth=15) + 
  
              stat_bin(aes(y=..count.., label=..count..), 
                       geom="text",  
                       vjust=-.5, 
                       binwidth=15)
  
  binned_df <- last_plot()  %>% layer_data(1) 
  binned_df$pct<- binned_df$ndensity/sum(binned_df$ndensity) 
  binned_list[[n]] <- binned_df
  a <- dim(binned_df)[1]%/%2
  b <- dim(binned_df)[1]
  
  p1<- binned_df[1:a, ]$ndensity %>% sum() 
  p2 <-binned_df[(b-a):b, ]$ndensity %>% sum()
  polarity_list[[n]]<- p2/p1
  
  roseplot_list[[n]] <- ggplot(binned_df, aes(x = x, y = ndensity)) +
  coord_polar(theta = "x", 
              start = 0, # Offset of starting point from 12 o'clock in radians. Offset is applied clockwise or anticlockwise depending on value of direction.
              direction = -1) + # 1, clockwise; -1, anticlockwise
  geom_bar(stat = "identity") +
  #geom_text(aes(y = y+5, label = y)) +
  theme(text = element_text(size=20),
        axis.text.x = element_text(angle=0, hjust=5)) +
  scale_x_continuous(limits=c(0,360), breaks=seq(0, 360, 30)) # scale_x_continuous(limits=c(-90,90),breaks=seq(-90, 90, 45))
  
}

names(polarity_list) <- colnames(df)
names(histplot_list) <- colnames(df)
names(roseplot_list) <- colnames(df)
names(binned_list) <- colnames(df)
```

```{r}
polarity_list %>% as.matrix() %>% clipr::write_clip()
```

```{r}
binned_stack_list <- list()
for (n in 1:length(binned_list)){
  
  binned_df<- binned_list[[n]][,c("x", "pct", "ndensity")] 
  binned_df$names <-  binned_list[n] %>% names()
  binned_stack_list[[n]] <-   binned_df
}
```

```{r}
binned_stack <- data.table::rbindlist(binned_stack_list )

WT_stack <- binned_stack[1:104, ]
WT_stack$Sample<- "WT"
KO_stack <- binned_stack[105:277, ]
KO_stack$Sample<- "KO"

binned_stack <- rbind(WT_stack , KO_stack)

WT_stack_wide <- WT_stack %>% pivot_wider(names_from = names, values_from = c("pct", "ndensity"), values_fill = 0) 
KO_stack_wide <- KO_stack %>% pivot_wider(names_from = names, values_from = c("pct", "ndensity"), values_fill = 0)
```

```{r}

KO_stack %>% 
  ggplot(aes(x = x, y = ndensity, fill = names)) +
    geom_area(alpha=0.5 , size=.5, colour="white") +
    scale_fill_viridis(discrete = T) + # scale_fill_brewer(palette = "Set2") +
    theme_ipsum(grid = F, ticks = T) 
```

```{r}
pct_WT <-  data.frame(Angle=WT_stack_wide$x,
                         Mean=tapply(WT_stack$pct, WT_stack$x, mean), 
                         sd=tapply(WT_stack$pct, WT_stack$x, sd), 
                          n=tapply(WT_stack$names, WT_stack$x, length),
                         Sample="WT"
                         )

pct_KO <-  data.frame(Angle=KO_stack_wide$x,
                         Mean=tapply(KO_stack$pct, KO_stack$x, mean), 
                         sd=tapply(KO_stack$pct, KO_stack$x, sd), 
                          n=tapply(KO_stack$names, KO_stack$x, length),
                         Sample="KO"
                         )

pct_combined <- rbind(pct_WT, pct_KO)
pct_combined $sem <- pct_combined $sd/sqrt(pct_combined $n-1)

ndensity_WT <-  data.frame(Angle=WT_stack_wide$x,
                         Mean=tapply(WT_stack$ndensity, WT_stack$x, mean), 
                         sd=tapply(WT_stack$ndensity, WT_stack$x, sd), 
                          n=tapply(WT_stack$names, WT_stack$x, length),
                         Sample="WT"
                         )

ndensity_KO <-  data.frame(Angle=KO_stack_wide$x,
                         Mean=tapply(KO_stack$ndensity, KO_stack$x, mean), 
                         sd=tapply(KO_stack$ndensity, KO_stack$x, sd), 
                          n=tapply(KO_stack$names, KO_stack$x, length),
                         Sample="KO"
                         )

ndensity_combined <- rbind(ndensity_WT, ndensity_KO)
ndensity_combined $sem <- ndensity_combined $sd/sqrt(ndensity_combined $n-1)

```

```{r, fig.width=10}
p1<-ggplot(ndensity_combined, aes(x=Angle, y=Mean, group=Sample, colour=Sample, fill=Sample)) +
   geom_line(size=1, alpha=1) +
   geom_ribbon(aes(ymin=Mean-sd, ymax=Mean+sd),  alpha=1, linetype=0) + 
  scale_x_continuous(name="Angle", limits=c(0, 360), breaks=c(0,60,120,180,240,300,360))+ 
  theme(
  # Hide panel borders and remove grid lines
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  # Change axis line
  #axis.line = element_line(colour = "black")
  axis.text.x = element_text(colour="red"), axis.text.y = element_text(colour="black")
  )+ 
  theme_classic() +
  scale_color_manual(values=c('red','black'))+
  scale_fill_manual(values=c("red", "black"))

p2 <- ggplot(pct_combined, aes(x=Angle, y=Mean, group=Sample, colour=Sample, fill=Sample)) +
   geom_line(size=1, alpha=1) +
   geom_ribbon(aes(ymin=Mean-sd, ymax=Mean+sd),  alpha=1, linetype=0) + 
  scale_x_continuous(name="Angle", limits=c(0, 360), breaks=c(0,60,120,180,240,300,360)) +theme_classic()

cowplot::plot_grid(p1, p2, ncol=2, labels=LETTERS[1:2])
```
```{r}
ggsave(
  "Angle_distribution.eps",
  plot = last_plot(),
  device = "eps")
```

```{r}
ndensity_combined %>% write_csv("angle_distribution.csv")
```

```{r}
# histplot_list[[1]] 
roseplot_list[[13]] 
```

```{r}
ggsave(
  "Angle_rose_KO_8.pdf",
  plot = last_plot(),
  device = "pdf")
```

## Import angle distribution derived from the analyisis with OrientationJ in ImageJ
```{r,  message = FALSE, warning=FALSE, fig.width=5}
fig_list <- list()

for (n in 1:length(tab_list)){
fig_list[[n]] <- ggplot(tab_list[[n]], aes(x = Orientation, y = Slice1)) + 
  geom_area(color="darkblue", fill="lightblue") +
  
  coord_polar(theta = "x", 
              start = pi, # Offset of starting point from 12 o'clock in radians. 
                         # Offset is applied clockwise or anticlockwise depending on value of direction.
              direction = 1) + # 1, clockwise; -1, anticlockwise
 
  theme(text = element_text(size=12), axis.text.x = element_text(angle=0, hjust=5)) + 
  
  scale_x_continuous(limits=c(-180,180), breaks=seq(-90, 90, 30),
                   labels=c(expression(-90^degree), expression(-60^degree), expression(-30^degree), expression(0^degree), 
                              expression(30^degree), expression(60^degree), expression(90^degree))) # scale_x_continuous(limits=c(-90,90),breaks=seq(-90, 90, 45))

}

ggpubr::ggarrange(fig_list[[1]], fig_list[[2]], fig_list[[3]], fig_list[[4]], fig_list[[5]], fig_list[[6]],
                  labels = c("A", "B", "C", "D", "E", "F"),
                  ncol = 3, nrow = 2,  legend="none")
```

```{r,  message = FALSE, warning=FALSE, fig.width=5}

for (n in 1:length(tab_list)){
  
  df <- tab_list[[n]] %>% transform(Discrete=cut(tab_list[[n]]$Orientation, breaks =seq(-97.5, 97.5, 15), include.lowest=T))
  
  df2 <- aggregate(df$Slice1, by=list(Category=df$Discrete), FUN=sum)
  
  df2$Category <- seq(-90, 90, by = 15)
  df2$ndensity<- df2$x/max(df2$x)
  
  fig_list[[n]] <- ggplot(df2, aes(x = Category, y = ndensity)) +
  coord_polar(theta = "x", 
              start = pi, # Offset of starting point from 12 o'clock in radians. Offset is applied clockwise or anticlockwise depending on value of direction.
              direction = 1) + # 1, clockwise; -1, anticlockwise
  geom_bar(stat = "identity",  fill="lightblue") +
  #geom_text(aes(y = y+5, label = y)) +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=0, hjust=5)) +
  scale_x_continuous(limits=c(-180,180), breaks=seq(-90, 90, 30),
                     labels=c(expression(-90^degree), expression(-60^degree), expression(-30^degree), expression(0^degree), 
                              expression(30^degree), expression(60^degree), expression(90^degree))) # scale_x_continuous(limits=c(-90,90),breaks=seq(-90, 90, 45))

}

ggpubr::ggarrange(fig_list[[1]], fig_list[[2]], fig_list[[3]], fig_list[[4]], fig_list[[5]], fig_list[[6]],
                  labels = c("A", "B", "C", "D", "E", "F"),
                  ncol = 3, nrow = 2,  legend="none")

```

## Branching point dsitribution
```{r}

df <- WT_TS_01_branches
sp <- ggplot(df, aes(x=V1, y=V2)) + geom_point() +theme_classic()

sp + geom_density_2d()

sp + stat_density_2d(aes(fill = ..level..), geom="polygon", contour_var = "ndensity")+
  scale_fill_gradient(low="blue", high="red")
```

```{r,  message = FALSE, warning=FALSE, fig.width=5}
# Marginal density plot of x (top panel)
xdensity <- ggplot(df, aes(x=V1)) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position = "none")
xdensity
# Marginal density plot of y (right panel)
ydensity <- ggplot(df, aes(y=V2)) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position = "none")
ydensity

library("gridExtra")
grid.arrange(xdensity, blankPlot, scatterPlot, ydensity, 
        ncol=2, nrow=2, widths=c(4, 1.4), heights=c(1.4, 4))
```

```{r,  message = FALSE, warning=FALSE, fig.width=5}
# Scatter plot of x and y variables and color by groups 
scatterPlot <- ggplot(df, aes(x=V1, y=V2)) + 
  theme_classic()+
  ylim(0, 4000)+
  geom_point(color="green4", size=1)  + 
  theme(legend.position=c(0,1), legend.justification=c(0,1), axis.title.x=element_blank(), axis.title.y=element_blank()) 
# Marginal density plot of x (top panel) 
xdensity <- ggplot(df, aes(x=V1)) + geom_density(aes(y=..ndensity..)) + theme(legend.position = "none", axis.title.x=element_blank(), axis.title.y=element_blank())
# Marginal density plot of y (right panel)  
ydensity <- ggplot(df, aes(y=V2)) + geom_density(aes(x=..ndensity..)) +  theme(legend.position = "none", axis.title.x=element_blank(), axis.title.y=element_blank()) 

blankPlot <- ggplot()+geom_blank()+ theme_void()

library("gridExtra") 
grid.arrange(xdensity, blankPlot, scatterPlot, ydensity, ncol=2, nrow=2, widths=c(4, 1.4), heights=c(1.4, 4))
```

```{r,  message = FALSE, warning=FALSE, fig.width=3, fig.height=3}
ggMarginal(scatterPlot, 
           type ="densigram", 
           #binwidth = 100, 
           colour = "black", 
           #fill="lightblue", 
           xparams = list(size=0.5), 
           yparams = list(size=0.5))
```

```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "actin_fila_quant.RData",  sep = "/")
save.image(file=saved_path)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 
