---
title: "Fig 3n + ED Fig 5b - PA-Rac1 T17N precursor (Pre/Act/Post)"
output: html_notebook
params:
  variant: "T17N"
  pdf_name: "T17N_length.eps"
  length_y_limit: !r c(-20, 30)
---

<!--
================================================================================
  STAGE-1 PRECURSOR wrapper for Fig 3n + ED Fig 5b, PA-Rac1 T17N variant
  (dominant-negative Rac1 — blocks endogenous Rac1 signaling).
================================================================================
  Variant-specific data loading happens HERE; the deterministic
  pivot / save / plot pipeline is delegated to:
      ../../templates/fig03n_parac1_precursor.Rmd

  Body N = 20 ROIs. Published Fig 3n N = 20 (exact match). ED Fig 5b
  reports N = 18 (2 outliers removed in the downstream consolidator).

  Pipeline -> downstream consolidator
      figures/Fig03_Arp3/fig03n_parac1_gc_soma_length.Rmd
  which reads:
      T17N_actin_int.rds  (produced here)
      T17N_length.rds     (produced here)

  NB: this wrapper is the SIMPLEST of the 3 PA-Rac1 precursors because
  `neurite_length` (the wrapper-built single source) carries BOTH the
  length and the actin-mean ("Mean") columns. KO and C450M needed to
  combine multiple tables; T17N's RData is already single-table.

  Essentiality: Essential. Pairs with the KO and C450M sibling precursors.

  Original 170-line monolithic wrapper refactored 2026-05-12: variant-
  specific RData load + single-table pivot kept inline; the generic
  pivot/save/plot pipeline moved to the canonical template.
  Original G:\ source: G:/Figure 3_Arp3/Fig. 3j-n_PARac1/Andor/PARac1_T17N/Neurite_length.Rmd
-->

## Step 1: libraries + load saved environment

```{r message=FALSE, warning=FALSE, include=FALSE}
library(pracma)
library(quantmod)
library(ggpmisc)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggfortify)
library(forecast)
library(tsibble)
library(broom)
library(tidyquant)
library(tsbox)
library("pheatmap")
library("RColorBrewer")
library(viridis)
library(hrbrthemes)
library(data.table)
library(superheat)
```

```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "T17N_length.RData", sep = "/")
load(file = saved_path)
```

## Step 2: `length_input` is the loaded `neurite_length` (already long-format)

For the T17N RData, `neurite_length` has columns `roi`, `Image`, `Site`,
`exp`, `Series`, `Length`, `Mean` — i.e., both length and actin
intensity in one table. The wrapper just aliases it; the template will
pivot from `length_input` for the length panel and from `actin_input`
for the intensity panel.

```{r}
length_input <- neurite_length
```

## Step 3: `actin_input` uses the same single source

The Series + Mean columns of `neurite_length` ARE the per-ROI Pre/Act/Post
actin-intensity table. We rename `Mean` -> `Mean` (already correct) and
pass through.

```{r}
actin_input <- neurite_length
```

## Step 4: include canonical pivot/save/plot template

```{r child = "../../templates/fig03n_parac1_precursor.Rmd"}
```

## Step 5: persist working environment

```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "T17N_length.RData", sep = "/")
save.image(file = saved_path)
```
