---
title: "R Notebook"
output: html_notebook
---

<!--
# Microperfusion of CK-666 or para-aminoblebbistatin: actin intensity + neurite-length response timecourse

## What this file does

Reads per-experiment ROI intensity CSVs (`*int.csv` via `list.files` at line
60, loaded line 63) and per-experiment neurite-length CSVs (`length.csv` via
`list.files` at line 333, loaded line 336) for local perfusion of CK-666 or
para-aminoblebbistatin at the soma. For each event the actin / GFP intensity
and neurite length are sampled across Pre / Act / Post timepoints, plotted
as line plots vs time with the perfusion-window band shaded, and pooled into
violin / boxplot summaries. Six PDFs are written (`ggsave` at lines 241,
433, 478, 543, 617, 701).

## Manuscript panels

- **Fig 3r**: "Percentage of neurons perfused with DMSO or CK-666 showing
  retrograde actin wave."
- **ED Fig 4b** / **ED Fig 4c** - CK-666 microperfusion extension /
  retraction duration and velocity.
- **ED Fig 4h** - longest-neurite length after CK-666 + washout.
- **ED Fig 5i**: "Length changes upon micro-perfusion of CK-666 at the
  soma."
- **ED Fig 7f**: "Length changes upon local perfusion of 40 uM
  para-aminoblebbistatin at the soma."

Source Data: `Fig 3r`, `ED Fig 4b/c/h`, `ED Fig 5i`, `ED Fig 7f`.

## Outputs

- Six PDFs (`ggsave` at lines 241, 433, 478, 543, 617, 701).

## Original G:\ source

G:/Figure 3_Arp3/Fig. 3o-r_local_perfusion/R_scripts/Neurite activity corr.Rmd
-->

```{r message=FALSE, warning=FALSE, include=FALSE}
library(pracma)
#library(peakPick)
library(quantmod)
library(ggpmisc)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggfortify)
library(forecast)
library(tsibble)
library(broom)
library(tidyquant)
library(tsbox)
library("pheatmap")
library("RColorBrewer")
library(ggpmisc)
library(ggfortify)
library(viridis) # Default Color Maps from 'matplotlib' 
library(hrbrthemes)
library(data.table)
library(superheat)
library(colorspace)
library(ggpubr)
```

## load existing/saved environment (.RData)
```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(Path_1)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
load(file=saved_path)
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
```

# save environment (.RData)
```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
save.image(file=saved_path)
```

# read csv files from the parent folder
```{r}
library(readr)  # For faster CSV reading
library(dplyr)  # For easy data manipulation (optional)

# Function to import all CSVs from a folder and its subfolders
import_csv_files <- function(directory) {
  # Get list of all CSV files in the folder and subfolders
  csv_files <- list.files(directory, pattern = "\\int.csv$", full.names = TRUE, recursive = TRUE)
  
  # Read each file into a list of data frames
  data_list <- lapply(csv_files, read_csv, show_col_types = FALSE)
  
  # Name the list elements with the corresponding file names (without the full path)
  names(data_list) <- dirname(csv_files) %>% basename()
  
  return(data_list)
}

# Set the parent directory containing CSV files
parent_directory <- Path_1 

# Import all CSV files
csv_data_list <- import_csv_files(parent_directory)
```

# filter for analyzed cells and relevant columns 
```{r}
library(readxl)
cell_meta <- read_excel("Microperfusion_12_TC.xlsx")  %>% as.data.table()
cell_meta <- cell_meta[analyzed == "Yes"]
cell_meta$perfused_compartment %>% table()
```

```{r}
cell_meta_count <- cell_meta[ , c("DMSO_CK666", "Cell", "perfused_compartment", "Proximal_GC", "Distal_GC", "Proximal_growth", "Distal_growth")]

summary_tbl <- cell_meta_count %>%                                    
  pivot_longer(cols = c("Proximal_GC", "Distal_GC", "Proximal_growth", "Distal_growth"),                   
               names_to  = "source_col",                 
               values_to = "category") %>%               
  group_by(DMSO_CK666, perfused_compartment, source_col, category) %>%                         
  summarise(n = n(), .groups = "drop") %>% as.data.table()

summary_tbl[source_col == "Distal_growth" & category == "shorter"]

summary_tbl$group <- paste(summary_tbl$DMSO_CK666, summary_tbl$perfused_compartment, sep = "_")

```

# consider only the first frame of in_perfusion, last frame of in_perfusion, and the last frame of post_perfusion
```{r}

data_norm_list_2 <- list()

for (n in 1: length (data_norm_list)){
  
    df <- data_norm_list[[n]]
    df_in <- df[period=="2_in_perfusion"]
    
    df_in_max <- df_in[Frame == max(df_in$Frame)]
    df_in_max$period <- "2_in_perfusion"
    
    df_in_min <- df_in[Frame == min(df_in$Frame)]
    df_in_min$period <- "1_pre_perfusion"
    
    df_post <- df[period=="3_post_perfusion"]
    df_post_max <- df_post[Frame == (max(df_post$Frame)-12)]
    
    data_norm_list_2[[n]] <- rbind(df_in_max, df_in_min, df_post_max) %>% as.data.table()
    
}
```

```{r, message = FALSE, warning=FALSE, fig.width=10}
n =5

meta <- cell_meta[Cell== names(csv_data_list)[n]]

x_start <- meta$perfusion_start %>% as.numeric()
x_end <- meta$perfusion_end  %>% as.numeric()

data_norm_list[[n]] %>%
ggplot(aes(x = Frame, y = int)) + 
  geom_line(aes(color = Roi),  size=2)+
  
  #scale_x_continuous( expand = c(0, 0), limits = c(450, 744), breaks = scales::pretty_breaks(n = 7))+
  #scale_y_continuous( expand = c(0, 0), limits = c(0.5, 2), breaks = scales::pretty_breaks(n = 7))+
  
  geom_vline(xintercept=x_start, linetype="dashed", color = "red")+
  geom_vline(xintercept=x_end, linetype="dashed", color = "red")+
  
  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold")) #for centering the text 
```

```{r}
int_df <- data_norm_list %>% bind_rows() %>% as.data.table()

int_df$exp <- "CK-666" 
int_df[ grepl("Control", int_df$cell), ]$exp <- "DMSO"

df_summary <- int_df %>% 
  group_by(period, perfusion_site,  cell, perfusion_compartment, exp) %>%  
  rstatix::get_summary_stats(int) %>% 
  as.data.table()

df_summary$group <- paste(df_summary$exp, df_summary$perfusion_compartment, df_summary$perfusion_site, df_summary$period, sep = "_")

```

```{r}
outliers <- df_summary %>% 
  group_by(group) %>% 
 rstatix::identify_outliers(mean) 

#df_summary <- df_summary [! (group %in% outliers$group) ]

df_summary <- df_summary [! (cell %in% "20241112_2") ]

#df_summary[period== "2_in_perfusion" & perfusion_compartment =="neurite"]$cell %>% table()
```

```{r}
all_df <- data_norm_list_2 %>% bind_rows() %>% as.data.table()

df_summary_2 <- all_df %>% group_by(period, perfusion_site,  cell, perfusion_compartment) %>%  rstatix::get_summary_stats(int) %>% as.data.table()
df_summary_2$group <- paste(df_summary_2$cell, df_summary_2$period, sep = "_")
df_summary_2$exp_site_period <- paste(df_summary_2$exp, df_summary_2$perfusion_site, df_summary_2$period, sep = "_")
df_summary_2$cell_site_period <- paste(df_summary_2$cell, df_summary_2$site_period, sep = "_")

outliers <- df_summary_2 %>% 
  group_by(site_period) %>% 
 rstatix::identify_outliers(mean) 

df_summary_2 <- df_summary_2 [! (cell_site_period %in% outliers$cell_site_period) ]

```

```{r message = FALSE, warning=FALSE, fig.width=10}
df_summary$site_period <- paste(df_summary$perfusion_site, df_summary$period, sep = "_")

my_comparisons <- list(c("perfused_1_pre_perfusion", "perfused_2_in_perfusion"), c("perfused_1_pre_perfusion", "perfused_3_post_perfusion"), c("non-perfused_1_pre_perfusion", "non-perfused_2_in_perfusion"), c("non-perfused_1_pre_perfusion", "non-perfused_3_post_perfusion"))

pj <- position_jitter(seed = 1, width = .1, height = 0)

# define group order
perfusion_site_order <- c("perfused", "non-perfused")
period_order <- c("1_pre_perfusion",   "2_in_perfusion", "3_post_perfusion")

 site_period_order <- c("non-perfused_1_pre_perfusion", "non-perfused_2_in_perfusion", "non-perfused_3_post_perfusion", "perfused_1_pre_perfusion", "perfused_2_in_perfusion", "perfused_3_post_perfusion")
#site_period_order <- c("non-perfused_pre_perfusion", "perfused_pre_perfusion", "non-perfused_in_perfusion", "perfused_in_perfusion", "non-perfused_post_perfusion",   "perfused_post_perfusion")
```

```{r message = FALSE, warning=FALSE, fig.width=10}
df_summary[perfusion_compartment == "soma" & exp == "CK-666"] %>%  #mutate(exp_site_period = factor(exp_site_period, levels=site_period_order))  %>%
  
ggplot(aes(x=site_period, y=mean, fill=perfusion_site))+
  #geom_point()
  theme_classic()+
 geom_violin(aes(), trim=F,  kernel ="gaussian", adjust=0.8, scale = "width")  +
 geom_boxplot(aes(), width=0.2, fill="white", outlier.colour = NA)+ 
  
 # geom_jitter(alpha=1, aes(color = period), position = position_jitterdodge(jitter.width=.1)) + 
  
  geom_point(position = pj) +
 
 #geom_path(aes(group = cell),alpha = 0.5, position = pj)+
  stat_compare_means(comparisons = my_comparisons, test = "wilcox.test", paired=F)+
 # scale_color_continuous_sequential(palette = "Heat", begin = 0.3, end = 1)+ 
 #scale_y_continuous(expand = c(0, 0), limits = c(0, 4.5), breaks = scales::pretty_breaks(n = 10))+ 

  theme_ipsum(grid = F, ticks = T) +
 theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold"), axis.text.x = element_text(angle = 0)) #for centering the text 
```

```{r}
ggsave(
  "neurite_avg.pdf",
  plot = last_plot(),
  device = "pdf")

```

```{r}
all_df <- data_norm_list %>% bind_rows() %>% as.data.frame()  %>% as.data.table()

df <- all_df[period == "2_in_perfusion" & perfusion_site == "perfused"]

cell_meta$perfused_compartment %>% table()
```

```{r message = FALSE, warning=FALSE, fig.width=10}
library(colorspace)
df  %>% 
ggplot(aes(x=new_time, y=int, color=cell)) +
  #geom_point()
  theme_classic()+

 # geom_boxplot(fill="bisque",color="black",alpha=0.5, outliers = F) +
  geom_jitter(alpha=1) + 
 # scale_color_continuous_sequential(palette = "Heat", begin = 0.3, end = 1)+ 
 # scale_y_continuous(expand = c(0, 0), limits = c(0, 1.5), breaks = scales::pretty_breaks(n = 10))+ 
   scale_x_continuous(expand = c(0, 0), limits = c(0, 6000), breaks = scales::pretty_breaks(n = 10))+ 
  theme_ipsum(grid = F, ticks = T) +
 theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold")) #for centering the text 
```

```{r}
outliers <- df %>% 
  group_by(new_time) %>% 
 rstatix::identify_outliers(int) 

df <- df [! ((cell %in% outliers$cell) &  (time %in% outliers$time))]

```

```{r}
df_summary_in_perfusion <- df %>%  
  group_by(new_time) %>% rstatix::get_summary_stats(int) %>% as.data.table()
  
```

```{r  message = FALSE, warning=FALSE, fig.width=8}

df_summary_in_perfusion[n>1] %>% 
  
ggplot(aes(x=new_time, y=mean)) +

  geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd), fill="lightgrey", alpha=1)+
  geom_line(size=1) +
  
 # geom_vline(xintercept=0, linetype="dashed", color = "black")+
 # geom_hline(yintercept=0, linetype="dashed", color = "black")+
  
 # stat_peaks(colour = "red", ignore_threshold = 0.9) +
 # stat_peaks(geom = "text", colour = "red", vjust = -0.5, ignore_threshold = 0.9) +
  
#  stat_valleys(colour = "blue", ignore_threshold = 0.9, span = 11)+
 # stat_valleys(geom = "text", colour = "blue", angle = 45, vjust = 1.5, hjust = 1, ignore_threshold = 0.9, span = 11)+
  
  scale_y_continuous(expand = c(0, 0), limits = c(-0.4, 2), breaks = scales::pretty_breaks(n = 7))+
 #scale_x_continuous(expand = c(0, 0), limits = c(0, 5000), breaks = scales::pretty_breaks(n = 8)) + 
  
  theme_ipsum(grid = F, ticks = T)+ 
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text
```

## Analysis of neuite length change
# read csv files of length from the parent folder
```{r message=FALSE, warning=FALSE}
library(readr)  # For faster CSV reading
library(dplyr)  # For easy data manipulation (optional)

# Function to import all CSVs from a folder and its subfolders
import_csv_files <- function(directory) {
  # Get list of all CSV files in the folder and subfolders
  csv_files <- list.files(directory, pattern = "length.csv$", full.names = TRUE, recursive = TRUE)
  
  # Read each file into a list of data frames
  data_list <- lapply(csv_files, read_csv, show_col_types = FALSE)
  
  # Name the list elements with the corresponding file names (without the full path)
  names(data_list) <- dirname(csv_files) %>% basename()
  
  return(data_list)
}

# Set the parent directory containing CSV files
parent_directory <- Path_1 

# Import all CSV files
length_data_list <- import_csv_files(parent_directory)

```

# filter for analyzed cells and relevant columns 
```{r}
library(readxl)
cell_meta <- read_excel("Microperfusion_12_TC.xlsx")  %>% as.data.table()
cell_meta <- cell_meta[analyzed == "Yes"]
cell_meta$perfused_compartment %>% table()
```

# add metadata and define the perfusion window

```{r}
length_df <- length_norm_list %>% rbindlist() %>% as.data.table()
length_df$exp <- "CK-666" 
length_df[ grepl("Control", length_df$cell), ]$exp <- "DMSO"
 
length_df$mean <- rowMeans(length_df[,c(5:7)])
length_df$Diff<- length_df$`2_in_perfusion` - length_df$`1_pre_perfusion`
    
length_df$`1_pre_perfusion_norm` <- length_df$`1_pre_perfusion` - length_df$mean
length_df$`2_in_perfusion_norm` <- length_df$`2_in_perfusion` - length_df$mean
length_df$`3_post_perfusion_norm` <- length_df$`3_post_perfusion` - length_df$mean

#saveRDS(length_df, file="Control_GC_neuriteLength.rds")
```

# get normalized values (raw - mean) and transform data structure 
```{r}
length_df_long<- pivot_longer(length_df, cols=c(10:12), names_to = "Series", values_to="length")  %>% as.data.table()
length_df_long$cell_neurite <- paste(length_df_long$cell, length_df_long$neurite, sep = "_")

#length_df_long$Series <- factor(length_df_long$Series, levels=unique(length_df_long$Series))    
```

```{r}
# split into subgroups based on perfused compartment 
length_df_long_soma <- length_df_long [perfusion_compartment == "soma"]
length_df_long_neurite_1  <- length_df_long [perfusion_compartment == "neurite" & neurite == "Neurite_1"]
length_df_long_neurite_rest  <- length_df_long [perfusion_compartment == "neurite" & neurite != "Neurite_1"]
```

# remove outliers
```{r}
outliers <- length_df_long_soma %>% 
  group_by(Series, exp) %>% 
 rstatix::identify_outliers(mean) 

length_df_long_soma <-  length_df_long_soma [! (cell_neurite %in% outliers$cell_neurite), ]
```

```{r, message = FALSE, warning=FALSE, fig.width=8}
library(ggpubr)

my_comparisons <- length_df_long$Series %>% unique() %>% as.character() %>% combn(2, simplify = FALSE) 

#neurite_length_long<- neurite_length_long%>% subset(neurite_length_long$Series != "Post")

length_df_long_soma %>% 
ggplot(aes(x=Series, y=length, fill=Series)) + 
  geom_violin(trim=FALSE, 
              kernel ="gaussian", 
              adjust=1.3)+
  #geom_boxplot(width=2, fill="white")  +
  geom_point(aes(fill=Series, group=cell_neurite), position = position_dodge(0.05))+
  geom_line(aes(group=cell_neurite), position = position_dodge(0.05)) +
  stat_compare_means(comparisons = my_comparisons, paired=T, method="wilcox.test")+
  #scale_y_continuous( expand = c(0, 0), limits = c(-25, 50), breaks = scales::pretty_breaks(n = 7))+
  
  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
  
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold")) #for centering the text 

#plotly::ggplotly(p)
```

```{r}
ggsave(
  "neurite_1_CK666_length_paired.pdf",
  plot = last_plot(),
  device = "pdf")
```

# Calculate the mean length change per cell
```{r}
df_summary <- length_df_long %>% 
  group_by(Series,  cell, perfusion_compartment, exp) %>%  
  rstatix::get_summary_stats(length) %>% 
  as.data.table()

```

```{r, message = FALSE, warning=FALSE, fig.width=8}
library(ggpubr)

my_comparisons <- length_df_long$Series %>% unique() %>% as.character() %>% combn(2, simplify = FALSE) 

#neurite_length_long<- neurite_length_long%>% subset(neurite_length_long$Series != "Post")

df_summary [perfusion_compartment == "soma"] %>% 
  
ggplot(aes(x=Series, y=mean, fill=Series)) + 
  geom_violin(trim=FALSE, kernel ="gaussian", adjust=1.5)+
 # geom_boxplot(width=2, fill="white")  +
  geom_point(aes(fill=Series, group=cell), position = position_dodge(0.05))+
  geom_line(aes(group=cell), position = position_dodge(0.05)) +
  stat_compare_means(comparisons = my_comparisons, paired=T, method="wilcox.test")+
 # scale_y_continuous( expand = c(0, 0), limits = c(-15, 15), breaks = scales::pretty_breaks(n = 7))+
  
  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
  
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold")) #for centering the text 

```

```{r}
ggsave(
  "soma_CK666_length_paired.pdf",
  plot = last_plot(),
  device = "pdf")
```

# Direct comparison of length difference before and after perfusion
# compare the length change between DMSO and CK-666 (soma)
```{r}
df_summary_diff <- length_df %>% 
  group_by(perfusion_compartment, exp, cell) %>%  
  rstatix::get_summary_stats(Diff) %>% 
  as.data.table()

df_summary_diff_soma <- df_summary_diff[perfusion_compartment == "soma"]

df_summary_diff_max <- df_summary_diff[max > 0]
df_summary_diff_min <- df_summary_diff[min < 0]
```

# remove outliers
```{r}
outliers <- df_summary_diff_soma  %>% 
  group_by(exp) %>% 
 rstatix::identify_outliers(mean) 

  df_summary_diff_soma <-  df_summary_diff_soma [! (cell %in% outliers$cell), ]
  
```

```{r}
outliers <-  df_summary_diff_max %>% 
  group_by(exp) %>% 
 rstatix::identify_outliers(max) 

  df_summary_diff_max <-  df_summary_diff_max [! (cell %in% outliers$cell), ]
  
```

```{r message = FALSE, warning=FALSE, fig.width=10}
my_comparisons <- list(c("DMSO", "CK666"), c("DMSO", "para-aminoblebb"))

 df_summary_diff_soma %>% mutate(exp = factor(exp, levels=c("DMSO", "CK666", "para-aminoblebb")))  %>%
  
 ggplot(aes(x=exp, y=mean, fill=exp)) + 
  geom_violin(trim=FALSE, kernel ="gaussian", adjust=1)+
  geom_boxplot(width=0.5, fill="white", outlier.shape = NA)  +
  geom_point(aes(fill=exp, group=cell), position = position_jitter(0.1))+
 # geom_line(aes(), position = position_dodge(0.05)) +
  stat_compare_means(comparisons = my_comparisons, paired=F, method="t.test")+
 #scale_y_continuous( expand = c(0, 0), limits = c(-25, 5), breaks = scales::pretty_breaks(n = 7))+
  
  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
  
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold")) #for centering the text 

 #plotly::ggplotly(p)
```
```{r}
ggsave(
  "CK666_DMSO_soma_length.pdf",
  plot = last_plot(),
  device = "pdf")

dev.off()
```

# split the soma data into perfused and non-perfused neurites based on neurite length
```{r}
length_df_soma <- length_df[perfusion_compartment == "soma"]

length_df_soma$site <- "perfused"
length_df_soma[`1_pre_perfusion` >40 ]$site <- "non-perfused"

length_df_soma$group <- paste(length_df_soma$exp, length_df_soma$site, sep = "_")
length_df_soma$cell_neurite <- paste(length_df_soma$cell, length_df_soma$neurite, sep = "_")

length_df_soma_1 <- length_df_soma
```

# remove outliers
```{r}
outliers <- length_df_soma_1  %>% 
  group_by(group) %>% 
 rstatix::identify_outliers(mean) 

length_df_soma_1 <-  length_df_soma_1[! (cell_neurite %in% outliers$cell_neurite), ]
#length_df_soma_1 <-  length_df_soma_1[! (cell %in% "Control_20250313_2_ext_ret"), ]

length_df_soma_1$group %>% table

df <- length_df_soma_1 %>% group_by(cell, exp) %>% rstatix::get_summary_stats(mean)

df %>% group_by(exp)%>% rstatix::get_summary_stats(mean)

length_df_soma_1[exp != "para-aminoblebb"] %>% write_csv("length_df_soma_1_CK666.csv")
```

```{r message = FALSE, warning=FALSE, fig.width=10}
my_comparisons <- list(
                       c("DMSO_perfused", "DMSO_non-perfused"), 
                       c("CK666_perfused", "CK666_non-perfused"), 
                       c("DMSO_perfused", "CK666_perfused"), 
                       c("DMSO_non-perfused", "CK666_non-perfused"))
                       #c("para-aminoblebb_perfused", "para-aminoblebb_non-perfused"), 
                       #c("DMSO_perfused", "para-aminoblebb_perfused"), 
                     # c("DMSO_non-perfused", "para-aminoblebb_non-perfused"))

 length_df_soma_1[exp != "para-aminoblebb"] %>% # mutate(exp = factor(exp, levels=c("DMSO", "CK-666")))  %>%
  
 ggplot(aes(x=group, y=Diff, fill=exp)) + 
  geom_violin(trim=FALSE, kernel ="gaussian", adjust=1)+
  geom_boxplot(width=0.5, fill="white", outlier.shape = NA)  +
  geom_point(aes(fill=exp, group=cell), position = position_jitter(0.1))+
 # geom_line(aes(), position = position_dodge(0.05)) +
  stat_compare_means(comparisons = my_comparisons, paired=F, method="t.test")+
scale_y_continuous( expand = c(0, 0), 
                    limits = c(-50, 100), 
                    breaks = scales::pretty_breaks(n = 8))+
  
  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
  
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold")) #for centering the text 

 #plotly::ggplotly(p)
```
```{r}
ggsave(
  "CK666_DMSO_soma_length_50_neuriteLevel.pdf",
  plot = last_plot(),
  device = "pdf")

#dev.off()
```

```{r}
library(rstatix); library(ggpubr); library(tidyverse)

# 1. Check assumptions
# Normality per cell (Shapiro-Wilk; ignore if n is large in each cell)
length_df_soma_1 %>%
  group_by(exp, site) %>%
  shapiro_test(Diff)

# Homogeneity of variance (Levene's test)
length_df_soma_1 %>% levene_test(Diff ~ exp * site)

# 2. Two-way ANOVA
anova_result <- length_df_soma_1 %>%
  anova_test(Diff ~ exp * site)

print(anova_result)
# Output: Effect (treatment, perfusion, treatment:perfusion),
#         DFn, DFd, F, p, p<.05, ges (generalized eta squared)

posthoc <- length_df_soma_1[exp != "CK666"] %>%
  tukey_hsd(Diff ~ exp * site)

posthoc <- length_df_soma_1[exp != "CK666"] %>% dunn_test(Diff ~ group, p.adjust.method = "bonferroni")

print(posthoc)

```

# analyze at cell level
```{r}
df_summary_diff_soma <- length_df_soma %>% 
  group_by(exp, site, cell, group) %>%  
  rstatix::get_summary_stats(Diff) %>% 
  as.data.table()
```

# remove outliers
```{r}
outliers <- df_summary_diff_soma  %>% 
  group_by(group) %>% 
 rstatix::identify_outliers(mean) 

  df_summary_diff_soma <-  df_summary_diff_soma [! (cell %in% outliers$cell), ]
  
df_summary_diff_soma$group %>% table()

df <- length_df_soma %>% group_by(cell, exp)  %>% rstatix::get_summary_stats(Diff)
```

```{r message = FALSE, warning=FALSE, fig.width=10}
my_comparisons <- list(c("DMSO_perfused", "DMSO_non-perfused"), c("CK-666_perfused", "CK-666_non-perfused"), c("DMSO_perfused", "CK-666_perfused"), c("DMSO_non-perfused", "CK-666_non-perfused"))

 length_df_soma %>% #mutate(exp = factor(exp, levels=c("DMSO", "CK-666")))  %>%
  
 ggplot(aes(x=group, y=Diff, fill=exp)) + 
  geom_violin(trim=FALSE, kernel ="gaussian", adjust=1)+
  geom_boxplot(width=0.5, fill="white", outlier.shape = NA)  +
  geom_point(aes(fill=exp, group=cell), position = position_jitter(0.1))+
 # geom_line(aes(), position = position_dodge(0.05)) +
  stat_compare_means(comparisons = my_comparisons, paired=F, method="t.test")+
 #scale_y_continuous( expand = c(0, 0), limits = c(-25, 5), breaks = scales::pretty_breaks(n = 7))+
  
  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
  
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold")) #for centering the text 

 #plotly::ggplotly(p)
```
```{r}
ggsave(
  "CK666_blebb_DMSO_soma_length_40.pdf",
  plot = last_plot(),
  device = "pdf")

dev.off()
```

# compare the length change between perfused and non-perfused neurite (GC), CK-666 vs DMSO
```{r}
 length_df_neurite <- length_df[perfusion_compartment == "neurite" && exp == c("DMSO","CK666")]
 length_df_neurite$perfusion <- "non-perfused" 
 length_df_neurite[ grepl("Neurite_1",  length_df_neurite$neurite), ]$perfusion <- "perfused"
 
 df_summary_diff <-  length_df_neurite %>% 
  group_by(perfusion, cell, exp) %>%  
  rstatix::get_summary_stats(Diff) %>% 
  as.data.table()
 
 df_summary_diff$group <- paste(df_summary_diff$exp, df_summary_diff$perfusion, sep = "_")
```

# remove outliers
```{r}
outliers <-  df_summary_diff %>% 
  group_by(group) %>% 
 rstatix::identify_outliers(mean) 

df_summary_diff <-  df_summary_diff [! (cell %in% outliers$cell), ]
  
df_summary_diff$group %>% table()  
  
df_summary_diff %>% group_by(group) %>%  
  rstatix::get_summary_stats(mean)

```

```{r message = FALSE, warning=FALSE, fig.width=10}
my_comparisons <- list(c("CK-666_non-perfused", "CK-666_perfused"), c("DMSO_non-perfused", "DMSO_perfused"), c("CK-666_non-perfused", "DMSO_non-perfused"), c("CK-666_perfused", "DMSO_perfused"))

  df_summary_diff %>% 
  
ggplot(aes(x=group, y=mean, fill=perfusion)) + 
  geom_violin(trim=FALSE, kernel ="gaussian", adjust=1) +
  geom_boxplot(width=0.2, fill="white", outlier.shape = NA) +
  geom_point(aes(fill=perfusion, group=cell), position = position_dodge(0.1)) +
# geom_line(aes(group=cell), position = position_dodge(0.1)) +
  stat_compare_means(comparisons = my_comparisons, paired=F, method="t.test") +
 scale_y_continuous( expand = c(0, 0), 
                     limits = c(-50, 40), 
                     breaks = scales::pretty_breaks(n = 8))+
  
  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
  
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold")) #for centering the text 
```

```{r}
ggsave(
  "CK666_DMSO_neurite_length.pdf",
  plot = last_plot(),
  device = "pdf")

#dev.off()
```
