---
title: "R Notebook"
output: html_notebook
---

<!--
# Arp3 / F-actin growth-cone Airyscan line-profile colocalization

## What this file does

Reads paired Airyscan growth-cone line-profile CSVs (`profiles.csv` and
`profiles2.csv` per neuron, lines 53, 58, 70, 74) for the Arp3-A488 +
phalloidin-Rhodamine staining cohort, computes per-profile correlation
between Arp3 and F-actin intensities, and produces the violin-plot
comparison (`ggsave` at line 190).

## Manuscript panel

**ED Fig 6g**: "Colocalization of actin patches with Arp3 and NMIIb patches.
N = 24 neurons for Arp3 and 31 neurons for NMIIb from 3 independent
experiments. ***p < 0.001 by Two-tailed Mann-Whitney test."

Source Data: `ED Fig 6g`.

## Outputs

- One violin-plot PDF (`ggsave` at line 190).

## Original G:\ source

G:/Figure S6_myosin/Fig. S6g/201002_ICC_extFig. 6g/Arp3A488_PhallRho/WT/Airyscan/Profiles/Patch_profile_Corr.Rmd
-->

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

## Package installation
```{r message=FALSE, warning=FALSE, include=FALSE}
library(pracma)
library(peakPick)
library(quantmod)
library(ggpmisc)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggfortify)
library(forecast)
library(tsibble)
library(broom)
library(tidyquant)
library(tsbox)
library("pheatmap")
library("RColorBrewer")
library(ggpmisc)
library(ggfortify)
library(viridis) # Default Color Maps from 'matplotlib' 
library(hrbrthemes)
library(data.table)
library(superheat)
```

## load existing/saved environment (.RData)
```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
load(file=saved_path)

```

## Read .csv files
```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Arp3_line_profile", subfolder_list) ]

```

```{r}
Arp3_csv_list <- list.files(subfolder_list,pattern = "profiles.csv", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file

Actin_csv_list <- list.files(subfolder_list,pattern = "profiles2.csv", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
```

```{r}
Arp3_list <- list()
Actin_list <- list()
  
  for (n in 1:length(Arp3_csv_list))  {

    df1<- read.csv(Arp3_csv_list[n], header = TRUE)  
    df1[ df1 == 0 ] <- NA
    Arp3_list[[n]]  <- df1 #%>% MTS::diffM(d=1)  %>% forecast::ma(5) 
    
    df1<- read.csv(Actin_csv_list[n], header = TRUE)  
    df1[ df1 == 0 ] <- NA
    Actin_list[[n]]  <- df1 #%>% MTS::diffM(d=1)  %>% forecast::ma(5) 
    
  }
  
names(Arp3_list) <- Arp3_csv_list %>% fs::path_file()
names(Actin_list) <- Actin_csv_list %>% fs::path_file()
```

# cross-correlation with neurite length at individual neurite level

```{r message = FALSE, warning=FALSE, fig.width=5}

Arp3Actin_ccf_list <- list()

for (n in 1:length(Arp3_list)){
  
roi_list <- list()
  
  for (m in 2:ncol(Arp3_list[[n]])){

   ccf1 <-ccf (Arp3_list[[n]][m],  
               Actin_list[[n]][m], 
               type = "correlation", 
               na.action = na.pass, plot=F,
               lag.max=0) 
   
   roi_list[[m-1]] <- ccf1[["acf"]]
  
  }

    df <- roi_list %>% unlist() %>% as.data.frame()
    df$img<- names(Arp3_list)[n]
    Arp3Actin_ccf_list[[n]] <- df
}   

names(Arp3Actin_ccf_list) <- names(Arp3_list) 

Arp3Actin_ccf <- Arp3Actin_ccf_list %>% reduce(bind_rows)  %>% as.data.frame()

names(Arp3Actin_ccf)[1] <- "ccf"

Arp3Actin_ccf$Col <- "Arp3"

```

# check point distribution
```{r}
library(easyGgplot2)
  ggplot2.violinplot(data = Arp3Actin_ccf, 
                     yName="ccf", xName='Col', groupName='Col', 
                     addDot=TRUE, dotSize=2,dotPosition="jitter", jitter=0.2, 
                     trim=F,  kernel ="gaussian", adjust=2)+
    
  geom_boxplot(width=0.1, fill="white")  +
  #theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold")) #for centering the text 
```
# Data points at cell level
```{r}
df_summary <-Arp3Actin_ccf %>% group_by(img) %>% rstatix::get_summary_stats(ccf)
```

```{r message = FALSE, warning=FALSE, fig.width=8}
library(easyGgplot2)
  ggplot2.violinplot(data = df_summary, yName="mean", xName='variable', groupName='variable', 
                     addDot=TRUE, dotSize=2, dotPosition="jitter", jitter=0.2, 
                     trim=F,  kernel ="gaussian", adjust=5)+
    
  geom_boxplot(width=0.1, fill="white")  +
  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold")) #for centering the text 

```

```{r message = FALSE, warning=FALSE, fig.width=5}
library(ggpubr)

#my_comparisons <- list( c("Pre", "Act") )

#neurite_length_long<- neurite_length_long%>% subset(neurite_length_long$Series != "Post")

ggplot(Arp3Actin_ccf, aes(x=Col, y=ccf)) + 
  geom_violin(trim=FALSE, kernel ="gaussian", adjust=1)+
  geom_boxplot(width=0.5, fill="white")  +
  #geom_point(aes(), position = position_dodge(0.5))+
  #geom_line(aes(), position = position_dodge(0.2)) +
  #stat_compare_means(comparisons = my_comparisons, paired=T, method="wilcox.test")+
  scale_y_continuous( expand = c(0, 0), breaks = scales::pretty_breaks(n = 5))+
  
  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
  
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold")) #for centering the text 

# plotly::ggplotly(p)
```
```{r}
ggsave(
  "Arp3Actin_Coloc.pdf",
  plot = last_plot(),
  device = "pdf")
```

```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
save.image(file=saved_path)

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
