---
title: "R Notebook"
output: html_notebook
---

<!--
# Lag cross-correlation of neurite-tip velocity vs soma actin (2D culture + cortical slice)

## What this file does

Reads kymograph TIFFs and per-cell neurite-tip / soma-actin CSVs (`list.files`
patterns at lines 59, 89, 94) from 2D-culture and cortical-slice DIV-1
movies of unpolarized neurons. For each neuron the per-neurite extension /
retraction profile is computed and the lag cross-correlation between
neurite-tip velocity and the change of soma actin intensity is pooled across
the cohort. Six PDFs are written (`ggsave` at lines 203, 251, 374, 448, 612,
736) including the CCF-curve PDFs for Fig 1e/f and the extending-neurite
distribution PDFs.

## Manuscript panels

- **Fig 1e**: "Neurite extension-retraction cross-correlation in cultured
  neurons. Values are plotted as mean +/- SD. N = 101 neurons from 4
  experiments."
- **Fig 1f**: "Neurite extension-retraction cross-correlation in acute
  cortical slice culture. Values are plotted as mean +/- SD. N = 59 neurons
  from 3 experiments."
- **Fig 1k**, **Fig 1l**, **ED Fig 1b**, **ED Fig 1c** - additional panels
  fed by this multi-section file.

Source Data: `Fig 1e`, `Fig 1f`, `Fig 1k`, `Fig 1l`, `ED Fig 1b`, `ED Fig 1c`.

## Outputs

- Six PDFs (`ggsave` at lines 203, 251, 374, 448, 612, 736).

## Original G:\ source

G:/Figure 1_coordination/Fig. 1f/WT/R_scripts/Neurite activity corr.Rmd
-->

## Input kymograph tiff files
```{r}
# BiocManager::install("bioimagetools")
# BiocManager::install("EBImage")
# devtools::install_github("BodenmillerGroup/cytomapper", build_vignettes = TRUE)

library(magick)
library("bioimagetools")
library("EBImage")

saved_img <- paste(Path_1, "kymogram_0.tif",  sep = "/")
cell <- bioimagetools::readTIF(saved_img)
par(pty="s")
bioimagetools::img(cell, z=1 ,col="b")

y = readImage(saved_img)
display(y)
```

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

```{r message=FALSE, warning=FALSE, include=FALSE}
library(pracma)
library(peakPick)
library(quantmod)
library(ggpmisc)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggfortify)
library(forecast)
library(tsibble)
library(broom)
library(tidyquant)
library(tsbox)
library("pheatmap")
library("RColorBrewer")
library(ggpmisc)
library(ggfortify)
library(viridis)
library(hrbrthemes)
library(data.table)
```

## load existing/saved environment (.RData)
```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(Path_1)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
 load(file=saved_path)
```

File sorting into folders
```{r}
file_list <- list.files(path =Path_1, full.names = TRUE, recursive = TRUE)
# subfolder_list<- subfolder_list [ grepl("Kymo_", subfolder_list) ]

  library(filesstrings)
for (n in 11:13)  {
 
   x <- paste("Cell_", n,  sep = "")
  dir.create(x)
  path<- paste(Path_1, "/", x, sep = "")
 file_tif <- file_list[grep(n,  file_list), drop = FALSE]

file.move(file_tif, path)
  
}

```

## Read .csv files
```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list<- subfolder_list [ grepl("Cell", subfolder_list) ]

csv_list<- list()
df_list<- list()

pixel <- 0.3243227

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list[[n]] <- list.files(subfolder_list[n],pattern = "(.*)csv$", 
                     recursive = TRUE,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list[[n]], read.csv, header = TRUE))
  df <-df[, grep(c("Tip.point|Length"),  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  colnames(df)<- paste("Length", seq( from = 1, to = ncol(df) ),  sep = ".")
  
  df_list[[n]]<- df
 
}

names(df_list)<- subfolder_list
```

## Import time-series neurite length data
```{r}

# multivariate time series: multiple neurites
mts_list<- list()
mts_avg_list <- list()
mts_stationary_list <- list()

freq <- 0.2 

for (n in 1:length(df_list))  {
 
  length_mat<-df_list[[n]] %>% as.data.frame()
  length_mat$sum_length <- length_mat %>% rowSums() 
  
  mts_list[[n]]<- length_mat %>% ts(frequency = freq, start = c(0, 1)) 
  
  # Moving Average to smoothen out outliers/noise
  mts_avg_list[[n]] <- length_mat %>% forecast::ma(3) %>% ts(frequency = freq, start = c(0, 1)) 
  colnames(mts_avg_list[[n]])<-colnames(length_mat)
  
  # Differential 
  mts_stationary_list[[n]] <- mts_avg_list[[n]] %>% MTS::diffM(d=1) %>% ts(frequency = freq, start = c(0, 1))
  
}

names(mts_list)<- names(df_list)
names(mts_avg_list)<- names(df_list)
names(mts_stationary_list)<- names(df_list)

```

## Plotting the time series as curves
```{r}
library(ggthemes)

mts_profile_list <- list()
avg_profile_list <- list()
sta_profile_list <- list()
thres= 0.3 # change threshold for stationary profile here

for (n in 1:length(mts_avg_list))  {
  
 plot_title<- paste(names(df_list)[n], sep = "")
 selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(mts_avg_list[[n]]))

 theme_set(theme_bw())
 avg_profile_list[[n]] <- mts_avg_list[[n]][, selected_data] %>%  autoplot() +
  #ggtitle (plot_title) +
   theme_linedraw()+
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 45))+
  geom_line(aes(),size=1)  +
   scale_color_viridis(discrete=T)+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_text(size=12)) #for centering the text
 
 mts_profile_list[[n]] <- mts_list[[n]][, selected_data] %>%  autoplot() +
 # ggtitle (plot_title) +
    theme_linedraw()+
  geom_line(aes(),size=1)  +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold")) #for centering the text
 
 sta_profile_list[[n]] <- mts_stationary_list[[n]][, selected_data] %>%  autoplot() +
  #ggtitle (plot_title) +
    theme_linedraw()+
  geom_line(aes(),size=0.25)  +
   geom_hline(yintercept=0, linetype="dashed", color = "red")+
   geom_hline(yintercept=thres, linetype="dashed", color = "blue")+
   geom_hline(yintercept=-thres, linetype="dashed", color = "blue")+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold")) #for centering the text
 
}

names(mts_profile_list) <- names(df_list)
names(avg_profile_list) <- names(df_list)
names(sta_profile_list) <- names(df_list)
```

## Visualization of the selected curves
```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=10}
fig_A <- mts_profile_list[[2]]
fig_B <- avg_profile_list[[2]]
fig_C <- sta_profile_list[[2]]

ggpubr::ggarrange(fig_B, fig_C,
                  labels = c("B", "C"),
                  ncol = 1, nrow = 2,  legend="none")

```

```{r}
ggsave(
  "avg_2.eps",
  plot = last_plot(),
  device = "eps")
```

## We convert the multiple times series data to appropriate dataframe for stack area plot
```{r}

selected_list <-mts_stationary_list
stacked_avg_list <- list()

for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  
  colNames <- colnames(selected_list[[m]][ ,selected_data])
  length_mat<- selected_list[[m]][, selected_data] %>%  forecast::ma(3) %>% as.data.frame() # add forecast::ma(5) if smoothening is wanted
  colnames(length_mat) <- colNames 
  
  mts_dim<- dim(length_mat)
  length_mat$Time <- as.numeric(rep(seq(0,mts_dim[1]-1)* 1/freq), times=mts_dim[2] )
  
 stacked_avg_list [[m]] <- length_mat %>% gather(Neurite, Length, -Time) %>% as.data.table()
  
}
```

```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=10}
stacked_avg_list[[2]] %>% 
  ggplot(aes(x = Time, y = Length, fill = Neurite)) +
    geom_area(alpha=1 ) +
  theme_linedraw()+
  
    scale_fill_viridis(discrete = T) + 
  scale_y_continuous( limits = c(-1.2, 1.8), breaks = scales::pretty_breaks(n = 7))+
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 

  # scale_fill_brewer(palette = "Set2") +
   # theme_ipsum(grid = F, ticks = T) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none", 
        axis.text=element_text(size=12)) #for centering the text
```

```{r}
ggsave(
  "avg_2.eps",
  plot = last_plot(),
  device = "eps")
```

## Calculate neurite outgrowth/retraction events from the stationary/differential curves

```{r}
selected_list <- mts_stationary_list # mts_stationary_list or mts_avg_list

pos_neg_thres_list <- list()
# thres= 0.02 # change threshold here

for (n in 1: 10){
  
  thres= n*0.05
  pos_neg_list <- list()
  
  for (m in 1:length(selected_list)) {
    
    selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
    length_mat<- selected_list[[m]][, selected_data] %>% as.data.frame()
    
    length_mat2 <- length_mat
    neurite_num <- ncol(length_mat)
  
    tab1 <- apply(length_mat, 1, function(x) sum(x > thres, na.rm = TRUE)) %>%  table() %>% as.data.table()
    tab1$act<- "extension"
    names(tab1)[1]<- "Num"
    tab1$Num <- tab1$Num %>% as.numeric()
    tab1$pct<- tab1$N/sum(tab1$N)
    tab1$rel_n <- tab1$Num/neurite_num
    tab1$Cell<- names(df_list)[m]
    
    tab2 <- apply(length_mat, 1, function(x) sum(x < -thres, na.rm = TRUE)) %>%  table() %>% as.data.table()
    tab2$act<- "retraction"
    names(tab2)[1]<- "Num"
    tab2$Num <- tab2$Num %>% as.numeric()
    tab2$pct<- tab2$N /sum(tab2$N)
    tab2$rel_n <- tab2$Num/neurite_num
    tab2$Cell<- names(df_list)[m]
  
    tab3 <- apply(length_mat, 1, function(x) sum(x < thres & x > -thres, na.rm = TRUE)) %>%  table() %>% as.data.table()
    tab3$act<- "pause"
    names(tab3)[1]<- "Num"
    tab3$Num <- tab3$Num %>% as.numeric()
    tab3$pct<- tab3$N/sum(tab3$N)
    tab3$rel_n <- tab3$Num/neurite_num
    tab3$Cell<- names(df_list)[m]
    
    tab4 <- rbind(tab1, tab2, tab3)
    tab4$Thres<- thres
    
    pos_neg_list[[m]] <- tab4
  }

  names(pos_neg_list)<- names(df_list)
  
  pos_neg_thres_list[[n]]<- assign(paste0("Thres_", thres, sep=""),pos_neg_list)
  names(pos_neg_thres_list)[n]<- paste0("Thres_", thres, sep="")
  
}
```

```{r message=FALSE, warning=FALSE}
summary_list <- list()

for (n in 1: 10){
  
  pos_neg_list <- pos_neg_thres_list[[n]]
  
  pos_neg_tab_all <- pos_neg_list[[1]] 
  
  for (m in 2:length(pos_neg_list)) {
    
    df2 <- pos_neg_list[[m]] 
    pos_neg_tab_all <- rbind(pos_neg_tab_all, df2) 
  
  }

  tab <- pos_neg_tab_all[act == "extension"] %>% group_by(Num) %>% rstatix::get_summary_stats(pct)
  tab$Thres<- n*0.05
  summary_list [[n]] <- tab 

}

names(summary_list) <-  names(pos_neg_thres_list)

df_summary_all <- summary_list [[1]]

for (n in 2: 10){
  df_summary_all <-  rbind (df_summary_all, summary_list [[n]])
  
}

```

```{r  message = FALSE, warning=FALSE, fig.width=10}

df_summary_all <- df_summary_all %>% as.data.table()
ggplot(df_summary_all, aes(x=Num, y=mean, group=Thres)) +
  
  geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd, fill=Thres),   alpha=0.05) + 
  geom_line(aes(colour=Thres), size=1, alpha=1) +
  
  scale_y_continuous(  breaks = scales::pretty_breaks(n = 6))+
  scale_x_continuous(name="Neurite_No", limits=c(0,4), expand = c(0, 0), breaks = c(0,1,2,3,4,5,6))+
  
  scale_fill_viridis(discrete = F) + # scale_fill_brewer(palette = "Set2") +
  scale_color_viridis(discrete = F)+

  theme_linedraw()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12), 
        legend.position = "none") #for centering the text
```

```{r}
ggsave(
  "extension_freq.pdf",
  plot = last_plot(),
  device = "pdf")
```

```{r}
pos_neg_list <- pos_neg_thres_list$Thres_0.5

pos_neg_tab_all <- pos_neg_list[[1]] 

for (m in 2:length(pos_neg_list)) {
  
  df2 <- pos_neg_list[[m]] 
  pos_neg_tab_all <- rbind(pos_neg_tab_all, df2) 

}

tab <- pos_neg_tab_all[act == "extension"]

df_summary <- tab %>% group_by(Num) %>% rstatix::get_summary_stats(pct)

```

```{r}
tab %>% write_csv("extension_freq.csv")
```

```{r  message = FALSE, warning=FALSE, fig.width=10}

ggplot(df_summary, aes(x=Num, y=mean)) +
  
  geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd), fill="lightgrey", alpha=1)+

  geom_line(size=1, alpha=1) +
  scale_y_continuous(  breaks = scales::pretty_breaks(n = 6))+
  scale_x_continuous(name="Neurite_No", limits=c(0,4), expand = c(0, 0), breaks = c(0,1,2,3,4,5,6))+
  
  scale_fill_viridis(discrete = F) + # scale_fill_brewer(palette = "Set2") +
  scale_color_viridis(discrete = F)+

  theme_linedraw()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12), 
        legend.position = "none") #for centering the text
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
# remove cells in "without any outgrowth" state more than 60% of the time 
removed <- pos_neg_tab_all[act == "extension" & Num == 0 & pct > 0.5] 
removed_cell<- removed$Cell
pos_neg_tab_all_removed<- pos_neg_tab_all[grep(paste(removed_cell,collapse="|"), pos_neg_tab_all$Cell , invert=TRUE) ,]
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
ggplot(data=pos_neg_tab_all_removed, aes(x=Num, y=pct, group=Cell)) +
  geom_line((aes(color=Cell)))+
  geom_point((aes(color=Cell)))+ 
scale_x_continuous(name="Neurite_No", limits=c(0, 6), breaks = c(0,1,2,3,4,5,6))+
theme_linedraw()+
  
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12), 
        legend.position = "none") #for centering the text
```

```{r}
ggsave(
  "WT_neuriteNumFreq_avg.eps",
  plot = last_plot(),
  device = "eps")
```

## We smoothen the stationary profiles first, separete outrgwoth and retraction events, and sum the changes of all neurites. (Smoothening -> separate -> sum)
```{r warning=FALSE}

# Smoothen the stationary (diff) profiles and optionally filter out small length change with threshold

selected_list <- mts_stationary_list # mts_stationary_list or mts_avg_list

mts_stationary_pos_list <- list()
mts_stationary_neg_list <- list()
thres= 0.0 # change threshold here

# Positive peaks only (outgrowth event)
for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  length_mat<- selected_list[[m]][, selected_data ]
  
  length_mat_avg<- length_mat %>% forecast::ma(3) %>% as.ts() # Average smoothening
  colnames(length_mat_avg)<-colnames(length_mat)

  mts_dim<- dim(length_mat_avg)
    for (n in 1:mts_dim[2]){
      length_mat_avg[, n][length_mat_avg[ , colnames(length_mat_avg)[n]] < thres] <- 0 
    }
  
  length_mat_avg <- length_mat_avg %>% as.data.frame() 
  length_mat_avg$sum <- length_mat_avg  %>% rowSums()  # We sum up the extent of outgrowth at each time point
  mts_stationary_pos_list[[m]] <- length_mat_avg %>% ts(frequency = freq, start = c(0, 1)) 

}

# Negative peaks only (retraction event)
for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  length_mat<- selected_list[[m]][, selected_data ]
  length_mat_avg<- length_mat %>% forecast::ma(3) %>% as.ts() # Average smoothening
  colnames(length_mat_avg)<-colnames(length_mat)
  
  mts_dim<- dim(length_mat_avg)
    for (n in 1:mts_dim[2]){
      length_mat_avg[, n][length_mat_avg[ , colnames(length_mat_avg)[n]] > - thres] <- 0 
    }
  length_mat_avg <- length_mat_avg %>% as.data.frame() 
  length_mat_avg$sum <- length_mat_avg  %>% rowSums()  # We sum up the extent of retraction at each time point
  mts_stationary_neg_list[[m]] <- length_mat_avg %>% ts(frequency = freq, start = c(0, 1)) 
}

```

```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=10}
selected_stationary <- mts_stationary_pos_list[[15]]
selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_stationary))

figA <- selected_stationary[ ,selected_data] %>%  autoplot() +
  geom_line(aes(),size=0.25)  +
   geom_hline(yintercept=0, linetype="dashed", color = "red")+
   geom_hline(yintercept=thres, linetype="dashed", color = "blue")+
   geom_hline(yintercept=-thres, linetype="dashed", color = "blue")+
   scale_color_viridis(discrete=T)+
   theme_minimal()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold"))

selected_stationary <- mts_stationary_neg_list[[15]]

figB <- selected_stationary[ ,selected_data] %>%  autoplot() +
   geom_line(aes(),size=0.25)  +
   geom_hline(yintercept=0, linetype="dashed", color = "red")+
   geom_hline(yintercept=thres, linetype="dashed", color = "blue")+
   geom_hline(yintercept=-thres, linetype="dashed", color = "blue")+
   scale_color_viridis(discrete=T)+
   theme_minimal()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold")) 
 
 ggpubr::ggarrange(figA, figB, 
                  labels = c("A", "B"),
                  ncol = 1, nrow = 2,  legend="none")
 
```

## We convert the multiple times series data to appropriate dataframe for stack area plot
```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=10}

# Positive peaks only (outgrowth event)
selected_list <- mts_stationary_pos_list
stacked_stationary_pos_list <- list()

for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  length_mat<- selected_list[[m]][, selected_data] %>% as.data.frame() 
  mts_dim<- nrow(length_mat)-1
  length_mat$Time <- as.numeric(rep(seq(0,mts_dim)*2))
  
  stacked_stationary_pos_list[[m]] <- length_mat %>% gather(Neurite, Length, -Time) %>% as.data.table()

}

# Negtive peaks only (retraction event)
selected_list <- mts_stationary_neg_list
stacked_stationary_neg_list <- list()

for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  length_mat<- selected_list[[m]][, selected_data] %>% as.data.frame() 
   mts_dim<- nrow(length_mat)-1
  length_mat$Time <-  as.numeric(rep(seq(0,mts_dim)*2))
  
stacked_stationary_neg_list[[m]] <- length_mat %>% gather(Neurite, Length, -Time) %>% as.data.table()
}

```

```{r warning=FALSE}
figA <-  stacked_stationary_pos_list[[2]] %>% 
  ggplot(aes(x = Time, y = Length, fill = Neurite)) +
    geom_area(alpha=0.5 , size=.5, colour="white") +
    scale_fill_viridis(discrete = T) + # scale_fill_brewer(palette = "Set2") +
    theme_ipsum(grid = F, ticks = T) 

figB <- stacked_stationary_neg_list[[2]] %>% 
  ggplot(aes(x = Time, y = Length, fill = Neurite)) +
    geom_area(alpha=0.5 , size=.5, colour="white") +
    scale_fill_viridis(discrete = T) + # scale_fill_brewer(palette = "Set2") +
    theme_ipsum(grid = F, ticks = T) 
 
 ggpubr::ggarrange(figA, figB, 
                  labels = c("A", "B"),
                  ncol = 1, nrow = 2,  legend="none")
 
```

# We plot the overall outgrowth/retraction level over time
```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=10}
selected_stationary_pos <- mts_stationary_pos_list[[2]][, "sum"]
selected_stationary_neg <- mts_stationary_neg_list[[2]][, "sum"]
ts.union(selected_stationary_pos, selected_stationary_neg) %>%  autoplot() +
  geom_line(aes(),size=0.25)  +
   geom_hline(yintercept=0, linetype="dashed", color = "black")+
   #geom_hline(yintercept=thres, linetype="dashed", color = "blue")+
   #geom_hline(yintercept=-thres, linetype="dashed", color = "blue")+
   scale_y_continuous( limits = c(-1.2, 1.8), breaks = scales::pretty_breaks(n = 7))+
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text
```

```{r}
ggsave(
  "avg_2.eps",
  plot = last_plot(),
  device = "eps")

```

```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=10}
ccf(selected_stationary_pos, selected_stationary_neg,  type = "correlation", na.action = na.pass, lag.max=100) 
# The lines give the values beyond which the autocorrelations are (statistically) significantly different from zero.

```

## Here we separate the outrgwoth and retraction events first, sum the changes of all neurites, and then do smoothening. (separate -> sum -> smoothening)
```{r}

selected_list <- mts_stationary_list 
mts_stationary_sum_list <- list()

thres= 0.0 # change threshold here

# Positive peaks only (outgrowth event)
for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]])) # Neurite length ("Length") only or include actin intensity (Integerated.intensity)
  length_mat_pos<- selected_list[[m]][, selected_data ]
  length_mat_neg<- selected_list[[m]][, selected_data ]
  
  mts_dim<- dim(length_mat_pos)
  
    for (n in 1:mts_dim[2]){
      length_mat_pos[, n][length_mat_pos[ , colnames(length_mat_pos)[n]] < 0] <- 0 
      length_mat_neg[, n][length_mat_neg[ , colnames(length_mat_neg)[n]] > 0] <- 0 
    }
  
  length_mat_pos <- as.data.frame(length_mat_pos)
  length_mat_neg <- as.data.frame(length_mat_neg)
   
  length_mat_pos$sum <- length_mat_pos  %>% rowSums()  
  length_mat_neg$sum <- length_mat_neg  %>% rowSums()  
  
  length_mat_pos <- length_mat_pos %>% ts(frequency = freq, start = c(0, 1))
  length_mat_neg <- length_mat_neg %>% ts(frequency = freq, start = c(0, 1))
  
  length_mat_sum <- ts.union(length_mat_pos[ ,"sum"], length_mat_neg[ ,"sum"]) 
  
  length_mat_sum <- length_mat_sum %>% forecast::ma(3) %>% as.ts() # Average smoothening
  colnames(length_mat_sum)<-c("Pos", "Neg")
  
  length_mat_sum [, "Pos"][length_mat_sum [, "Pos"] < thres] <- 0 
  length_mat_sum [, "Neg"][length_mat_sum [, "Neg"] > -thres] <- 0 
  
  mts_stationary_sum_list[[m]] <- length_mat_sum

}

names(mts_stationary_sum_list)<- names(mts_stationary_list )

```

```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=10}
selected_stationary_sum <- mts_stationary_sum_list[[2]]

 selected_stationary_sum %>%  autoplot() +
  geom_line(aes(),size=0.25)  +
   geom_hline(yintercept=0, linetype="dashed", color = "red")+
   geom_hline(yintercept=thres, linetype="dashed", color = "blue")+
   geom_hline(yintercept=-thres, linetype="dashed", color = "blue")+
   theme_ipsum(grid = F, ticks = T) + 
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold"),
        legend.position = "none")  #for centering the text
 # stat_peaks(colour = "red", ignore_threshold = 0.1) +
 # stat_peaks(geom = "text", colour = "red", vjust = -0.5) +
 # stat_valleys(colour = "blue", ignore_threshold = 0.1) + 
 # stat_valleys(geom = "text", colour = "blue", angle = 45, vjust = 1.5, hjust = 1)
```

# Time-lagged cross-corrleation of outgrowth and retraction profiles
```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=10}

length_ccf_list <- list()

for (m in 1:length(mts_stationary_sum_list)) {
  
  selected_length_sta_sum <- mts_stationary_sum_list[[m]]
  
 # The lines give the values beyond which the autocorrelations are (statistically) significantly different from zero.

  length_ccf_plot <-ccf (selected_length_sta_sum[, "Pos"],  
                            -selected_length_sta_sum[, "Neg"], 
                            type = "correlation", 
                            na.action = na.pass, plot=F,
                            lag.max=20) 
  
  length_ccf_list[[m]] <- length_ccf_plot 
  
}

names(length_ccf_list) <-  names(mts_stationary_sum_list)

length_ccf <- length_ccf_list[[1]][["acf"]] %>% as.data.frame()
for (m in 2:length(length_ccf_list)) {
  
  length_ccf <- cbind(length_ccf , length_ccf_list[[m]][["acf"]])
}

colnames(length_ccf) <- names(length_ccf_list)

length_ccf$lag<- c(-20:20)

```

```{r  message = FALSE, warning=FALSE, fig.width=10}
selected_ccf <- length_ccf %>% as.data.table()
selected_ccf$lag<- selected_ccf$lag/freq 

stacked_ccf_mat <- selected_ccf %>% gather(Cell, ccf, -lag) # select which CCF of measurements to show
df_summary <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf)
```

```{r}
ggsave(
  "length_ccf.pdf",
  plot = last_plot(),
  device = "pdf")

```
```{r}
selected_ccf2 %>%  write_csv("length_ccf.csv")
```
```{r}
ggsave(
  "length_ccf.eps",
  plot = last_plot(),
  device = "eps")

```
# Time-lagged cross-corrleation of outgrowth and retraction profiles
```{r warning=FALSE}
# constructing partial dependence (i.e., marginal effect) plots 
library(pdp)
theme_set(theme_bw())
mts_avg_list[[2]][ ,c("Length.2", "Length.4")] %>%  autoplot() +
  ylim(0, 45) +
  ggtitle (plot_title) +
  geom_line(aes(),size=1)  +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold")) #for centering the text
```

```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
save.image(file=saved_path)

```

# Identify peak and valley

## moving/rolling window cross-correlation
```{r warning=FALSE}
save.image(saved_path)
```

## windowed Time Lagged Cross Correlation (wTLCC) 
```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
save.image(saved_path)
```

## Package installation
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
