---
title: "R Notebook"
output: html_notebook
---

<!--
# Lag cross-correlation between neurite-tip velocity and tip actin at polarization transition

## What this file does

Reads per-neurite tip-actin and neurite-actin CSVs (`TipActin*.csv`,
`NeuriteActin*.csv` from per-cell subfolders; lines 65, 97, 130) for neurons
at the transition to polarization (canonical published cohort). For each
neurite the local-detrended growth velocity is cross-correlated against the
change of actin intensity at the tip. CCF curves are pooled to a mean +/- SD
across the cohort. Six PDFs are written (`ggsave` at lines 877, 1614, 1928,
2251, 2558, 2606).

## Manuscript panel

**Fig 2l**: "Cross-correlation between neurite growth velocity and the change
of actin intensity at neurite tips in the neurons at the transition to
polarization. Data source is the same as (j)."

Source Data: `Fig 2l`, `ED Fig 2l`.

## Outputs

- Six PDFs (`ggsave` at lines 877, 1614, 1928, 2251, 2558, 2606).

## Original G:\ source

G:/Figure 2_actin wave/Fig. 2g/S2S3_transition/Neurite activity corr.Rmd
-->

```{r message=FALSE, warning=FALSE, include=FALSE}
library(pracma)
#library(peakPick)
library(quantmod)
library(ggpmisc)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggfortify)
library(forecast)
library(tsibble)
library(broom)
library(tidyquant)
library(tsbox)
library("pheatmap")
library("RColorBrewer")
library(ggpmisc)
library(ggfortify)
library(viridis) # Default Color Maps from 'matplotlib' 
library(hrbrthemes)
library(data.table)
library(ggpubr)
library(superheat)
```

## load existing/saved environment (.RData)
```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
load(file=saved_path)
```

# save environment (.RData)
```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
save.image(file=saved_path)
```

# Read .csv files

## 170905_Max
```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("170905", subfolder_list) ]

length_list_170905_Max <- list()
freq <- 0.5 # set frequency (1/interval)
pixel <- 0.1609 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  
 csv_list <- list.files(subfolder_list [n],pattern = "^TipActin(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Length|Tip",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  colnames(df)<- paste("Length", seq( from = 1, to = ncol(df) ),  sep = ".")
  
  length_list_170905_Max[[n]] <- df %>% as.data.table()
}

subfolder_name_list <- subfolder_list %>% fs::path_file()

names(length_list_170905_Max) <- gsub("Kymo_", "", subfolder_name_list)
```

```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("170905", subfolder_list) ]

Tip_actin_list_170905_Max <- list()
freq <- 0.5 # set frequency (1/interval)
pixel <- 0.1609 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  
 csv_list <- list.files(subfolder_list [n],pattern = "^TipActin(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, 1/10000)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  colnames(df)<- paste("Tip_actin", seq( from = 1, to = ncol(df) ),  sep = ".")
  
  Tip_actin_list_170905_Max[[n]] <- df %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()

names(Tip_actin_list_170905_Max) <- gsub("Kymo_", "", subfolder_name_list)
```

```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("170905", subfolder_list) ]

Neurite_actin_list_170905_Max <- list()
freq <- 0.5 # set frequency (1/interval)
pixel <- 0.1609 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  
 csv_list <- list.files(subfolder_list [n],pattern = "^NeuriteActin(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, 1/10000)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  colnames(df)<- paste("Neurite_actin", seq( from = 1, to = ncol(df) ),  sep = ".")
  
  Neurite_actin_list_170905_Max[[n]] <- df %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()

names(Neurite_actin_list_170905_Max) <- gsub("Kymo_", "", subfolder_name_list)
```

## 170921_DVCore
```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("170921", subfolder_list) ]

length_list_170921_DVCore <- list()
freq <- 0.5 # set frequency (1/interval)
pixel <- 0.1609 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Actin_tip_(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
   df <-df[, grep("Length|Tip",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  colnames(df)<- paste("Length", seq( from = 1, to = ncol(df)  ),  sep = ".")
  
  length_list_170921_DVCore[[n]] <-df %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(length_list_170921_DVCore) <- gsub("Kymo_", "", subfolder_name_list)
```

```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("170921", subfolder_list) ]

Tip_actin_list_170921_DVCore <- list()
freq <- 0.5 # set frequency (1/interval)
pixel <- 0.1609 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Actin_tip_(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, 1/10000)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  colnames(df)<- paste("Tip_actin", seq( from = 1, to = ncol(df)  ),  sep = ".")
  
  Tip_actin_list_170921_DVCore[[n]] <-df %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(Tip_actin_list_170921_DVCore) <- gsub("Kymo_", "", subfolder_name_list)
```

```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("170921", subfolder_list) ]

Neurite_actin_list_170921_DVCore <- list()
freq <- 0.5 # set frequency (1/interval)
pixel <- 0.1609 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Actin_neurite_(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, 1/10000)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  colnames(df)<- paste("Neurite_actin", seq( from = 1, to = ncol(df)  ),  sep = ".")
  
  Neurite_actin_list_170921_DVCore[[n]] <-df %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(Neurite_actin_list_170921_DVCore) <- gsub("Kymo_", "", subfolder_name_list)
```

## 180129
```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("180129", subfolder_list) ]

length_list_180129 <- list()
freq <- 0.5 # set frequency (1/interval)
pixel <- 0.0648 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "Result_(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Length|Tip",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  colnames(df)<- paste("Length", seq( from = 1, to = ncol(df)  ),  sep = ".")
   
  length_list_180129 [[n]] <-df %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(length_list_180129) <- gsub("Kymo_", "", subfolder_name_list)
```

```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("180129", subfolder_list) ]

Tip_actin_list_180129 <- list()
freq <- 0.5 # set frequency (1/interval)
pixel <- 0.0648 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "Result_(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, 1/10000)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  colnames(df)<- paste("Tip_actin", seq( from = 1, to = ncol(df)  ),  sep = ".")
   
  Tip_actin_list_180129 [[n]] <-df %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(Tip_actin_list_180129) <- gsub("Kymo_", "", subfolder_name_list)
```

```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("180129", subfolder_list) ]

Neurite_actin_list_180129 <- list()
freq <- 0.5 # set frequency (1/interval)
pixel <- 0.0648 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "Actin_neurite_(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, 1/10000)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  colnames(df)<- paste("Neurite_actin", seq( from = 1, to = ncol(df)  ),  sep = ".")
   
  Neurite_actin_list_180129 [[n]] <-df %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(Neurite_actin_list_180129) <- gsub("Kymo_", "", subfolder_name_list)
```

## 180205
```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("180205", subfolder_list) ]

length_list_180205 <- list()
freq <- 0.25# set frequency (1/interval)
pixel <- 0.0648 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Result_(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Length|Tip",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  df <-pracma::hampel(df, k=3)$y  %>% as.data.table() # remove jumping spikes
  colnames(df)<- paste("Length", seq( from = 1, to = ncol(df)  ),  sep = ".")
  
  # Interpolation of time series data to match temporal resolution
  df$time<- (0:(nrow(df)-1)*4)
  newTime <- seq(0,960, by=2) %>% as.data.frame()
  names(newTime) <- "time"
  # Replace NA by Interpolation
  df <- left_join(x=newTime,y=df,by=c("time")) %>% zoo::na.approx() %>% as.data.table()
  
  length_list_180205[[n]] <-df[ ,-1] %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(length_list_180205) <- gsub("Kymo_", "", subfolder_name_list)
```

```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("180205", subfolder_list) ]

Tip_actin_list_180205 <- list()
freq <- 0.25# set frequency (1/interval)
pixel <- 0.0648 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Result_(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, 1/10000)
  df <-pracma::hampel(df, k=3)$y  %>% as.data.table() # remove jumping spikes
  colnames(df)<- paste("Tip_actin", seq( from = 1, to = ncol(df)  ),  sep = ".")
  
  # Interpolation of time series data to match temporal resoultion
  df$time<- (0:(nrow(df)-1)*4)
  newTime <- seq(0,960, by=2) %>% as.data.frame()
  names(newTime) <- "time"
  # Replace NA by Interpolation
  df <- left_join(x=newTime,y=df,by=c("time")) %>% zoo::na.approx() %>% as.data.table()
  
  Tip_actin_list_180205[[n]] <-df[ ,-1] %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(Tip_actin_list_180205) <- gsub("Kymo_", "", subfolder_name_list)
```

```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("180205", subfolder_list) ]

Neurite_actin_list_180205 <- list()
freq <- 0.25# set frequency (1/interval)
pixel <- 0.0648 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Actin_neurite_(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, 1/10000)
  df <-pracma::hampel(df, k=3)$y  %>% as.data.table() # remove jumping spikes
  colnames(df)<- paste("Neurite_actin", seq( from = 1, to = ncol(df)  ),  sep = ".")
  
  # Interpolation of time series data to match temporal resolution
  df$time<- (0:(nrow(df)-1)*4)
  newTime <- seq(0,960, by=2) %>% as.data.frame()
  names(newTime) <- "time"
  # Replace NA by Interpolation
  df <- left_join(x=newTime,y=df,by=c("time")) %>% zoo::na.approx() %>% as.data.table()
  
  Neurite_actin_list_180205[[n]] <-df[ ,-1] %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(Neurite_actin_list_180205) <- gsub("Kymo_", "", subfolder_name_list)
```

## 180207
```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("180207", subfolder_list) ]

length_list_180207 <- list()
freq <- 0.5# set frequency (1/interval)
pixel <- 0.0648 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Result_(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Length|Tip",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  colnames(df)<- paste("Length", seq( from = 1, to = ncol(df)  ),  sep = ".")
  
  length_list_180207[[n]] <-df %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(length_list_180207) <- gsub("Kymo_", "", subfolder_name_list)
```

```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("180207", subfolder_list) ]

Tip_actin_list_180207 <- list()
freq <- 0.5# set frequency (1/interval)
pixel <- 0.0648 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Result_(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, 1/10000)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  colnames(df)<- paste("Tip_actin", seq( from = 1, to = ncol(df)  ),  sep = ".")
  
  Tip_actin_list_180207[[n]] <-df %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(Tip_actin_list_180207) <- gsub("Kymo_", "", subfolder_name_list)
```

```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("180207", subfolder_list) ]

Neurite_actin_list_180207 <- list()
freq <- 0.5# set frequency (1/interval)
pixel <- 0.0648 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Actin_neurite_(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, 1/10000)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  colnames(df)<- paste("Neurite_actin", seq( from = 1, to = ncol(df)  ),  sep = ".")
  
  Neurite_actin_list_180207[[n]] <-df %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(Neurite_actin_list_180207) <- gsub("Kymo_", "", subfolder_name_list)
```

## 240604
```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("240604", subfolder_list) ]

length_list_240604 <- list()
freq <- 1 # set frequency (1/interval)
pixel <- 0.1625 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Tip_actin(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Length|Tip",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  
    row_num <- nrow(df)
    df <- df[seq(0, row_num , by=2), ] # define the subsampling freq
    
  colnames(df)<- paste("Length", seq( from = 1, to = ncol(df)  ),  sep = ".")
  
  length_list_240604[[n]] <-df %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(length_list_240604) <- gsub("Kymo_", "", subfolder_name_list)
```

```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("240604", subfolder_list) ]

Tip_actin_list_240604 <- list()
freq <- 1 # set frequency (1/interval)
pixel <- 0.1625 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Tip_actin(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, 1/10000)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  
    row_num <- nrow(df)
    df <- df[seq(0, row_num , by=2), ] # define the subsampling freq
    
  colnames(df)<- paste("Tip_actin", seq( from = 1, to = ncol(df)  ),  sep = ".")
  
  Tip_actin_list_240604[[n]] <-df %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(Tip_actin_list_240604) <- gsub("Kymo_", "", subfolder_name_list)
```

```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("240604", subfolder_list) ]

Neurite_actin_list_240604 <- list()
freq <- 1 # set frequency (1/interval)
pixel <- 0.1625 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Neurite_actin(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, 1/10000)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  
    row_num <- nrow(df)
    df <- df[seq(0, row_num , by=2), ] # define the subsampling freq
    
  colnames(df)<- paste("Neurite_actin", seq( from = 1, to = ncol(df)  ),  sep = ".")
  
  Neurite_actin_list_240604[[n]] <-df %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(Neurite_actin_list_240604) <- gsub("Kymo_", "", subfolder_name_list)
```

## 240702
```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("240702", subfolder_list) ]

length_list_240702 <- list()
freq <- 1 # set frequency (1/interval)
pixel <- 0.1625 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Actin_tip(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Length|Tip",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  
    row_num <- nrow(df)
    df <- df[seq(0, row_num , by=2), ] # define the subsampling freq
    
  colnames(df)<- paste("Length", seq( from = 1, to = ncol(df)  ),  sep = ".")
  
  length_list_240702[[n]] <-df %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(length_list_240702) <- gsub("Kymo_", "", subfolder_name_list)
```

```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("240702", subfolder_list) ]

Tip_actin_list_240702 <- list()
freq <- 1 # set frequency (1/interval)
pixel <- 0.1625 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Actin_tip(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, 1/10000)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  
    row_num <- nrow(df)
    df <- df[seq(0, row_num , by=2), ] # define the subsampling freq
    
  colnames(df)<- paste("Tip_actin", seq( from = 1, to = ncol(df)  ),  sep = ".")
  
  Tip_actin_list_240702[[n]] <-df %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(Tip_actin_list_240702) <- gsub("Kymo_", "", subfolder_name_list)
```

```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo_", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("240702", subfolder_list) ]

Neurite_actin_list_240702 <- list()
freq <- 1 # set frequency (1/interval)
pixel <- 0.1625 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list <- list.files(subfolder_list[n],pattern = "^Actin_neurite(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = T, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, 1/10000)
  df <-pracma::hampel(df, k=3)$y # remove jumping spikes
  
    row_num <- nrow(df)
    df <- df[seq(0, row_num , by=2), ] # define the subsampling freq
    
  colnames(df)<- paste("Neurite_actin", seq( from = 1, to = ncol(df)  ),  sep = ".")
  
  Neurite_actin_list_240702[[n]] <-df %>% as.data.table()

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(Neurite_actin_list_240702) <- gsub("Kymo_", "", subfolder_name_list)
```

```{r}
length_list <- c(length_list_170905_Max, length_list_170921_DVCore, length_list_180129, length_list_180205, length_list_180207, length_list_240604, length_list_240702)
Tip_actin_list <- c(Tip_actin_list_170905_Max, Tip_actin_list_170921_DVCore, Tip_actin_list_180129, Tip_actin_list_180205, Tip_actin_list_180207, Tip_actin_list_240604, Tip_actin_list_240702)
Neurite_actin_list <- c(Neurite_actin_list_170905_Max, Neurite_actin_list_170921_DVCore, Neurite_actin_list_180129, Neurite_actin_list_180205, Neurite_actin_list_180207, Neurite_actin_list_240604, Neurite_actin_list_240702)
```

# Actin density of the whole neurite ((tip + neurite shaft)/neurite length))
```{r}
Density_actin_list<- list()

for (n in 1:length(length_list))  {

  df <- (Neurite_actin_list[[n]])/(length_list[[n]])
  colnames(df)<- paste("density_actin", seq( from = 1, to = ncol(df) ),  sep = ".")
  df[sapply(df, is.infinite)] <- NA
  Density_actin_list[[n]]<- df
  
}

names(Density_actin_list) <- names(length_list)
```

```{r}
all_list <- list(length = length_list, 
                 tip_actin = Tip_actin_list,  
                 neurite_actin = Neurite_actin_list, 
                 Density_actin = Density_actin_list)
```

# read soma actin
```{r}
# set the directory where the CSV files are located
folder_path <- Path_1

# recursively find & read all CSV files in the directory and subdirectories
soma_list <- list.files(path = folder_path, pattern = "^soma_actin.csv$", recursive = TRUE, full.names = TRUE) %>% map(read.csv) %>% map(as.data.table)

names(soma_list) <- dirname(list.files(path = folder_path, pattern = "^soma_actin.csv$", recursive = TRUE, full.names = TRUE)) %>% basename()

names(soma_list$'170905_Max')[2:ncol(soma_list$'170905_Max')] <-  gsub("X", "", names(soma_list$'170905_Max')[2:ncol(soma_list$'170905_Max')])
names(soma_list$'170905_Max')[2:ncol(soma_list$'170905_Max')] <-  gsub("d.l", "d-l", names(soma_list$'170905_Max')[2:ncol(soma_list$'170905_Max')])
```

# upsampling or downsmapling soma data
```{r}
# upsampling soma data of "180205"
df <- soma_list$'180205'

# Interpolation of time series data to match temporal resoultion
df$time<- (0:(nrow(df)-1)*4)
newTime <- seq(0,960, by=2) %>% as.data.frame()
names(newTime) <- "time"
# Replace NA by Interpolation
df <- left_join(x=newTime,y=df,by=c("time")) %>% zoo::na.approx() %>% as.data.table()
df <-  df[,-(1:2)]
df$X <- 1:nrow(df)
soma_list$'180205' <- df
 
# downsampling soma data of "240604" 
df <- soma_list$'240604'
 
row_num <- nrow(df)
df <- df[seq(0, row_num , by=2), ] # define the subsampling freq
df$X <- 1:nrow(df)
soma_list$'240604' <- df

# downsampling soma data of "240702" 
df <- soma_list$'240702_DIV1'
 
row_num <- nrow(df)
df <- df[seq(0, row_num , by=2), ] # define the subsampling freq
df$X <- 1:nrow(df)
soma_list$'240702_DIV1' <- df

df <- soma_list$'240702_DIV2'
 
row_num <- nrow(df)
df <- df[seq(0, row_num , by=2), ] # define the subsampling freq
df$X <- 1:nrow(df)
soma_list$'240702_DIV2' <- df
```

```{r}
soma_actin <- soma_list %>% reduce(full_join, by = "X") %>% as.data.frame()
soma_actin <- soma_actin[ , colnames(soma_actin) %in% names(polarity_list)] %>% as.data.table()
```

```{r}
saveRDS(soma_list, "soma_list.rds")
```

```{r}
freq_soma <- 1
soma_ts <- soma_list$`240604`$DIV1_LifeactmSC_02_01_R3D_D3D.tif %>%
  
    forecast::ma(5) %>%                                   
    MTS::diffM(d=1) %>% 
    forecast::ma(3) %>% 
    ts(frequency = freq_soma, start = c(0, 1))

soma_ts %>%  autoplot() +
  geom_line(aes(),size=0.25)  +
   geom_hline(yintercept=0, linetype="dashed", color = "black")+
   #geom_hline(yintercept=thres, linetype="dashed", color = "blue")+
   #geom_hline(yintercept=-thres, linetype="dashed", color = "blue")+
   scale_y_continuous( breaks = scales::pretty_breaks(n = 7))+
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA), breaks = scales::pretty_breaks(n = 8)) + 
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text

```

```{r}
ggsave(
  "soma_profile.pdf",
   width=14, height=7, 
  plot = last_plot(),
  device = "pdf")
```

# retrive neurite length and calculate the longest and the 2nd longest neurite length
```{r}
library(matrixStats)

length_list_2 <- length_list %>% map(as.data.table)

length_list_2_MA <- list()

# Moving average smoothing
for (n in 1:length(length_list_2)){
  
  df <- length_list_2[[n]] %>% as.ts()
  length_list_2_MA[[n]]  <- df %>% forecast::ma(20) %>% as.data.table()
  
}

length_list_2 <- length_list_2_MA
names(length_list_2) <- names(length_list)

# remove cells with max neurite length < 100 um
# length_list_2 <- length_list_2 %>% discard (function(x) max(x, na.rm=T)<70)

# create function to get the 2nd longest neurite length
testFunc <- function(x) x %>% as.matrix() %>% Rfast::nth(2, descending = T)
polarity_list<- list()

for (n in 1:length(length_list_2)){
  
  df <- length_list_2[[n]] 
  max <- apply(df, 1, max) # the longest neurite length
  max_2nd <-  apply(df, 1, function(x) testFunc(x)) # the 2nd longest neurite length

  df$max <- max
  df$max_2nd <- max_2nd
  df$polarity <- max/max_2nd #%>% forecast::ma(10)
  df$Cell <- names(length_list_2)[n]
  df$Time <- seq( from = 0, to =  (nrow(df)-1) ) *2
  polarity_list [[n]] <- df
} 

names(polarity_list)<- names(length_list_2)
```

# filter for neurons showing S2-S3 transition
```{r}
# remove cells with the 2nd longest neurite longer than 50 um
#polarity_list <- polarity_list %>% discard (function(x) min(x$max_2nd, na.rm = TRUE)  > 50)

# remove cells with minimal polarity index larger than 1.5 (already polarized)
#polarity_list <- polarity_list %>% discard (function(x) min(x$polarity, na.rm = TRUE)  > 1.5)

# remove cells with maximal polarity index smaller than 1.1 (did not polarized)
#polarity_list <- polarity_list %>% discard (function(x) max(x$polarity, na.rm = TRUE)  > 20)
```

# get the time point of polarization
```{r}
polarityPoint_list <- list.files(path = folder_path, pattern = "^polarityPoints.csv$", recursive = TRUE, full.names = TRUE) %>% map(read.csv)
names(polarityPoint_list) <- dirname(list.files(path = folder_path, pattern = "^soma_actin.csv$", recursive = TRUE, full.names = TRUE)) %>% basename()

polarityPoint_list$'170905_Max'$Time  <- (polarityPoint_list$'170905_Max'$Frame -1)*2
polarityPoint_list$'170921_DVcore'$Time  <- (polarityPoint_list$'170921_DVcore'$Frame -1)*2
polarityPoint_list$'180129'$Time  <- (polarityPoint_list$'180129'$Slice -1)*2
polarityPoint_list$'180205'$Time  <- (polarityPoint_list$'180205'$Slice -1)*4
polarityPoint_list$'180207'$Time  <- (polarityPoint_list$'180207'$Slice -1)*2
polarityPoint_list$'240604'$Time  <- (polarityPoint_list$'240604'$Slice -1)*1
polarityPoint_list$'240702_DIV1'$Time  <- (polarityPoint_list$'240702_DIV1'$Frame -1)*1
polarityPoint_list$'240702_DIV2'$Time  <- (polarityPoint_list$'240702_DIV2'$Frame -1)*1

for (n in 1: length(polarityPoint_list)){
  
  df <- polarityPoint_list[[n]]
  df$Exp <- names(polarityPoint_list)[n]
  polarityPoint_list[[n]] <- df
  
}

polarityPoint_df <- polarityPoint_list %>% reduce(bind_rows) %>% as.data.table()
#polarityPoint_df$Frame <- coalesce(polarityPoint_df$Frame, polarityPoint_df$Slice)
polarityPoint_df$Image <- polarityPoint_df$Label %>% gsub(".* - ","",.)
```

```{r}

polarityPoint_df$Image <- names(polarity_list)

polarity_list_2 <- list()

for (n in 1:length(polarity_list)){
  
  imgName <- names(polarity_list)[n]
  
  df <- polarity_list[[n]] %>% as.data.table()
  df <- df[ ,.(Time, max, max_2nd, polarity, Cell)] 
  df$polarized <- "polarized"
  df[Time <  polarityPoint_df[Image == imgName]$Time]$polarized <- "not_polarized"
  
  df$Exp <- polarityPoint_df[Image == imgName]$Exp
  
  polarity_list_2[[n]] <- df
  
}
names(polarity_list_2) <- names(polarity_list)

```

```{r}
# extract the data after the time point of polarization
polarity_list_2_polarized <- list()

for (n in 1:length(polarity_list_2)){
  
  df <- polarity_list_2[[n]] 
  df <- df[polarized == "polarized"]
  df$Time <- seq( from = 0, to =  (nrow(df)-1) ) *2
  
  polarity_list_2_polarized[[n]] <- df
  
}
names(polarity_list_2_polarized) <- names(polarity_list)

# extract the data before the time point of polarization
polarity_list_2_Notpolarized <- list()

for (n in 1:length(polarity_list)){
  
  df <- polarity_list_2[[n]] 
  df  <- df[polarized == "not_polarized"]
  df$Time <- seq( from = 0, to =  (nrow(df)-1) ) *2
  
  polarity_list_2_Notpolarized[[n]] <- df
  
}
names(polarity_list_2_Notpolarized) <- names(polarity_list)
```

```{r}
polarized_df <- polarity_list_2_polarized %>% reduce(bind_rows) %>% as.data.table()
Notpolarized_df <- polarity_list_2_Notpolarized %>% reduce(bind_rows) %>% as.data.table()
```

```{r}
library(plotly)
p <- polarity_df %>% 
ggplot(aes(x=Time, y=max, group=Cell)) +
  geom_line((aes(color=Cell)))+
 theme_ipsum(grid = F, ticks = T)+ 

  #scale_y_continuous( breaks = scales::pretty_breaks(n = 8), limits=c(0,5))+
 # scale_x_continuous(expand = c(0, 0),  breaks = scales::pretty_breaks(n = 8), limits=c(0,NA)) +theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text

ggplotly(p)
```

# calculate the the average growth profile of axon before and after polarization
```{r}
df_summary_polarized <- polarized_df %>% group_by(Time) %>% rstatix::get_summary_stats(max, type = "common")
df_summary_Notpolarized <- Notpolarized_df %>% group_by(Time) %>% rstatix::get_summary_stats(max, type = "common")
```

```{r  message = FALSE, warning=FALSE, fig.width=8}
ggplot(df_summary_Notpolarized, aes(x=Time, y=mean)) +
  
   geom_ribbon(aes(ymin=mean-se, ymax=mean+se), fill="lightgrey")+
   geom_line(size=1, alpha=1) +
  
  #stat_peaks(colour = "red", span = 21, ignore_threshold = 0.8) +
  #stat_peaks(geom = "text", colour = "red", span = 21, ignore_threshold = 0.8, hjust = 2, angle = 0) +
  #stat_valleys(colour = "lightblue", span = 21, ignore_threshold = 1) +
  #stat_valleys(geom = "text", colour = "blue", span = 21, ignore_threshold = 1, hjust = -0.5, angle = 0) +
  
  scale_y_continuous( breaks = scales::pretty_breaks(n = 6))+
  scale_x_continuous(expand = c(0, 0),  breaks = scales::pretty_breaks(n = 8)) +
  
  theme_ipsum(grid = F, ticks = T)+ 
  theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text
```

# The neurite that has the longest neurite at the last frame is the axon
# filter for the max final neurite length (axon)
```{r}
length_axon_list <- list()
tip_actin_axon_list <- list()
neurite_actin_axon_list <- list()
density_actin_axon_list <- list()

selected_list <- length_list

for (n in 1:length(selected_list)){

 df <- selected_list[[n]] %>% as.data.table()
 
 selected_columns <- df[nrow(df), ] %>% max.col()
 selected_columns <- colnames(df)[selected_columns]
 length_axon_list[[n]] <- df[,..selected_columns]
 
 df <- Tip_actin_list[[n]] %>% as.data.table()
 modified_vector <- gsub("Length", "Tip_actin",  selected_columns)
 tip_actin_axon_list [[n]] <- df[, ..modified_vector]
 
 df <- Neurite_actin_list[[n]] %>% as.data.table()
 modified_vector <- gsub("Length", "Neurite_actin",  selected_columns)
 neurite_actin_axon_list[[n]] <- df[,..modified_vector]
 
 df <- Density_actin_list[[n]] %>% as.data.table()
 modified_vector <- gsub("Length", "density_actin",  selected_columns)
 density_actin_axon_list[[n]] <- df[,..modified_vector]
}
 
names(length_axon_list) <- names(length_list) 
names(tip_actin_axon_list) <- names(length_list) 
names(neurite_actin_axon_list) <- names(length_list) 
names(density_actin_axon_list) <- names(length_list) 

all_axon_list <- list(length = length_axon_list, 
                 tip_actin = tip_actin_axon_list,  
                 neurite_actin = neurite_actin_axon_list, 
                 Density_actin = density_actin_axon_list)
```

# filter for the minor neurites
```{r}

length_minor_list <- list()
tip_actin_minor_list <- list()
neurite_actin_minor_list <- list()
density_actin_minor_list <- list()

selected_list <- length_list

for (n in 1:length(selected_list)){

 df <- selected_list[[n]] %>% as.data.table()
 
 selected_columns <- df[nrow(df), ] %>% max.col()
 selected_columns <- colnames(df)[-selected_columns]
 length_minor_list[[n]] <- df[,..selected_columns]
 
 df <- Tip_actin_list[[n]] %>% as.data.table()
 modified_vector <- gsub("Length", "Tip_actin",  selected_columns)
 tip_actin_minor_list [[n]] <- df[, ..modified_vector]
 
 df <- Neurite_actin_list[[n]] %>% as.data.table()
 modified_vector <- gsub("Length", "Neurite_actin",  selected_columns)
 neurite_actin_minor_list[[n]] <- df[,..modified_vector]
 
 df <- Density_actin_list[[n]] %>% as.data.table()
 modified_vector <- gsub("Length", "density_actin",  selected_columns)
 density_actin_minor_list[[n]] <- df[,..modified_vector]
}
 
names(length_minor_list) <- names(length_list) 
names(tip_actin_minor_list) <- names(length_list) 
names(neurite_actin_minor_list) <- names(length_list) 
names(density_actin_minor_list) <- names(length_list) 

all_minor_list <- list(length = length_minor_list, 
                 tip_actin = tip_actin_minor_list,  
                 neurite_actin = neurite_actin_minor_list, 
                 Density_actin = density_actin_minor_list)
```

# Create stationary (differential) curve of all measurements
```{r}
all_sta_list <- list()
freq <- 0.5 # set frequency (1/interval)

for (m in 1:length(all_list)) {
  
  selected_list <- all_list[[m]]
  sta_list <- list()

  for (n in 1:length(selected_list))  {
 
  sta_list[[n]]<- selected_list[[n]] %>% 
    forecast::ma(5) %>%                                    # smoothing #1
    MTS::diffM(d=1) %>% 
    forecast::ma(3) %>% 
    ts(frequency = freq, start = c(0, 1))
  
  colnames(sta_list[[n]])<-colnames(selected_list[[n]])
  
  }
names(sta_list)<- names(selected_list)
all_sta_list[[m]]<- sta_list
}

names(all_sta_list)<- names(all_list )
```

# Create stationary (differential) curve of axon and minor neurites, separately. 
```{r}

# Axon

all_axon_sta_list <- list()

freq <- 0.5 # set frequency (1/interval)

for (m in 1:length(all_axon_list)) {
  
  selected_list <- all_axon_list[[m]]

  sta_list <- list()

  for (n in 1:length(selected_list))  {
 
  sta <- selected_list[[n]] %>% 
    forecast::ma(5) %>%                                    # smoothing #1
    MTS::diffM(d=1) %>% 
    forecast::ma(3) %>% 
    as.data.frame()
  
    colnames(sta) <- colnames(selected_list[[n]])
   sta_list [[n]] <- sta   %>%  ts(frequency = freq, start = c(0, 1))
  
  }

names(sta_list)<- names(selected_list)
  
all_axon_sta_list[[m]]<- sta_list
  
}

names(all_axon_sta_list)<- names(all_list)

# Minor neurites

all_minor_sta_list <- list()

freq <- 0.5 # set frequency (1/interval)

for (m in 1:length(all_minor_list)) {
  
  selected_list <- all_minor_list[[m]]

  sta_list <- list()

  for (n in 1:length(selected_list))  {
 
  sta <- selected_list[[n]] %>% 
    forecast::ma(5) %>%                                    # smoothing #1
    MTS::diffM(d=1) %>% 
    forecast::ma(3) %>% 
    as.data.frame()
  
    colnames(sta) <- colnames(selected_list[[n]])
   sta_list [[n]] <- sta   %>%  ts(frequency = freq, start = c(0, 1))
  
  }

names(sta_list)<- names(selected_list)
  
all_minor_sta_list[[m]]<- sta_list
  
}

names(all_minor_sta_list)<- names(all_list)
```

# Prepare data for cross-correlation analysis at individual neurite level (axon + minor neurite)
```{r}

library("data.table")

all_sta_cell_list <- list()

for (k in 1:length(length_list)){
  
neurite_list <- list()

dt1 <- all_sta_list[[1]][[k]] %>% as.data.table()
NC <- dt1 %>% ncol()

    for (n in 1:NC){
      
     dt <- dt1[,..n]
     
          for (m in 2:4){
            
            dt2 <- all_sta_list[[m]][[k]] %>% as.data.table()
            dt<- cbind(dt, dt2[, ..n])
              
          }
    
    neurite_list[[n]] <- dt
    
    }

names(neurite_list)<- paste("Neurite_", 1:NC, sep = "")

all_sta_cell_list[[k]]<- neurite_list

}

names(all_sta_cell_list) <- names(all_sta_list[[1]])

```

# Separate the data before and after the time point of polarization (Axon + minor)
```{r}
#######################################################################################
# label the data before and after the time point of polarization
all_sta_cell_list_2 <- all_sta_cell_list

for (n in 1:length(all_sta_cell_list_2)){
  
  imgName <- names(all_sta_cell_list_2)[n]
  
  df <- all_sta_cell_list_2[[n]][[1]] %>% as.data.table()
 # df <- df[ ,.(Time, max, max_2nd, polarity, Cell)] 
  
  df$Time <- seq( from = 0, to =  (nrow(df)-1) ) *2
  df$polarized <- "polarized"
  
  df[Time <  polarityPoint_df[Image == imgName]$Time]$polarized <- "not_polarized"
  
  #df$Exp <- polarityPoint_df[Image == imgName]$Exp
  
  all_sta_cell_list_2[[n]][[1]] <- df
  
}

#######################################################################################
# extract the data after the time point of polarization
all_sta_cell_list_2_polarized <- all_sta_cell_list_2 

for (n in 1:length(all_sta_cell_list_2)){
  
  df <- all_sta_cell_list_2[[n]][[1]] 
  df <- df[polarized == "polarized"]
 # df$Time <- seq( from = 0, to =  (nrow(df)-1) ) *2
  
  all_sta_cell_list_2_polarized[[n]][[1]] <- df[ ,c("Time","polarized"):=NULL]
  
}
names(all_sta_cell_list_2_polarized) <- names(all_sta_cell_list_2)

all_sta_cell_list_2_polarized <- purrr::keep(all_sta_cell_list_2_polarized, ~ nrow(.x[[1]]) >50 ) # keep data with frames more than 50, which is required for correlation analysis
#######################################################################################
# extract the data before the time point of polarization
all_sta_cell_list_2_Notpolarized <- all_sta_cell_list_2 

for (n in 1:length(all_sta_cell_list_2)){
  
  df <- all_sta_cell_list_2[[n]][[1]] 
  df  <- df[polarized == "not_polarized"]
#  df$Time <- seq( from = 0, to =  (nrow(df)-1) ) *2
  
  all_sta_cell_list_2_Notpolarized[[n]][[1]]  <- df[ ,c("Time","polarized"):=NULL]
  
}
names(all_sta_cell_list_2_Notpolarized) <- names(all_sta_cell_list_2)

all_sta_cell_list_2_Notpolarized <- purrr::keep(all_sta_cell_list_2_Notpolarized, ~ nrow(.x[[1]]) >50 )

```

```{r}
all_sta_Notpolarized <- reorganise_nested(all_sta_cell_list_2_Notpolarized)
all_sta_polarized <- reorganise_nested(all_sta_cell_list_2_polarized)
```

# Prepare data for cross-correlation analysis at individual neurite level (Axon)
```{r}
all_axon_sta_cell_list <- list()

for (k in 1:length(length_list)){
  
  neurite_list <- list()
dt1 <- all_axon_sta_list[[1]][[k]] %>% as.data.table()
NC <- dt1 %>% ncol()

    for (n in 1:NC){
      
     dt <- dt1[,..n]
     
          for (m in 2:4){
            
            dt2 <- all_axon_sta_list[[m]][[k]] %>% as.data.table()
            dt<- cbind(dt, dt2[, ..n])
          }
    
    neurite_list[[n]] <- dt
    }

names(neurite_list) <- gsub ( "Length.", "Neurite_", names(length_axon_list[[k]]))
#names(neurite_list)<- paste("Neurite_", 1:ncol(dt1), sep = "")

all_axon_sta_cell_list[[k]]<- neurite_list

}
names(all_axon_sta_cell_list) <- names(all_axon_sta_list[[1]])

```

# Separate the data before and after the time point of polarization (Axon)
```{r}
#######################################################################################
# label the data before and after the time point of polarization
all_axon_sta_cell_list_2 <- all_axon_sta_cell_list

for (n in 1:length(all_axon_sta_cell_list_2)){
  
  imgName <- names(all_axon_sta_cell_list_2)[n]
  
  df <- all_axon_sta_cell_list_2[[n]][[1]] %>% as.data.table()
 # df <- df[ ,.(Time, max, max_2nd, polarity, Cell)] 
  
  df$Time <- seq( from = 0, to =  (nrow(df)-1) ) *2
  df$polarized <- "polarized"
  
  df[Time <  polarityPoint_df[Image == imgName]$Time]$polarized <- "not_polarized"
  
  #df$Exp <- polarityPoint_df[Image == imgName]$Exp
  
  all_axon_sta_cell_list_2[[n]][[1]] <- df
  
}

#######################################################################################
# extract the data after the time point of polarization
all_axon_sta_cell_list_2_polarized <- all_axon_sta_cell_list_2 

for (n in 1:length(all_axon_sta_cell_list_2)){
  
  df <- all_axon_sta_cell_list_2[[n]][[1]] 
  df <- df[polarized == "polarized"]
 # df$Time <- seq( from = 0, to =  (nrow(df)-1) ) *2
  
  all_axon_sta_cell_list_2_polarized[[n]][[1]] <- df[ ,c("Time","polarized"):=NULL]
  
}
names(all_axon_sta_cell_list_2_polarized) <- names(all_axon_sta_cell_list_2)

all_axon_sta_cell_list_2_polarized <- purrr::keep(all_axon_sta_cell_list_2_polarized, ~ nrow(.x[[1]]) >50 )
#######################################################################################
# extract the data before the time point of polarization
all_axon_sta_cell_list_2_Notpolarized <- all_axon_sta_cell_list_2 

for (n in 1:length(all_axon_sta_cell_list_2)){
  
  df <- all_axon_sta_cell_list_2[[n]][[1]] 
  df  <- df[polarized == "not_polarized"]
#  df$Time <- seq( from = 0, to =  (nrow(df)-1) ) *2
  
  all_axon_sta_cell_list_2_Notpolarized[[n]][[1]]  <- df[ ,c("Time","polarized"):=NULL]
  
}
names(all_axon_sta_cell_list_2_Notpolarized) <- names(all_axon_sta_cell_list_2)

all_axon_sta_cell_list_2_Notpolarized <- purrr::keep(all_axon_sta_cell_list_2_Notpolarized, ~ nrow(.x[[1]]) >50 )

```

# Prepare data for cross-correlation analysis at individual neurite level (minor neurites)
```{r}
all_minor_sta_cell_list <- list()

for (k in 1:length(length_minor_list)){
  
neurite_list <- list()
dt1 <- all_minor_sta_list[[1]][[k]] %>% as.data.table()

    for (n in 1:ncol(dt1)){
      
     dt <- dt1[,..n]
          for (m in 2:4){
            
            dt2 <- all_minor_sta_list[[m]][[k]] %>% as.data.table()
            dt<- cbind(dt, dt2[, ..n])
          }
    
    neurite_list[[n]] <- dt
    }

# names(neurite_list)<- paste("Neurite_", 1:ncol(dt1), sep = "")
 names(neurite_list) <- gsub ( "Length.", "Neurite_", names(length_minor_list[[k]]))
all_minor_sta_cell_list[[k]]<- neurite_list

}
names(all_minor_sta_cell_list) <- names(all_minor_sta_list[[1]])

```

# Separate the data before and after the time point of polarization (Minor neurites)
```{r}
#######################################################################################
# label the data before and after the timpoint of polarization

all_minor_sta_cell_list_2 <- all_minor_sta_cell_list

for (n in 1:length(all_minor_sta_cell_list_2)){
  
  imgName <- names(all_minor_sta_cell_list_2)[n]
  neurite_list <- all_minor_sta_cell_list_2[[n]]
  neurite_num <- length(neurite_list)
  
  for (m in 1: neurite_num){
    
    df <- neurite_list [[m]] %>% as.data.table()
 # df <- df[ ,.(Time, max, max_2nd, polarity, Cell)] 
  
  df$Time <- seq( from = 0, to =  (nrow(df)-1) ) *2
  df$polarized <- "polarized"
  
  df[Time <  polarityPoint_df[Image == imgName]$Time]$polarized <- "not_polarized"
    
    neurite_list[[m]] <- df
    
  }
  
  #df$Exp <- polarityPoint_df[Image == imgName]$Exp
  
  all_minor_sta_cell_list_2[[n]] <- neurite_list
  
}
names(all_minor_sta_cell_list_2) <- names(all_minor_sta_cell_list)

#######################################################################################
# extract the data after the time point of polarization
all_minor_sta_cell_list_2_polarized <- all_minor_sta_cell_list_2

for (n in 1:length(all_minor_sta_cell_list_2)){
  
  neurite_list <- all_minor_sta_cell_list_2[[n]]
  neurite_num <- length(neurite_list)
  
  for (m in 1: neurite_num){
    
    df <- neurite_list [[m]] %>% as.data.table()
    df <- df[polarized == "polarized"]
    neurite_list[[m]] <- df[ ,c("Time","polarized"):=NULL]
    
  }
  
  all_minor_sta_cell_list_2_polarized[[n]] <- neurite_list
  
}
names(all_minor_sta_cell_list_2_polarized) <- names(all_minor_sta_cell_list_2)

all_minor_sta_cell_list_2_polarized <- purrr::keep(all_minor_sta_cell_list_2_polarized, ~ nrow(.x[[1]]) >50 )
#######################################################################################
# extract the data before the time point of polarization
all_minor_sta_cell_list_2_Notpolarized <- list()

for (n in 1:length(all_minor_sta_cell_list_2)){
  
  neurite_list <- all_minor_sta_cell_list_2[[n]]
  neurite_num <- length(neurite_list)
  
  for (m in 1: neurite_num){
    
    df <- neurite_list [[m]] %>% as.data.table()
    df <- df[polarized == "not_polarized"]
    neurite_list[[m]] <- df[ ,c("Time","polarized"):=NULL]
    
  }
  
  all_minor_sta_cell_list_2_Notpolarized[[n]] <- neurite_list
  
}
names(all_minor_sta_cell_list_2_Notpolarized) <- names(all_minor_sta_cell_list_2)

all_minor_sta_cell_list_2_Notpolarized <- purrr::keep(all_minor_sta_cell_list_2_Notpolarized, ~ nrow(.x[[1]]) >50 )
```

# cross-correlation with neurite length at individual neurite level
```{r warning=FALSE}
ccf_neurite_length_list <- list()
selected_list <- all_axon_sta_cell_list_2_Notpolarized # all_sta_cell_list or all_axon_sta_cell_list or all_minor_sta_cell_list
selected_list <- unlist(selected_list, recursive=FALSE)

for (n in 2:4){
  
    ccfPlot <- ccf (selected_list[[1]][ ,1],  
                      selected_list[[1]][ ,..n], 
                            type = "correlation", 
                            na.action = na.pass,
                            plot=F, 
                            lag.max=50) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf (selected_list[[m]][ ,1],  
                      selected_list[[m]][ ,..n], 
                            type = "correlation", 
                            na.action = na.pass,
                            plot=F, 
                            lag.max=50) 
  
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-50:50)
  ccf_neurite_length_list[[n-1]] <- ccf
}

names(ccf_neurite_length_list)<- c("Tip_actin",  "Neurite_actin",  "density_actin")

```

# select the corr for summarization and plotting graph. 
```{r}
selected_ccf <- ccf_neurite_length_list$Tip_actin
selected_ccf <- selected_ccf %>% as.data.table() 
selected_ccf$lag<- c(-50:50)/freq
stacked_ccf_mat <- selected_ccf %>% gather(Neurite, ccf, -lag) # select which CCF of measurements to show
df_summary <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf)
```

```{r}
ggsave(
  "axon_Notpolarized.pdf",
  plot = last_plot(),
  device = "pdf")
```
# function to reorganize data structure 
```{r}
library(data.table)

## helper ---------------------------------------------------------------
## given   : list of neurite data.tables for *one* Cell
## returns : a single data.table that holds *one* measure
##           ‑ one column per neurite (Length.3, Length.5, …)
merge_neurites_by_measure <- function(neurite_list, measure) {
  # extract the one column that starts with measure (e.g. "^Length\\.")
  cols <- lapply(neurite_list, \(dt) {
    cnam <- grep(paste0("^", measure, "\\."), names(dt), value = TRUE,
                 ignore.case = TRUE)
    if (length(cnam) == 0) return(NULL)               # neurite missing that measure
    out <- dt[, ..cnam]
    setnames(out, cnam, cnam)                         # keep original name
    out[, rowid := .I]                                # helper index for merges
  })

  cols <- Filter(Negate(is.null), cols)               # drop neurites w/o the measure
  if (!length(cols)) return(data.table())             # nothing found

  # column‑bind via successive left joins on rowid (works with unequal row counts)
  res <- Reduce(function(a, b) merge(a, b, by = "rowid", all = TRUE), cols)
  res[, rowid := NULL]
}

## main -----------------------------------------------------------------
reorganise_nested <- function(x) {

  measure_map <- c(
    length        = "Length",
    tip_actin     = "Tip_actin",
    neurite_actin = "Neurite_actin",
    Density_actin = "density_actin"
  )

  # build skeleton of result list
  out <- setNames(vector("list", length(measure_map)), names(measure_map))

  for (m in names(measure_map)) {
    meas <- measure_map[[m]]
    out[[m]] <- lapply(
      x,
      merge_neurites_by_measure,
      measure = meas
    )
    names(out[[m]]) <- names(x)     # Cell_1, Cell_2, …
  }

  out
}

```

```{r}
all_minor_sta_Notpolarized <- reorganise_nested(all_minor_sta_cell_list_2_Notpolarized)
all_minor_sta_polarized <- reorganise_nested(all_minor_sta_cell_list_2_polarized)
```

# Here we separate the outrgwoth and retraction events first, sum the changes of all neurites, and then do smoothening. (separate -> sum -> smoothening)

```{r}
selected_list <- all_sta_Notpolarized  # all_sta_list or all_minor_sta_list

thres= 0.0 # change threshold here

all_sta_PN_list <- list()

for (k in 1:length(selected_list )){  
  
  list_1 <- selected_list [[k]]
  sta_PN_list <- list()
  
  for (m in 1:length(list_1)) {
    
    DT <- list_1[[m]] %>% as.data.frame()
   
    DT_pos<- DT
    DT_neg<- DT
  
    for (n in 1:dim(DT)[2]){
        DT_pos[, n][DT_pos[, n] < 0] <- 0 
        DT_neg[, n][DT_neg[, n] > 0] <- 0 
    }
    
    mat_sum <- data.frame(rowSums(DT_pos), -rowSums(DT_neg)) # add neg sign
    
   # mat_sum [, 1][mat_sum [, 1] < thres] <- 0 
   # mat_sum [, 2][mat_sum [, 2] > -thres] <- 0
    
    mat_sum <- mat_sum %>% forecast::ma(5)                    # smoothing #2
  
    colnames(mat_sum)<- c(paste("Pos_", names(all_list)[k], sep = ""), paste("Neg_", names(all_list)[k], sep = "")) # create column names
    sta_PN_list[[m]] <- mat_sum
  
  }
  
  names(sta_PN_list)<-  names(list_1)
  all_sta_PN_list[[k]]<- sta_PN_list
  
}

names(all_sta_PN_list)<- names(all_list )
```

# Combine all stationary PN data of each videos, and add stationary soma data
```{r}
comb_sta_list<- list()

k <- all_sta_PN_list[[1]]%>% length()

for (m in 1:k){
  
  mat <- all_sta_PN_list[[1]][m] %>% as.data.frame()
  colnames(mat)<- sub('.*\\.', '', colnames(mat))
  
  for (n in 2:length(all_sta_PN_list)){
    
    mat2 <- all_sta_PN_list[[n]][m] %>% as.data.table()
    colnames(mat2)<- sub('.*\\.', '', colnames(mat2))
    
    mat <- data.frame(mat, mat2)
  }
  
  Dim <- mat %>% nrow()
  NM <- all_sta_PN_list[[1]][m] %>% names()
  mat3 <- mts_stationary_list[["minor_notpolarized"]][[NM]][1:Dim, c(2, 3)] %>% as.data.table() # soma data; "minor_notpolarized" or "minor_polarized"
  mat <- data.frame(mat, mat3)
  
 comb_sta_list[[m]]<- mat %>% ts(frequency = freq, start = c(0, 1)) 

}

names(comb_sta_list)<-   all_sta_PN_list[[1]] %>% names()
```

# cross-correlation of stationary PN data

```{r warning=FALSE}
selected_list <- comb_sta_list

ccf_PN_list<- vector(mode = "list", length = length(all_list)+1)

names(ccf_PN_list)<- c(names(all_list), "soma")

for (n in 1:5){
  
    ccfPlot <- ccf (selected_list[[1]][,(n*2-1)],  
                            selected_list[[1]][,(n*2)], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=50) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf (selected_list[[m]][,(n*2-1)],  
                            selected_list[[m]][,(n*2)], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=50) 
                            
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-50:50)
  ccf_PN_list[[n]] <- ccf
}

```

```{r warning=FALSE}
# cross-correlation with stationary Pos_length data

ccf_Poslen_list<- vector(mode = "list", length = dim(selected_list[[1]])[2]-2)

names(ccf_Poslen_list)<- colnames(selected_list[[1]])[-c(1,2)] 

for (n in 3:10){
  
    ccfPlot <- ccf (selected_list[[1]][,c("Pos_length")],  
                            selected_list[[1]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=50) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf (selected_list[[m]][,c("Pos_length")],  
                            selected_list[[m]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=50) 
  
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-50:50)
  ccf_Poslen_list[[n-2]] <- ccf
}
```

```{r warning=FALSE}
# cross-correlation with stationary Neg_length data

ccf_Neglen_list<- vector(mode = "list", length = dim(selected_list[[1]])[2]-2)

names(ccf_Neglen_list)<- colnames(selected_list[[1]])[-c(1,2)] 

for (n in 3:10){
  
    ccfPlot <- ccf (selected_list[[1]][,c("Neg_length")],  
                            selected_list[[1]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=50) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf (selected_list[[m]][,c("Neg_length")],  
                            selected_list[[m]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=50) 
  
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-50:50) # corresponds to "lag.max=50" set in ccf
  ccf_Neglen_list[[n-2]] <- ccf
}
```

```{r warning=FALSE}
# cross-correlation with soma actin

ccf_Actsoma_list<- vector(mode = "list", length = 8)

names(ccf_Actsoma_list)<- c(paste("Pos_", c("length", "tip_actin", "neurite_actin", "Density_actin"), sep=""), paste("Neg_", c("length","tip_actin", "neurite_actin", "Density_actin"), sep=""))

for (n in 1:8){
  
  df1 <- selected_list[[1]][ ,"soma_actin"]
  df2 <- selected_list[[1]][,grep(names(ccf_Actsoma_list)[n],  colnames(selected_list[[n]]))] 
  
    ccfPlot <- ccf (df1,  
                      df2, 
                      type = "correlation", 
                      na.action = na.pass,  plot=F, 
                      lag.max=50) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
     df1 <- selected_list[[m]][ ,"soma_actin"]
     df2 <- selected_list[[m]][,grep(names(ccf_Actsoma_list)[n],  colnames(selected_list[[n]]))]  
  
      ccfPlot_2 <- ccf (df1,  
                          df2, 
                          type = "correlation", 
                          na.action = na.pass,  plot=F, 
                          lag.max=50) 
      
      ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
      
      ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-50:50)
  ccf_Actsoma_list[[n]] <- ccf
}

ccf_all_list<-list(PN=ccf_PN_list, Poslen=ccf_Poslen_list, Neglen=ccf_Neglen_list, somaActin=ccf_Actsoma_list)

```

```{r}
selected_ccf <- ccf_all_list$PN$length

selected_ccf <-  selected_ccf %>% as.data.table()  # filter for DIV3 or DIV4 neuron only
  
selected_ccf$lag<- c(-50:50)/freq

stacked_ccf_mat <- selected_ccf %>% gather(Neurite, ccf, -lag) # select which CCF of measurements to show

df_summary <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf)
```

```{r}
ggsave(
  "PosNeg_velocity_corr_afterPolarization.pdf",
  plot = p,
  device = "pdf")
```
# analysis at whole-cell level
# Create time-series data of total neurite length, and actin intensity/density at neurite tip, shaft, and soma. 
```{r}
mts_list<- list()
mts_avg_list <- list()
mts_stationary_list <- list()

freq <- 0.5 # set frequency (1/interval)

for (n in 1:length(length_list))  {
 
  mat <- length_list[[n]] %>% rowSums() %>% as.data.table() # total neurite length
  colnames(mat) <- "sum_length"
  
  # get soma intensity
  ColName <- names(Tip_actin_list)[n]
  Dim <- nrow(Tip_actin_list[[n]])
  mat$soma_actin <- soma_actin[1:Dim, ..ColName ] 
  
  # 
  mat$sum_tip_actin <- Tip_actin_list[[n]] %>% rowSums()
  mat$sum_actin <-  Neurite_actin_list[[n]] %>% rowSums() # total actin intensity at the whole neurite
  mat$sum_neurite_actin <- Neurite_actin_list[[n]] %>% rowSums() 
  mat$sum_actin_density <- mat$sum_neurite_actin/mat$sum_length # actin density at the whole neurite

  mat <- mat %>% ts(frequency = freq, start = c(0, 1)) 
  mts_list[[n]]<- mat
   
  mat <- mat %>% forecast::ma(7) %>% as.ts() # Moving Average #1
  colnames(mat)<-colnames(mts_list[[n]])
  mts_avg_list[[n]] <- mat
   
  mts_stationary_list[[n]] <- mts_avg_list[[n]] %>% 
    MTS::diffM(d=1) %>% 
    forecast::ma(7) %>%                                             # Moving Average #2
    ts(frequency = freq, start = c(0, 1)) 
  
  colnames(mts_stationary_list[[n]])<-colnames(mts_list[[n]])
  
}

names(mts_list)<- names(Tip_actin_list)
names(mts_avg_list)<- names(Tip_actin_list)
names(mts_stationary_list)<- names(Tip_actin_list)
```

# Axon 
```{r}
mts_axon_list<- list()
mts_avg_axon_list <- list()
mts_stationary_axon_list <- list()

freq <- 0.5 # set frequency (1/interval)

for (n in 1:length(all_axon_list$length))  {
 
  mat <- all_axon_list$length[[n]] %>% rowSums() %>% as.data.table() # total neurite length
  colnames(mat) <- "sum_length"
  
  # get soma intensity
  ColName <- names(Tip_actin_list)[n]
  Dim <- nrow(Tip_actin_list[[n]])
  mat$soma_actin <- soma_actin[1:Dim, ..ColName ] 
  
  # get neurite shaft and tip intensity
  mat$sum_tip_actin <- all_axon_list$tip_actin[[n]] %>% rowSums()
  mat$sum_actin <-  all_axon_list$neurite_actin[[n]] %>% rowSums() # total actin intensity at the whole neurite
  mat$sum_neurite_actin <- all_axon_list$neurite_actin[[n]]  %>% rowSums() 
  mat$sum_actin_density <- mat$sum_neurite_actin/mat$sum_length # actin density at the whole neurite

  mat <- mat %>% ts(frequency = freq, start = c(0, 1)) 
  mts_axon_list[[n]]<- mat
   
  mat <- mat %>% forecast::ma(7) %>% as.ts() # Moving Average #1
  colnames(mat)<-colnames(mts_axon_list[[n]])
  mts_avg_axon_list[[n]] <- mat
   
  mat<- mat %>% MTS::diffM(d=1) %>% forecast::ma(7) %>%                                             # Moving Average #2
  ts(frequency = freq, start = c(0, 1)) 
  colnames(mat)<-colnames(mts_axon_list[[n]])
  mts_stationary_axon_list[[n]] <- mat
  
}

names(mts_axon_list)<- names(tip_actin_axon_list)
names(mts_avg_axon_list)<- names(tip_actin_axon_list)
names(mts_stationary_axon_list)<- names(tip_actin_axon_list)
```

# Separate the data before and after the time point of polarization (Axon)
```{r}
mts_stationary_axon_list_polarized <- mts_stationary_axon_list
mts_stationary_axon_list_Notpolarized <- mts_stationary_axon_list

for (n in 1:length(mts_stationary_axon_list)){
  
  imgName <- names(mts_stationary_axon_list)[n]
  
  # label the data before and after the time point of polarization
  df <- mts_stationary_axon_list[[n]] %>% as.data.table()
  df$Time <- seq( from = 0, to =  (nrow(df)-1) ) *2
  df$polarized <- "polarized"
  df[Time <  polarityPoint_df[Image == imgName]$Time]$polarized <- "not_polarized"
  
  # extract the data before and after the time point of polarization
  df1 <- df[polarized == "polarized"]
  mts_stationary_axon_list_polarized[[n]] <- df1[ ,c("Time","polarized"):=NULL]
  
  df2  <- df[polarized == "not_polarized"]
  mts_stationary_axon_list_Notpolarized[[n]]  <- df2[ ,c("Time","polarized"):=NULL]
  
}

names(mts_stationary_axon_list_polarized) <- names(mts_stationary_axon_list)
mts_stationary_axon_list_polarized <- purrr::keep(mts_stationary_axon_list_polarized, ~ nrow(.x) >25 )

names(mts_stationary_axon_list_Notpolarized) <- names(all_axon_sta_cell_list)
mts_stationary_axon_list_Notpolarized<- purrr::keep(mts_stationary_axon_list_Notpolarized, ~ nrow(.x) >25 )
```

# Minor neurites
```{r}
mts_minor_list<- list()
mts_avg_minor_list <- list()
mts_stationary_minor_list <- list()

freq <- 0.5 # set frequency (1/interval)

for (n in 1:length(all_minor_list$length))  {
 
  mat <- all_minor_list$length[[n]] %>% rowSums() %>% as.data.table() # total neurite length
  colnames(mat) <- "sum_length"
  
  # get soma intensity
  ColName <- names(Tip_actin_list)[n]
  Dim <- nrow(Tip_actin_list[[n]])
  mat$soma_actin <- soma_actin[1:Dim, ..ColName ] 
  
  # get neurite shaft and tip intensity
  mat$sum_tip_actin <- all_minor_list$tip_actin[[n]] %>% rowSums()
  mat$sum_actin <-  all_minor_list$neurite_actin[[n]] %>% rowSums() # total actin intensity at the whole neurite
  mat$sum_neurite_actin <- all_minor_list$neurite_actin[[n]]  %>% rowSums() 
  mat$sum_actin_density <- mat$sum_neurite_actin/mat$sum_length # actin density at the whole neurite

  mat <- mat %>% ts(frequency = freq, start = c(0, 1)) 
  mts_minor_list[[n]]<- mat
   
  mat <- mat %>% forecast::ma(7) %>% as.ts() # Moving Average #1
  colnames(mat)<-colnames(mts_minor_list[[n]])
  mts_avg_minor_list[[n]] <- mat
   
  mat<- mat %>% MTS::diffM(d=1) %>% forecast::ma(7) %>%                                             # Moving Average #2
  ts(frequency = freq, start = c(0, 1)) 
  colnames(mat)<-colnames(mts_minor_list[[n]])
  mts_stationary_minor_list[[n]] <- mat
  
}

names(mts_minor_list)<- names(tip_actin_minor_list)
names(mts_avg_minor_list)<- names(tip_actin_minor_list)
names(mts_stationary_minor_list)<- names(tip_actin_minor_list)
```

# Separate the data before and after the time point of polarization (Minor neurites)
```{r}
mts_stationary_minor_list_polarized <- mts_stationary_minor_list
mts_stationary_minor_list_Notpolarized <- mts_stationary_minor_list

for (n in 1:length(mts_stationary_minor_list)){
  
  imgName <- names(mts_stationary_minor_list)[n]
  
  # label the data before and after the time point of polarization
  df <- mts_stationary_minor_list[[n]] %>% as.data.table()
  df$Time <- seq( from = 0, to =  (nrow(df)-1) ) *2
  df$polarized <- "polarized"
  df[Time <  polarityPoint_df[Image == imgName]$Time]$polarized <- "not_polarized"
  
  # extract the data before and after the time point of polarization
  df1 <- df[polarized == "polarized"]
  mts_stationary_minor_list_polarized[[n]] <- df1[ ,c("Time","polarized"):=NULL]
  
  df2  <- df[polarized == "not_polarized"]
  mts_stationary_minor_list_Notpolarized[[n]]  <- df2[ ,c("Time","polarized"):=NULL]
  
}

names(mts_stationary_minor_list_polarized) <- names(mts_stationary_minor_list)
mts_stationary_minor_list_polarized <- purrr::keep(mts_stationary_minor_list_polarized, ~ nrow(.x) >25 )

names(mts_stationary_minor_list_Notpolarized) <- names(all_minor_sta_cell_list)
mts_stationary_minor_list_Notpolarized<- purrr::keep(mts_stationary_minor_list_Notpolarized, ~ nrow(.x) >25 )
```

```{r}
mts_stationary_list <- list(axon_polarized = mts_stationary_axon_list_polarized, 
                            axon_notpolarized = mts_stationary_axon_list_Notpolarized, 
                            minor_polarized = mts_stationary_minor_list_polarized,
                            minor_notpolarized = mts_stationary_minor_list_Notpolarized)
```

# cross-correlation with sum neurite length
```{r warning=FALSE}

selected_list <- mts_stationary_list$axon_polarized# mts_stationary_list or mts_50_stationary_list

ccf_sum_list<- vector(mode = "list", length = ncol(selected_list[[1]])-1)

names(ccf_sum_list)<- colnames(selected_list[[1]])[-1]

for (n in 2:6){
  
    ccfPlot <- ccf(selected_list[[1]][,1],  
                            selected_list[[1]][,..n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=25) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf(selected_list[[m]][,1],  
                            selected_list[[m]][,..n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=25) 
  
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-25:25)
  ccf_sum_list[[n-1]] <- ccf
}
```

# cross-correlation with soma actin
```{r warning=FALSE}

selected_list <- mts_stationary_list$axon_polarized # mts_stationary_list or mts_50_stationary_list

ccf_sumActsoma_list<- vector(mode = "list", length = ncol(selected_list[[1]]))

names(ccf_sumActsoma_list)<- colnames(selected_list[[1]])

for (n in 1:6){
  
    ccfPlot <- ccf(selected_list[[1]][,2],  
                            selected_list[[1]][,..n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=25) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf(selected_list[[m]][,2],  
                            selected_list[[m]][,..n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=25) 
  
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-25:25)
  ccf_sumActsoma_list[[n]] <- ccf
}
```

```{r}
ccf_sum_all_list<-list(sumlength=ccf_sum_list,  somaActin=ccf_sumActsoma_list)

```

```{r}
selected_ccf <- ccf_sum_all_list$somaActin$sum_tip_actin
selected_ccf <- selected_ccf %>% as.data.table() 
selected_ccf$lag<- c(-25:25)/freq
stacked_ccf_mat <- selected_ccf %>% gather(Neurite, ccf, -lag) 
df_summary <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf, type = "common")
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
ggplot(df_summary, aes(x=lag, y=mean)) +
   geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd), fill="lightgrey")+
   geom_line(size=1, alpha=1) +
  
   stat_peaks(colour = "red", span = 15, ignore_threshold = 0.5) +
  stat_peaks(geom = "text", colour = "red", span = 15, ignore_threshold = 0.6, hjust = 2, angle = 0) +
  stat_valleys(colour = "lightblue", span = 13, ignore_threshold = 0.9) +
 stat_valleys(geom = "text", colour = "blue", span = 13, ignore_threshold = 0.9, hjust = -0.5, angle = 0) +
  
   theme_ipsum(grid = F, ticks = T)+ 
   geom_vline(xintercept=0, linetype="dashed", color = "black")+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
   scale_y_continuous(expand = c(0, 0), limits = c(-0.5 ,0.8),  breaks = scales::pretty_breaks(n = 6))+
   scale_x_continuous(expand = c(0, 0),  breaks = scales::pretty_breaks(n = 8), limits = c(-50,50)) +
  
  theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text
```

```{r}
ggsave(
  "somactin_tipactin_axon_polarized_2.pdf",
  plot = last_plot(),
  device = "pdf")
```

## We smoothen the stationary profiles first, separete outrgwoth and retraction events, and sum the changes of all neurites. (Smoothening -> separate -> sum)
```{r warning=FALSE}

# Smoothen the stationary (diff) profiles and optionally filter out small length change with threshold

selected_list <- all_sta_list$length 

mts_stationary_pos_list <- list()
mts_stationary_neg_list <- list()
thres= 0.0 # change threshold here

# Positive peaks only (outgrowth event)
for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  length_mat<- selected_list[[m]][, selected_data ]
  
  length_mat_avg<- length_mat %>% forecast::ma(3) %>% as.ts() # smoothening 1
  colnames(length_mat_avg)<-colnames(length_mat)

  mts_dim<- dim(length_mat_avg)
    for (n in 1:mts_dim[2]){
      length_mat_avg[, n][length_mat_avg[ , colnames(length_mat_avg)[n]] < thres] <- 0 
    }
  
  length_mat_avg <- length_mat_avg %>% as.data.frame() 
  length_mat_avg$sum <- length_mat_avg  %>% rowSums()  # We sum up the extent of outgrowth at each time point
  mts_stationary_pos_list[[m]] <- length_mat_avg %>% ts(frequency = freq, start = c(0, 1)) 

}

names(mts_stationary_pos_list) <- names(length_list)

# Negative peaks only (retraction event)
for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  length_mat<- selected_list[[m]][, selected_data ]
  length_mat_avg<- length_mat %>% forecast::ma(3) %>% as.ts() # smoothening 1
  colnames(length_mat_avg)<-colnames(length_mat)
  
  mts_dim<- dim(length_mat_avg)
    for (n in 1:mts_dim[2]){
      length_mat_avg[, n][length_mat_avg[ , colnames(length_mat_avg)[n]] > - thres] <- 0 
    }
  length_mat_avg <- length_mat_avg %>% as.data.frame() 
  length_mat_avg$sum <- length_mat_avg  %>% rowSums()  # We sum up the extent of retraction at each time point
  mts_stationary_neg_list[[m]] <- length_mat_avg %>% ts(frequency = freq, start = c(0, 1)) 
}

names(mts_stationary_neg_list) <- names(length_list)

```

# Separate the extension/retraction and not polarized/not polarized
```{r}
mts_stationary_pos_list_polarized <- mts_stationary_pos_list
mts_stationary_pos_list_Notpolarized <- mts_stationary_pos_list

for (n in 1:length(mts_stationary_pos_list)){
  
  imgName <- names(mts_stationary_pos_list)[n]
  
  # label the data before and after the time point of polarization
  df <- mts_stationary_pos_list[[n]] %>% as.data.table()
  df$Time <- seq( from = 0, to =  (nrow(df)-1) ) *2
  df$polarized <- "polarized"

  df[Time <  polarityPoint_df[Image == imgName]$Time]$polarized <- "not_polarized"
  
  # extract the data before and after the time point of polarization
  df1 <- df[polarized == "polarized"]
  mts_stationary_pos_list_polarized[[n]] <- df1[ ,c("Time","polarized"):=NULL]
  
  df2  <- df[polarized == "not_polarized"]
  mts_stationary_pos_list_Notpolarized[[n]]  <- df2[ ,c("Time","polarized"):=NULL]
  
}

names(mts_stationary_pos_list_polarized) <- names(mts_stationary_pos_list)
#mts_stationary_pos_list_polarized <- purrr::keep(mts_stationary_pos_list_polarized, ~ nrow(.x) >25) # remove cells w less than 25 time points

names(mts_stationary_pos_list_Notpolarized) <- names(mts_stationary_pos_list)
#mts_stationary_pos_list_Notpolarized<- purrr::keep(mts_stationary_pos_list_Notpolarized, ~ nrow(.x) >25) # remove cells w less than 25 time points

mts_stationary_neg_list_polarized <- mts_stationary_neg_list
mts_stationary_neg_list_Notpolarized <- mts_stationary_neg_list

for (n in 1:length(mts_stationary_neg_list)){
  
  imgName <- names(mts_stationary_neg_list)[n]
  
  # label the data before and after the time point of polarization
  df <- mts_stationary_neg_list[[n]] %>% as.data.table()
  df$Time <- seq( from = 0, to =  (nrow(df)-1) ) *2
  df$polarized <- "polarized"

  df[Time <  polarityPoint_df[Image == imgName]$Time]$polarized <- "not_polarized"
  
  # extract the data before and after the time point of polarization
  df1 <- df[polarized == "polarized"]
  mts_stationary_neg_list_polarized[[n]] <- df1[ ,c("Time","polarized"):=NULL]
  
  df2  <- df[polarized == "not_polarized"]
  mts_stationary_neg_list_Notpolarized[[n]]  <- df2[ ,c("Time","polarized"):=NULL]
  
}

names(mts_stationary_neg_list_polarized) <- names(mts_stationary_neg_list)
#mts_stationary_neg_list_polarized <- purrr::keep(mts_stationary_neg_list_polarized, ~ nrow(.x) >25 ) # remove cells w less than 25 time points

names(mts_stationary_neg_list_Notpolarized) <- names(mts_stationary_neg_list)
#mts_stationary_neg_list_Notpolarized<- purrr::keep(mts_stationary_neg_list_Notpolarized, ~ nrow(.x) >25 ) # remove cells w less than 25 time points
```

```{r}
ext_ret_list <- list(Ext_Notpolarized = mts_stationary_pos_list_Notpolarized, Ext_polarized = mts_stationary_pos_list_polarized, Ret_Notpolarized = mts_stationary_neg_list_Notpolarized, Ret_polarized = mts_stationary_neg_list_polarized)

# remove cell with no pre-polarization time points
ext_ret_list[["Ret_Notpolarized"]][["DIV1_LifeactmSC_02_14_R3D_D3D.tif"]] <- NULL
ext_ret_list[["Ext_Notpolarized"]][["DIV1_LifeactmSC_02_14_R3D_D3D.tif"]] <- NULL

ext_ret_list[["Ret_Notpolarized"]][["WT_DIV1_LAmSC_01_13_R3D_D3D.tif"]] <- NULL
ext_ret_list[["Ext_Notpolarized"]][["WT_DIV1_LAmSC_01_13_R3D_D3D.tif"]] <- NULL

ext_ret_list[["Ret_Notpolarized"]][["WT_DIV1_LAmSC_01_16_R3D_D3D-2.tif"]] <- NULL
ext_ret_list[["Ext_Notpolarized"]][["WT_DIV1_LAmSC_01_16_R3D_D3D-2.tif"]] <- NULL

ext_ret_list[["Ret_Notpolarized"]][["WT_DIV2_LAmSC_02_06_R3D_D3D.tif"]] <- NULL
ext_ret_list[["Ext_Notpolarized"]][["WT_DIV2_LAmSC_02_06_R3D_D3D.tif"]] <- NULL

ext_ret_list[["Ret_Notpolarized"]][["WT_DIV2_LAmSC_02_24_R3D_D3D.tif"]] <- NULL
ext_ret_list[["Ext_Notpolarized"]][["WT_DIV2_LAmSC_02_24_R3D_D3D.tif"]] <- NULL

ext_ret_list[["Ret_Notpolarized"]][["WT_DIV2_LAmSC_02_36_R3D_D3D.tif"]] <- NULL
ext_ret_list[["Ext_Notpolarized"]][["WT_DIV2_LAmSC_02_36_R3D_D3D.tif"]] <- NULL

ext_ret_list[["Ret_Notpolarized"]][["WT_DIV2_LAmSC_02_38_R3D_D3D.tif"]] <- NULL
ext_ret_list[["Ext_Notpolarized"]][["WT_DIV2_LAmSC_02_38_R3D_D3D.tif"]] <- NULL

ext_ret_list[["Ret_Notpolarized"]][["WT_DIV2_LAmSC_02_40_R3D_D3D.tif"]] <- NULL
ext_ret_list[["Ext_Notpolarized"]][["WT_DIV2_LAmSC_02_40_R3D_D3D.tif"]] <- NULL
```

## extension and retraction lifetime (duration)
```{r}
thres <- 0.00

 list_duration <- list()  

for (k in 1:length (ext_ret_list )){

 list_1 <- list()    
 list_k <- ext_ret_list[[k]]
  
  for (m in 1:length(list_k)){
  
  df <- list_k[[m]] %>% as.data.table()
  df <- df[, !"sum"] %>% abs() %>% as.data.frame()
  df<- df[, (colSums(df, na.rm = T) != 0)] %>% as.data.table()
  
  list_2 <-  list()
  
   for (n in 1: ncol(df)){
    
    df2 <- df[ ,..n]
    df2$time <- seq(0, (nrow(df2)-1), 1)
    df2 <- df2[complete.cases(df2), ] 
    
    df2$inside <- df2[ , 1] > thres
    df2$start  <- ! c(FALSE, df2$inside[-nrow(df2)]) & df2$inside
    df2$end    <- df2$inside & ! c(df2$inside[-1], FALSE)  # Not used
    df2 <- df2[ df2$inside, ]
    df2$group  <- cumsum(df2$start)  # Number the spells
    colnames(df2)[1] <- "velocity"
    
    df3 <- plyr::ddply( df2, "group", summarize,
                        start  = min(time),
                        end    = max(time),
                        length = length(velocity),
                        mean   = mean(velocity))
   
    df3$neurite <- names(df)[n]
    
    # df3$neurite_length <- length_list[[m]][1,n] 
    list_2[[n]] <- df3
   }
  
  df3<- list_2 %>% data.table::rbindlist()
  df3$cell<- names(ext_ret_list[[k]])[m]
  df3$growth_polarization <- names(ext_ret_list)[k]

  list_1[[m]] <- df3
  }
 
 duration_df <- list_1 %>% data.table::rbindlist()

 list_duration [[k]] <-  duration_df
}
```

```{r}
duration_df <- list_duration  %>% data.table::rbindlist()
duration_df$length <- duration_df$length/freq
duration_df$cell_neurite <- paste(duration_df$cell, duration_df$neurite, sep = "_")
duration_df$axon_minor <- "Minor"

max_length_neurite_list <- list()
for (m in 1: length(length_list)){
  
   df1 <- length_list[[m]] 
   k <- df1 %>% tail(1)  %>% max.col() # filter for the column with the max final length as the axon
   df <- data.frame( neurite= names(df1)[[k]],  cell=names(length_list)[m])
   df$cell_neurite<- paste(df$cell, df$neurite, sep = "_")
  max_length_neurite_list[[m]] <- df %>% as.data.table()
}

max_length_neurite <- max_length_neurite_list %>% data.table::rbindlist()
duration_df[cell_neurite %in% max_length_neurite$cell_neurite]$axon_minor <- "Axon"
```

# Data points at cell level
```{r}
df_summary_duration <-duration_df %>% 
 group_by(cell, growth_polarization, axon_minor) %>% 
  rstatix::get_summary_stats(length) %>% as.data.table()
```

```{r}
# add sample labeling for grouping and batch control
df_summary_duration$ID<- paste (df_summary_duration$cell, df_summary_duration$growth_polarization, df_summary_duration$axon_minor,   sep = "_")
df_summary_duration$group<- paste (df_summary_duration$axon_minor, df_summary_duration$growth_polarization,  sep = "_")
```

```{r}
df_summary_duration$batch<- "_"
df_summary_duration[grep("170905", df_summary_duration$cell), ]$batch <- "170905"
df_summary_duration[grep("DIV1_LifeactmSC", df_summary_duration$cell), ]$batch <- "240604"
df_summary_duration[grep("Long2", df_summary_duration$cell), ]$batch <- "180109"
df_summary_duration[grep("Long_LifeactmScar_01", df_summary_duration$cell), ]$batch <- "180207"

df_summary_duration[grep("Long_LifeactmScar_03", df_summary_duration$cell), ]$batch <- "170921"
df_summary_duration[grep("Long_LifeactmScar_02", df_summary_duration$cell), ]$batch <- "180205"

df_summary_duration[grep("WT_DIV1", df_summary_duration$cell), ]$batch <- "240702_DIV1"
df_summary_duration[grep("WT_DIV2", df_summary_duration$cell), ]$batch <- "240702_DIV2"
```

```{r}
outliers <- df_summary_duration %>% 
  group_by(axon_minor, growth_polarization) %>% 
 rstatix::identify_outliers(mean) 

df_summary_duration<- df_summary_duration[! (ID %in% outliers$ID)]
```

```{r}
duration_df %>% group_by(cell, growth_polarization, neurite) %>% dplyr::summarise(n = n())

table(df_summary_duration$cell)
```

# generate pairwise comparison list
```{r}
my_comparisons <- df_summary_duration$group %>% unique() %>% as.character() %>% combn(2, simplify = FALSE) 

my_comparisons <- list(c("Axon_Ext_Notpolarized",  "Minor_Ext_Notpolarized"), c("Axon_Ext_polarized",  "Minor_Ext_polarized"), c("Axon_Ret_Notpolarized",  "Minor_Ret_Notpolarized"), c("Axon_Ret_polarized",  "Minor_Ret_polarized"), c("Axon_Ext_Notpolarized","Axon_Ext_polarized"), c("Minor_Ext_Notpolarized","Minor_Ext_polarized"), c("Axon_Ret_Notpolarized","Axon_Ret_polarized"), c("Minor_Ret_Notpolarized","Minor_Ret_polarized"))
```

```{r message = FALSE, warning=FALSE, fig.width=10}
# my_comparisons <- list(c("axon", "minor"))
#pj <- position_jitter(seed = 1, width = .1, height = 0)

df_summary_duration  %>% mutate(group = fct_relevel(group,  "Axon_Ext_Notpolarized",  "Minor_Ext_Notpolarized", "Axon_Ext_polarized",     "Minor_Ext_polarized" ,   "Axon_Ret_Notpolarized",  "Minor_Ret_Notpolarized", "Axon_Ret_polarized","Minor_Ret_polarized")) %>% 
  
ggplot ( aes(x=group, y=mean, fill =axon_minor) )+ 
  scale_fill_brewer(palette="Dark2") + 
  theme_classic()+
geom_violin(aes(), trim=F,  kernel ="gaussian", adjust=1)  +
 geom_boxplot(aes(), width=0.2, outlier.colour = NA, fill="white")+ 
# geom_jitter(aes(), shape=16, position=position_jitter(0.1))+ #aes(color = batch)
  
  # geom_line(aes(group = Image))+
   geom_point(size=1) +
 geom_path(aes(group = cell),alpha = 0.2)+
    
 scale_y_continuous(expand = c(0, 0), limits = c(0, 70), breaks = scales::pretty_breaks(n = 5))+ 
# geom_signif(comparisons = my_comparisons ,test = "wilcox.test",  color="black")+
 stat_compare_means(comparisons = my_comparisons, paired=F, method="wilcox.test", 
                     #label.y = c(7, 7.5, 8, 8.5, 9, 9.5, 10, 10.5), 
                     label = "{p.adj.format}{p.adj.signif}") +

  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold", angle = 90)) #for centering the text 
```

```{r}
ggsave(
  "ext_ret_duration_thres_0.pdf",
  plot = last_plot(),
  device = "pdf")
```

# merge pre-axon and minor of unpolarized neurons into one group
```{r}
df_summary_duration$group_2 <- df_summary_duration$group
df_summary_duration[grepl("Notpolarized", growth_polarization)]$group_2 <- df_summary_duration[grepl("Notpolarized", growth_polarization)]$growth_polarization

my_comparisons <- list( c("Axon_Ext_polarized",  "Minor_Ext_polarized"),  c("Axon_Ret_polarized",  "Minor_Ret_polarized"), c("Ext_Notpolarized","Axon_Ext_polarized"), c("Ext_Notpolarized","Minor_Ext_polarized"), c("Ret_Notpolarized","Axon_Ret_polarized"), c("Ret_Notpolarized","Minor_Ret_polarized"))
```

```{r message = FALSE, warning=FALSE, fig.width=10}
# my_comparisons <- list(c("axon", "minor"))
#pj <- position_jitter(seed = 1, width = .1, height = 0)

df_summary_duration  %>% mutate(group_2 = fct_relevel(group_2,  "Ext_Notpolarized",  "Axon_Ext_polarized",     "Minor_Ext_polarized" ,   "Ret_Notpolarized", "Axon_Ret_polarized","Minor_Ret_polarized")) %>% 
  
ggplot ( aes(x=group_2, y=mean) )+ 
  scale_fill_brewer(palette="Dark2") + 
  theme_classic()+
geom_violin(aes(), trim=F,  kernel ="gaussian", adjust=1, scale="area")  +
 geom_boxplot(aes(), width=0.2, outlier.colour = NA, fill="white")+ 
geom_jitter(aes(), shape=16, position=position_jitter(0.1))+ #aes(color = batch)
  
  # geom_line(aes(group = Image))+
 #  geom_point(size=1) +
 #geom_path(aes(group = cell),alpha = 0.2)+
    
 scale_y_continuous(expand = c(0, 0), limits = c(0, 50), breaks = scales::pretty_breaks(n = 5))+ 
# geom_signif(comparisons = my_comparisons ,test = "wilcox.test",  color="black")+
 stat_compare_means(comparisons = my_comparisons, paired=F, method="wilcox.test", 
                     label.y = c(42, 42, 42, 45, 42, 45), 
                     label = "{p.adj.format}{p.adj.signif}") +

  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold", angle = 90)) #for centering the text 
```

```{r}
ggsave(
  "ext_ret_duration_2.pdf",
  plot = last_plot(),
  device = "pdf")
```

# extension and retraction velocity
```{r}
ext_ret_list_2 <- ext_ret_list
for (n in 1: length(ext_ret_list)){
  
  list_1 <- ext_ret_list[[n]]
  sample_name <- names(list_1)
  
  for (m in 1:length(list_1)){
  
    df <- list_1[[m]] %>% as.data.frame()
    df [ df == 0 ] <- NA # replace zero with NA to avoid zero inflation
    df1 <- colMeans(df, na.rm = T)  %>% as.data.frame() # calculate the average velocity through the observation window. Or matrixStats::colMedians
    colnames(df1) <- "mean_vel"
    df1$median_vel <- df %>% as.matrix() %>% matrixStats::colMedians(na.rm = T)  # calculate the average velocity through the observation window. Or matrixStats::colMedians
    
    df1$cond <- "WT"
    df1$cell <- names(list_1)[m]
    df1$neurite <- rownames(df1)
    df1 <- df1 %>% subset (df1$neurite !="sum")
    
    df1$axon_minor <- "Minor"
    k <- length_list[[names(list_1)[m]]] %>% tail(1)  %>% max.col() # filter for the column with the max final length as the axon
    df1$axon_minor[k]  <- "Axon" 
    df1$growth_polarization <- names(ext_ret_list)[[n]]

   list_1 [[m]] <- df1 %>% as.data.table()
      
    }
  
  names(list_1)<- sample_name

  ext_ret_list_2 [[n]] <- list_1 %>% reduce(bind_rows)  %>% as.data.frame()
}

velocity_comb_df <- ext_ret_list_2 %>% reduce(bind_rows)  %>% as.data.table()

# label extension and retraction 
velocity_comb_df$Growth<- "Extension"
velocity_comb_df[mean_vel <0]$Growth <- "Retraction"

# convert retraction velocity to positive value
velocity_comb_df$mean_vel <- velocity_comb_df$mean_vel %>% abs()
velocity_comb_df$median_vel <- velocity_comb_df$median_vel %>% abs()

velocity_comb_df[mean_vel == "NaN"]$mean_vel <- 0
velocity_comb_df[median_vel == "NaN"]$median_vel <- 0
```

```{r}
velocity_comb_df[Growth=="Retraction"] %>% ggplot(aes(x=median_vel, y=length) ) +
  scale_x_continuous( expand = c(0, 0),  limits = c(0, 1),  breaks = scales::pretty_breaks(n = 5)) +
  geom_point()
```

# Data points at cell level (axon vs minor neurites; extension vs retraction: not polarized vs polarized)
```{r}
df_summary_vel <-velocity_comb_df %>% 
  group_by(cell, growth_polarization, axon_minor) %>% 
  rstatix::get_summary_stats(median_vel, mean_vel) %>% as.data.table()

df_summary_vel$ID<- paste (df_summary_vel$cell, df_summary_vel$axon_minor, df_summary_vel$growth_polarization, sep = "_")
df_summary_vel$group<- paste (df_summary_vel$axon_minor, df_summary_vel$growth_polarization, sep = "_")
```

```{r}
outliers <- df_summary_vel %>% 
  group_by(growth_polarization, axon_minor, variable) %>% 
 rstatix::identify_outliers(mean) 

df_summary_vel <- df_summary_vel[! (ID %in% outliers$ID)]
```

```{r}
my_comparisons <- df_summary_median_vel$group %>% unique() %>% as.character() %>% combn(2, simplify = FALSE) 
my_comparisons <- list(c("Axon_Ext_Notpolarized",  "Minor_Ext_Notpolarized"), c("Axon_Ext_polarized",  "Minor_Ext_polarized"), c("Axon_Ret_Notpolarized",  "Minor_Ret_Notpolarized"), c("Axon_Ret_polarized",  "Minor_Ret_polarized"), c("Axon_Ext_Notpolarized","Axon_Ext_polarized"), c("Minor_Ext_Notpolarized","Minor_Ext_polarized"), c("Axon_Ret_Notpolarized","Axon_Ret_polarized"), c("Minor_Ret_Notpolarized","Minor_Ret_polarized"))
```

```{r message = FALSE, warning=FALSE, fig.width=10}
#pj <- position_jitter(seed = 1, width = .1, height = 0)

df_summary_vel[variable == "mean_vel"]  %>%  
  
  mutate(group = fct_relevel(group,  "Axon_Ext_Notpolarized",  
                             "Minor_Ext_Notpolarized", 
                             "Axon_Ext_polarized",     
                             "Minor_Ext_polarized" ,   
                             "Axon_Ret_Notpolarized",  
                             "Minor_Ret_Notpolarized", 
                             "Axon_Ret_polarized",
                             "Minor_Ret_polarized")) %>% 
  
ggplot ( aes(x=group, y=mean, fill =axon_minor))+ 
  
  scale_fill_brewer(palette="Dark2") + 

 geom_violin(aes(), trim=F,  kernel ="gaussian", adjust=1)  +
 geom_boxplot(aes(), width=0.2, outlier.colour = NA, fill="white")+ 
 geom_jitter(aes(), shape=16, position=position_jitter(0.1))+
  
 # geom_line(aes(group = Image))+
 # geom_point(size=1) +
 # geom_path(aes(group = cell),alpha = 0.5)+
    
 scale_y_continuous(expand = c(0, 0), limits = c(-0.2, 3), breaks = scales::pretty_breaks(n = 5))+ 
# geom_signif(comparisons = my_comparisons ,test = "wilcox.test",  color="black")+
  stat_compare_means(comparisons = my_comparisons, paired=F, method="wilcox.test", 
                     #label.y = c(1.15, 1.2, 1.25, 1.3, 1.35, 1.4), 
                     label = "{p.adj.format}{p.adj.signif}") +
  
  theme_classic() + theme_ipsum(grid = F, ticks = T) + theme_linedraw()+
  
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold", angle = 90)) #for centering the text 
```

```{r}
table(df_summary_median_vel$cell)
```

```{r}
ggsave(
  "Axon_minor_velocities_2.pdf",
  plot = last_plot(),
  device = "pdf")
```

```{r}
df_summary_vel$group_2 <- df_summary_vel$group
df_summary_vel[grepl("Notpolarized", growth_polarization)]$group_2 <- df_summary_vel[grepl("Notpolarized", growth_polarization)]$growth_polarization

my_comparisons <- list( c("Axon_Ext_polarized",  "Minor_Ext_polarized"),  c("Axon_Ret_polarized",  "Minor_Ret_polarized"), c("Ext_Notpolarized","Axon_Ext_polarized"), c("Ext_Notpolarized","Minor_Ext_polarized"), c("Ret_Notpolarized","Axon_Ret_polarized"), c("Ret_Notpolarized","Minor_Ret_polarized"))
```

```{r message = FALSE, warning=FALSE, fig.width=10}
# my_comparisons <- list(c("axon", "minor"))
#pj <- position_jitter(seed = 1, width = .1, height = 0)

df_summary_vel[variable == "median_vel"]  %>% mutate(group_2 = fct_relevel(group_2,  "Ext_Notpolarized",  "Axon_Ext_polarized",     "Minor_Ext_polarized" ,   "Ret_Notpolarized", "Axon_Ret_polarized","Minor_Ret_polarized")) %>% 
  
ggplot ( aes(x=group_2, y=mean) )+ 
  scale_fill_brewer(palette="Dark2") + 
  theme_classic()+
geom_violin(aes(), trim=F,  kernel ="gaussian", adjust=1, scale="area")  +
 geom_boxplot(aes(), width=0.2, outlier.colour = NA, fill="white")+ 
geom_jitter(aes(), shape=16, position=position_jitter(0.1))+ #aes(color = batch)
  
  # geom_line(aes(group = Image))+
 #  geom_point(size=1) +
 #geom_path(aes(group = cell),alpha = 0.2)+
    
 scale_y_continuous(expand = c(0, 0), limits = c(-0.2, 1.5), breaks = scales::pretty_breaks(n = 8))+ 
# geom_signif(comparisons = my_comparisons ,test = "wilcox.test",  color="black")+
 stat_compare_means(comparisons = my_comparisons, paired=F, method="wilcox.test", 
                     label.y = c(1.2, 1.2, 1.2, 1.3, 1.2, 1.3), 
                     label = "{p.adj.format}{p.adj.signif}") +

  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold")) #for centering the text 
```

```{r}
ggsave(
  "Axon_minor_velocities_3.pdf",
  plot = last_plot(),
  device = "pdf")
```

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
