---
title: "R Notebook"
output: html_notebook
---

<!--
# Lag cross-correlation of somatic actin intensity change vs binarized actin-wave occurrence

## What this file does

Reads per-cell somatic actin-intensity CSVs (`soma_actin*.csv`) and per-wave
occurrence CSVs (`wave*.csv`) from per-cell subfolders (lines 58, 63, 74,
82). For each cell the local-detrended change of soma actin intensity is
cross-correlated against the binarized actin-wave occurrence trace. CCF
curves are pooled to a mean +/- SD across the cohort (N = 23 neurons from 3
experiments). Six PDFs are written (`ggsave` at lines 144, 208, 390, 474,
542, 801).

## Manuscript panel

**Fig 2c**: "Cross-correlation between the changes of soma actin intensity
and the binarized occurrence of actin waves. Values are plotted as mean +/-
SD. N = 23 neurons from 3 experiments."

Source Data: `Fig 2c`.

## Outputs

- Six PDFs (`ggsave` at lines 144, 208, 390, 474, 542, 801).

## Original G:\ source

G:/Figure 2_actin wave/Fig. 2c/Actin_wave_Fig. 2di_ExtFig.2a/R_script/Neurite activity corr.Rmd
-->

```{r message=FALSE, warning=FALSE, include=FALSE}
library(pracma)
library(peakPick)
library(quantmod)
library(ggpmisc)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggfortify)
library(forecast)
library(tsibble)
library(broom)
library(tidyquant)
library(tsbox)
library("pheatmap")
library("RColorBrewer")
library(ggpmisc)
library(ggfortify)
library(viridis) # Default Color Maps from 'matplotlib' 
library(hrbrthemes)
library(data.table)
library(superheat)
```

## load existing/saved environment (.RData)
```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
load(file=saved_path)
```

```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
save.image(file=saved_path)

```

# Soma actin and actin wave frame
```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = F)
somaActin_list<- list()
wave_list<- list()
freq <- 0.5 # set frequency (1/interval)

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  soma_csv_list <- list.files(subfolder_list[n],pattern = "^soma_actin(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
   wave_csv_list <- list.files(subfolder_list[n],pattern = "wave(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
   
   folder_wave_list<- list()
   
   # Reading actin wave data
   for (k in 1:length(wave_csv_list))  
        {
     
     df <- wave_csv_list[k] %>% read.csv(header = TRUE) 
     folder_wave_list[[k]] <-  df[, grep("Slice|Frame",  colnames(df)), drop = FALSE] 
     colnames(folder_wave_list[[k]])[1]<- "Frame"
     names(folder_wave_list)[k]<- wave_csv_list[k] %>% basename() %>% gsub("_wave.csv", "", .)
   
   }
   
   # Reading soma actin data
  somaActin_list[[n]]  <- soma_csv_list[1] %>% read.csv(header = TRUE)
  wave_list[[n]]<- folder_wave_list

}

names(somaActin_list) <- subfolder_list %>% basename()
names(wave_list) <- subfolder_list %>% basename()
```

# count average wave count per hour
```{r}
wave_list_2 <- wave_list 
  # Reading actin wave data
   for (k in 1:length(wave_list_2))  
        {
     
    list_k <- wave_list_2[[k]]
    
    for (n in 1:length(list_k)){
      
      df <- list_k[[n]]
      df$Cell <- names(list_k)[[n]]
      df$Batch <- names(wave_list)[[k]]
      list_k[[n]]<- df
      
    }
    
   df_2 <-  list_k %>% reduce(bind_rows)
   wave_list_2[[k]] <- df_2

   }

wave_df <- wave_list_2 %>% reduce(bind_rows)
wave_summary <- wave_df %>% group_by(Cell, Batch) %>% rstatix::get_summary_stats(Frame)
wave_summary$wave_freq_hour <- wave_summary$n/((wave_summary$max-wave_summary$min)*2) * 60
```

```{r message = FALSE, warning=FALSE, fig.width=8}
library(colorspace)
wave_summary  %>% 
ggplot(aes(x=variable, y=wave_freq_hour, fill = variable)) +
  #geom_point()
  theme_classic()+
  geom_violin(trim=FALSE, kernel ="gaussian", adjust=1)+
  geom_boxplot(width=0.4, fill="white",color="black",alpha=1, outliers = F) +
  geom_jitter(alpha=1, width=0.1) + 
 # scale_color_continuous_sequential(palette = "Heat", begin = 0.3, end = 1)+ 
 # scale_y_continuous(expand = c(0, 0), limits = c(0, 1.5), breaks = scales::pretty_breaks(n = 10))+ 
 # stat_compare_means(comparisons = my_comparisons, test = "t.test", paired=F)+
  theme_ipsum(grid = F, ticks = T) +
 theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold")) #for centering the text 

#plotly::ggplotly(p)
```

```{r}
ggsave(
  "wave_count.pdf",
  plot = last_plot(),
  device = "pdf")
#dev.off()
```

```{r}
wave_summary %>% write_csv("DIV1_wavecount.csv")
```

```{r}
wave_summary$polarization <- "S2"
wave_summary<- wave_summary %>% as.data.table()

S2S3_wave_count <- readRDS("S2S3_Total_wave_3.rds")%>% as.data.table()
names(S2S3_wave_count)[[9]]<- "wave_freq_hour"
names(S2S3_wave_count)[[2]]<- "Cell"
S2S3_wave_count <- S2S3_wave_count[polarization == "before_wave_hour"]
S2S3_wave_count$polarization <- "S2S3"

S3_wave_count <- readRDS("wave_count_2.rds") %>% as.data.table()
S3_wave_count$polarization <- "S3"
names(S3_wave_count)[[15]] <- "wave_freq_hour"
names(S3_wave_count)[[1]]<- "Cell"

wave_count_S2_S2S3_S3 <- rbind(wave_summary[,.(Cell, polarization,wave_freq_hour) ], S2S3_wave_count[,.(Cell, polarization,wave_freq_hour) ], S3_wave_count[,.(Cell, polarization,wave_freq_hour) ]) 
```

```{r}
outliers <- wave_count_S2_S2S3_S3 %>% 
  group_by(polarization) %>% 
 rstatix::identify_outliers(wave_freq_hour) 

wave_count_S2_S2S3_S3 <- wave_count_S2_S2S3_S3 [! (Cell %in% outliers$Cell) ]

wave_count_S2_S2S3_S3$polarization %>% table()
```

```{r message = FALSE, warning=FALSE, fig.width=8}
library(ggpubr)
my_comparisons <- list(c("S2",   "S2S3"), c("S2S3",   "S3"),  c("S2",   "S3"))

wave_count_S2_S2S3_S3  %>%  mutate(polarization = factor(polarization, levels=c("S2", "S2S3", "S3"))) %>% 
ggplot(aes(x=polarization, y=wave_freq_hour, fill = polarization)) +
  #geom_point()
  theme_classic()+
  geom_violin(trim=FALSE, kernel ="gaussian", adjust=1)+
  geom_boxplot(width=0.2, fill="white",color="black",alpha=1, outliers = F) +
  geom_jitter(alpha=1, width=0.1) + 
 # scale_color_continuous_sequential(palette = "Heat", begin = 0.3, end = 1)+ 
  scale_y_continuous(expand = c(0, 0), limits = c(-2, 14), breaks = scales::pretty_breaks(n = 8))+ 
  stat_compare_means(comparisons = my_comparisons, test = "wilcox.test", paired=F)+
  theme_ipsum(grid = F, ticks = T) +
 theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold")) #for centering the text 

```
```{r}
ggsave(
  "wave_count.pdf",
  plot = last_plot(),
  width = 9,
  height = 6,
  device = "pdf")
#dev.off()
```

```{r}
wave_count_S2_S2S3_S3 %>% write_csv("S2S3_wavecount.csv")
```

# subset soma actin data matching actin wave data
```{r}
for (n in 1:length(somaActin_list))  
{
  somaActin_list[[n]]<- somaActin_list[[n]][, names(somaActin_list[[n]]) %in% names(wave_list[[n]])]
  
}
```

# Create dataframe and assign logic value (0 for no wave or 1 for wave emerging at the soma) 
```{r}

wave_frame_list <- list()

for (n in 1:length(wave_list)){
  
  list_1 <- wave_list[[n]]
  list_2 <- list()
  
  for (k in 1:length(list_1)){
    
    df <- seq(1:481) %>% as.data.frame()
    names(df)<- "Time"
    df$Wave <- 0
    df[df$Time %in% list_1[[k]]$Frame ,]$Wave <- 1
    list_2[[k]] <- df$Wave %>% ts(frequency = freq, start = c(0, 1))
    
  }
 names(list_2)<- names(list_1)
 wave_frame_list[[n]] <- list_2
  
}

names(wave_frame_list) <- names(wave_list)

```

```{r}

wave_frame_list_2<- wave_frame_list %>%  unlist(recursive=FALSE)
somaActin_list_2<- somaActin_list %>%  unlist(recursive=FALSE)

somaActin_list_2<- somaActin_list_2 [names(wave_frame_list_2)]

```

# Create stationary (differential) curve of all measurements
```{r}
freq <- 0.5

selected_list <- somaActin_list_2
somaActin_sta_list <- list()

  for (n in 1:length(selected_list))  {
 
  somaActin_sta_list[[n]]<- selected_list[[n]] %>% 
    forecast::ma(2) %>%                                    # smoothing #1
    MTS::diffM(d=1) %>% 
    forecast::ma(2) %>% 
    ts(frequency = freq, start = c(0, 1))
  
  colnames(somaActin_sta_list[[n]])<-colnames(selected_list[[n]])
  
  }

names(somaActin_sta_list)<- names(selected_list)

```

# Create data frame for visualization of selected data
```{r  message = FALSE, warning=FALSE, fig.width=8}
df_1 <- somaActin_sta_list[[14]] %>% as.data.frame()
df_1$x<- df_1$x %>%  as.numeric()
df_1[nrow(df_1)+1, ]<- "NA"
#df_1[nrow(df_1)+1, ]<- "NA"
df_1$x<- df_1$x %>%  as.numeric()

df_2 <- wave_frame_list_2[[14]] %>% as.data.frame()
df_1$wave_frame<- df_2$x
df_1$Time<- (0:480)*2

df_mark<- df_1 %>% subset(df_1$wave_frame == 1)
```

# check the occurance of actin waves of selected data
```{r}
ggplot(df_mark, aes(x=Time)) + 
 geom_histogram(aes(), colour="black", fill="white", binwidth = 60)
```
# check the curve of selected stationary data
```{r  message = FALSE, warning=FALSE, fig.width=8}
ggplot(df_1, aes(x=Time, y=x)) +

  #geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd), fill="lightgrey", alpha=1)+
  geom_line(size=1) +
  
  #geom_vline(xintercept=0, linetype="dashed", color = "black")+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  
  #stat_peaks(colour = "red", ignore_threshold = 0.9) +
  #stat_peaks(geom = "text", colour = "red", vjust = -0.5, ignore_threshold = 0.9) +
  
  #stat_valleys(colour = "blue", ignore_threshold = 0.9, span = 11)+
  #stat_valleys(geom = "text", colour = "blue", angle = 45, vjust = 1.5, hjust = 1, ignore_threshold = 0.9, span = 11)+
  
  #scale_y_continuous(expand = c(0, 0), limits = c(-0.7, 0.7), breaks = scales::pretty_breaks(n = 7))+
  #scale_x_continuous(expand = c(0, 0), limits = c(-50, 50), breaks = scales::pretty_breaks(n = 8)) + 
  geom_point(data=df_mark, 
             aes(x=Time,y=x), 
             color='red',
             size=2)+

  theme_ipsum(grid = F, ticks = T)+ 
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text
```

```{r}
binned_list<-list()
for (n in 1: length(df_mark_list)){
  p <- df_mark_list[[n]] %>% ggplot(aes(x=Time)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="white", binwidth = 30)+
    geom_density(alpha=.2, fill="#FF6666") 
 
  binned_df <- p %>% layer_data(1)
  binned_df$pct <- binned_df$ndensity/sum(binned_df$ndensity) 
  binned_df$img_name <- names(df_mark_list)[n]
  binned_list[[n]] <- binned_df
  
}

names(binned_list) <- names(somaActin_sta_list)

binned_df <- binned_list[[1]]
for (n in 2: length(binned_list)){
 binned_df <- rbind(binned_df, binned_list[[n]])
}

df_summary <- binned_df %>% group_by(x) %>% rstatix::get_summary_stats(count)
```

```{r  message = FALSE, warning=FALSE, fig.width=8}
ggplot(df_summary, aes(x=x, y=mean)) +

  geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd), fill="lightgrey", alpha=1)+
  geom_line(size=1) +
  
  #geom_vline(xintercept=0, linetype="dashed", color = "black")+
  #geom_hline(yintercept=0, linetype="dashed", color = "black")+
  
  scale_y_continuous(expand = c(0, 0), limits = c(-2, 6), breaks = scales::pretty_breaks(n = 8)) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 960), breaks =  seq(0, 960, by = 60)) + 
  
  theme_ipsum(grid = F, ticks = T)+ 
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text
```

```{r}
ggsave(
  "wave_count.eps",
  plot = last_plot(),
  device = "eps")
```

#calculate the percentage of waves occuring at increasing or decreasing soma actin
```{r}

somaActin_sta_wave <- list()
df_mark_list <- list()
for (n in 1: length(somaActin_sta_list)){
  
    df_1 <- somaActin_sta_list[[n]] %>% as.data.frame()
    df_1$x<- df_1$x %>%  as.numeric()
    df_1[nrow(df_1)+1, ]<- "NA"
    df_1$x<- df_1$x %>%  as.numeric()
    
    df_2 <- wave_frame_list_2[[n]] %>% as.data.frame()
    df_1$wave_frame<- df_2$x
    df_1$Time<- (0:480)*2
    
    df_mark<- df_1 %>% subset(df_1$wave_frame == 1) %>% as.data.table()
    
    df_mark$change <- "Negative"
    df_mark[df_mark$x >0 ,]$change <- "Positive"
    df_mark_list [[n]] <- df_mark
    somaActin_sta_wave [[n]]<- df_mark$change %>% table() %>% as.data.frame() %>% t()
    
}

names(df_mark_list)<- names(somaActin_sta_list)

# combine the counts of Neg and Pos from all cells as single data frame
df_1 <- somaActin_sta_wave[[1]] %>% as.data.frame()
df_1 <- df_1[2, ]

for (n in 2: length(somaActin_sta_wave)){
  
    df_2 <- somaActin_sta_wave[[n]] %>% as.data.frame()
    
    df_1 <- rbind (df_1, df_2[2, ])
}

colnames(df_1)<- c("Negative", "Positive")

df_1$img_name<- names(somaActin_sta_list)
df_1$Negative<- df_1$Negative %>% as.numeric()
df_1$Positive<- df_1$Positive %>% as.numeric()

df_1$Neg_pct<- df_1$Negative/(df_1$Negative+df_1$Positive)
df_1$Pos_pct<- df_1$Positive/(df_1$Negative+df_1$Positive)

df_1_long <- df_1 %>% pivot_longer(cols=c("Neg_pct", "Pos_pct"), values_to = "pct", names_to="change")
```

```{r, message = FALSE, warning=FALSE, fig.width=5}
library("ggpubr")
my_comparisons <- list( c("Neg_pct", "Pos_pct") )

ggplot(df_1_long, aes(x=change, y=pct, fill=change)) + 
  geom_violin(trim=FALSE, 
              kernel ="gaussian", 
              adjust=1.3)+
  geom_boxplot(width=0.5, fill="white")  +
  geom_point(aes(fill=change,group=img_name), position = position_dodge(0.2))+
  geom_line(aes(group=img_name), position = position_dodge(0.2)) +
  stat_compare_means(comparisons = my_comparisons, paired=T, method="wilcox.test")+
  scale_y_continuous( expand = c(0, 0), limits = c(-0.2, 1.2), breaks = scales::pretty_breaks(n = 7))+
  
  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
  
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold")) #for centering the text 

# plotly::ggplotly(p)
```

```{r}
ggsave(
  "awave_pos_neg.eps",
  plot = last_plot(),
  device = "eps")
```

# cross-correlation with neurite length at individual neurite level
```{r warning=FALSE}

df_1 <- somaActin_sta_list[[1]] %>% as.data.frame()
df_1$x<- df_1$x %>%  as.numeric()
df_1[nrow(df_1)+1, ]<- "NA"
#df_1[nrow(df_1)+1, ]<- "NA"
df_1$x<- df_1$x %>%  as.numeric()

df_2 <- wave_frame_list_2[[1]] %>% as.data.frame()
df_1$wave_frame<- df_2$x %>%  as.numeric()
df_1$Time<- (0:480)*2
  
ccfPlot <- ccf (df_1$x,  
                df_1$wave_frame, 
                type = "correlation", 
                na.action = na.pass,
                plot=F, 
                lag.max=50) 
    
ccf <- ccfPlot[["acf"]] %>% as.data.frame()

for (n in 2:length(somaActin_sta_list)){
  
    df_1 <- somaActin_sta_list[[n]] %>% as.data.frame()
    df_1$x<- df_1$x %>%  as.numeric()
    df_1[nrow(df_1)+1, ]<- "NA"
    #df_1[nrow(df_1)+1, ]<- "NA"
    df_1$x<- df_1$x %>%  as.numeric()
    
    df_2 <- wave_frame_list_2[[n]] %>% as.data.frame()
    df_1$wave_frame<- df_2$x %>%  as.numeric()
    df_1$Time<- (0:480)*2
      
    ccfPlot_2 <- ccf (df_1$x,  
                    df_1$wave_frame, 
                    type = "correlation", 
                    na.action = na.pass,
                    plot=F, 
                    lag.max=50) 
      
      ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
      
      ccf <- cbind(ccf, ccf_2)
  
}

colnames(ccf)<- names(somaActin_sta_list)
ccf$lag<- c(-50:50)

```

```{r}
selected_ccf <- ccf %>% as.data.table()

selected_ccf$lag<- selected_ccf$lag/freq

stacked_ccf_mat <- selected_ccf %>% gather(Neurite, ccf, -lag) # select which CCF of measurements to show
df_summary <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf)
```

```{r}
ggsave(
  "somaActin_wave.eps",
  plot = last_plot(),
  device = "eps")
```
```{r}
df_summary %>% write_csv("somaactin_wave_corr.csv")
```
# cross-correlation of stationary PN data

```{r warning=FALSE}
selected_list <- comb_sta_list
ccf_PN_list<- vector(mode = "list", length = length(all_list)+1)

names(ccf_PN_list)<- c(names(all_list), "soma")

for (n in 1:8){
  
    ccfPlot <- ccf (selected_list[[1]][,(n*2-1)],  
                            selected_list[[1]][,(n*2)], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=50) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf (selected_list[[m]][,(n*2-1)],  
                            selected_list[[m]][,(n*2)], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=50) 
                            
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-50:50)
  ccf_PN_list[[n]] <- ccf
}

```

```{r warning=FALSE}
# cross-correlation with stationary Pos_length data

ccf_Poslen_list<- vector(mode = "list", length = dim(selected_list[[1]])[2]-2)

names(ccf_Poslen_list)<- colnames(selected_list[[1]])[-c(1,2)] 

for (n in 3:16){
  
    ccfPlot <- ccf (selected_list[[1]][,c("Pos_length")],  
                            selected_list[[1]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=50) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf (selected_list[[m]][,c("Pos_length")],  
                            selected_list[[m]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=50) 
  
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-50:50)
  ccf_Poslen_list[[n-2]] <- ccf
}
```

```{r warning=FALSE}
# cross-correlation with stationary Neg_length data

ccf_Neglen_list<- vector(mode = "list", length = dim(selected_list[[1]])[2]-2)

names(ccf_Neglen_list)<- colnames(selected_list[[1]])[-c(1,2)] 

for (n in 3:16){
  
    ccfPlot <- ccf (selected_list[[1]][,c("Neg_length")],  
                            selected_list[[1]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=50) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf (selected_list[[m]][,c("Neg_length")],  
                            selected_list[[m]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=50) 
  
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-50:50)
  ccf_Neglen_list[[n-2]] <- ccf
}
```

```{r warning=FALSE}
# cross-correlation of actin-Lyn data

ccf_ActLyn_list<- vector(mode = "list", length = 7)

names(ccf_ActLyn_list)<- c(paste("Pos_", c("tip", "neurite", "Density"), sep=""), paste("Neg_", c("tip", "neurite", "Density"), sep=""), "soma")

for (n in 1:7){
  
  pair_df<- selected_list[[1]][,grep(names(ccf_ActLyn_list)[n],  colnames(selected_list[[1]]))] 
  
    ccfPlot <- ccf (pair_df[ ,1],  
                      pair_df[ ,2], 
                      type = "correlation", 
                      na.action = na.pass,  plot=F, 
                      lag.max=50) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
    
     pair_df2<- selected_list[[m]][,grep(names(ccf_ActLyn_list)[n],  colnames(selected_list[[m]]))] 
  
      ccfPlot_2 <- ccf (pair_df2[ ,1],  
                          pair_df2[ ,2], 
                          type = "correlation", 
                          na.action = na.pass,  plot=F, 
                          lag.max=50) 
      
      ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
      
      ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-50:50)
  ccf_ActLyn_list[[n]] <- ccf
}
```

```{r warning=FALSE}
# cross-correlation with soma actin

ccf_Actsoma_list<- vector(mode = "list", length = 8)

names(ccf_Actsoma_list)<- c(paste("Pos_", c("length", "tip_actin", "neurite_actin", "Density_actin"), sep=""), paste("Neg_", c("length","tip_actin", "neurite_actin", "Density_actin"), sep=""))

for (n in 1:8){
  
  df1 <- selected_list[[1]][ ,"soma_actin"]
  df2 <- selected_list[[1]][,grep(names(ccf_Actsoma_list)[n],  colnames(selected_list[[n]]))] 
  
    ccfPlot <- ccf (df1,  
                      df2, 
                      type = "correlation", 
                      na.action = na.pass,  plot=F, 
                      lag.max=50) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
     df1 <- selected_list[[m]][ ,"soma_actin"]
     df2 <- selected_list[[m]][,grep(names(ccf_Actsoma_list)[n],  colnames(selected_list[[n]]))]  
  
      ccfPlot_2 <- ccf (df1,  
                          df2, 
                          type = "correlation", 
                          na.action = na.pass,  plot=F, 
                          lag.max=50) 
      
      ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
      
      ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-50:50)
  ccf_Actsoma_list[[n]] <- ccf
}
```

```{r warning=FALSE}
# cross-correlation with soma Lyn

ccf_Lynsoma_list<- vector(mode = "list", length = 8)

names(ccf_Lynsoma_list)<- c(paste("Pos_", c("length", "tip_Lyn", "neurite_Lyn", "Density_Lyn"), sep=""), paste("Neg_", c("length","tip_Lyn", "neurite_Lyn", "Density_Lyn"), sep=""))

for (n in 1:8){
  
  df1 <- selected_list[[1]][ ,"soma_Lyn"]
  df2 <- selected_list[[1]][,grep(names(ccf_Lynsoma_list)[n],  colnames(selected_list[[n]]))] 
  
    ccfPlot <- ccf (df1,  
                      df2, 
                      type = "correlation", 
                      na.action = na.pass,  plot=F, 
                      lag.max=50) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
     df1 <- selected_list[[m]][ ,"soma_Lyn"]
     df2 <- selected_list[[m]][,grep(names(ccf_Lynsoma_list)[n],  colnames(selected_list[[n]]))]  
  
      ccfPlot_2 <- ccf (df1,  
                          df2, 
                          type = "correlation", 
                          na.action = na.pass,  plot=F, 
                          lag.max=50) 
      
      ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
      
      ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-50:50)
  ccf_Lynsoma_list[[n]] <- ccf
}

ccf_all_list<-list(PN=ccf_PN_list, Poslen=ccf_Poslen_list, Neglen=ccf_Neglen_list, ActLyn=ccf_ActLyn_list, somaActin=ccf_Actsoma_list, somaLyn=ccf_Lynsoma_list)

```

```{r}
selected_ccf <- ccf_all_list$Poslen$Pos_tip_Lyn  %>% as.data.table()

selected_ccf$lag<- selected_ccf$lag/freq

stacked_ccf_mat <- selected_ccf %>% gather(Neurite, ccf, -lag) # select which CCF of measurements to show
df_summary <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf)
```

```{r}
ggsave(
  "avg_2.eps",
  plot = last_plot(),
  device = "eps")
```
# Create time-series data of total neurite length, Lyn and actin intensity/density at neurite tip, shaft, and soma. 
```{r}
library(dplyr)
mts_list<- list()
mts_avg_list <- list()
mts_stationary_list <- list()

for (n in 1:length(length_list))  {
 
  mat <- length_list[[n]] %>% rowSums() %>% as.data.table() # total neurite length
  colnames(mat) <- "sum_length"
  
  # get soma intensity
  ColName <- names(length_list)[n]
  
  Dim <- nrow(Tip_Lyn_list[[n]])
  
  soma_actin_1<- soma_actin[1:Dim, ] %>% select(all_of(ColName)) %>%  as.matrix() %>% pracma::hampel(k=5)
  mat$soma_actin <-  soma_actin_1$y
  
  soma_Lyn_1<- soma_Lyn[1:Dim, ] %>% select(all_of(ColName)) %>%  as.matrix() %>% pracma::hampel(k=5)
  mat$soma_Lyn <- soma_Lyn_1$y
  
  mat$sum_tip_actin <- Tip_actin_list[[n]] %>% rowSums()
  mat$sum_tip_Lyn <- Tip_Lyn_list[[n]] %>% rowSums() 
  
  mat$sum_neurite_actin <-  Neurite_actin_list[[n]] %>% rowSums() # total actin intensity at the whole neurite
  mat$sum_neurite_Lyn <-  Neurite_Lyn_list[[n]] %>% rowSums() # total Lyn intensity at the whole neurite
  
  mat$sum_actin <- (Neurite_actin_list[[n]]- Tip_actin_list[[n]]) %>% rowSums() 
  mat$sum_Lyn <- (Neurite_Lyn_list[[n]]- Tip_Lyn_list[[n]]) %>% rowSums()  
  
  mat$sum_actin_density <- mat$sum_neurite_actin/mat$sum_length # actin density at the whole neurite
  mat$sum_Lyn_density <- mat$sum_neurite_Lyn/mat$sum_length # Lyn density at the whole neurite

  mts_list[[n]]<- mat 
  
}

names(mts_list)<- names(Tip_Lyn_list)

```

```{r}

selected_list <- mts_list
mts_stationary_list <- list()

  for (n in 1:length(selected_list))  {
 
  mts_stationary_list[[n]]<- selected_list[[n]] %>% 
    forecast::ma(5) %>%                                    # smoothing #1
    MTS::diffM(d=1) %>% 
    forecast::ma(5) %>%                                    # smoothing #2                                   
    ts(frequency = freq, start = c(0, 1))
  
  colnames(mts_stationary_list[[n]])<-colnames(selected_list[[n]])
  
  }

names(mts_stationary_list)<- names(selected_list)

```

# cross-correlation with sum neurite length
```{r warning=FALSE}

selected_list <- mts_stationary_list

ccf_sum_list<- vector(mode = "list", length = ncol(selected_list[[1]])-1)

names(ccf_sum_list)<- colnames(selected_list[[1]])[-1]

for (n in 2:11){
  
    ccfPlot <- ccf(selected_list[[1]][,1],  
                            selected_list[[1]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=80) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf(selected_list[[m]][,1],  
                            selected_list[[m]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=80) 
  
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-80:80)
  ccf_sum_list[[n-1]] <- ccf
}

```

# cross-correlation of actin-Lyn data
```{r warning=FALSE}

selected_list <-mts_stationary_list

ccf_sumActLyn_list<- vector(mode = "list", length = 4)

names(ccf_sumActLyn_list)<- c("soma", "sum_tip", "sum_neurite", "density")

for (n in 1:4){
  
  pair_df<- selected_list[[1]][,grep(names(ccf_sumActLyn_list)[n],  colnames(selected_list[[1]]))] 
  
    ccfPlot <- ccf(pair_df[ ,1],  
                      pair_df[ ,2], 
                      type = "correlation", 
                      na.action = na.pass,  plot=F, 
                      lag.max=80) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
    
     pair_df2<- selected_list[[m]][,grep(names(ccf_sumActLyn_list)[n],  colnames(selected_list[[m]]))] 
  
      ccfPlot_2 <- ccf(pair_df2[ ,1],  
                          pair_df2[ ,2], 
                          type = "correlation", 
                          na.action = na.pass,  plot=F, 
                          lag.max=80) 
      
      ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
      
      ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-80:80)
  ccf_sumActLyn_list[[n]] <- ccf
}

```

# cross-correlation with soma actin
```{r warning=FALSE}

selected_list <-mts_stationary_list

ccf_sumActsoma_list<- vector(mode = "list", length = ncol(selected_list[[1]]))

names(ccf_sumActsoma_list)<- colnames(selected_list[[1]])

for (n in 1:11){
  
    ccfPlot <- ccf(selected_list[[1]][,2],  
                            selected_list[[1]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=80) 
    
    ccf <- ccfPlot[["acf"]] %>% as.data.frame()
  
  for (m in 2:length(selected_list)) {
  
  ccfPlot_2 <- ccf(selected_list[[m]][,2],  
                            selected_list[[m]][,n], 
                            type = "correlation", 
                            na.action = na.pass,  plot=F, 
                            lag.max=80) 
  
  ccf_2 <- ccfPlot_2[["acf"]] %>% as.data.frame()
  
  ccf <- cbind(ccf, ccf_2)
  
  }
    
  colnames(ccf)<- names(selected_list)
  ccf$lag<- c(-80:80)
  ccf_sumActsoma_list[[n]] <- ccf
}

```

```{r}
ccf_sum_all_list<-list(sumlength=ccf_sum_list, ActLyn= ccf_sumActLyn_list, somaActin=ccf_sumActsoma_list, somaLyn=ccf_sumLynsoma_list)

```

```{r}
selected_ccf <- ccf_sum_all_list$sumlength$soma_actin %>% as.data.table()

selected_ccf$lag<- selected_ccf$lag/freq

stacked_ccf_mat <- selected_ccf %>% gather(Neurite, ccf, -lag) 
df_summary <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf, type = "common")
```

# Significance of the correlation function
The variance of the cross-correlation coefficient under the null hypothesis of zero correlation is approx 1/n where n is the length of the series. The coefficients are also asymptotically normal. So approximate critical values (at the 5% level) are ±2/√n

To reduce the effect of noise, the correlation functions for all sampling windows were averaged by a spline fit (Fig. S4b). To test whether variable relationships also contributed significantly to the window-to-window variation of the correlation function we calculated confidence intervals for the average correlation function of a particular cell in two ways: i) The 95% confidence  level  of  Pearson’s  correlation  coefficient  is  defined  by  1.96N,  with  N  denoting  the  number of samples in the time course (blue lines in Fig. S4b). ii) Variation of the correlation functions about  the  spline  fit  was  calculated  from  bootstrapping  2000  samples  from  the  residuals  to  the  spline  fit8.  

The   correlation   function   for   a   cell   was   considered   insignificant   if   either   the   correlation   maximum/minimum  did  not  reach  above/below  the  95%  confidence  levels  of  Pearson’s  correlation  coefficient,  or  the  bootstrapped  95%  confidence  interval  about  the  fitted  spline  included  0  at  the  position    of    the    maximum/minimum.    Cells    without    as    significant    correlation    coefficient    maximum/minimum were eliminated from further analysis. 
```{r  message = FALSE, warning=FALSE, fig.width=10}
ggplot(df_summary, aes(x=lag, y=mean)) +
   geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd), fill="lightgrey")+
   geom_line(size=1, alpha=1) +
  
   stat_peaks(colour = "red", span = 15, ignore_threshold = 0.5) +
  stat_peaks(geom = "text", colour = "red", span = 15, ignore_threshold = 0.6, hjust = 2, angle = 0) +
  stat_valleys(colour = "lightblue", span = 13, ignore_threshold = 0.9) +
 stat_valleys(geom = "text", colour = "blue", span = 13, ignore_threshold = 0.9, hjust = -0.5, angle = 0) +
  
   theme_ipsum(grid = F, ticks = T)+ 
   geom_vline(xintercept=0, linetype="dashed", color = "black")+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
   scale_y_continuous(expand = c(0, 0), limits = c(-0.5 ,0.5),  breaks = scales::pretty_breaks(n = 6))+
   scale_x_continuous(expand = c(0, 0),  breaks = scales::pretty_breaks(n = 8), limits = c(-50,50)) +
  
  theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text
```

```{r}
ggsave(
  "sumLength_somaLyn.eps",
  plot = last_plot(),
  device = "eps")
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
fig_A <- stacked_ccf_mat %>% 
ggplot(aes(x=lag, y=ccf)) + geom_point(alpha=0.3, size=0.5) +
   geom_line(aes(color=Cell), size=0.8)+ theme(legend.position = "none") 

fig_B <-stacked_ccf_mat %>% 
  ggplot(aes(x = lag, y = ccf, fill=Cell)) +
    geom_area(alpha=0.5 , size=.5, colour="grey") +
    scale_fill_viridis(discrete = T) +  scale_fill_brewer(palette = "Set2") +
    theme_ipsum(grid = F, ticks = T)+ theme(legend.position = "none") 

fig_C <-ggplot(df_summary, aes(x=Time, y=mean)) +
   geom_line(size=1, alpha=0.8) +
   geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd), fill="blue", alpha=0.2)

ggpubr::ggarrange(fig_A,fig_B,fig_C,
                  labels = LETTERS[1:3],
                  ncol = 2, nrow = 2,  legend="none")

```

# Time-lagged cross-corrleation of outgrowth and retraction profiles
```{r warning=FALSE}
# constructing partial dependence (i.e., marginal effect) plots 
library(pdp)
theme_set(theme_bw())
mts_avg_list[[2]][ ,c("Length.2", "Length.4")] %>%  autoplot() +
  ylim(0, 45) +
  ggtitle (plot_title) +
  geom_line(aes(),size=1)  +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold")) #for centering the text
```

```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
save.image(file=saved_path)

```

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

## Package installation
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
