---
title: "ED Fig 11i — Somatic EB3 intensity timecourse during washout"
output: html_notebook
---

<!--
# What this file does

Computes the **time-course of somatic EB3 fluorescence intensity** during
the CK666 → washout recovery window. For each cell, intensity is binned
into 25-frame time windows, normalized to that cell's own mean (so each
cell starts around 1.0), then aggregated to a population mean ± SE
ribbon over time. A paired test compares bin 0 vs bin 4 (early vs late
washout).

This file feeds **ED Fig 11i** (Source Data sheet `ED Fig 11i`,
columns `Batch | bin | cell | Intensity_norm`) — the integrated EB3
line graph over time during the washout vs CK666 recovery experiment.

## NOT to be confused with the sibling file `edfig11f_eb3_tip_integrated_intensity.Rmd`

These two files look similar at first read but compute different things:

| Aspect | This file (`edfig11i_*`) | Sibling (`edfig11f_*`) |
|---|---|---|
| ROI | **Soma** (`soma_EB3.csv`) | **Neurite tip** (`Tip_EB3.csv`) |
| Time treatment | **Binned timecourse** (25-frame bins) | **Single timepoint** per condition |
| Comparison | Bin 0 vs bin 4 (paired *t*-test) | Before vs CK666 vs Washout (Wilcoxon) |
| Source Data | ED Fig 11i (`Batch \| bin \| cell \| Intensity_norm`) | ED Fig 11f (`Cell_id \| IntDen \| Condition \| Batch \| Label`) |
| Output PDFs | `soma_EB3_timelapse.pdf`, `soma_EB3.pdf` | `TipEB3_CK666.pdf` |

If you're tracking ED Fig 11i (the line graph over time): this file.
If you're tracking ED Fig 11f (the boxplot at three discrete conditions):
the sibling.

## Inputs

- `df_list.RData` (cached environment, optional).
- `soma_EB3.csv` files in subfolders (one per acquisition session).
  Each CSV has time as the first column and one column per cell, with
  intensity values per frame.
- The folder structure under `Path_1` is expected to include a path
  segment matching `CK666_LA_EB3/<batch_id>/`; the regex
  `sub(".*/CK666_LA_EB3/([^/]+).*", "\\1", ...)` extracts the batch ID
  for each CSV.

## Outputs (written next to this Rmd)

- `soma_EB3_timelapse.pdf` — population-level line plot of normalized
  somatic EB3 intensity vs time bin (mean ± SE ribbon).
- `soma_EB3.pdf` — violin + paired-line plot comparing bin 0 to bin 4
  per cell, with a paired *t*-test.
- Updated `df_list.RData` (re-saved at end with post-analysis state).

## Beginner notes

- **EB3-SnAP-SiR** = EB3 (a microtubule plus-end-tracking protein)
  fused to a SnAP-tag and labeled with a silicon-rhodamine (SiR)
  far-red fluorophore. Brighter intensity = more polymerizing
  microtubule plus-ends in the ROI.
- **Why bin into 25-frame windows?** Acquisition is at high temporal
  resolution (one frame per few seconds); 25-frame bins (~1-2 minutes
  each) reduce noise while preserving the slow recovery dynamics
  during washout.
- **Why normalize each cell to its own mean?** Different cells have
  different baseline EB3 expression levels. Per-cell normalization
  (each cell's mean = 1) puts every cell on a comparable scale before
  population averaging, so a cell with 2× expression doesn't dominate
  the mean.
- **Outlier removal at bin 3** uses `rstatix::identify_outliers` (Q1 −
  1.5·IQR / Q3 + 1.5·IQR rule) on bin 3 specifically (mid-recovery)
  and drops the affected cells from all bins for consistency.
- **Why paired *t*-test bin 0 vs bin 4?** Each cell appears in both
  bins (paired observations); the test asks whether per-cell intensity
  differs between early and late recovery.

## Original G:\ source

`G:/Figure S11_Microtubules/Fig. S11f/R_scripts/Neurite activity corr.Rmd`
-->

```{r message=FALSE, warning=FALSE, include=FALSE}
# Library imports. Commented-out libraries (pracma, peakPick, quantmod,
# pheatmap, viridis, hrbrthemes, superheat) were in the original but
# were either disabled by the author or are unused by this notebook.
# Left commented for traceability with the G:\ source.
#library(pracma)
#library(peakPick)
#library(quantmod)
library(ggpmisc)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggfortify)
library(forecast)
library(tsibble)
library(broom)
library(tidyquant)
library(tsbox)
#library("pheatmap")
library("RColorBrewer")
library(ggpmisc)
library(ggfortify)
#library(viridis)
#library(hrbrthemes)
library(data.table)
#library(superheat)
```

## Load existing/saved environment (.RData)

```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list.RData", sep = "/")
load(file = saved_path)
```

## Save environment (utility — re-saves intermediate state)

```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list.RData", sep = "/")
save.image(file = saved_path)
```

## Locate and read all `soma_EB3.csv` files (one per acquisition session)

```{r}
library(readr)
library(purrr)

# 1) Set the root folder you want to search.
root <- Path_1

# 2) Find matching files (recursively).
files <- list.files(root,
                    pattern = "^soma_EB3\\.csv$",
                    recursive = TRUE,
                    full.names = TRUE)

if (length(files) == 0) stop("No files named 'soma_EB3.csv' were found under: ", root)

# 3) Extract a batch ID for each CSV from the path
#    (e.g., .../CK666_LA_EB3/240312_DIV1/soma_EB3.csv -> 240312_DIV1).
obj_names <- sub(".*/CK666_LA_EB3/([^/]+).*", "\\1", gsub("\\\\", "/", files))

# 4) Read into a named list, key = batch ID.
soma_list <- set_names(
  map(files, ~ read_csv(.x, show_col_types = FALSE)),
  obj_names
)
```

## Bin into 25-frame time windows + per-cell normalization

For each session-level CSV: bin the rows into groups of 25 (approx 1-2
minutes per bin given the acquisition rate), average within each bin,
then normalize every cell column by its own mean (so each cell's series
centers around 1.0). Reshape to long format for plotting and concatenate
all sessions.

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

soma_list_2 <- list()

for (n in 1:length(soma_list)) {
  df <- soma_list[[n]]
  # First column is time; all others are per-cell intensities.
  bin_size <- 25
  time_var <- names(df)[1]

  # Average rows within each bin of `bin_size` frames.
  df_avg25 <- df %>%
    mutate(bin = ceiling(row_number() / bin_size)) %>%
    group_by(bin) %>%
    summarise(
      !!time_var := mean(.data[[time_var]], na.rm = TRUE),     # midpoint time of the bin
      across(-all_of(time_var), ~ mean(.x, na.rm = TRUE)),
      .groups = "drop"
    )

  # Normalize each cell column by its own mean so comparisons are scale-invariant.
  df_norm <- df_avg25[-2] %>%
    mutate(across(-1, ~ {
      m <- mean(., na.rm = TRUE)
      if (is.finite(m) && m != 0) . / m else NA_real_
    }))

  df_norm$bin <- df_norm$bin - 1   # make bins 0-indexed for plotting

  # Long format for plotting.
  df_long <- df_norm %>%
    pivot_longer(-1, names_to = "cell", values_to = "value_norm")

  soma_list_2[[n]] <- df_long
}

names(soma_list_2) <- names(soma_list)

# Concatenate all sessions into one long table; keep batch-id `source` column.
combined <- bind_rows(soma_list_2, .id = "source") %>%
  relocate(source, .before = 1) %>% as.data.table()
```

## Outlier removal (cells whose bin-3 normalized value is an outlier)

```{r}
outliers <- combined[bin == 3] %>%
  group_by(bin) %>%
  rstatix::identify_outliers(value_norm)

# Drop the outlier cells from ALL bins so the per-cell traces stay intact.
combined_2 <- combined[!(cell %in% outliers$cell)]
```

```{r}
summary_2 <- combined_2 %>%
  group_by(bin) %>%
  rstatix::get_summary_stats(value_norm) %>% as.data.table()
```

## Plot — per-cell traces over time

```{r}
combined_2 %>%
  ggplot(aes(x = bin, y = value_norm, color = cell, group = cell)) +
  geom_line(alpha = 0.9, linewidth = 0.5) +
  labs(x = "bin",
       y = "Normalized value (value / column mean)",
       color = "Cell") +
  theme_minimal() +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 5),
                     breaks = scales::pretty_breaks(n = 5)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0.4, 1.8),
                     breaks = scales::pretty_breaks(n = 5)) +
  theme(legend.position = "none")
```

## Plot — population mean ± SE ribbon over time (THIS IS ED FIG 11i)

```{r message=FALSE, warning=FALSE, fig.width=10}
library(hrbrthemes)
summary_2 %>%
  ggplot(aes(x = bin, y = mean)) +
  geom_ribbon(aes(ymin = mean - se, ymax = mean + se), fill = "lightgrey") +
  geom_line(size = 1, alpha = 1) +
  theme_ipsum(grid = FALSE, ticks = TRUE) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  scale_y_continuous(expand = c(0, 0), limits = c(0.9, 1.1),
                     breaks = scales::pretty_breaks(n = 6)) +
  scale_x_continuous(expand = c(0, 0),
                     breaks = scales::pretty_breaks(n = 5),
                     limits = c(0, 5)) +
  theme_linedraw() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text = element_text(size = 12, face = "bold"))
```

```{r}
ggsave("soma_EB3_timelapse.pdf", plot = last_plot(), device = "pdf")
```

## Paired comparison — bin 0 vs bin 4 (early vs late washout)

```{r message=FALSE, warning=FALSE, fig.width=10}
library(ggpubr)

my_comparisons <- list(c("0", "4"))
pj <- position_jitter(seed = 1, width = 0.1, height = 0)

combined_2[bin == 4 | bin == 0] %>%
  mutate(bin = as.character(bin)) %>%
  ggplot(aes(x = bin, y = value_norm)) +
  theme_classic() +
  geom_violin(trim = FALSE, kernel = "gaussian", adjust = 1, scale = "area") +
  geom_boxplot(width = 0.2, outlier.colour = NA, fill = "white") +
  geom_jitter(shape = 16, position = pj) +
  geom_path(aes(group = cell), alpha = 0.5, position = pj) +
  stat_compare_means(comparisons = my_comparisons,
                     paired = TRUE, method = "t.test",
                     label = "{p.adj.format}{p.adj.signif}") +
  theme_ipsum(grid = FALSE, ticks = TRUE) +
  theme_linedraw() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text = element_text(size = 12, face = "bold", angle = 0))
```

```{r}
combined_2[bin == 4 | bin == 0]$bin %>% table()
ggsave("soma_EB3.pdf", plot = last_plot(), device = "pdf")
```
