---
title: "Canonical template: Fig 4o brightfield CCF (Lyn-live) for one genotype"
output: html_notebook
---

<!--
================================================================================
  CANONICAL TEMPLATE - DO NOT RENDER DIRECTLY
================================================================================
  This Rmd is INCLUDED by 2 thin wrappers in figures/Fig04_Arp3KO/:

    fig04o_lyn_live_neurite_xcorr_wt.Rmd  (WT genotype, pixel=0.3243227, freq=0.2)
    fig04o_lyn_live_neurite_xcorr_ko.Rmd  (Arp3 KO, pixel=0.1081075, freq=0.8)

  Each wrapper sets params and includes this file via:
      ```{r child = "../../templates/fig04o_lyn_live_neurite_xcorr.Rmd"}
      ```

  -------- PARAMETERS expected --------
    params$genotype          "WT" or "KO" (metadata only)
    params$pixel             pixel-to-micrometer calibration (numeric)
    params$freq              imaging frame frequency in Hz (numeric)
    params$cell_range_start  first cell number (integer)
    params$cell_range_end    last cell number (integer)
    params$cell_prefix       "Cell_" (WT) or "Cell" (KO) - folder-name prefix

  -------- PANEL TARGETS (Lin et al., Nature 2026) --------
  Fig 4o: brightfield neurite extension-retraction CCF for in-utero
          electroporation cohorts.  WT vs Arp3 KO neurons.

  Extracted: 2026-05-12 from 2 sibling 2000+-line wrappers.
  Essentiality: Template (canonical parameterized analysis).
================================================================================
-->

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

<!--
================================================================================
  GUIDED TOUR — read this before diving into the chunks below
================================================================================
  This 2000-line template is the canonical brightfield-CCF analysis for Fig 4o.
  It reads per-cell time-series of neurite lengths and computes cross-
  correlation functions between neurites within a cell. There are ~80 chunks,
  but they fall into the following logical phases:

    PHASE A — SETUP (chunks 1-3)
       Library imports, env restore, one-time file sorting helper.

    PHASE B — DATA INGEST (chunks 4-7, ~lines 92-220)
       Read per-cell CSV outputs from the Bradke-lab "Neurite length"
       ImageJ macro. Each cell has multiple neurites tracked over time;
       stack them into multivariate time-series (`mts_*_list`).

    PHASE C — EXTENSION/RETRACTION PHASE DETECTION (chunks 8-15, lines ~280-700)
       Convert each neurite's length trace to its derivative (`stationary`),
       smooth with `forecast::ma`, identify extension events (peaks) and
       retraction events (valleys) by thresholding the derivative.

    PHASE D — POPULATION SUMMARIES + PLOTS (chunks 16-25, lines ~700-960)
       Per-cell summaries (sum/count of extension/retraction events,
       integrated outgrowth) and per-population summary plots.

    PHASE E — CROSS-CORRELATION ANALYSIS (chunks 26-50, lines ~970-1700)
       The headline analysis: pairwise CCFs between neurites within each
       cell. Builds `ccf_*_list`, computes lag-zero correlations, aggregates
       across cells. Outputs the ccf-curve PDFs that feed Fig 4o.

    PHASE F — PEAK/VALLEY DETECTION + CHANGEPOINT ANALYSES (chunks ~50-65,
              lines ~1700-2000)
       Several alternative peak-finding implementations (peakPick, pracma,
       changepoint, etc.) explored interactively — only one is used in the
       headline output; others are kept as exploratory.

    PHASE G — TEARDOWN (last chunks)
       Save the working environment back to df_list.RData for the next render.

  -------- KEY VARIABLES TO KNOW --------
    Path_1               wrapper directory (chosen at render time)
    file_list            absolute paths of every file recursively under Path_1
    subfolder_list       per-cell folders (one per cell)
    csv_list             per-cell CSV file paths
    df_list              raw per-cell wide tables (one element per cell)
    mts_avg_list         multivariate time-series per cell (per-neurite length)
    mts_stationary_list  first-difference (delta) of each mts_avg_list element
    selected_list        whichever list is fed into the CCF step
    ccf_*_list           CCF results per neurite-pair per cell
    rollccf_list         rolling-window CCF (alternative to full-window CCF)

  -------- KEY PARAMETERS (set by the wrapper YAML) --------
    params$genotype          metadata only
    params$pixel             µm-per-pixel calibration (drives length scaling)
    params$freq              imaging frequency in Hz (drives time axis)
    params$cell_range_start  first cell number (for file sorting helper)
    params$cell_range_end    last cell number
    params$cell_prefix       "Cell_" or "Cell" (folder name convention)

  -------- COMMON IDIOM: stationary curves --------
    "stationary" in this codebase means "first-differenced" — i.e., the
    time-derivative computed by diff() / forecast::stationarity tests.
    A neurite's length is non-stationary (drifts over time); its DERIVATIVE
    (extension velocity) is roughly stationary, which is what CCF analyses
    require to avoid spurious lag-0 correlations from shared trends.
================================================================================
-->

## Step 1: Load libraries

A larger superset than the smaller templates because the CCF pipeline
uses time-series + peak-detection + multi-cell aggregation packages.
See annotations on the small templates for which packages each chunk uses.

```{r message=FALSE, warning=FALSE, include=FALSE}
library(pracma)        # findpeaks(), gradient(), rolling stats
library(peakPick)      # alternative peak-detection algorithm
library(quantmod)      # findPeaks/findValleys aliases
library(ggpmisc)       # stat_peaks/stat_valleys for ggplot
library(tidyverse)     # dplyr, tidyr, purrr, ggplot, readr
library(magrittr)      # %>%, %<>% pipe operators
library(ggplot2)
library(ggfortify)     # autoplot.ts() for one-line ggplot of mts objects
library(forecast)      # ma() moving average + time-series utilities
library(tsibble)       # tidy time-series tibble
library(broom)         # tidy()/glance() for stat models
library(tidyquant)
library(tsbox)         # cross-format ts conversions
library("pheatmap")
library("RColorBrewer")
library(ggpmisc)
library(ggfortify)
library(viridis)
library(hrbrthemes)
library(data.table)
```

## Step 2: Load saved environment (.RData) — fast path

Same fast-load convention as the other templates: on subsequent renders,
this restores all named objects from a prior successful run.
`setwd(Path_1)` is set here so file-write operations in subsequent chunks
land in the wrapper's directory.

```{r}
Path_1     <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(Path_1)
saved_path <- paste(Path_1, "df_list.RData", sep = "/")
load(file = saved_path)
```

## Step 3 (one-time, manual): File sorting into per-Cell folders

When raw data first arrives, individual .tif files are scattered at the top
level. This helper sorts them into `Cell_<n>/` (or `Cell<n>/`,
depending on `params$cell_prefix`) sub-folders so every downstream step
can iterate over a clean per-cell structure.

`grep(x, file_list)` searches for the literal cell folder name (e.g.
"Cell_11") inside each path. Matching .tif files are moved into that
folder. This is **destructive** (moves, not copies) but idempotent — if
the files are already sorted, the grep returns nothing and nothing moves.

```{r}
file_list <- list.files(path = Path_1, full.names = TRUE, recursive = TRUE)

library(filesstrings)   # provides file.move()

# Loop over the cell-index range supplied by the wrapper
# (e.g. WT: 11..13; KO: 10..18). Each iteration creates a destination
# folder and moves matching .tif files into it.
for (n in params$cell_range_start:params$cell_range_end)  {

   # Build the cell folder name; params$cell_prefix supplies "Cell_" or "Cell".
   x    <- paste(params$cell_prefix, n,  sep = "")
   dir.create(x)
   path <- paste(Path_1, "/", x, sep = "")

   # Find files whose paths contain this cell name; move them.
   file_tif <- file_list[grep(x, file_list), drop = FALSE]
   file.move(file_tif, path)
}
```

## Read .csv files
```{r}
subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list<- subfolder_list [ grepl("Cell", subfolder_list) ]

csv_list<- list()
df_list<- list()

pixel <- params$pixel

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  csv_list[[n]] <- list.files(subfolder_list[n],pattern = "(.*)csv$", 
                     recursive = TRUE,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list[[n]], read.csv, header = TRUE))
  df <-df[, grep(c("Tip.point|Length"),  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  colnames(df)<- paste("Length", seq( from = 1, to = ncol(df) ),  sep = ".")
  
  df_list[[n]]<- df
 
}

names(df_list)<- subfolder_list
```

## Import time-series neurite length data
```{r}

# multivariate time series: multiple neurites
mts_list<- list()
mts_avg_list <- list()
mts_stationary_list <- list()

freq <- params$freq

for (n in 1:length(df_list))  {
 
  length_mat<-df_list[[n]] %>% as.data.frame()
  length_mat$sum_length <- length_mat %>% rowSums() 
  
  mts_list[[n]]<- length_mat %>% ts(frequency = freq, start = c(0, 1)) 
  
  # Moving Average to smoothen out outliers/noise
  mts_avg_list[[n]] <- length_mat %>% forecast::ma(3) %>% ts(frequency = freq, start = c(0, 1)) 
  colnames(mts_avg_list[[n]])<-colnames(length_mat)
  
  # Differential 
  mts_stationary_list[[n]] <- mts_avg_list[[n]] %>% MTS::diffM(d=1) %>% ts(frequency = freq, start = c(0, 1))
  
}

names(mts_list)<- names(df_list)
names(mts_avg_list)<- names(df_list)
names(mts_stationary_list)<- names(df_list)

```

## Plotting the time series as curves
```{r}
library(ggthemes)

mts_profile_list <- list()
avg_profile_list <- list()
sta_profile_list <- list()
thres= 0.3 # change threshold for stationary profile here

for (n in 1:length(mts_avg_list))  {
  
 plot_title<- paste(names(df_list)[n], sep = "")
 selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(mts_avg_list[[n]]))

 theme_set(theme_bw())
 avg_profile_list[[n]] <- mts_avg_list[[n]][, selected_data] %>%  autoplot() +
  #ggtitle (plot_title) +
   theme_linedraw()+
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 45))+
  geom_line(aes(),size=1)  +
   scale_color_viridis(discrete=T)+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_text(size=12)) #for centering the text
 
 mts_profile_list[[n]] <- mts_list[[n]][, selected_data] %>%  autoplot() +
 # ggtitle (plot_title) +
    theme_linedraw()+
  geom_line(aes(),size=1)  +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold")) #for centering the text
 
 sta_profile_list[[n]] <- mts_stationary_list[[n]][, selected_data] %>%  autoplot() +
  #ggtitle (plot_title) +
    theme_linedraw()+
  geom_line(aes(),size=0.25)  +
   geom_hline(yintercept=0, linetype="dashed", color = "red")+
   geom_hline(yintercept=thres, linetype="dashed", color = "blue")+
   geom_hline(yintercept=-thres, linetype="dashed", color = "blue")+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold")) #for centering the text
 
}

names(mts_profile_list) <- names(df_list)
names(avg_profile_list) <- names(df_list)
names(sta_profile_list) <- names(df_list)
```

## Visualization of the selected curves
```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=10}
fig_A <- mts_profile_list[[2]]
fig_B <- avg_profile_list[[2]]
fig_C <- sta_profile_list[[2]]

ggpubr::ggarrange(fig_B, fig_C,
                  labels = c("B", "C"),
                  ncol = 1, nrow = 2,  legend="none")

```

```{r}
ggsave(
  "avg_2.eps",
  plot = last_plot(),
  device = "eps")
```

## We convert the multiple times series data to appropriate dataframe for stack area plot
```{r}

selected_list <-mts_stationary_list
stacked_avg_list <- list()

for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  
  colNames <- colnames(selected_list[[m]][ ,selected_data])
  length_mat<- selected_list[[m]][, selected_data] %>%  forecast::ma(3) %>% as.data.frame() # add forecast::ma(5) if smoothening is wanted
  colnames(length_mat) <- colNames 
  
  mts_dim<- dim(length_mat)
  length_mat$Time <- as.numeric(rep(seq(0,mts_dim[1]-1)* 1/freq), times=mts_dim[2] )
  
 stacked_avg_list [[m]] <- length_mat %>% gather(Neurite, Length, -Time) %>% as.data.table()
  
}
```

```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=10}
stacked_avg_list[[2]] %>% 
  ggplot(aes(x = Time, y = Length, fill = Neurite)) +
    geom_area(alpha=1 ) +
  theme_linedraw()+
  
    scale_fill_viridis(discrete = T) + 
  scale_y_continuous( limits = c(-1.2, 1.8), breaks = scales::pretty_breaks(n = 7))+
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 

  # scale_fill_brewer(palette = "Set2") +
   # theme_ipsum(grid = F, ticks = T) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none", 
        axis.text=element_text(size=12)) #for centering the text
```

```{r}
ggsave(
  "avg_2.eps",
  plot = last_plot(),
  device = "eps")
```

## Calculate neurite outgrowth/retraction events from the stationary/differential curves

```{r}
selected_list <- mts_stationary_list # mts_stationary_list or mts_avg_list

pos_neg_thres_list <- list()
# thres= 0.02 # change threshold here

for (n in 1: 10){
  
  thres= n*0.05
  pos_neg_list <- list()
  
  for (m in 1:length(selected_list)) {
    
    selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
    length_mat<- selected_list[[m]][, selected_data] %>% as.data.frame()
    
    length_mat2 <- length_mat
    neurite_num <- ncol(length_mat)
  
    tab1 <- apply(length_mat, 1, function(x) sum(x > thres, na.rm = TRUE)) %>%  table() %>% as.data.table()
    tab1$act<- "extension"
    names(tab1)[1]<- "Num"
    tab1$Num <- tab1$Num %>% as.numeric()
    tab1$pct<- tab1$N/sum(tab1$N)
    tab1$rel_n <- tab1$Num/neurite_num
    tab1$Cell<- names(df_list)[m]
    
    tab2 <- apply(length_mat, 1, function(x) sum(x < -thres, na.rm = TRUE)) %>%  table() %>% as.data.table()
    tab2$act<- "retraction"
    names(tab2)[1]<- "Num"
    tab2$Num <- tab2$Num %>% as.numeric()
    tab2$pct<- tab2$N /sum(tab2$N)
    tab2$rel_n <- tab2$Num/neurite_num
    tab2$Cell<- names(df_list)[m]
  
    tab3 <- apply(length_mat, 1, function(x) sum(x < thres & x > -thres, na.rm = TRUE)) %>%  table() %>% as.data.table()
    tab3$act<- "pause"
    names(tab3)[1]<- "Num"
    tab3$Num <- tab3$Num %>% as.numeric()
    tab3$pct<- tab3$N/sum(tab3$N)
    tab3$rel_n <- tab3$Num/neurite_num
    tab3$Cell<- names(df_list)[m]
    
    tab4 <- rbind(tab1, tab2, tab3)
    tab4$Thres<- thres
    
    pos_neg_list[[m]] <- tab4
  }

  names(pos_neg_list)<- names(df_list)
  
  pos_neg_thres_list[[n]]<- assign(paste0("Thres_", thres, sep=""),pos_neg_list)
  names(pos_neg_thres_list)[n]<- paste0("Thres_", thres, sep="")
  
}
```

```{r message=FALSE, warning=FALSE}
summary_list <- list()

for (n in 1: 10){
  
  pos_neg_list <- pos_neg_thres_list[[n]]
  
  pos_neg_tab_all <- pos_neg_list[[1]] 
  
  for (m in 2:length(pos_neg_list)) {
    
    df2 <- pos_neg_list[[m]] 
    pos_neg_tab_all <- rbind(pos_neg_tab_all, df2) 
  
  }

  tab <- pos_neg_tab_all[act == "extension"] %>% group_by(Num) %>% rstatix::get_summary_stats(pct)
  tab$Thres<- n*0.05
  summary_list [[n]] <- tab 

}

names(summary_list) <-  names(pos_neg_thres_list)

df_summary_all <- summary_list [[1]]

for (n in 2: 10){
  df_summary_all <-  rbind (df_summary_all, summary_list [[n]])
  
}

```

```{r  message = FALSE, warning=FALSE, fig.width=10}
ggplot(df_summary_all, aes(x=Num, y=mean, group=Thres)) +
  
  geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd, fill=Thres),   alpha=0.05) + 
  geom_line(aes(colour=Thres), size=1, alpha=1) +
  
  scale_y_continuous(  breaks = scales::pretty_breaks(n = 6))+
  scale_x_continuous(name="Neurite_No", limits=c(0,4), expand = c(0, 0), breaks = c(0,1,2,3,4,5,6))+
  
  scale_fill_viridis(discrete = F) + # scale_fill_brewer(palette = "Set2") +
  scale_color_viridis(discrete = F)+

  theme_linedraw()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12), 
        legend.position = "none") #for centering the text
```

```{r}
ggsave(
  "extension_freq.pdf",
  plot = last_plot(),
  device = "pdf")
```

```{r}
pos_neg_list <- pos_neg_thres_list$Thres_0.5

pos_neg_tab_all <- pos_neg_list[[1]] 

for (m in 2:length(pos_neg_list)) {
  
  df2 <- pos_neg_list[[m]] 
  pos_neg_tab_all <- rbind(pos_neg_tab_all, df2) 

}

tab <- pos_neg_tab_all[act == "extension"]

df_summary <- tab %>% group_by(Num) %>% rstatix::get_summary_stats(pct)

```

```{r  message = FALSE, warning=FALSE, fig.width=10}
# remove cells in "without any outgrowth" state more than 60% of the time 
removed <- pos_neg_tab_all[act == "extension" & Num == 0 & pct > 0.5] 
removed_cell<- removed$Cell
pos_neg_tab_all_removed<- pos_neg_tab_all[grep(paste(removed_cell,collapse="|"), pos_neg_tab_all$Cell , invert=TRUE) ,]
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
ggplot(data=tab, aes(x=Num, y=pct, group=Cell)) +
  geom_line((aes(color=Cell)))+
  geom_point((aes(color=Cell)))+ 
scale_x_continuous(name="Neurite_No", limits=c(0, 6), breaks = c(0,1,2,3,4,5,6))+
theme_linedraw()+
  
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12), 
        legend.position = "none") #for centering the text
```

```{r}
ggsave(
  "WT_neuriteNumFreq_avg.eps",
  plot = last_plot(),
  device = "eps")
```

## We smoothen the stationary profiles first, separete outrgwoth and retraction events, and sum the changes of all neurites. (Smoothening -> separate -> sum)
```{r warning=FALSE}

# Smoothen the stationary (diff) profiles and optionally filter out small length change with threshold

selected_list <- mts_stationary_list # mts_stationary_list or mts_avg_list

mts_stationary_pos_list <- list()
mts_stationary_neg_list <- list()
thres= 0.0 # change threshold here

# Positive peaks only (outgrowth event)
for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  length_mat<- selected_list[[m]][, selected_data ]
  
  length_mat_avg<- length_mat %>% forecast::ma(3) %>% as.ts() # Average smoothening
  colnames(length_mat_avg)<-colnames(length_mat)

  mts_dim<- dim(length_mat_avg)
    for (n in 1:mts_dim[2]){
      length_mat_avg[, n][length_mat_avg[ , colnames(length_mat_avg)[n]] < thres] <- 0 
    }
  
  length_mat_avg <- length_mat_avg %>% as.data.frame() 
  length_mat_avg$sum <- length_mat_avg  %>% rowSums()  # We sum up the extent of outgrowth at each time point
  mts_stationary_pos_list[[m]] <- length_mat_avg %>% ts(frequency = freq, start = c(0, 1)) 

}

# Negative peaks only (retraction event)
for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  length_mat<- selected_list[[m]][, selected_data ]
  length_mat_avg<- length_mat %>% forecast::ma(3) %>% as.ts() # Average smoothening
  colnames(length_mat_avg)<-colnames(length_mat)
  
  mts_dim<- dim(length_mat_avg)
    for (n in 1:mts_dim[2]){
      length_mat_avg[, n][length_mat_avg[ , colnames(length_mat_avg)[n]] > - thres] <- 0 
    }
  length_mat_avg <- length_mat_avg %>% as.data.frame() 
  length_mat_avg$sum <- length_mat_avg  %>% rowSums()  # We sum up the extent of retraction at each time point
  mts_stationary_neg_list[[m]] <- length_mat_avg %>% ts(frequency = freq, start = c(0, 1)) 
}

```

```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=10}
selected_stationary <- mts_stationary_pos_list[[15]]
selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_stationary))

figA <- selected_stationary[ ,selected_data] %>%  autoplot() +
  geom_line(aes(),size=0.25)  +
   geom_hline(yintercept=0, linetype="dashed", color = "red")+
   geom_hline(yintercept=thres, linetype="dashed", color = "blue")+
   geom_hline(yintercept=-thres, linetype="dashed", color = "blue")+
   scale_color_viridis(discrete=T)+
   theme_minimal()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold"))

selected_stationary <- mts_stationary_neg_list[[15]]

figB <- selected_stationary[ ,selected_data] %>%  autoplot() +
   geom_line(aes(),size=0.25)  +
   geom_hline(yintercept=0, linetype="dashed", color = "red")+
   geom_hline(yintercept=thres, linetype="dashed", color = "blue")+
   geom_hline(yintercept=-thres, linetype="dashed", color = "blue")+
   scale_color_viridis(discrete=T)+
   theme_minimal()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold")) 
 
 ggpubr::ggarrange(figA, figB, 
                  labels = c("A", "B"),
                  ncol = 1, nrow = 2,  legend="none")
 
```

## We convert the multiple times series data to appropriate dataframe for stack area plot
```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=10}

# Positive peaks only (outgrowth event)
selected_list <- mts_stationary_pos_list
stacked_stationary_pos_list <- list()

for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  length_mat<- selected_list[[m]][, selected_data] %>% as.data.frame() 
  mts_dim<- nrow(length_mat)-1
  length_mat$Time <- as.numeric(rep(seq(0,mts_dim)*2))
  
  stacked_stationary_pos_list[[m]] <- length_mat %>% gather(Neurite, Length, -Time) %>% as.data.table()

}

# Negtive peaks only (retraction event)
selected_list <- mts_stationary_neg_list
stacked_stationary_neg_list <- list()

for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
  length_mat<- selected_list[[m]][, selected_data] %>% as.data.frame() 
   mts_dim<- nrow(length_mat)-1
  length_mat$Time <-  as.numeric(rep(seq(0,mts_dim)*2))
  
stacked_stationary_neg_list[[m]] <- length_mat %>% gather(Neurite, Length, -Time) %>% as.data.table()
}

```

```{r warning=FALSE}
figA <-  stacked_stationary_pos_list[[2]] %>% 
  ggplot(aes(x = Time, y = Length, fill = Neurite)) +
    geom_area(alpha=0.5 , size=.5, colour="white") +
    scale_fill_viridis(discrete = T) + # scale_fill_brewer(palette = "Set2") +
    theme_ipsum(grid = F, ticks = T) 

figB <- stacked_stationary_neg_list[[2]] %>% 
  ggplot(aes(x = Time, y = Length, fill = Neurite)) +
    geom_area(alpha=0.5 , size=.5, colour="white") +
    scale_fill_viridis(discrete = T) + # scale_fill_brewer(palette = "Set2") +
    theme_ipsum(grid = F, ticks = T) 
 
 ggpubr::ggarrange(figA, figB, 
                  labels = c("A", "B"),
                  ncol = 1, nrow = 2,  legend="none")
 
```

# We plot the overall outgrowth/retraction level over time
```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=10}
selected_stationary_pos <- mts_stationary_pos_list[[2]][, "sum"]
selected_stationary_neg <- mts_stationary_neg_list[[2]][, "sum"]
ts.union(selected_stationary_pos, selected_stationary_neg) %>%  autoplot() +
  geom_line(aes(),size=0.25)  +
   geom_hline(yintercept=0, linetype="dashed", color = "black")+
   #geom_hline(yintercept=thres, linetype="dashed", color = "blue")+
   #geom_hline(yintercept=-thres, linetype="dashed", color = "blue")+
   scale_y_continuous( limits = c(-1.2, 1.8), breaks = scales::pretty_breaks(n = 7))+
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text
```

```{r}
ggsave(
  "avg_2.eps",
  plot = last_plot(),
  device = "eps")

```

## Here we separate the outrgwoth and retraction events first, sum the changes of all neurites, and then do smoothening. (separate -> sum -> smoothening)
```{r}

selected_list <- mts_stationary_list 
mts_stationary_sum_list <- list()

thres= 0.0 # change threshold here

# Positive peaks only (outgrowth event)
for (m in 1:length(selected_list)) {
  selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]])) # Neurite length ("Length") only or include actin intensity (Integerated.intensity)
  length_mat_pos<- selected_list[[m]][, selected_data ]
  length_mat_neg<- selected_list[[m]][, selected_data ]
  
  mts_dim<- dim(length_mat_pos)
  
    for (n in 1:mts_dim[2]){
      length_mat_pos[, n][length_mat_pos[ , colnames(length_mat_pos)[n]] < 0] <- 0 
      length_mat_neg[, n][length_mat_neg[ , colnames(length_mat_neg)[n]] > 0] <- 0 
    }
  
  length_mat_pos <- as.data.frame(length_mat_pos)
  length_mat_neg <- as.data.frame(length_mat_neg)
   
  length_mat_pos$sum <- length_mat_pos  %>% rowSums()  
  length_mat_neg$sum <- length_mat_neg  %>% rowSums()  
  
  length_mat_pos <- length_mat_pos %>% ts(frequency = freq, start = c(0, 1))
  length_mat_neg <- length_mat_neg %>% ts(frequency = freq, start = c(0, 1))
  
  length_mat_sum <- ts.union(length_mat_pos[ ,"sum"], length_mat_neg[ ,"sum"]) 
  
  length_mat_sum <- length_mat_sum %>% forecast::ma(3) %>% as.ts() # Average smoothening
  colnames(length_mat_sum)<-c("Pos", "Neg")
  
  length_mat_sum [, "Pos"][length_mat_sum [, "Pos"] < thres] <- 0 
  length_mat_sum [, "Neg"][length_mat_sum [, "Neg"] > -thres] <- 0 
  
  mts_stationary_sum_list[[m]] <- length_mat_sum

}

names(mts_stationary_sum_list)<- names(mts_stationary_list )

```

```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=10}
selected_stationary_sum <- mts_stationary_sum_list[[2]]

 selected_stationary_sum %>%  autoplot() +
  geom_line(aes(),size=0.25)  +
   geom_hline(yintercept=0, linetype="dashed", color = "red")+
   geom_hline(yintercept=thres, linetype="dashed", color = "blue")+
   geom_hline(yintercept=-thres, linetype="dashed", color = "blue")+
   theme_ipsum(grid = F, ticks = T) + 
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold"),
        legend.position = "none")  #for centering the text
 # stat_peaks(colour = "red", ignore_threshold = 0.1) +
 # stat_peaks(geom = "text", colour = "red", vjust = -0.5) +
 # stat_valleys(colour = "blue", ignore_threshold = 0.1) + 
 # stat_valleys(geom = "text", colour = "blue", angle = 45, vjust = 1.5, hjust = 1)
```

# Time-lagged cross-correlation of outgrowth and retraction profiles
<!--
================================================================================
  HEADLINE CCF ANALYSIS — this is where Fig 4o's source data is computed.
================================================================================
  For each cell, we have two time-series:
     selected_pos  -> per-frame "outgrowth/extension" signal (sum across
                      all neurites of POSITIVE length-derivatives)
     selected_neg  -> per-frame "retraction" signal (sum of |NEGATIVE|
                      length-derivatives)

  The cross-correlation function (CCF) at lag k tells us the linear
  correlation between outgrowth at time t and retraction at time t+k,
  scanned across many k values. A negative-lag peak means outgrowth leads
  retraction; a positive-lag peak means retraction leads outgrowth.

  Below: one CCF curve per cell (`length_ccf_list[[m]]`), then aggregate
  across cells into the mean +/- SD ribbon plot that appears in Fig 4o.

  -------- the `ccf()` call signature --------
    ccf(x, y, lag.max = NULL, plot = TRUE, type = "correlation")
      x, y   : the two time-series to correlate
      lag.max: number of lags to compute (auto-chosen if NULL)
      plot   : suppress the auto-plot here (we collect values, plot later)
      type   : "correlation" -> Pearson r at each lag

  The returned object has $acf (correlation values) and $lag (the lag axis,
  in time-series step units; multiply by 1/freq to get seconds).
================================================================================
-->
```{r  message = FALSE, warning=FALSE, fig.width=10}
selected_ccf <- length_ccf %>% as.data.table()
selected_ccf$lag<- selected_ccf$lag/freq 

stacked_ccf_mat <- selected_ccf %>% gather(Cell, ccf, -lag) # select which CCF of measurements to show
df_summary <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf)
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
ggplot(df_summary, aes(x=lag, y=mean)) +
  
   geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd), fill="lightgrey", alpha=1)+
   geom_line(size=1, alpha=1) +
  
  stat_peaks(colour = "red", span = 15, ignore_threshold = 0.4) +
  stat_peaks(geom = "text", colour = "red", span = 15, ignore_threshold = 0.4, hjust = -0.1, angle = 90) +
  stat_valleys(colour = "lightblue", span = 21, ignore_threshold = 0.7) +
  stat_valleys(geom = "text", colour = "blue", span = 21, ignore_threshold = 0.7, hjust = -0.1, angle = 90) +
  
  theme_ipsum(grid = F, ticks = T)+ 
   geom_vline(xintercept=0, linetype="dashed", color = "black")+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
   scale_y_continuous( breaks = scales::pretty_breaks(n = 7))+
  scale_x_continuous(expand = c(0, 0), limits = c(-80, 80), breaks = scales::pretty_breaks(n = 8)) + 
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text

```

```{r}
ggsave(
  "length_ccf.pdf",
  plot = last_plot(),
  device = "pdf")

```

```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=10}
ggplot(stacked_ccf_mat, aes(x=lag, y=ccf)) + geom_point(alpha=0.3, size=0.5) +
   geom_line(aes(color=Cell), size=0.8) + theme(legend.position = "none")
```

```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=10}
stacked_ccf_mat %>% 
  ggplot(aes(x = lag, y = ccf, fill=Cell)) +
    geom_area(alpha=1 , size=.1, colour="white") +
    #scale_fill_viridis(discrete = T) + 
  # scale_fill_brewer(palette = "Set2") +
   theme_ipsum(grid = F, ticks = T)+
  theme_linedraw()+
  scale_x_continuous(expand = c(0, 0), limits = c(-100, 100), breaks = scales::pretty_breaks(n = 9)) + 

  # scale_fill_brewer(palette = "Set2") +
   # theme_ipsum(grid = F, ticks = T) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none", 
        axis.text=element_text(size=18)) #for centering the text
```

```{r}
ggsave(
  "length_ccf.eps",
  plot = last_plot(),
  device = "eps")

```

```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
save.image(saved_path)
```

## Package installation
