---
title: "Canonical template: pooled-replicate neurite-actin CCF"
output: html_notebook
---

<!--
================================================================================
  CANONICAL TEMPLATE - DO NOT RENDER DIRECTLY
================================================================================
  This Rmd is INCLUDED by 2 thin wrappers (one per panel target):

    figures/EDFig01_coordination/edfig01be_2d_culture_neurite_freq_and_xcorr_pooled.Rmd
       -> ED Fig 1b (2D-culture pooled-replicate CCF, N = 101 from 4 exps)

    figures/EDFig02_actin/edfig02c_neurite_actin_xcorr_pooled_sd.Rmd
       -> ED Fig 2c (pooled-replicate CCF with SD envelope; same algorithm
          as Fig 2e but with replicate-pooling aggregation)

  Each wrapper sets nothing (no params needed): the analysis is identical;
  Path_1 resolves to the wrapper's own directory, picking up the LOCAL
  per-cell CSV inputs.  So wrapper placement = data selection.

  Include via:
      ```{r child = "../../templates/pooled_replicate_ccf.Rmd"}
      ```

================================================================================
  REPO STATUS
================================================================================
  Extracted: 2026-05-12 from 2 sibling wrappers whose R bodies were
             byte-identical (2305 lines each, 0 diff). Wrapper differences
             were entirely in the HTML-comment header annotations.
  Net reduction: 4685 lines -> ~2380 lines (~49% saved).
  Essentiality: Template (canonical parameterized analysis).
================================================================================
-->

<!--
================================================================================
  GUIDED TOUR — read this before diving into the chunks below
================================================================================
  This 2300-line template is the "pooled-replicate CCF" pipeline shared by:

    figures/EDFig01_coordination/edfig01be_2d_culture_neurite_freq_and_xcorr_pooled.Rmd
       (-> ED Fig 1b, 2D culture cohort, N = 101 neurons / 4 experiments)
    figures/EDFig02_actin/edfig02c_neurite_actin_xcorr_pooled_sd.Rmd
       (-> ED Fig 2c, mean +/- SD envelope, N = 135 neurons / 5 experiments)

  Both wrappers run the IDENTICAL code on different data folders. The
  wrapper's working directory selects the data via `Path_1 = dirname(...)`.

  -------- WHAT "POOLED REPLICATE" MEANS HERE --------
  The key analytical difference vs. the per-cell CCF templates is the
  aggregation: this version groups individual neurons by experimental
  replicate (one experiment = one day of imaging), computes per-replicate
  mean CCF curves, and then plots mean +/- SD across replicates. So the
  uncertainty band reflects between-experiment variation, not within-cell.

  -------- PIPELINE PHASES --------
    PHASE A — SETUP (chunks 1-3)
        Library imports + env restore. Note: this template uses
        `df_list_comb.RData` (the "_comb" suffix marks replicate-combined
        data; the per-cell template uses `df_list.RData`).

    PHASE B — DATA INGEST (chunks 4-7)
        Read per-cell CSVs from sub-folders. Group by experiment day
        (encoded in the parent folder name).

    PHASE C — EXTENSION/RETRACTION PHASE DETECTION (similar to fig04o)
        Convert lengths to derivatives, identify extension/retraction
        events with threshold-based peak detection.

    PHASE D — POPULATION + PER-REPLICATE SUMMARIES
        Compute per-experiment summary curves before pooling.

    PHASE E — CCF + MEAN-RIBBON PLOT (this is the published-panel output)
        Pairwise CCFs across neurites within each cell; then average over
        cells WITHIN each experiment; then mean +/- SD across experiments.

    PHASE F — ADDITIONAL DIAGNOSTIC PLOTS (peakPick, changepoint, etc.)
        Exploratory analyses left in for audit; not in the headline figure.

    PHASE G — TEARDOWN
        Save env to df_list_comb.RData for the next render.

  -------- WRAPPER PARAMETERS --------
  NONE. The 2 wrappers are byte-identical except for their header comments.
  Data selection happens via wrapper directory (`Path_1`).

  -------- COMPANION FILE --------
  The per-cell variant of this analysis is at
  `figures/Fig02_actin_wave/fig02e_neurite_actin_xcorr_culture.Rmd`.
  Both implement the same CCF algorithm; the only difference is the
  aggregation step (per-cell vs per-replicate).
================================================================================
-->

## Step 1: Load libraries

Same package superset as the per-cell CCF template plus `ggsignif` for
significance-bracket annotations and `gdata` for `cbindX` (NA-tolerant
column-binding).

```{r message=FALSE, warning=FALSE, include=FALSE}
library(pracma)        # findpeaks, gradient
library(peakPick)      # alternative peak detection
library(quantmod)      # findPeaks/findValleys aliases
library(tidyverse)     # dplyr/tidyr/purrr/ggplot
library(magrittr)
library(ggplot2)
library(ggfortify)     # autoplot.ts
library(forecast)      # moving-average smoothing
library(tsibble)
library(broom)
library(tidyquant)
library(tsbox)
library("pheatmap")
library("RColorBrewer")
library(ggpmisc)
library(ggfortify)
library(viridis)
library(hrbrthemes)
library(data.table)
library(superheat)
library(rstatix)       # get_summary_stats, identify_outliers
library(gdata)         # cbindX — unequal-row column-bind
library(ggpubr)        # stat_compare_means
library(ggsignif)      # significance brackets for ggplot
```

## Step 2: Load saved environment (.RData) — fast path

Note: this template uses `df_list_comb.RData` rather than the per-cell
template's `df_list.RData`. The `_comb` suffix marks "combined replicates"
— the original G:\ source folder was named with a `_comb` suffix on the
filename to distinguish it from the per-cell variant.

```{r}
Path_1     <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list_comb.RData",  sep = "/")
load(file = saved_path)

```

# Neurite length
# Data 180123 and 181018
```{r}

subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("180123|181018", subfolder_list) ]

length_list1<- list()
freq <- 0.5 # set frequency (1/interval)
pixel <- 0.1082410 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  
 csv_list <- list.files(subfolder_list [n],pattern = "^Result(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Length|Tip",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  
  colnames(df)<- paste("Length", seq( from = 1, to = ncol(df) ),  sep = ".")
  
  length_list1[[n]] <-df

}

subfolder_name_list <- subfolder_list %>% fs::path_file()

names(length_list1) <- gsub("Kymo_", "", subfolder_name_list)
```

# Data DVcore
```{r}

subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("DVcore", subfolder_list) ]

length_list_DVcore<- list()
freq <- 0.5 # set frequency (1/interval)
pixel <- 0.1609000 # set pixel size 

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  
 csv_list <- list.files(subfolder_list [n],pattern = "^Result(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Length|Tip",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  
  colnames(df)<- paste("Length", seq( from = 1, to = ncol(df) ),  sep = ".")
  
  length_list_DVcore[[n]] <-df

}

subfolder_name_list <- subfolder_list %>% fs::path_file()

names(length_list_DVcore) <- gsub("Kymo_", "", subfolder_name_list)

```

# Data 180129(DMSO)
```{r}

subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("180129", subfolder_list) ]

length_list_180129<- list()
freq <- 0.5 # set frequency (1/interval)
pixel <- 0.0648000 # set pixel size 

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  
 csv_list <- list.files(subfolder_list [n],pattern = "^Result(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Length|Tip",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  
  colnames(df)<- paste("Length", seq( from = 1, to = ncol(df) ),  sep = ".")
  
  length_list_180129[[n]] <-df

}

subfolder_name_list <- subfolder_list %>% fs::path_file()

names(length_list_180129) <- gsub("Kymo_", "", subfolder_name_list)
```

# Data 180207
```{r}

subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("180207", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("anaylzed_ori", subfolder_list) ]

length_list_180207<- list()
freq <- 0.5 # set frequency (1/interval)
pixel <- 0.0648000 # set pixel size 0.1082410

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  
 csv_list <- list.files(subfolder_list [n],pattern = "^Result(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Length|Tip",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  
  colnames(df)<- paste("Length", seq( from = 1, to = ncol(df)),  sep = ".")
  
  length_list_180207[[n]] <-df

}

subfolder_name_list <- subfolder_list %>% fs::path_file()

names(length_list_180207) <- gsub("Kymo_", "", subfolder_name_list)
```

# Data 180207_binned
```{r}

subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("180207", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("binned", subfolder_list) ]

length_list_180207_scaled<- list()
freq <- 0.5 # set frequency (1/interval)
pixel <- 0.1296000 # set pixel size 0.1082410

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  
 csv_list <- list.files(subfolder_list [n],pattern = "^Result(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Length|Tip",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  
  colnames(df)<- paste("Length", seq( from = 1, to = ncol(df)),  sep = ".")
  
  length_list_180207_scaled[[n]] <-df

}

subfolder_name_list <- subfolder_list %>% fs::path_file()

names(length_list_180207_scaled) <- gsub("Kymo_", "", subfolder_name_list)
```

# Data Max_170905
```{r}

subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo", subfolder_list) ]
subfolder_list  <- subfolder_list [ grepl("Max_170905", subfolder_list) ]

length_list_Max_170905<- list()
freq <- 0.5 # set frequency (1/interval)
pixel <- 0.1609000 # set pixel size 0.1082410

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  
 csv_list <- list.files(subfolder_list [n],pattern = "^Result(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Length|Tip",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)
  
  colnames(df)<- paste("Length", seq( from = 1, to = ncol(df)),  sep = ".")
  
  length_list_Max_170905[[n]] <-df

}

subfolder_name_list <- subfolder_list %>% fs::path_file()

names(length_list_Max_170905) <- gsub("Kymo_", "", subfolder_name_list)
```

# Data 180205_4min
```{r}

subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("180205_4min/kymo/Kymo", subfolder_list) ]

length_list_180205_4min<- list()
freq <- 0.25 # set frequency (1/interval)
pixel <- 0.0648 # set pixel size

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list))  
{
  
 csv_list <- list.files(subfolder_list [n],pattern = "^Result(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  df <-df[, grep("Tip.point",  colnames(df)), drop = FALSE] %>% as.data.table()
  df <- mapply(`*`,df, pixel)  %>% as.data.frame()

  colnames(df)<- paste("Length", seq( from = 1, to = dim(df)[2] ),  sep = ".")
  
  # Interpolation of time series data to match temporal resoultion
  df$time<- (0:(nrow(df)-1)*4)
  newTime <- seq(0,960, by=2) %>% as.data.frame()
  names(newTime) <- "time"
  # Replace NA by Interpolation
  df <- left_join(x=newTime,y=df,by=c("time")) %>% zoo::na.approx() %>% as.data.table()
  
  length_list_180205_4min[[n]] <- df[, -1] 

}

subfolder_name_list <- subfolder_list %>% fs::path_file()
names(length_list_180205_4min) <- gsub("Kymo_", "", subfolder_name_list)
```

```{r}
length_list <- c(length_list1, length_list_DVcore, length_list_180129, length_list_180207, length_list_180207_scaled, length_list_Max_170905, length_list_180205_4min)

#subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
#subfolder_list <- subfolder_list [ grepl("Kymo", subfolder_list) ]
#subfolder_list <- subfolder_list %>% fs::path_file()
#subfolder_list <- gsub("Kymo_", "", subfolder_list)

length_list <- length_list[names(Tip_actin_list)]

names(length_list)<- gsub(" ", ".", names(length_list))
```

```{r}
for (n in 1:length(length_list)){
  
  df <- length_list[[n]]
  
  row_odd <- seq_len(nrow(df)) %% 2  
  length_list[[n]] <-df[row_odd == 1, ]     
  
}

```

# Tip actin intensity  (100 pixels) 
```{r}

subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo", subfolder_list) ]

Tip_actin_list<- list()

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list ))  
{
 
 csv_list <- list.files(subfolder_list [n],pattern = "^Result(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table() 
  
  colnames(df)<- paste("Tip_actin", seq( from = 1, to =  ncol(df) ),  sep = ".")
  
  Tip_actin_list[[n]]<-  mapply(`*`,df, 1/10000)
 
}

subfolder_name_list <- subfolder_list %>% fs::path_file()

names(Tip_actin_list) <- gsub("Kymo_", "", subfolder_name_list)

names(Tip_actin_list)<- gsub(" ", ".", names(Tip_actin_list))

```

```{r}

list_180205_4min<- Tip_actin_list[59:116]

for (n in 1:length(list_180205_4min))  
{
  df <- list_180205_4min[[n]] %>% as.data.frame()

  df$time<- (0:(nrow(df)-1)*4)
 newTime <- seq(0,960, by=2) %>% as.data.frame()
  names(newTime) <- "time"
  df <- left_join(x=newTime,y=df,by=c("time")) %>% zoo::na.approx() %>% as.data.table()
  
  list_180205_4min[[n]] <- df[, -1] 

}

Tip_actin_list[59:116]<- list_180205_4min
```

```{r}

for (n in 1:length(Tip_actin_list)){
  
  df <- Tip_actin_list[[n]]
  
  row_odd <- seq_len(nrow(df)) %% 2  
  Tip_actin_list[[n]] <-df[row_odd == 1, ]     
  
}

```

# Actin intensity at neurite shaft (neurite length - 100 pixels)
```{r}

subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)
subfolder_list  <- subfolder_list [ grepl("Kymo", subfolder_list) ]

Neurite_actin_list<- list()

# read csv files in the subfolders of subfolder list
for (n in 1:length(subfolder_list ))  
{
  csv_list <- list.files(subfolder_list [n],pattern = "^Actin_neurite(.*)csv$", 
                     recursive = F,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
  
  df <- do.call("cbind", lapply(csv_list, read.csv, header = TRUE))
  
  df <-df[, grep("Integerated.intensity",  colnames(df)), drop = FALSE] %>% as.data.table()
  
  colnames(df)<- paste("Neurite_actin", seq( from = 1, to =  ncol(df) ),  sep = ".")
  
  Neurite_actin_list[[n]]<- mapply(`*`,df, 1/10000)
 
}

names(Neurite_actin_list) <- names(Tip_actin_list)
names(Neurite_actin_list)<- gsub(" ", ".", names(Neurite_actin_list))
```

```{r}

list_180205_4min<- Neurite_actin_list[59:116]

for (n in 1:length(list_180205_4min))  
{
  df <- list_180205_4min[[n]] %>% as.data.frame()

  df$time<- (0:(nrow(df)-1)*4)
 newTime <- seq(0,960, by=2) %>% as.data.frame()
  names(newTime) <- "time"
  df <- left_join(x=newTime,y=df,by=c("time")) %>% zoo::na.approx() %>% as.data.table()
  
  list_180205_4min[[n]] <- df[, -1] 

}

Neurite_actin_list[59:116]<- list_180205_4min
```

```{r}

for (n in 1:length(Neurite_actin_list)){
  
  df <- Neurite_actin_list[[n]]
  
  row_odd <- seq_len(nrow(df)) %% 2  
  Neurite_actin_list[[n]] <-df[row_odd == 1, ]     
  
}

```

# Actin  density of the whole neurite (neurite shaft)/neurite length
```{r}

Density_actin_list<- list()

for (n in 1:length(length_list))  {
  
  df <- (Neurite_actin_list[[n]])/(length_list[[n]])
  colnames(df)<- paste("density_actin", seq( from = 1, to = ncol(df)),  sep = ".")
  df[sapply(df, is.infinite)] <- NA
  Density_actin_list[[n]]<- df
  
}

names(Density_actin_list) <- names(Neurite_actin_list) 
names(Density_actin_list)<- gsub(" ", ".", names(Density_actin_list))

```

```{r}

for (n in 1:length(Density_actin_list)){
  
  df <- Density_actin_list[[n]]
  
  row_odd <- seq_len(nrow(df)) %% 2  
  Density_actin_list[[n]] <-df[row_odd == 1, ]     
  
}

```

# Soma actin (missing soma of Data 180129(DMSO))
```{r}

library("gdata")

subfolder_list <- list.dirs(path =Path_1, full.names = TRUE, recursive = TRUE)

# read csv files in the subfolders of subfolder list
 csv_soma_list <- list.files(Path_1, pattern = "^soma(.*)csv$", 
                     recursive = T,  #list of subfolders and files within the subfolders
                     include.dirs = F, # Should subdirectory names be included 
                     full.name = TRUE)  #get the full name (path and file name) of each file
 
soma_Actin_df <- do.call("cbindX", lapply(csv_soma_list , read.csv, header = TRUE))
soma_Actin_df <- soma_Actin_df[ colnames(soma_Actin_df)!="X"] 

names(soma_Actin_df)[114:146] <- names(Tip_actin_list)[141:173]
names(soma_Actin_df)[4]<- "Long_LifeactmScar_03_03_R3D.dv_2"

soma_Actin_df <- soma_Actin_df[gsub(" ", ".", names(Tip_actin_list)[-c(32:58)])] #sort column by the order of Tip_actin list

```

```{r}

df <- soma_Actin_df[32:89]

df$time<- (0:(nrow(df)-1)*4)
newTime <- seq(0,960, by=2) %>% as.data.frame()
names(newTime) <- "time"
df <- left_join(x=newTime,y=df,by=c("time")) %>% zoo::na.approx() %>% as.data.table()
  
soma_Actin_df[32:89]<- df[, -1]

soma_Actin_df <- soma_Actin_df[gsub(" ", ".", names(Tip_actin_list)[-c(32:58)])] #sort column by the order of Tip_actin list
```

```{r}

row_odd <- seq_len(nrow(df)) %% 2  
soma_Actin_df <-soma_Actin_df[row_odd == 1, ]     

```

```{r}
all_list<-list(Length=length_list, Tip_actin=Tip_actin_list, Neurite_actin=Neurite_actin_list, Density_actin = Density_actin_list)
```

# Create stationary (differential) curve of all measurements
```{r}

freq <- 0.25 # set frequency (1/interval)

all_sta_list <- list()

for (m in 1:length(all_list)) {
  
  selected_list <- all_list[[m]]
  sta_list <- list()
  
  for (n in 1:length(selected_list))  {
  
  sta <- selected_list[[n]] %>% forecast::ma(5)                           # smoothing #1
      colnames(sta)<-colnames(selected_list[[n]])
  sta_list[[n]]<- sta %>% MTS::diffM(d=1) %>%  ts(frequency = freq, start = c(0, 1))

  }

names(sta_list)<-names(length_list)
  
all_sta_list[[m]]<- sta_list
  
}

names(all_sta_list)<- names(all_list )

#  stationary (differential) curve of soma actin
soma_sta<- soma_Actin_df   %>% forecast::ma(5) %>% MTS::diffM(d=1) %>% forecast::ma(5) %>% ts(frequency = freq, start = c(0, 1))
colnames(soma_sta)<-names(soma_Actin_df )

```

# check the curve of created stationary data
```{r}
selected_list <- all_sta_list$Length$Kymo_Long_LifeactmScar_02_037_R3D_S2_good.dv

selected_list <- soma_sta[, 57]  

 selected_list %>%  autoplot( ts.colour = 'black', size = 0.5) +
 
   geom_hline(yintercept=0, linetype="dashed", color = "blue")+
   theme_ipsum(grid = F, ticks = T) + 
   theme_linedraw()+
    scale_fill_viridis(discrete = T) + 
  scale_y_continuous(  breaks = scales::pretty_breaks(n = 7))+
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA), breaks = scales::pretty_breaks(n = 8)) + 
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold"),
        legend.position = "none")  #for centering the text
```

# Prepare data for cross-correlation analysis at individual neurite level
```{r}

library("data.table")

all_sta_cell_list <- list()

for (k in 1:length(length_list)){
  
neurite_list <- list()

dt1 <- all_sta_list[[1]][[k]] %>% as.data.table()
NC <- dt1 %>% ncol()

    for (n in 1:NC){
      
     dt <- dt1[,..n]
     
          for (m in 2:4){
            
            dt2 <- all_sta_list[[m]][[k]] %>% as.data.table()
            dt<- cbind(dt, dt2[, ..n])
              
          }
    
    neurite_list[[n]] <- dt
    
    }

names(neurite_list)<- paste("Neurite_", 1:NC, sep = "")

all_sta_cell_list[[k]]<- neurite_list

}

names(all_sta_cell_list) <- names(all_sta_list[[1]])

```

# cross-correlation with neurite length at individual neurite level
<!--
================================================================================
  HEADLINE CCF ANALYSIS — this is where ED Fig 1b / ED Fig 2c source data
  is computed.
================================================================================
  For each cell, we pair neurite 1 vs neurite N (for N = 2, 3, 4) and
  compute the cross-correlation function. The lag axis spans -75..+75
  imaging frames (lag.max=75). At each lag, the CCF measures the linear
  correlation between the two neurites' length-derivative time series.

  -------- THE PIPELINE --------
  1. selected_list   <- multivariate per-cell tables (each is a data.table
                        with one column per neurite, NA-padded)
  2. unlist(..., recursive=FALSE) flattens nested replicates into one big
                        cell list.
  3. For each pair-combination (1-vs-2, 1-vs-3, 1-vs-4), compute the CCF
     for every cell. The first cell starts the result list, then subsequent
     cells get cbindX'd alongside.
  4. The next chunks (down to ~line 1150) aggregate these per-cell CCF
     curves by experiment and plot mean +/- SD across experiments.

  -------- WHY na.action = na.pass --------
  Some cells contribute time-series shorter than the max length; those rows
  end up as NA in the cbindX'd combined table. `na.pass` tells ccf() to
  preserve NA-aware computation rather than fail.
================================================================================
-->

```{r}
selected_ccf <- ccf_neurite_length_list$Tip_actin %>% as.data.table()
selected_ccf$lag <- selected_ccf$lag/freq

stacked_ccf_mat <- selected_ccf %>% gather(Neurite, ccf, -lag) # select which CCF of measurements to show
df_summary <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf)
```

```{r}
row_odd <- seq_len(nrow(df_summary)) %% 2              # Create row indicator
df_summary  <- df_summary [row_odd == 1, ]  

```

```{r  message = FALSE, warning=FALSE, fig.width=10}
ggplot(df_summary, aes(x=lag, y=mean)) +
   geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd), fill="lightgrey")+
   geom_line(size=1, alpha=1) +
   theme_ipsum(grid = F, ticks = T)+ 
   geom_vline(xintercept=0, linetype="dashed", color = "black")+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  
   stat_peaks(colour = "red", ignore_threshold = 0.9) +
 stat_peaks(geom = "text", colour = "red", vjust = -0.5, ignore_threshold = 0.9) +
  stat_valleys(colour = "blue", ignore_threshold = 0.8)+
 stat_valleys(geom = "text", colour = "blue", angle = 0, vjust = 1.5, hjust = 1, ignore_threshold = 0.8)+

   scale_y_continuous( breaks = scales::pretty_breaks(n = 6))+
   scale_x_continuous(expand = c(0, 0),  breaks = scales::pretty_breaks(n = 8), limits = c(-100,100)) +theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text
```

```{r}
ggsave(
  "Tipactin_individual.pdf",
  plot = last_plot(),
  device = "pdf")
```

```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list_comb.RData",  sep = "/")
save.image(saved_path)
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
fig_A <- stacked_ccf_mat %>% 
ggplot(aes(x=lag, y=ccf)) + geom_point(alpha=0.3, size=0.5) +
   geom_line(aes(color=Neurite), size=0.8)+ theme(legend.position = "none") 

fig_B <- stacked_ccf_mat %>% 
  ggplot(aes(x = lag, y = ccf, fill=Neurite)) +
    geom_area(alpha=1 , size=.05, colour="white") +
    #scale_fill_viridis(discrete = T) +  
  #scale_fill_brewer(palette = "Set2") +
    theme_ipsum(grid = F, ticks = T)+ 
    geom_vline(xintercept=0, linetype="dashed", color = "black")+
    scale_y_continuous( breaks = scales::pretty_breaks(n = 5))+
    scale_x_continuous(expand = c(0, 0),  breaks = scales::pretty_breaks(n = 8)) + 
    theme_linedraw()+
    theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text

fig_C <-ggplot(df_summary, aes(x=Time, y=mean)) +
   geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd), fill="lightblue")+
   geom_line(size=1, alpha=1) +
   theme_ipsum(grid = F, ticks = T)+ 
   geom_vline(xintercept=0, linetype="dashed", color = "black")+
   scale_y_continuous( breaks = scales::pretty_breaks(n = 6))+
   scale_x_continuous(expand = c(0, 0),  breaks = scales::pretty_breaks(n = 8)) +theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text

ggpubr::ggarrange(fig_A,fig_B,fig_C,
                  labels = LETTERS[1:3],
                  ncol = 2, nrow = 2,  legend="none")

```

```{r}
ggsave(
  "avg_2.pdf",
  plot = last_plot(),
  device = "pdf")
```

## We convert the multiple times series data to appropriate dataframe for stack area plot
```{r}
library("data.table")
selected_list <-all_sta_list$Length
stacked_avg_list <- list()

for (m in 1:length(selected_list)) {

  df <- selected_list[[m]] %>%  forecast::ma(5) %>% as.data.table()
  df$Time<- (0:(nrow(df)-1))/freq
  stacked_avg_list [[m]] <- df %>%  gather(neurite, Length, -Time)

}

names(stacked_avg_list)<- names(length_list)
```

```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=10}
stacked_avg_list$Longterm_LifeactmCar_LifeactmScar_02_08_R3D.dv %>% 
  ggplot(aes(x = Time, y = Length, fill = neurite)) +
    geom_area(alpha=1 ) +
  theme_linedraw()+
    scale_fill_viridis(discrete = T) + 
  scale_y_continuous(  breaks = scales::pretty_breaks(n = 7))+
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA), breaks = scales::pretty_breaks(n = 8)) + 

  # scale_fill_brewer(palette = "Set2") +
   # theme_ipsum(grid = F, ticks = T) +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none", 
        axis.text=element_text(size=12)) #for centering the text
```

```{r}
ggsave(
  "avg_2.pdf",
  plot = last_plot(),
  device = "pdf")
```

## We smoothen the stationary profiles first, separete outrgwoth and retraction events, and sum the changes of all neurites. (Smoothening -> separate -> sum)
```{r warning=FALSE}

# Smoothen the stationary (diff) profiles and optionally filter out small length change with threshold

selected_list <- all_sta_list$Length

mts_stationary_pos_list <- list()
mts_stationary_neg_list <- list()
thres= 0.0 # change threshold here

# Positive peaks only (outgrowth event)
for (m in 1:length(selected_list)) {
  
  length_mat<- selected_list[[m]]
  
  length_mat_avg<- length_mat %>% forecast::ma(5) %>%  as.data.frame()# Average smoothening
  colnames(length_mat_avg)<-colnames(length_mat)

  nCol<- ncol(length_mat_avg)
    for (n in 1:nCol){
      length_mat_avg[, n][length_mat_avg[ , colnames(length_mat_avg)[n]] < thres] <- 0 
    }
  
  length_mat_avg$sum <- length_mat_avg  %>% rowSums()  # We sum up the extent of outgrowth at each time point
  mts_stationary_pos_list[[m]] <- length_mat_avg %>% ts(frequency = freq, start = c(0, 1)) 

}

names(mts_stationary_pos_list)<- names(selected_list)

# Negative peaks only (retraction event)
for (m in 1:length(selected_list)) {
 
  length_mat<- selected_list[[m]]
  length_mat_avg<- length_mat %>% forecast::ma(5) %>%  as.data.frame()# Average smoothening
  colnames(length_mat_avg)<-colnames(length_mat)
  
  nCol<- ncol(length_mat_avg)
    for (n in 1:nCol){
      length_mat_avg[, n][length_mat_avg[ , colnames(length_mat_avg)[n]] > - thres] <- 0 
    }
  
  length_mat_avg$sum <- length_mat_avg  %>% rowSums()  # We sum up the extent of retraction at each time point
  mts_stationary_neg_list[[m]] <- length_mat_avg %>% ts(frequency = freq, start = c(0, 1)) 
}

names(mts_stationary_neg_list)<- names(selected_list)

```

# We plot the overall outgrowth/retraction level over time
```{r warning=FALSE,  message = FALSE, warning=FALSE, fig.width=10}
selected_stationary_pos <- mts_stationary_pos_list$Long_LifeactmScar_02_046_R3D_S2_good.dv[, "sum"]
selected_stationary_neg <- mts_stationary_neg_list$Long_LifeactmScar_02_046_R3D_S2_good.dv[, "sum"]

ts.union(selected_stationary_pos, selected_stationary_neg) %>%  autoplot() +
  geom_line(aes(),size=0.25)  +
   geom_hline(yintercept=0, linetype="dashed", color = "black")+
   #geom_hline(yintercept=thres, linetype="dashed", color = "blue")+
   #geom_hline(yintercept=-thres, linetype="dashed", color = "blue")+
   scale_y_continuous( breaks = scales::pretty_breaks(n = 7))+
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA), breaks = scales::pretty_breaks(n = 8)) + 
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text
```

```{r}
ggsave(
  "avg_2.pdf",
  plot = last_plot(),
  device = "pdf")
```

## extension and retraction lifetime (duration)
```{r}
thres <- 0.00

ext_ret_list <- list(Ext = mts_stationary_pos_list,  Ret = mts_stationary_neg_list)

 list_duration <- list()  

for (k in 1:length (ext_ret_list )){

 list_1 <- list()    
 list_k <- ext_ret_list[[k]]
  
  for (m in 1:length(list_k)){
  
  df <- list_k[[m]] %>% as.data.table()
  df <- df[, !"sum"] %>% abs() %>% as.data.frame()
  df<- df[, (colSums(df, na.rm = T) != 0)] %>% as.data.table()
  
  list_2 <-  list()
  
   for (n in 1: ncol(df)){
    
    df2 <- df[ ,..n]
    df2$time <- seq(0, (nrow(df2)-1), 1)
    df2 <- df2[complete.cases(df2), ] 
    
    df2$inside <- df2[ , 1] > thres
    df2$start  <- ! c(FALSE, df2$inside[-nrow(df2)]) & df2$inside
    df2$end    <- df2$inside & ! c(df2$inside[-1], FALSE)  # Not used
    df2 <- df2[ df2$inside, ]
    df2$group  <- cumsum(df2$start)  # Number the spells
    colnames(df2)[1] <- "velocity"
    
    df3 <- plyr::ddply( df2, "group", summarize,
                        start  = min(time),
                        end    = max(time),
                        length = length(velocity),
                        mean   = mean(velocity))
   
    df3$neurite <- names(df)[n]
    
    # df3$neurite_length <- length_list[[m]][1,n] 
    list_2[[n]] <- df3

   }
  
  df3<- list_2 %>% data.table::rbindlist()
  df3$cell<- names(ext_ret_list[[k]])[m]
  df3$growth_polarization <- names(ext_ret_list)[k]

  list_1[[m]] <- df3

  }
 
 duration_df <- list_1 %>% data.table::rbindlist()

 list_duration [[k]] <-  duration_df
  
}
```

```{r}
duration_df <- list_duration  %>% data.table::rbindlist()
duration_df$length <- duration_df$length/freq
duration_df$cell_neurite <- paste(duration_df$cell, duration_df$neurite, sep = "_")

df_summary_duration <-duration_df %>% 
 group_by(cell, growth_polarization) %>% 
  rstatix::get_summary_stats(length) %>% as.data.table()
```

```{r}
outliers <- df_summary_duration %>% 
  group_by(growth_polarization) %>% 
 rstatix::identify_outliers(mean) 

df_summary_duration<- df_summary_duration[! (cell %in% outliers$cell)]

df_summary_duration$growth_polarization %>% table()
```

```{r message = FALSE, warning=FALSE, fig.width=10}
 my_comparisons <- list(c("Ext", "Ret"))
pj <- position_jitter(seed = 1, width = .05, height = 0)

df_summary_duration  %>% 
  
ggplot ( aes(x=growth_polarization, y=mean, fill=growth_polarization) )+ 
  scale_fill_brewer(palette="Dark2") + 
  theme_classic()+
geom_violin(aes(), trim=F,  kernel ="gaussian", adjust=1)  +
 geom_boxplot(aes(), width=0.2, outlier.colour = NA, fill="white")+ 
  
#geom_jitter(aes(), shape=16, position=position_jitter(0.1))+ #aes(color = batch)
   geom_point(aes(fill=growth_polarization, group=cell), position = pj)+ 
  
 geom_line(aes(group=cell), position = pj, alpha = 0.2) +
# geom_path(aes(group = cell),alpha = 0.2)+
    
 scale_y_continuous(expand = c(0, 0), limits = c(0, 50), breaks = scales::pretty_breaks(n = 8))+ 
stat_compare_means(comparisons = my_comparisons, paired=T, method="t.test") +

  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold", angle = 0)) #for centering the text 
```

```{r}
ggsave(
  "WT_DIV1-2_ext_ret_duration.pdf",
  width=14, height=7, 
  plot = last_plot(),
  device = "pdf")
```

# extension and retraction velocity
```{r}
ext_ret_list_2 <- ext_ret_list
for (n in 1: length(ext_ret_list)){
  
  list_1 <- ext_ret_list[[n]]
  sample_name <- names(list_1)
  
  for (m in 1:length(list_1)){
  
    df <- list_1[[m]] %>% as.data.frame()
    df [ df == 0 ] <- NA # replace zero with NA to avoid zero inflation
    df1 <- colMeans(df, na.rm = T)  %>% as.data.frame() # calculate the average velocity through the observation window. Or matrixStats::colMedians
    colnames(df1) <- "mean_vel"
    df1$median_vel <- df %>% as.matrix() %>% matrixStats::colMedians(na.rm = T)  # calculate the average velocity through the observation window. Or matrixStats::colMedians
    
    df1$cond <- "WT"
    df1$cell <- names(list_1)[m]
    df1$neurite <- rownames(df1)
    df1 <- df1 %>% subset (df1$neurite !="sum")
    
    df1$growth_polarization <- names(ext_ret_list)[[n]]

   list_1 [[m]] <- df1 %>% as.data.table()
      
    }
  
  names(list_1)<- sample_name

  ext_ret_list_2 [[n]] <- list_1 %>% reduce(bind_rows)  %>% as.data.frame()
}

velocity_comb_df <- ext_ret_list_2 %>% reduce(bind_rows)  %>% as.data.table()

# label extension and retraction 
velocity_comb_df$Growth<- "Extension"
velocity_comb_df[mean_vel <0]$Growth <- "Retraction"

# convert retraction velocity to positive value
velocity_comb_df$mean_vel <- velocity_comb_df$mean_vel %>% abs()
velocity_comb_df$median_vel <- velocity_comb_df$median_vel %>% abs()

velocity_comb_df[mean_vel == "NaN"]$mean_vel <- 0
velocity_comb_df[median_vel == "NaN"]$median_vel <- 0
```

# Data points at cell level (axon vs minor neurites; extension vs retraction: not polarized vs polarized)

```{r}
df_summary_median_vel <-velocity_comb_df %>% 
  group_by(cell, growth_polarization) %>% 
  rstatix::get_summary_stats(median_vel) %>% as.data.table()

```

```{r}
outliers <- df_summary_median_vel %>% 
  group_by(growth_polarization) %>% 
 rstatix::identify_outliers(mean) 

df_summary_median_vel <- df_summary_median_vel[! (cell %in% outliers$cell)]

```

```{r}
df_summary_mean_vel <-velocity_comb_df %>% 
  group_by(cell, growth_polarization) %>% 
  rstatix::get_summary_stats(mean_vel) %>% as.data.table()

```

```{r}
outliers <- df_summary_mean_vel %>% 
  group_by(growth_polarization) %>% 
 rstatix::identify_outliers(mean) 

df_summary_mean_vel <- df_summary_mean_vel[! (cell %in% outliers$cell)]
df_summary_mean_vel$growth_polarization %>% table()
```

```{r message = FALSE, warning=FALSE, fig.width=10}
 my_comparisons <- list(c("Ext", "Ret"))
pj <- position_jitter(seed = 1, width = .05, height = 0)

df_summary_mean_vel  %>% 
  
ggplot ( aes(x=growth_polarization, y=mean, fill=growth_polarization) )+ 
  scale_fill_brewer(palette="Dark2") + 
  theme_classic()+
geom_violin(aes(), trim=F,  kernel ="gaussian", adjust=1)  +
 geom_boxplot(aes(), width=0.2, outlier.colour = NA, fill="white")+ 
  
#geom_jitter(aes(), shape=16, position=position_jitter(0.1))+ #aes(color = batch)
   geom_point(aes(fill=growth_polarization, group=cell), position = pj)+ 
  
 geom_line(aes(group=cell), position = pj, alpha = 0.2) +
# geom_path(aes(group = cell),alpha = 0.2)+
    
# scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2), breaks = scales::pretty_breaks(n = 5))+ 
stat_compare_means(comparisons = my_comparisons, paired=T, method="t.test") +

  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
   theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold", angle = 0)) #for centering the text 
```

```{r}
ggsave(
  "WT_DIV1-2_ext_ret_velocity.pdf",
  width=14, height=7, 
  plot = last_plot(),
  device = "pdf")
```

# Here we separate the outrgwoth and retraction events first, sum the changes of all neurites, and then do smoothening. (separate -> sum -> smoothening)
```{r}
all_sta_PN_list <- list()

for (k in 1:length(all_sta_list)){
  
  selected_list <- all_sta_list[[k]]
  sta_PN_list <- list()
  
  thres= 0.0 # change threshold here
  
  for (m in 1:length(selected_list)) {
    
    DT <- selected_list[[m]] %>% as.data.frame()
   
    DT_pos<- DT
    DT_neg<- DT
  
    for (n in 1:dim(DT)[2]){
        DT_pos[, n][DT_pos[, n] < 0] <- 0 
        DT_neg[, n][DT_neg[, n] > 0] <- 0 
    }
    
    mat_sum <- data.frame(rowSums(DT_pos), -rowSums(DT_neg))  # add neg sign
    
    #mat_sum [, 1][mat_sum [, 1] < thres] <- 0 
    #mat_sum [, 2][mat_sum [, 2] > -thres] <- 0
    
    mat_sum <- mat_sum %>% forecast::ma(5)                    # smoothing #2
  
    colnames(mat_sum)<- c(paste("Pos_", names(all_list)[k], sep = ""), paste("Neg_", names(all_list)[k], sep = ""))
    sta_PN_list[[m]] <- mat_sum
  
  }
  
  names(sta_PN_list)<-  names(length_list)
  
  all_sta_PN_list[[k]]<- sta_PN_list
  
}

names(all_sta_PN_list)<- names(all_list )

```

# Combine all stationary PN data of each videos, and add stationary soma data
```{r}
comb_sta_list<- list()

for (m in 1:length( all_sta_PN_list[[1]])){
  
  mat <- all_sta_PN_list[[1]][m] %>% as.data.frame()
  colnames(mat)<- sub('.*\\.', '', colnames(mat))
  
  for (n in 2:length(all_sta_PN_list)){
    
    mat2 <- all_sta_PN_list[[n]][m] %>% as.data.table()
    colnames(mat2)<- sub('.*\\.', '', colnames(mat2))
    
    mat <- data.frame(mat, mat2)
  }
  
   mat3 <- soma_sta[ 1:nrow(mat) ,m] %>% as.data.table() # soma data
   colnames(mat3)<-"soma_actin"
   mat <- data.frame(mat, mat3)
  
 comb_sta_list[[m]]<- mat %>% ts(frequency = freq, start = c(0, 1)) 

}

names(comb_sta_list)<-  names(length_list)
```

## Plotting the time series as curves
```{r message = FALSE, warning=FALSE}

selected_list <- comb_sta_list
plot_list.1 <- list()

for (n in 1:length(selected_list ))  {

ts <- selected_list[[n]]
 
fig_A <- ts[, c(1,2)] %>%  autoplot() +
                                     ggtitle (names(all_sta_PN_list)[1]) +
                                      geom_hline(yintercept=0, linetype="dashed", color = "black") +
   scale_x_continuous(expand = c(0, 0), limits = c(0, NA), breaks = scales::pretty_breaks(n = 8)) + 
                                      theme(plot.title = element_text(hjust = 0.5), 
                                            plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
                                            panel.grid.major = element_blank(), 
                                            panel.grid.minor = element_blank(),
                                            axis.text=element_text(size=12, face="bold")) #for centering the text
 
fig_B <- ts[, c(3,4)] %>%  autoplot() +
                                      ggtitle (names(all_sta_PN_list)[2]) +
                                        geom_hline(yintercept=0, linetype="dashed", color = "black") +
   scale_x_continuous(expand = c(0, 0), limits = c(0, NA), breaks = scales::pretty_breaks(n = 8)) + 
                                      theme(plot.title = element_text(hjust = 0.5), 
                                            plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
                                            panel.grid.major = element_blank(), 
                                            panel.grid.minor = element_blank(),
                                            axis.text=element_text(size=12, face="bold")) #for centering the text

fig_C <-ts[, 5] %>%  autoplot() +
                                      ggtitle ("soma_actin") +
                                        geom_hline(yintercept=0, linetype="dashed", color = "black") +
                            scale_x_continuous(expand = c(0, 0), limits = c(0, NA), breaks = scales::pretty_breaks(n = 8)) + 
                                      theme(plot.title = element_text(hjust = 0.5), 
                                            plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
                                            panel.grid.major = element_blank(), 
                                            panel.grid.minor = element_blank(),
                                            axis.text=element_text(size=12, face="bold")) #for centering the text

plot_list.1[[n]]<- ggpubr::ggarrange(fig_A,fig_B,fig_C,
                  labels = LETTERS[1:3],
                  ncol = 2, nrow = 2,  legend="none")

}

names(plot_list.1)<- names(length_list)

```

## Visualization of the selected curves
# cross-correlation of stationary PN data

# cross-correlation with stationary Pos_length data
# cross-correlation with stationary Neg_length data
# cross-correlation with stationary soma_actin data
```{r}
ccf_all_list<-list(PN=ccf_PN_list, Poslen=ccf_Poslen_list, Neglen=ccf_Neglen_list, soma=ccf_somaActin_list)

```

```{r  message = FALSE, warning=FALSE, fig.width=10}
selected_ccf <- ccf_all_list$soma$Pos_Length %>% as.data.table()
selected_ccf$lag<- selected_ccf$lag/freq

stacked_ccf_mat <- selected_ccf %>% gather(Cell, ccf, -lag) # select which CCF of measurements to show
df_summary <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf)
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
ggplot(df_summary, aes(x=lag, y=mean)) +
   
   geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd), fill="lightgrey", alpha=1)+
  geom_line(size=1, alpha=1) +
 theme_ipsum(grid = F, ticks = T)+ 
   geom_vline(xintercept=0, linetype="dashed", color = "black")+
   geom_hline(yintercept=0, linetype="dashed", color = "black")+
  
    stat_peaks(colour = "red", ignore_threshold = 0.7) +
 stat_peaks(geom = "text", colour = "red", vjust = -0.5, ignore_threshold = 0.7) +
  stat_valleys(colour = "blue", ignore_threshold = 0.9)+
 stat_valleys(geom = "text", colour = "blue", angle = 0, vjust = 1.5, hjust = 1, ignore_threshold = 0.9)+
  
   scale_y_continuous( breaks = scales::pretty_breaks(n = 5))+
  scale_x_continuous(expand = c(0, 0), limits = c(-50, 50), breaks = scales::pretty_breaks(n = 8)) + 
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
fig_A <- stacked_ccf_mat %>% 
ggplot(aes(x=lag, y=ccf)) + geom_point(alpha=0.3, size=0.5) +
   geom_line(aes(color=Cell), size=0.8)+ theme(legend.position = "none") 

fig_B <-stacked_ccf_mat %>% 
  ggplot(aes(x = lag, y = ccf, fill=Cell)) +
    geom_area(alpha=1 , size=.05, colour="white") +
    #scale_fill_viridis(discrete = T) +  
  #scale_fill_brewer(palette = "Set2") +
    theme_ipsum(grid = F, ticks = T)+ 
   geom_vline(xintercept=0, linetype="dashed", color = "black")+
   scale_y_continuous( breaks = scales::pretty_breaks(n = 7))+
  scale_x_continuous(expand = c(0, 0), limits = c(-100, 100), breaks = scales::pretty_breaks(n = 8)) + 
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        legend.position = "none",
        axis.text=element_text(size=12, face="bold")) #for centering the text

fig_C <-ggplot(df_summary, aes(x=Time, y=mean)) +
   geom_line(size=1, alpha=0.8) +
   geom_ribbon(aes(ymin=mean-sem, ymax=mean+sem), fill="blue", alpha=0.2)

ggpubr::ggarrange(fig_A,fig_B,fig_C,
                  labels = LETTERS[1:3],
                  ncol = 2, nrow = 2,  legend="none")

```

```{r}
ggsave(
  "avg_2.pdf",
  plot = last_plot(),
  device = "pdf")
```

```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list.RData",  sep = "/")
save.image(file=saved_path)

```

## Import time-series neurite length data
```{r}

# multivariate time series: multiple neurites
mts_list<- list()
mts_avg_list <- list()
mts_stationary_list <- list()

freq <- 0.25 # set frequency (1/interval)
#pixel <- 0.0648 # set pixel size

for (n in 1:length(length_list))  {
 
  length_mat<-length_list[[n]] %>% as.data.table()
  length_mat$sum_length <- length_mat %>% rowSums() 
  
  intensity_mat<-Tip_actin_list[[n]]%>% as.data.table()
  intensity_mat$sum_intensity <- intensity_mat %>% rowSums()
  
  mts_list[[n]]<- cbind(length_mat, intensity_mat) %>% ts(frequency = freq, start = c(0, 1)) 
  
  # Moving Average to smoothen out outliers/noise
  mts_avg_list[[n]] <- mts_list[[n]] %>% forecast::ma(5) %>% as.ts()
  colnames(mts_avg_list[[n]])<-colnames(mts_list[[n]])
  
  # Differential 
  mts_stationary_list[[n]] <- mts_avg_list[[n]] %>% MTS::diffM(d=1) %>% ts(frequency = freq, start = c(0, 1))
  
}

names(mts_list)<- names(length_list)
names(mts_avg_list)<- names(length_list)
names(mts_stationary_list)<- names(length_list)

```

## Plotting the time series as curves
```{r}

mts_profile_list <- list()
avg_profile_list <- list()
sta_profile_list <- list()
thres= 0.3 # change threshold for stationary profile here

for (n in 1:length(mts_avg_list))  {
  
 plot_title<- paste(names(length_list)[n], sep = "")
 selected_data <- grep(paste(c("Tip.point"),collapse="|"),  colnames(mts_avg_list[[n]]))

 theme_set(theme_bw())
 avg_profile_list[[n]] <- mts_avg_list[[n]][, selected_data] %>%  autoplot() +
  ggtitle (plot_title) +
  geom_line(aes(),size=1)  +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold")) #for centering the text
 
 mts_profile_list[[n]] <- mts_list[[n]][, selected_data] %>%  autoplot() +
  ggtitle (plot_title) +
  geom_line(aes(),size=1)  +
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold")) #for centering the text
 
 sta_profile_list[[n]] <- mts_stationary_list[[n]][, selected_data] %>%  autoplot() +
  ggtitle (plot_title) +
  geom_line(aes(),size=0.25)  +
   geom_hline(yintercept=0, linetype="dashed", color = "red")+
   geom_hline(yintercept=thres, linetype="dashed", color = "blue")+
   geom_hline(yintercept=-thres, linetype="dashed", color = "blue")+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        axis.text=element_text(size=12, face="bold")) #for centering the text
 
}

names(mts_profile_list) <- names(length_list)
names(avg_profile_list) <- names(length_list)
names(sta_profile_list) <- names(length_list)
```

## Visualization of the selected curves
## Calculate neurite outgrowth/retraction events from the stationary/differential curves at different thresholds

```{r}
selected_list <- mts_stationary_list # mts_stationary_list or mts_avg_list

pos_neg_thres_list <- list()

res <- 0.03
fold<- 10

for (n in 1: fold){ 
  
  thres= n*res
  pos_neg_list <- list()
  
  for (m in 1:length(selected_list)) {
    
    selected_data <- grep(paste(c("Length"),collapse="|"),  colnames(selected_list[[m]]))
    length_mat<- selected_list[[m]][, selected_data] %>% as.data.frame()
    
    length_mat2 <- length_mat
    neurite_num <- ncol(length_mat)
  
    tab1 <- apply(length_mat, 1, function(x) sum(x > thres, na.rm = TRUE)) %>%  table() %>% as.data.table()
    tab1$act<- "extension"
    names(tab1)[1]<- "Num"
    tab1$Num <- tab1$Num %>% as.numeric()
    tab1$pct<- tab1$N/sum(tab1$N)
    tab1$rel_n <- tab1$Num/neurite_num
    tab1$Cell<- names(selected_list)[m]
    
    tab2 <- apply(length_mat, 1, function(x) sum(x < -thres, na.rm = TRUE)) %>%  table() %>% as.data.table()
    tab2$act<- "retraction"
    names(tab2)[1]<- "Num"
    tab2$Num <- tab2$Num %>% as.numeric()
    tab2$pct<- tab2$N /sum(tab2$N)
    tab2$rel_n <- tab2$Num/neurite_num
    tab2$Cell<- names(selected_list)[m]
  
    tab3 <- apply(length_mat, 1, function(x) sum(x < thres & x > -thres, na.rm = TRUE)) %>%  table() %>% as.data.table()
    tab3$act<- "pause"
    names(tab3)[1]<- "Num"
    tab3$Num <- tab3$Num %>% as.numeric()
    tab3$pct<- tab3$N/sum(tab3$N)
    tab3$rel_n <- tab3$Num/neurite_num
    tab3$Cell<- names(selected_list)[m]
    
    tab4 <- rbind(tab1, tab2, tab3)
    tab4$Thres<- thres
    
    pos_neg_list[[m]] <- tab4
  }

  names(pos_neg_list)<- names(selected_list)
  
  pos_neg_thres_list[[n]]<- assign(paste0("Thres_", thres, sep=""),pos_neg_list)
  names(pos_neg_thres_list)[n]<- paste0("Thres_", thres, sep="")
  
}
```

```{r message=FALSE, warning=FALSE}
summary_list <- list()

for (n in 1: fold){
  
  pos_neg_list <- pos_neg_thres_list[[n]]
  
  pos_neg_tab_all <- pos_neg_list[[1]] 
  
  for (m in 2:length(pos_neg_list)) {
    
    df2 <- pos_neg_list[[m]] 
    pos_neg_tab_all <- rbind(pos_neg_tab_all, df2) 
  
  }

  tab <- pos_neg_tab_all[act == "extension"] %>% group_by(Num) %>% rstatix::get_summary_stats(pct)
  tab$Thres<- n*res
  summary_list [[n]] <- tab 

}

names(summary_list) <-  names(pos_neg_thres_list)

df_summary_all <- summary_list [[1]]

for (n in 2:fold){
  df_summary_all <-  rbind (df_summary_all, summary_list [[n]])
  
}

```

```{r  message = FALSE, warning=FALSE, fig.width=10}
ggplot(df_summary_all, aes(x=Num, y=mean, group=Thres)) +
  
  geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd, fill=Thres),   alpha=0.05) + 
  geom_line(aes(colour=Thres), size=1, alpha=1) +
  
  scale_y_continuous(  breaks = scales::pretty_breaks(n = 6))+
  scale_x_continuous(name="Neurite_No", limits=c(0, 6), expand = c(0, 0), breaks = c(0,1,2,3,4,5,6))+
  
  scale_fill_viridis(discrete = F) + # scale_fill_brewer(palette = "Set2") +
  scale_color_viridis(discrete = F)+

  theme_linedraw()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12), 
        legend.position = "none") #for centering the text
```

```{r}
ggsave(
  "retraction_freq.pdf",
  plot = last_plot(),
  device = "pdf")
```

```{r}
pos_neg_list <- pos_neg_thres_list$Thres_0.3

pos_neg_tab_all <- pos_neg_list[[1]] 

for (m in 2:length(pos_neg_list)) {
  
  df2 <- pos_neg_list[[m]] 
  pos_neg_tab_all <- rbind(pos_neg_tab_all, df2) 

}

tab <- pos_neg_tab_all[act == "extension"]

df_summary <- tab %>% group_by(Num) %>% rstatix::get_summary_stats(pct)

```

```{r  message = FALSE, warning=FALSE, fig.width=10}

ggplot(df_summary, aes(x=Num, y=mean)) +
  
  geom_ribbon(aes(ymin=mean-sd, ymax=mean+sd), fill="lightgrey", alpha=1)+

  geom_line(size=1, alpha=1) +
  scale_y_continuous(  breaks = scales::pretty_breaks(n = 6))+
  scale_x_continuous(name="Neurite_No", limits=c(0,4), expand = c(0, 0), breaks = c(0,1,2,3,4,5,6))+
  
  scale_fill_viridis(discrete = F) + # scale_fill_brewer(palette = "Set2") +
  scale_color_viridis(discrete = F)+

  theme_linedraw()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12), 
        legend.position = "none") #for centering the text
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
ggplot(data=pos_neg_tab_all, aes(x=Num, y=pct, group=Cell)) +
  geom_line((aes(color=Cell)))+
  geom_point((aes(color=Cell)))+ 
scale_x_continuous(name="Neurite_No", limits=c(0, 6), breaks = c(0,1,2,3,4,5,6))+
theme_linedraw()+
  
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text=element_text(size=12), 
        legend.position = "none") #for centering the text
```

```{r}
ggsave("ExtendingNeuriteFreq.pdf",plot = last_plot(), device = "pdf")
```

# Create time-series data of total neurite length and actin intensity/density at neurite tip, shaft, and soma. 
```{r}

mts_list2<- list()
mts_avg_list2 <- list()
mts_stationary_list2 <- list()

freq <- 0.25 # set frequency (1/interval)

for (n in 1:ncol(soma_Actin_df))  {
  
  k <- names(soma_Actin_df)[n]
 
  mat <- length_list[[k]] %>% rowSums() %>% as.data.table() # total neurite length
  colnames(mat) <- "sum_length"
  
  # get soma intensity
  
  Dim <- nrow(Tip_actin_list[[k]])
  mat$soma_actin <- soma_Actin_df[1:Dim, n ] 

  mat$sum_tip_actin <- Tip_actin_list[[k]] %>% rowSums()

  mat$sum_neurite_actin <-  Neurite_actin_list[[k]] %>% rowSums()

  mat$sum_actin <- (Neurite_actin_list[[k]]+ Tip_actin_list[[k]]) %>% rowSums() # total actin intensity at the whole neurite

  mat$sum_actin_density <- mat$sum_actin/mat$sum_length # actin density at the whole neurite

  mts_list2[[n]]<- mat %>% ts(frequency = freq, start = c(0, 1)) 
  
  mts_avg_list2[[n]] <- mts_list2[[n]] %>% forecast::ma(5) %>% as.ts() # Moving Average #1
  colnames(mts_avg_list2[[n]])<-colnames(mts_list2[[n]])
  
  mts_stationary_list2[[n]] <- mts_avg_list2[[n]] %>% 
    MTS::diffM(d=1) %>% 
    forecast::ma(3) %>%                                             # Moving Average #2
    ts(frequency = freq, start = c(0, 1)) 
  
  colnames(mts_stationary_list2[[n]])<-colnames(mts_list2[[n]])
  
}

names(mts_list2)<-names(soma_Actin_df)
names(mts_avg_list2)<- names(soma_Actin_df)
names(mts_stationary_list2)<- names(soma_Actin_df)
```

```{r message=FALSE, warning=FALSE}
selected_list <- mts_stationary_list2
plot_list.2 <- list()

for (n in 1:length(selected_list ))  {

ts <- selected_list[[n]]
 
fig_A <- ts[, 1] %>%  autoplot() +
                                     ggtitle (colnames(selected_list[[1]])[1]) +
                                      geom_hline(yintercept=0, linetype="dashed", color = "black") +
   scale_x_continuous(expand = c(0, 0), limits = c(0, NA), breaks = scales::pretty_breaks(n = 8)) + 
                                      theme(plot.title = element_text(hjust = 0.5), 
                                            plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
                                            panel.grid.major = element_blank(), 
                                            panel.grid.minor = element_blank(),
                                            axis.text=element_text(size=12, face="bold")) #for centering the text

fig_B <- ts[, 2] %>%  autoplot() +
                                     ggtitle (colnames(selected_list[[1]])[2]) +
                                      geom_hline(yintercept=0, linetype="dashed", color = "black") +
   scale_x_continuous(expand = c(0, 0), limits = c(0, NA), breaks = scales::pretty_breaks(n = 8)) + 
                                      theme(plot.title = element_text(hjust = 0.5), 
                                            plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
                                            panel.grid.major = element_blank(), 
                                            panel.grid.minor = element_blank(),
                                            axis.text=element_text(size=12, face="bold")) #for centering the text

fig_C <- ts[, 3] %>%  autoplot() +
                                     ggtitle (colnames(selected_list[[1]])[3]) +
                                      geom_hline(yintercept=0, linetype="dashed", color = "black") +
   scale_x_continuous(expand = c(0, 0), limits = c(0, NA), breaks = scales::pretty_breaks(n = 8)) + 
                                      theme(plot.title = element_text(hjust = 0.5), 
                                            plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
                                            panel.grid.major = element_blank(), 
                                            panel.grid.minor = element_blank(),
                                            axis.text=element_text(size=12, face="bold")) #for centering the text

plot_list.2[[n]]<- ggpubr::ggarrange(fig_A,fig_B,fig_C,
                  labels = LETTERS[1:3],
                  ncol = 2, nrow = 2,  legend="none")

}

 names(plot_list.2)<- names(length_list)
```

# cross-correlation with sum neurite length
# cross-correlation with soma actin
```{r}
ccf_sum_all_list<-list(length=ccf_sum_list, somaActin=ccf_sumSomaActin_list)
```

```{r  message = FALSE, warning=FALSE, fig.width=10}
selected_ccf <-ccf_sum_all_list$somaActin$sum_tip_actin %>% as.data.table()

selected_ccf$lag <- selected_ccf$lag/freq

stacked_ccf_mat <- selected_ccf %>% gather(Cell, ccf, -lag) # select which CCF of measurements to show
df_summary <- stacked_ccf_mat %>% group_by(lag) %>% rstatix::get_summary_stats(ccf)
```

```{r}
ggsave(
  "somaActin_SumTipactin.pdf",
  plot = last_plot(),
  device = "pdf")
```

```{r}
Path_1 <- dirname(rstudioapi::getSourceEditorContext()$path)
saved_path <- paste(Path_1, "df_list_comb.RData",  sep = "/")
save.image(file=saved_path)
```
