---
title: "R Notebook"
output: html_notebook
---

<!--
# Tau-positive axon polarization level vs neurite length (blebbistatin / cytoD)

## What this file does

Reads in-line per-neuron Tau-intensity and neurite-length tables for WT and
Arp3 KO neurons treated with DMSO or 20 uM blebbistatin. Neurons with at
least one neurite longer than 70 um are scored for asymmetric Tau
distribution (a polarization-level proxy). Per-neuron polarization scores
are pooled by genotype + treatment and plotted as violin + boxplot. Five
PDFs are written (`ggsave` at lines 106, 249, 304, 365, 437).

## Manuscript panel

**ED Fig 12h**: "Polarization level quantified by asymmetric Tau distribution.
Neurons with neurite length > 70 um were analyzed. N = 73 DMSO-treated WT,
180 blebbistatin treated WT, and 197 blebbistatin treated Arp3 KO neurons
from 3-4 independent experiments. ***p < 0.001; *p < 0.05 by Two-tailed
Kruskal-Wallis test followed by Dunn's post-hoc test with Bonferroni
correction. Exact P values are shown."

Source Data: `ED Fig 12h`.

## Outputs

- Five PDFs (`ggsave` at lines 106, 249, 304, 365, 437).

## Original G:\ source

G:/Figure S12_blebbistatin_CytoD/Fig. S12h/R_scripts/NeuriteLength_Tau.Rmd
-->

```{r message=FALSE, warning=FALSE}

library(pracma)
#library(peakPick)
library(quantmod)
library(ggpmisc)
library(tidyverse)
library(magrittr)
#library(ggplot2)
library(ggfortify)
library(forecast)
library(tsibble)
library(broom)
library(tidyquant)
library(tsbox)
#library("pheatmap")
#library("RColorBrewer")
library(ggpmisc)
library(ggfortify)
#library(viridis) # Default Color Maps from 'matplotlib' 
library(hrbrthemes)
library(circular)
library("gridExtra") 
library("ggExtra") 
library("ggpubr")
library(data.table)
```

```{r Load RData, message=FALSE, warning=FALSE}
dir  <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(dir)
saved_path <- paste(dir, "NeuriteLength_Tau.RData",  sep = "/")
load(file=saved_path)
#save.image(file=saved_path)
```


```{r}
Blebb_Tau_quant <- Blebb_Tau_quant %>% as.data.table()
Blebb_Tau_quant_2 <- Blebb_Tau_quant[`Neurite length` != 0]

Blebb_Tau_quant_Cell <- Blebb_Tau_quant_2 %>% 
group_by(Cell, Batch, Blebb, WT_KO) %>% rstatix::get_summary_stats(`Tau IntDen`) %>% as.data.table()

```


```{r}
Blebb_Tau_quant_Cell$max_2nd <- 0

for (n in 1:nrow(Blebb_Tau_quant_Cell))
{
  
  if(Blebb_Tau_quant_Cell$n[n] >= 2){
    
    x <- Blebb_Tau_quant_2[Cell == Blebb_Tau_quant_Cell$Cell[n] &  
                             Batch == Blebb_Tau_quant_Cell$Batch[n] & 
                             Blebb ==  Blebb_Tau_quant_Cell$Blebb[n] & 
                             WT_KO == Blebb_Tau_quant_Cell$WT_KO[n] ]$`Tau IntDen` %>% sort()
    
    Blebb_Tau_quant_Cell$max_2nd [n] <- x[length(x)-1]

    } else { 
      
     Blebb_Tau_quant_Cell$max_2nd [n] <- 0 
      
              }
}

Blebb_Tau_quant_Cell$max_z <- (Blebb_Tau_quant_Cell$max - Blebb_Tau_quant_Cell$mean)/Blebb_Tau_quant_Cell$sd
Blebb_Tau_quant_Cell$max_2nd_z <- (Blebb_Tau_quant_Cell$max_2nd - Blebb_Tau_quant_Cell$mean)/Blebb_Tau_quant_Cell$sd

Blebb_Tau_quant_Cell$mean_2 <- ((Blebb_Tau_quant_Cell$mean*Blebb_Tau_quant_Cell$n)-Blebb_Tau_quant_Cell$max)/(Blebb_Tau_quant_Cell$n-1)

Blebb_Tau_quant_Cell$max_z2 <- Blebb_Tau_quant_Cell$max/Blebb_Tau_quant_Cell$mean_2
Blebb_Tau_quant_Cell$max_2nd_z2 <- Blebb_Tau_quant_Cell$max_2nd/Blebb_Tau_quant_Cell$mean_2

```

# Check distribution of max and max_2nd (Tau)
```{r}
# Blebb_Tau_quant_Cell[WT_KO == "WT" & Blebb == 20] %>% 
#  pivot_longer(cols = c("max_z2", "max_2nd_z2"), values_to = "Tau_intensity", names_to = "Tau") %>%

Blebb_Tau_quant_Cell %>% 
  
ggplot( aes(x=WT_KO, y=max_2nd_z2, fill =factor (Blebb)) )+ 
   # geom_hline(yintercept = 1)+
  scale_fill_brewer(palette="Dark2") + 
  geom_hline(yintercept=2, linetype=3)+

  theme_classic()+
  geom_violin(aes(), trim=F, kernel ="gaussian", adjust=1, position=dodge)  +
 geom_boxplot(aes(), width=0.5, position=dodge)+
scale_y_continuous(expand = c(0, 0), limits = c(0, 5), breaks = scales::pretty_breaks(n = 7))

Blebb_Tau_quant_Cell$group_2 <- paste (Blebb_Tau_quant_Cell$WT_KO, Blebb_Tau_quant_Cell$Blebb, sep = "_")
Blebb_Tau_quant_Cell$group_2 %>% table()
```

```{r}
ggsave(
  "Tau_zScoreFold.pdf",
  plot = last_plot(),
  device = "pdf")
```

```{r}
Blebb_Tau_quant_Cell %>% write_csv("Blebb_Tau_quant_Cell_ExtFig.12i.csv")
```

```{r}
Blebb_Tau_quant_Cell$Tau_pos <- "No"

Blebb_Tau_quant_Cell[max_2nd_z2 > 2]$Tau_pos <- "Yes"

freq_df <- Blebb_Tau_quant_Cell%>% group_by(WT_KO, Blebb, Batch)%>%
    count(Tau_pos) %>% 
    mutate(percent = n/sum(n)) 

freq_df %>% group_by(WT_KO, Blebb, Tau_pos) %>% mutate(avg = n/sum(n)) 

```


```{r}
Tau_length_2 <- Tau_length[NeuriteNum >1]
Tau_length_2$TauDiff <- Tau_length_2$Tau_1 - Tau_length_2$Tau_2
Tau_length_2$Tau_1_Int <- Tau_length_2$Tau_1 * Tau_length_2$Length_1
Tau_length_2$Tau_2_Int <- Tau_length_2$Tau_2 * Tau_length_2$Length_2

```

# Check density distribution of Tau_1 and Tau_2
```{r}
df <- Tau_length_2[Batch == "Batch_5" & Var =="WT_0"] %>% 
  pivot_longer(cols = c("Tau_1_Int", "Tau_2_Int"), values_to = "Tau_intensity", names_to = "Tau") 

df %>% 
ggplot(aes(x=Tau, y=Tau_intensity)) + 
   # geom_hline(yintercept = 1)+
  scale_fill_brewer(palette="Dark2") + 
  theme_classic()+
  geom_violin(aes(), trim=F, kernel ="gaussian", adjust=1, position=dodge)  +
  geom_boxplot(aes(), width=0.1, position=dodge)+ 
 scale_y_continuous(expand = c(0, 0), limits = c(-500, 70000), breaks = scales::pretty_breaks(n = 7))


# histogram
df %>% 
  ggplot(aes(x=Tau_intensity, fill=Tau, color=Tau)) + 
  geom_histogram(binwidth=10, position="identity", alpha=0.5)

# cumulative histogram
df %>% 
ggplot(aes(x=Tau_intensity, 
           fill=Tau, 
           color=Tau,  
           y = ave(after_stat(density), group, FUN = cumsum)*after_stat(width),
                   group = Tau))+
  geom_histogram(aes(), binwidth=5, position="identity", alpha=0.5)+
       #stat_bin(aes(y=after_stat(cumsum(count / sum(count)))),geom="line",color="green") +
  geom_line(stat = "bin") +
  scale_y_continuous(labels = scales::percent_format())
```

# Calculate the 1st and the 3rd quantile of Tau_1 and Tau_2 in each batch of exp
```{r}
Tau_quantile <- Tau_length_2[Var =="WT_0"] %>% 
  pivot_longer(cols = c("Tau_1", "Tau_2"), values_to = "Tau_density", names_to = "Tau") %>% 
group_by(Batch, Tau) %>% rstatix::get_summary_stats(Tau_density)

Tau_length_2[Batch == "Batch_5" & Var =="WT_0"]$Tau_2 %>% quantile( 0.95)

# Tau_1
# Batch_3: q1 = 78.363; 
# Batch_4: q1 = 136.414; 
# Batch_5: q1 = 143.601; 

# Tau_2
# Batch_3: q3 = 108.623; 
# Batch_4: q3 = 182.333; 
# Batch_5: q3 = 208.173; 

# Threshold to be considered as axon
# Batch_3: q3 = 108.623; 
# Batch_4: q3 = 182.333; 
# Batch_5: q3 = 208.173; 
```

```{r}
Tau_length_2[Batch == "Batch_3"] %>% mutate(Var = factor(Var, levels=c("WT_0", "WT_20", "KO_0", "KO_20"))) %>%
ggplot(aes(x=Var, y=Tau_1)) + 
   # geom_hline(yintercept = 1)+
  scale_fill_brewer(palette="Dark2") + 
  theme_classic()+
  geom_violin(aes(), trim=F, kernel ="gaussian", adjust=1, position=dodge)  +
  geom_boxplot(aes(group=interaction(Blebb,WT_KO)), width=0.1, position=dodge)
  #scale_y_continuous(expand = c(0, 0), limits = c(0 ,25), breaks = scales::pretty_breaks(n = 5)) 
  #stat_compare_means(comparisons = my_comparisons, label.y = c(5, 6, 7, 8, 9)) # Add pairwise comparisons p-value
  #stat_compare_means(label.y = 22)     # Add global p-value 
```


# 2nd longest neurtie length
# Remove outliers
```{r}
Tau_length <- Tau_length %>% as.data.table()
measurement <- "Length_2"

df <-Tau_length [WT_KO=="WT" &  Blebb == 0]
outliers <-  boxplot(df[, ..measurement], plot=FALSE)$out
Tau_length_WT_0 <- df[-which(Length_2 %in% outliers)]

df <-Tau_length [WT_KO=="WT" &  Blebb == 20]
outliers <-  boxplot(df[, ..measurement], plot=FALSE)$out
Tau_length_WT_20<- df[-which(Length_2 %in% outliers)]

df <-Tau_length [WT_KO=="KO" &  Blebb == 0]
outliers <-  boxplot(df[, ..measurement], plot=FALSE)$out
Tau_length_KO_0 <- df[-which(Length_2 %in% outliers)]

df <-Tau_length [WT_KO=="KO" &  Blebb == 20]
outliers <-  boxplot(df[, ..measurement], plot=FALSE)$out
Tau_length_KO_20<- df[-which(Length_2 %in% outliers)]

Tau_length_OL <- rbind(Tau_length_WT_0, Tau_length_WT_20, Tau_length_KO_0, Tau_length_KO_20)
Tau_length_OL$Blebb <- as.factor(Tau_length_OL$Blebb)

Tau_length_OL$Var<- paste(Tau_length_OL$WT_KO, Tau_length_OL$Blebb, sep = "_" )
```

# Length distribution of the 2nd longest neurtie length
```{r message = FALSE, warning=FALSE, fig.width=5}
a <- Tau_length_OL %>% ggplot(aes(x = Length_2))

a + geom_density(aes(color = Var), alpha = 0.4) +
   theme_classic()+
  #geom_vline(aes(xintercept = mean(Length_2)), linetype = "dashed", size = 0.6) +
  #geom_histogram(aes(y = ..density.., color = Var, fill=Var),  alpha = 0.4, position = "identity")+
  scale_x_continuous(expand = c(0, 0), limits = c(0, 150), breaks = scales::pretty_breaks(n = 10)) 
```

```{r}
ggsave(
  "Tau_1.eps",
  plot = last_plot(),
  device = "eps")
```

```{r}
Tau_length_OL %>% write_csv("Tau_length_OL.csv")
```


# Polarization index 3 (highest Tau/2nd highest Tau)
# Remove outliers
```{r}
Tau_length_WT_0<- Tau_length %>% subset(WT_KO=="WT") %>% subset(Blebb == 0)
outliers_WT_0 <-  boxplot(Tau_length_WT_0$Polarization_3, plot=FALSE)$out

Tau_length_WT_20<- Tau_length %>% subset(WT_KO=="WT") %>% subset(Blebb == 20)
outliers_WT_20 <-  boxplot(Tau_length_WT_20$Polarization_3, plot=FALSE)$out

Tau_length_KO_0<- Tau_length %>% subset(WT_KO=="KO") %>% subset(Blebb == 0)
outliers_KO_0 <-  boxplot(Tau_length_KO_0$Polarization_3, plot=FALSE)$out

Tau_length_KO_20<- Tau_length %>% subset(WT_KO=="KO") %>% subset(Blebb == 20)
outliers_KO_20 <-  boxplot(Tau_length_KO_20$Polarization_3, plot=FALSE)$out

Tau_length_WT_0 <- Tau_length_WT_0[-which(Tau_length_WT_0$Polarization_3 %in% outliers_WT_0),]
Tau_length_WT_20 <- Tau_length_WT_20[-which(Tau_length_WT_20$Polarization_3 %in% outliers_WT_20),]
Tau_length_KO_0 <- Tau_length_KO_0[-which(Tau_length_KO_0$Polarization_3 %in% outliers_KO_0),]
Tau_length_KO_20 <- Tau_length_KO_20[-which(Tau_length_KO_20$Polarization_3 %in% outliers_KO_20),]

Tau_length_OL <- rbind(Tau_length_WT_0, Tau_length_WT_20, Tau_length_KO_0, Tau_length_KO_20)
Tau_length_OL$Blebb <- as.factor(Tau_length_OL$Blebb)

Tau_length_OL$Var<- paste(Tau_length_OL$WT_KO, Tau_length_OL$Blebb, sep = "_" )
```


```{r message = FALSE, warning=FALSE, fig.width=5}
ylim1 = boxplot.stats(Tau_length_OL$Polarization_3)$stats[c(1, 5)] #+ coord_cartesian(ylim = ylim1*1.05) 
my_comparisons <- list( c("KO_0", "KO_20"), c("WT_0", "WT_20"),c("KO_0", "WT_0"), c("KO_20", "WT_20"),  c("KO_20", "WT_0") )
dodge <- position_dodge(width = 1)

Tau_length_OL %>% mutate(Var = factor(Var, levels=c("WT_0", "WT_20", "KO_0", "KO_20"))) %>%
ggplot(aes(x=Var, y=Polarization_3)) + 
    geom_hline(yintercept = 1)+
  scale_fill_brewer(palette="Dark2") + 
  theme_classic()+
  geom_violin(aes(fill = Blebb), trim=F, kernel ="gaussian", adjust=1, position=dodge) +
  geom_boxplot(aes(group=interaction(Blebb,WT_KO)), width=0.1, position=dodge)+
  #scale_y_continuous(expand = c(0, 0), limits = c(0 ,25), breaks = scales::pretty_breaks(n = 5)) 
  stat_compare_means(comparisons = my_comparisons, label.y = c(5, 6, 7, 8, 9)) # Add pairwise comparisons p-value
  #stat_compare_means(label.y = 22)     # Add global p-value 
```
```{r}
ggsave(
  "Polarization_3.eps",
  plot = last_plot(),
  device = "eps")
```

```{r}
Tau_length_OL %>% write_csv("Tau_length_OL.csv")
```


# Polarization index 2 (longest neurite length/2nd longest neurite length)
# Remove outliers
```{r}
Tau_length_WT_0<- Tau_length %>% subset(WT_KO=="WT") %>% subset(Blebb == 0)
outliers_WT_0 <-  boxplot(Tau_length_WT_0$Polarization_2, plot=FALSE)$out

Tau_length_WT_20<- Tau_length %>% subset(WT_KO=="WT") %>% subset(Blebb == 20)
outliers_WT_20 <-  boxplot(Tau_length_WT_20$Polarization_2, plot=FALSE)$out

Tau_length_KO_0<- Tau_length %>% subset(WT_KO=="KO") %>% subset(Blebb == 0)
outliers_KO_0 <-  boxplot(Tau_length_KO_0$Polarization_2, plot=FALSE)$out

Tau_length_KO_20<- Tau_length %>% subset(WT_KO=="KO") %>% subset(Blebb == 20)
outliers_KO_20 <-  boxplot(Tau_length_KO_20$Polarization_2, plot=FALSE)$out

Tau_length_WT_0 <- Tau_length_WT_0[-which(Tau_length_WT_0$Polarization_2 %in% outliers_WT_0),]
Tau_length_WT_20 <- Tau_length_WT_20[-which(Tau_length_WT_20$Polarization_2 %in% outliers_WT_20),]
Tau_length_KO_0 <- Tau_length_KO_0[-which(Tau_length_KO_0$Polarization_2 %in% outliers_KO_0),]
Tau_length_KO_20 <- Tau_length_KO_20[-which(Tau_length_KO_20$Polarization_2 %in% outliers_KO_20),]

Tau_length_OL <- rbind(Tau_length_WT_0, Tau_length_WT_20, Tau_length_KO_0, Tau_length_KO_20)
Tau_length_OL$Blebb <- as.factor(Tau_length_OL$Blebb)

Tau_length_OL$Var<- paste(Tau_length_OL$WT_KO, Tau_length_OL$Blebb, sep = "_" )
```

# generate pairwise comparison list
```{r}
df <-Tau_length_OL$Var %>% table() %>% as.data.frame()
my_comparisons <- df$. %>% as.character() %>% combn(2, simplify = FALSE) 
```

```{r message = FALSE, warning=FALSE, fig.width=5}
ylim1 = boxplot.stats(Tau_length_OL$Polarization_2)$stats[c(1, 5)] #+ coord_cartesian(ylim = ylim1*1.05) 
#my_comparisons <- list( c("KO_0", "KO_20"), c("WT_0", "WT_20"),c("KO_0", "WT_0"), c("KO_20", "WT_20"),  c("KO_20", "WT_0") )
dodge <- position_dodge(width = 1)

Tau_length_OL %>% mutate(Var = factor(Var, levels=c("WT_0", "WT_20", "KO_0", "KO_20"))) %>%
ggplot(aes(x=Var, y=Polarization_2)) + 
  geom_hline(yintercept = 1)+
  scale_fill_brewer(palette="Dark2") + 
  theme_classic()+
  geom_violin(aes(fill = Blebb), trim=F, kernel ="gaussian", adjust=1, position=dodge) +
  geom_boxplot(aes(group=interaction(Blebb,WT_KO)), width=0.1, position=dodge)+
  #scale_y_continuous(expand = c(0, 0), limits = c(0 ,25), breaks = scales::pretty_breaks(n = 5)) 
  stat_compare_means(comparisons = my_comparisons, label.y = c(6, 7, 8, 9, 10)) # Add pairwise comparisons p-value
  #stat_compare_means(label.y = 22)     # Add global p-value 
```

```{r}
ggsave(
  "Polarization_2.eps",
  plot = last_plot(),
  device = "eps")
```

```{r}
Tau_length$Var<- paste(Tau_length$WT_KO, Tau_length$Blebb, sep = "_" )
Tau_length_polarized <- Tau_length [Length_1 >70]
```

# Remove outliers
```{r}

measurement <- "Polarization_3"

df <-Tau_length_polarized [WT_KO=="WT" &  Blebb == 0]
outliers <-  boxplot(df[, ..measurement], plot=FALSE)$out
Tau_length_WT_0 <- df[-which(Polarization_3 %in% outliers)]

df <-Tau_length_polarized [WT_KO=="WT" &  Blebb == 20]
outliers <-  boxplot(df[, ..measurement], plot=FALSE)$out
Tau_length_WT_20<- df[-which(Polarization_3 %in% outliers)]

df <-Tau_length_polarized [WT_KO=="KO" &  Blebb == 0]
#outliers <-  boxplot(df[, ..measurement], plot=FALSE)$out
Tau_length_KO_0 <- df#[-which(Polarization_3 %in% outliers)]

df <-Tau_length_polarized [WT_KO=="KO" &  Blebb == 20]
outliers <-  boxplot(df[, ..measurement], plot=FALSE)$out
Tau_length_KO_20<- df[-which(Polarization_3 %in% outliers)]

df_1 <- rbind(Tau_length_WT_0, Tau_length_WT_20, Tau_length_KO_20)
# Tau_length_OL$Blebb <- as.factor(Tau_length_OL$Blebb)
```

# generate pairwise comparison list
```{r}
df <-df_1$Var %>% table() %>% as.data.frame()
my_comparisons <- df$. %>% as.character() %>% combn(2, simplify = FALSE) 
```

```{r message = FALSE, warning=FALSE, fig.width=8}
library(easyGgplot2)
#my_comparisons <- list( c("WT_0", "WT_20"), c("KO_20", "WT_20"),  c("KO_20", "WT_0") )

df_1 %>% 
  mutate(Var = fct_relevel(Var, "WT_0", "WT_20", "KO_0", "KO_20")) %>% 
  ggplot2.violinplot(yName="Polarization_3", xName='Var', groupName='Var', 
                     addDot=TRUE, dotSize=1,dotPosition="jitter", jitter=0.2, 
                     trim=F,  kernel ="gaussian", adjust=0.5)+
    
  geom_boxplot(width=0.1, fill="white")  +
 # rotate_x_text(angle = 45)+  
  scale_y_continuous(expand = c(0, 0), limits = c(0 ,6), breaks = scales::pretty_breaks(n = 5)) +
  stat_compare_means(comparisons = my_comparisons, paired=F, method="wilcox.test", label.y = c(4.6, 4.8, 5, 5.2, 5.4, 5.6), label = "{p.adj.format}{p.adj.signif}") +
  
  geom_hline(yintercept = mean(df_1$Polarization_3, na.rm = T), linetype = 2)+ # Add horizontal line at base mean
  #stat_compare_means(label.y = 0.45) +        # Add global annova p-value
  #stat_compare_means(label = "p.signif", method = "wilcox.test", ref.group = ".all.") +
  
  theme_ipsum(grid = F, ticks = T) +
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5), 
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        panel.grid.major = element_blank(), S
        panel.grid.minor = element_blank(),
        legend.position = "right",
        axis.text=element_text(size=12, face="bold"))
   
```
```{r}
ggsave(
  "Polarization_3_longerThan70.eps",
  plot = last_plot(),
  device = "eps")
```

```{r, message=FALSE, warning=FALSE}
dir  <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(dir)
saved_path <- paste(dir, "NeuriteLength_Tau.RData",  sep = "/")
#load(file=saved_path)

save.image(file=saved_path)
```
